### 2014-08-04 asymptotic.tex
first draft

The impact of passing through resonance on the evolution can be modelled analytically using asymptotic expansions~\cite{Gair2012}. Solutions for the motion are constructed far away from resonance and these are matched to a transition region in the vicinity of resonance~\cite{Kevorkian1971,Bosley1992}. By comparing the matched solution, which incorporates the effects of resonance, with the results of an adiabatic evolution, it is possible to estimate the discrepancy in the orbital parameters. This determines the difference in the orbital phase between the two approaches. If this error is sufficiently small, then it is safe to ignore the effects of the resonance; however, only a small difference is needed to impact the subsequent waveform, since the error accumulates over the subsequent observation of $\sim \order{\eta^{-1}}$ cycles~\cite{Flanagan2012}. We derive formulae which can be used to calculate the discrepancy in the orbital parameters.

The following derivation is based upon the analysis of Kevorkian~\cite{Kevorkian1987}.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} Small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.

We model the system using action--angle variables. We are only concerned with the $r$ and $\theta$ motions, so we have a $2$-dimensional system. We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992}. This becomes one of the new angle variables, the other variable $q'$ can be either $q_r$ or $q_\theta$ (as, on resonance, varying one necessarily changes the other). We use $J$ as the conjugate action variable to $q$ and $\omega = n_r \omega_r - n_\theta \omega_\theta$ as its frequency. Similarly, we use $J'$ as the action variable conjugate to $q'$. The system evolves through resonance slowly, on an evolution time-scale, so we parameterize it in terms of a slow time parameter
\begin{equation}
\widetilde{\lambda} = \eta\lambda.
\end{equation}
The orbits of $q'$ proceed with the fast time $\lambda$; since this is much more rapid than the evolution we are interested in, it is safe to average over it. We are not interested in the fine-grained fast oscillations caused by changes in $q'$. For this analysis we consider the reduced problem of evolving $q$ and $J$.

At resonance $\widetilde{\lambda} = \widetilde{\lambda}_\star$ and $\omega\left(\widetilde{\lambda}_\star\right) = 0$. We assume that the frequency has a simple zero and can be expanded as
\begin{equation}
\omega\left(\widetilde{\lambda}\right) = \varpi_1\left(\widetilde{\lambda} - \widetilde{\lambda}_\star\right) + \varpi_2\left(\widetilde{\lambda} - \widetilde{\lambda}_\star\right)^2 + \ldots
\label{eq:omega-series}
\end{equation}
The frequency is actually a function of the angle variables, but since these evolve with $\widetilde{\lambda}$ it is safe to write it as a function of the slow time.\footnote{In effect we are defining $\omega\left(\tilde{\lambda}\right) \equiv \omega\left[J\left(\tilde{\lambda}\right)\right]$.}

Using the slow time, the equations of motion (\ref{eq:Mino-E-o-M}) become
\begin{subequations}
\begin{align}
\diff{q}{\widetilde{\lambda}} = {} & \dfrac{\omega(J)}{\eta} + \sum_s g_s^{(1)}(J)\exp(is q)  + \order{\eta}, \\
\diff{J}{\widetilde{\lambda}} = {} & \sum_s G_s^{(1)}(J)\exp(is q) + \order{\eta},
\end{align}
\end{subequations}
where we have rewritten the forcing terms as Fourier series and adapted the forcing functions to those appropriate for $q$ and $J$. We shall solve these before resonance and then match to solutions in the transition regime about resonance.

\subsection{Solution before resonance}\label{sec:before-res}

To find a solution away from the resonance, we decompose the problem to be a function of two time-scales~\cite{Kevorkian1971}. We use the slow time $\widetilde{\lambda}$ and, as a proxy for the fast time,
\begin{equation}
\Psi = \intd{0}{\lambda}{\omega(\eta\tau)}{\tau} = \recip{\eta}\intd{0}{\tilde{\lambda}}{\omega(\widetilde{\tau})}{\widetilde{\tau}}.
\end{equation}
From this
\begin{equation}
\omega = \diff{\Psi}{\lambda}.
\end{equation}
In terms of these two variables, we can build ansatz solutions
\begin{subequations}
\begin{align}
\label{eq:q-series}
q(\lambda;\,\eta) = {} & \Psi + q_0\left(\widetilde{\lambda}\right) + \eta q_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}, \\
J(\lambda;\,\eta) = {} & J_0\left(\widetilde{\lambda}\right) + \eta J_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}.
\label{eq:J-series}
\end{align}
\end{subequations}
We can also write a series expansion for the frequency, since it is affected by the self-force too,
\begin{equation}
\omega(\lambda;\,\eta) = \omega_0\left(\widetilde{\lambda}\right) + \eta \omega_1\left(\widetilde{\lambda}\right) + \order{\eta^2}.
\end{equation}
In the limit of $\eta \rightarrow 0$ we are left with a constant frequency $\omega_0(0)$. The higher order terms are identified below by matching terms in the series expansions of the equations of motion. Taking the two time-scales as independent, we may write the time derivative to $\order{\eta}$ as
\begin{equation}
\diffop{\lambda} = \omega_0\partialdiffop{\Psi} + \eta\omega_1\partialdiffop{\Psi} + \eta\partialdiffop{\widetilde{\lambda}}.
\end{equation}

Using the two time-scale decomposition to replace the time derivatives in the equations of motion, and substituting in the ansatz expansions gives, to first order,
\begin{widetext}
\begin{subequations}
\begin{align}
\label{eq:q-1}
\omega_0 + \eta\omega_1 + \eta\partialdiff{q_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{q_1}{\Psi} = {} & \omega(J_0) + \eta\diff{\omega}{J}J_1 + \eta \sum_s g_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right], \\
\eta\partialdiff{J_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{J_1}{\Psi} = {} & \eta \sum_s G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right].
\label{eq:J-1}
\end{align}
\end{subequations}
\end{widetext}
Averaging \eqnref{J-1} over $\Psi$ gives\footnote{The ansatz is constructed such that $J_0 \equiv \langle J_0\rangle_\Psi$ and $q_0 \equiv \langle q_0\rangle_\Psi$.}
\begin{equation}
\partialdiff{J_0}{\widetilde{\lambda}} = G_0^{(1)}(J_0).
\label{eq:J-ad}
\end{equation}
This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest order piece of) the adiabatic solution~\cite{Hinderer2008}. If we similarly average \eqnref{q-1} we find
\begin{equation}
\omega_0 + \eta\omega_1 +\eta\partialdiff{q_0}{\widetilde{\lambda}} = \omega(J_0) + \eta\partialdiff{\omega}{J}\langle J_1\rangle_\Psi + \eta g_0^{(1)}(J_0).
\end{equation}
From this we can identify the terms that originate from the frequency and, matching by order in $\eta$, obtain
\begin{subequations}
\begin{align}
\omega_0 = {} & \omega(J_0), \\
\omega_1 = {} & \partialdiff{\omega}{J}\langle J_1\rangle_\Psi.
\end{align}
\end{subequations}
This leaves
\begin{align}
\partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0), \\
q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau}.
\label{eq:q-0-sol}
\end{align}
We now have expressions for the lowest order terms in the expansions.

Subtracting the $s = 0$ components from \eqnref{J-1} leaves
\begin{equation}
\omega_0\partialdiff{J_1}{\Psi} = \sum_{s\,\neq\,0} G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right].
\end{equation}
This can be solved to give
\begin{equation}
J_1 = \langle J_1\rangle_\Psi + \recip{\omega_0}\sum_{s\,\neq\,0} \dfrac{G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right]}{is}.
\label{eq:J-1-sol}
\end{equation}
We can do the same for \eqnref{q-1} to obtain
\begin{equation}
q_1 = \langle q_1\rangle_\Psi + \recip{\omega_0}\sum_{s\,\neq\,0} \dfrac{g_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right]}{is}.
\label{eq:q-1-sol}
\end{equation}
To find the constants of integration, $\langle q_1\rangle_\Psi$ and $\langle J_1\rangle_\Psi$, it is necessary to extend the analysis to second order in $\eta$. This shows that $\langle J_1\rangle_\Psi$ is the first-order component of the adiabatic solution. We do not need explicit forms for later calculations, so we will not proceed further. We have successfully constructed the pre-resonance solution.

\subsection{Solution near resonance}\label{sec:interior-res}

Near to resonance, we consider an interior layer expansion~\cite{Kevorkian1971}. As explained in \secref{res-time}, evolution near resonance proceeds on a time-scale intermediate between the slow and fast times. We therefore introduce a rescaled time
\begin{equation}
\widehat{\lambda} = \dfrac{\widetilde{\lambda} - \widetilde{\lambda}_\star}{\eta^{1/2}} = \eta^{1/2}(\lambda - \lambda_\star).
\end{equation}
As for the before resonance solution, we can create a series expansion; however, now we expand in terms of $\eta^{1/2}$~\cite{Flanagan2012}
\begin{subequations}
\begin{align}
q\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{q}_0\left(\widehat{\lambda}\right) + \eta^{1/2} \widehat{q}_{1/2}\left(\widehat{\lambda}\right) + \order{\eta}, \\
J\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{J}_0 + \eta^{1/2} \widehat{J}_1\left(\widehat{\lambda}\right) + \order{\eta}.
\end{align}
\end{subequations}
The series expansion for the frequency, \eqnref{omega-series}, can be rewritten as
\begin{equation}
\omega\left(\widehat{\lambda}\right) = \eta^{1/2}\varpi_1\widehat{\lambda} + \eta\varpi_2\widehat{\lambda}^2 + \order{\eta^{3/2}}.
\label{eq:omega-hat}
\end{equation}
Proceeding to write the equations of motion in terms of the rescaled time gives
\begin{subequations}
\begin{align}
\diff{q}{\widehat{\lambda}} = {} & \varpi_1\widehat{\lambda} + \eta^{1/2}\varpi_2\widehat{\lambda}^2 \nonumber \\ 
 {} & + \left. \eta^{1/2}\sum_s g_s^{(1)}\left(\widehat{J}_0,\widetilde{\lambda}_\star\right)\exp(is \widehat{q}_0)  + \order{\eta}, \right. \\
\diff{J}{\widehat{\lambda}} = {} & \eta^{1/2}\sum_s G_s^{(1)}\left(\widehat{J}_0,\widetilde{\lambda}_\star\right)\exp(is \widehat{q}_0) + \order{\eta}.
\end{align}
\end{subequations}

From the equations of motion we can match terms by their order in $\eta^{1/2}$. At zeroth order we find
\begin{equation}
\widehat{J}_0 = \widehat{\varrho}_0
\end{equation}
is constant, and
\begin{equation}
\widehat{q}_0 = \widehat{\kappa}_0 + \dfrac{\varpi_1\widehat{\lambda}^2}{2}.
\end{equation}
Using these, we can build the next order terms
\begin{align}
\widehat{q}_{1/2} = {} & \widehat{\kappa}_{1/2} + \dfrac{\varpi_2\widehat{\lambda}^3}{3} + g_0^{(1)}(\widehat{\varrho}_0)\widehat{\lambda} \nonumber \\ 
 {} & + \sum_{s\,\neq\,0}g_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}, \\
\widehat{J}_{1/2} = {} & \widehat{\varrho}_{1/2} + G_0^{(1)}(\widehat{\varrho}_0)\widehat{\lambda} \nonumber \\
 {} & + \sum_{s\,\neq\,0}G_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}.
\end{align}
Both of these involve the complex Fresnel integral~\cite{Olver2010}, the details of which are given in the following section. We have now constructed the interior solution. % chapter 7

\subsection{The complex Fresnel integral}

The solution for the motion in the interior region near to resonance involves the integral
\begin{equation}
W\left(\widehat{\lambda}\right) = \intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}.
\end{equation}
The complex Fresnel integral is
\begin{equation}
\mathcal{Y}(z) = \intd{0}{z}{\exp\left(\dfrac{i\pi x^2}{2}\right)}{x} = \mathcal{C}(z) + i\mathcal{S}(z),
\end{equation}
where $\mathcal{C}(z)$ and $\mathcal{S}(z)$ are the cosine and sine Fresnel integrals~\cite{Olver2010}, and hence % [7.2.7, 7.2.8]
\begin{equation}
W\left(\widehat{\lambda}\right) = \sqrt{\dfrac{\pi}{s\varpi_1}}\mathcal{Y}\left(\sqrt{\dfrac{s\varpi_1}{\pi}}\widehat{\lambda}\right).
\end{equation}

We shall be interested in the asymptotic behaviour for $|\widehat{\lambda}| \rightarrow \infty$. The complex Fresnel integral has the limit~\cite{Olver2010} % [7.5.3, 7.5.4, 7.12.2, 7.12.3]
\begin{equation}
\lim_{|z|\,\rightarrow\,\infty} \mathcal{Y}(z) \sim \dfrac{\sgn z}{\sqrt{2}} \exp\left(\dfrac{i\pi}{4}\right) - \dfrac{i}{\pi z}\exp\left(-\dfrac{i\pi z^2}{2}\right),
\end{equation}
where 
\begin{equation}
\sgn z = \begin{cases}
1 & z > 0 \\
-1 & z < 0
\end{cases}\,.
\end{equation}
Hence,
\begin{align}
\lim_{|\widehat{\lambda}|\,\rightarrow\,\infty}W\left(\widehat{\lambda}\right) \sim {} & \dfrac{\sgn \widehat{\lambda}}{\sqrt{2}}\sqrt{\dfrac{\pi}{|s\varpi_1|}}\exp\left[\sgn(s\varpi_1)\dfrac{i\pi}{4}\right] \nonumber \\
  {} & + \recip{is\varpi_1}\exp\left(\dfrac{is \varpi_1 \widehat{\lambda}^2}{2}\right).
\label{eq:Fres-limit}
\end{align}

\subsection{Matching solutions}

To complete our solution we must match the pre-resonance solution of \secref{before-res} with the near resonance solution of \secref{interior-res}. This is achieved by rewriting the pre-resonance solution in terms of the rescaled time $\widehat{\lambda}$ and comparing this with the near resonance solution expanded in the limit of $\widehat{\lambda} \rightarrow -\infty$.

To rewrite the pre-resonance solution, we begin with the fast phase parameter
\begin{equation}
\Psi\left(\widehat{\lambda}\right) = \dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \eta^{1/2}\dfrac{\varpi_2\widehat{\lambda}^3}{3} + \order{\eta}.
\end{equation}
Using this together with equations (\ref{eq:q-0-sol}) and (\ref{eq:q-1-sol}) in \eqnref{q-series}, the angle variable is
\begin{widetext}
\begin{align}
q\left(\widehat{\lambda};\,\eta\right) = {} & \dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \kappa_\star + \eta^{1/2}\dfrac{\varpi_2\widehat{\lambda}^3}{3} + \eta^{1/2}g_0^{(1)}(J_\star)\widehat{\lambda} + \dfrac{\eta^{1/2}}{\varpi_1\widehat{\lambda}}\sum_{s\,\neq\,0}\recip{is}g_s^{(1)}(J_\star)\exp\left[is\left(\dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \kappa_\star\right)\right] + \order{\eta},
\end{align}
where we have defined $J_\star \equiv J_0\left(\widetilde{\lambda}_\star\right)$ and $\kappa_\star = \kappa_0 + \intd{0}{\tilde{\lambda}_\star}{g_0^{(1)}[J_0(\tau)]}{\tau}$, and used \eqnref{omega-hat} to substitute for $\omega$. The action variable is similarly determined by using equations (\ref{eq:J-ad}) and (\ref{eq:J-1-sol}) with \eqnref{J-series} to give
\begin{align}
J\left(\widehat{\lambda};\,\eta\right) = {} & J_\star + \eta^{1/2}G_0^{(1)}(J_\star)\widehat{\lambda} + \dfrac{\eta^{1/2}}{\varpi_1\widehat{\lambda}}\sum_{s\,\neq\,0}\recip{is}G_s^{(1)}(J_\star)\exp\left[is\left(\dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \kappa_\star\right)\right] + \order{\eta}.
\end{align}
\end{widetext}
We can now compare this to the near resonance expansion with the integral replaced by the limiting form given in \eqnref{Fres-limit}.

At zeroth order, we immediately obtain
\begin{align}
\widehat{\kappa}_0 = {} & \dfrac{\Psi_\star}{\eta} + \kappa_\star, \\
\widehat{\varrho}_0 = {} & J_\star.
\end{align}
These fix the integration constants. The more interesting result is now found by comparing the $\order{\eta^{1/2}}$ terms. Equating the angle variable expressions and cancelling terms gives
\begin{align}
%\widehat{\kappa}_{1/2} + \sum_{s\,\neq\,0}g_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\left\{\dfrac{\sgn \widehat{\lambda}}{\sqrt{2}}\sqrt{\dfrac{\pi}{|s\varpi_1}}\exp\left[\sgn(s\varpi_1)\dfrac{\pi}{4}\right] + \recip{is\varpi_1}\exp\left(\dfrac{is \varpi_1 \widehat{\lambda}^2}{2}\right)\right\} = {} & \recip{is\varpi_1\widehat{\lambda}}g_s^{(1)}(\widehat{\varrho}_0)\exp\left[is\left(\dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2}\right)\right] \nonumber \\
\widehat{\kappa}_{1/2} = {} & \sum_{s\,\neq\,0}g_s^{(1)}(\widehat{\varrho}_0)\sqrt{\dfrac{\pi}{2|s\varpi_1|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\varpi_1\right)\right].
\end{align}
Similarly, for the action variables
\begin{equation}
\widehat{\varrho}_{1/2} = \sum_{s\,\neq\,0}G_s^{(1)}(\widehat{\varrho}_0)\sqrt{\dfrac{\pi}{2|s\varpi_1|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\varpi_1\right)\right].
\label{eq:J-1/2}
\end{equation}
We now have a matched solution through resonance.

Having constructed the solution, we see that the lowest order evolution corresponds to the adiabatic solution; the deviations come in at the following order. When we switch from the pre-resonance solution to the post-resonance solution, there is a change in the sign of $\widehat{\lambda}$. Therefore, when matching the post-resonance solution $\widehat{\varrho}_{1/2}$ and $\widehat{\kappa}_{1/2}$ also change sign: there is a change of
\begin{align}
\Delta q = {} & 2 \eta^{1/2}\widehat{\kappa}_{1/2}, \\
\Delta J = {} & 2 \eta^{1/2}\widehat{\varrho}_{1/2}
\label{eq:jumps}
\end{align}
across the resonance~\cite{Kevorkian1987}. We are not particularly interested in the deviation in $J$, of greater concern is the change in the orbital parameters $\{E,L_z,Q\}$. Assuming that there is a smooth transformation that maps between $J$ and these, then, to lowest order, we can calculate the deviation relative to the adiabatic prescription by substituting the forcing functions $G^{(1)} \rightarrow G_a^{(1)}$, where $G_a^{(1)}$ describes the evolution of $\mathcal{I}_a$ through the effects of the self-force. This result is quoted by Flanagan and Hinderer~\cite{Flanagan2012}. The change in the orbital parameters is determined by the forcing functions, hence it is essential to have an accurate self-force model.

As a final step in understanding our result, we switch from Mino time to coordinate time. An appropriate redefinition of the forcing functions can be done by scaling by $\Gamma$, we define
\begin{equation}
F_a^{(1)} = \dfrac{G_a^{(1)}}{\Gamma},
\end{equation}
such that the equation of motion becomes
\begin{equation}
\left\langle \diff{\mathcal{I}_a}{t}\right\rangle_{q'} =  \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
\end{equation}
Here we have made the averaging over $q'$ explicit to show that the equation is only defined as an orbital average: not only does our asymptotic expansion average out oscillations over an orbit in $q'$, but in converting from $\lambda$ to $t$ we have used $\Gamma$ which is an orbital average.
From \eqnref{omega-series}, we recognise that
\begin{equation}
\varpi_1 = \partialdiff{\omega}{\widetilde{\lambda}} = \dfrac{\Gamma^2}{\eta}\left\langle\dot{\Omega}\right\rangle_{q'}.
\end{equation}
We have used the averaged form of $\dot{\Omega}(t)$ as this is appropriate. Using these to adapt equations (\ref{eq:J-1/2}) and (\ref{eq:jumps}), we obtain
\begin{align}
\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\sqrt{\dfrac{2\pi}{\left|s \langle\dot{\Omega}\rangle_{q'}\right|}} \nonumber \\*
 {} & \times \left.\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\dot{\Omega}\right)\right]\right. \\
 = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega} \right)\right],
\end{align}
using \eqnref{T-res-s} and representing the values on resonance of $E$, $L_z$ and $Q$ with $\boldsymbol{\mathcal{I}}_\star$.

### 2014-08-04 res_char.tex
first draft

\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,reprint,showpacs]{revtex4}
\usepackage{graphicx}
\usepackage{color}
%\usepackage{psfrag}
\usepackage{mathrsfs}
\usepackage{dcolumn}
\usepackage{hyperref}

\newcommand{\eqnref}[1]{(\ref{eq:#1})}
\newcommand{\figref}[1]{Fig.\ \ref{fig:#1}}
\newcommand{\Figref}[1]{Fig.\ \ref{fig:#1}}
\newcommand{\secref}[1]{Sec.\ \ref{sec:#1}}
\newcommand{\apref}[1]{Appendix \ref{sec:#1}}
\newcommand{\Tabref}[1]{Table \ref{tab:#1}}

\newcommand{\units}[1]{\ensuremath{~\mathrm{#1}}}

\newcommand{\sub}[1]{\ensuremath{_\text{#1}}}
\newcommand{\super}[1]{\ensuremath{^\text{#1}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\diff}[2]{\ensuremath{\dfrac{\dd {#1}}{\dd {#2}}}}
\newcommand{\linediff}[2]{\ensuremath{\dd {#1}/\dd {#2}}}
\newcommand{\diffop}[1]{\ensuremath{\dfrac{\dd}{\dd {#1}}}}
\newcommand{\difftwo}[2]{\ensuremath{\dfrac{\dd^2 {#1}}{\dd {#2}^2}}}
\newcommand{\linedifftwo}[2]{\ensuremath{\dd^2 {#1}/\dd {#2}^2}}
\newcommand{\difftwoop}[1]{\ensuremath{\dfrac{\dd^2}{\dd {#1}^2}}}
\newcommand{\diffn}[3]{\ensuremath{\dfrac{\dd^{#3} {#1}}{\dd {#2}^{#3}}}}
\newcommand{\linediffn}[3]{\ensuremath{\dd^{#3} {#1}/\dd {#2}^{#3}}}
\newcommand{\diffnop}[2]{\ensuremath{\dfrac{\dd^{#2}}{\dd {#1}^{#2}}}}
\newcommand{\partialdiff}[2]{\ensuremath{\dfrac{\partial {#1}}{\partial {#2}}}}
\newcommand{\linepartialdiff}[2]{\ensuremath{\partial {#1}/\partial {#2}}}
\newcommand{\partialdiffop}[1]{\ensuremath{\dfrac{\partial}{\partial {#1}}}}\newcommand{\recip}[1]{\ensuremath{\frac{1}{#1}}}
\newcommand{\grad}{\ensuremath{\boldsymbol{\nabla}}}
\newcommand{\intd}[4]{\ensuremath{\int_{#1}^{#2}{#3}\,\dd{#4}}}
\newcommand{\order}[1]{\ensuremath{\mathcal{O}({#1})}}

\newcommand{\overlap}[2]{\ensuremath{\left(#1\middle|#2\right)}}

\DeclareMathOperator{\sgn}{sgn}

\begin{document}

%\preprint{}

\title{Characterization of transient resonances in extreme-mass-ratio-inspirals}

\author{Robert H. Cole}
\email[]{rhc2@ast.cam.ac.uk}
\author{Christopher P.L. Berry}
\email[]{cplb2@ast.cam.ac.uk}
\author{Priscilla Ca\~{n}izares}
\email[]{pcm@ast.cam.ac.uk}
\author{Jonathan R. Gair}
\email[]{jgair@ast.cam.ac.uk}
\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}

\date{\today}

\begin{abstract}
The inspiral of stellar mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational field regime, via the emission of gravitational waves. In order to detect such signals and extract the desired information, it is necessary to possess a bank of accurate waveform templates. For computational efficiency, adiabatic templates are often used, but these neglect the effect of transient resonances, where the radial and poloidal fundamental frequencies become commensurate. During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object trajectory and ultimately the phase of the emitted radiation. An astrophysical population of detectable extreme-mass-ratio-inspirals is studied and it is found that a large proportion will encounter a low-order resonance in the later stages of inspiral. The resulting effect on SNR recovery is small due to the low eccentricity of the systems, but nevertheless at least 20\% of EMRIs may become undetectable if transient resonances aren't accounted for.
\end{abstract}

% 04.25.Nx 	Post-Newtonian approximation; perturbation theory; related approximations
% 04.30.-w 	Gravitational waves
% 04.70.-s 	Physics of black holes
% 98.62.Js 	Galactic nuclei (including black holes), circumnuclear matter, and bulges 
\pacs{04.25.Nx, 04.30.--w, 04.70.--s, 98.62.Js}

\maketitle

\section{Introduction}

In the prologue to his classic monograph, Chandrasekhar~\cite{Chandrasekhar1992} celebrates the simplicity of black holes (BHs). The Kerr solution is defined by just two parameters: mass and spin. Despite the baldness of the BH metrics, great intricacies manifest in their properties. This is made evident when a second body is introduced. The two-body problem in general relativity (GR) is well studied. It is of paramount importance for gravitational wave (GW) astronomy, where binary systems are the dominant source of radiation. Correctly modelling the dynamics of these systems is necessary to interpret and extract information from gravitational waveforms.

We have made progress in understanding the general relativistic two-body problem in recent years. Bodies of comparable mass can be studied using numerical relativity. Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}. These simulations shall allow us to understand BH--BH mergers. Stellar-mass BH mergers are targets for ground-based GW detectors, such as the currently mid-upgrade LIGO~\cite{Harry2010} and Virgo~\cite{Accadia2011}, and the in-construction Kagra~\cite{Kuroda2010}. Massive BH mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}. Systems of unequal masses are more challenging to evolve numerically as they complete a larger number of orbits and it is necessary to resolve two different scales. Calculations can instead be performed perturbatively. The paradigm unequal mass system has a stellar-mass BH orbiting a massive black hole, such as those expected to be found at the centre of most galaxies. These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors~\cite{Amaro-Seoane2007}. Despite being well-studied there still remain open questions.

In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}. In the test particle limit the smaller body follows an exact geodesic of the BH's spacetime. Including the effects of the smaller body's finite mass, the background spacetime is perturbed. The back-reaction from this deformation alters the small body's orbital trajectory, and can be modelled as a self-force that moves the body from its geodesic. The self-force is commonly divided into two pieces, dissipative and conservative. The former encapsulates the slow decay of the orbital energy and angular momentum, constants of the motion in the test particle limit, through radiation. The latter shifts the orbital phases inducing precession. The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase although this can accumulate over many orbits. Being able to accurately model the influence of the self-force shall allow us to create accurate waveform models.

Flanagan and Hinderer~\cite{Flanagan2012} highlighted a previously overlooked phenomenon that occurs in the general relativistic two-body problem, that of transient resonances. Geodesic orbits in GR have three associated frequencies: the radial frequency $\Omega_r$, the polar frequency $\Omega_\theta$ and the azimuthal frequency $\Omega_\phi$. The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}. % Section 10.6
In the weak-field limit these all tend towards the Keplerian frequency; in the strong-field $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly. For extreme-mass-ratio systems, the evolution time-scale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital time-scales. The evolution of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}. During this evolution, the frequencies may become commensurate. Resonances occur when the radial and polar frequencies are rational multiples of each other:
\begin{equation}
\nu = \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r},
\end{equation}
where $n_r$ and $n_\theta$ are integers (with no common factors). During resonance, terms in the self-force that usually average to zero can combine coherently, significantly impacting the orbital motion~\cite{Flanagan2012a}. Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.

Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
We shall consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$, and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}. We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$. These are some combination of the energy per unit mass $E$ and the axial angular momentum per unit mass $L_z$, which arise from isometries of the metric in $t$ and $\phi$, and the Carter constant per unit mass squared $Q$~\cite{Carter1968}, which is related to the separability of the equations of motion in $r$ and $\theta$. The system evolves following~\cite{Flanagan2012}
\begin{subequations}
\label{eq:Mino-E-o-M}
\begin{align}
\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta g_\alpha^{(1)}(q_r,q_\theta,\boldsymbol{J}) + \order{\eta^2}, \\
\diff{J_\alpha}{\lambda} = {} & \eta G_\alpha^{(1)}(q_r,q_\theta,\boldsymbol{J}) + \order{\eta^2},
\end{align}
\end{subequations}
where $\lambda$ is Mino time~\cite{Mino2003}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force. By working with $lambda$ instead of proper time $\tau$, the radial and polar motions decouple. At zeroth order in the mass ratio we recover the limit of purely geodesic motion: the integrals of the motion are actually constants and the angle variables evolve according to their associated frequencies $\omega_\alpha$.

The leading order dissipative correction to geodesic motion is calculated following the adiabatic prescription~\cite{Hinderer2008}: by dropping the forcing term $g_\alpha^{(1)}$ (and all higher order terms) and replacing the forcing term $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$ $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}. For most orbits this is sufficient, $G_\alpha^{(1)}$ is given by its average value plus a rapidly oscillating component~\cite{Arnold1988}. % Chapter 5, section 1
However, this averaging fails when the ratio of frequencies is the ratio of integers. In this case the trajectory does not ergodically fill the $2$-torus but instead traces out a $1$-dimensional subspace.\footnote{For illustrations, see Grossman, Levin and Perez-Giz~\cite{Grossman2012}.} There are then contributions to the self-force that no longer average out beyond $\langle G_A^{(1)}\rangle_{q_r,\,q_\theta}$. Intuitively, we expect that this effect is more important for ratios of small integers as when the integers are large, the orbit comes close to all points on the $2$-torus.

In this work we seek to characterise the importance of these resonances for the purposes of modelling EMRIs. The amplitude of expected signals will be below the level of noise in a space-based GW detector. However, systems remain in band for many hundreds of thousands of cycles and so may be detected using a matched filter, provided we have sufficiently accurate waveform templates. Ensuring the accuracy of EMRI templates requires calculating the impact that passing through a resonance has on the orbital evolution and discovering for which resonances this is significant.

We use geometric units with $G = c = 1$ throughout. We always use $M$ for the mass of the central massive BH and $a$ as the spin parameter. We also use the dimensionless spin $a_\ast \equiv a/M$; we take the convention that $0 \leq a_\ast < 1$.

\section{Problem formulation}

\subsection{Kerr geodesics}

The evolution of an extreme-mass-ratio ($\eta \ll 1$) system is slow. Instantaneously, the motion of the orbiting mass can be described as geodesic, with the integrals of the motion changing on time-scales of many orbital periods. We analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance. It is therefore necessary to develop an understanding of the Kerr geodesics.

The geodesic equations may be written as~\cite{Carter1968, Chandrasekhar1992} % section 62
\begin{subequations}
\begin{align}
\diff{t}{\lambda} = {} & a\left(L_z - aE\sin^2 \theta\right) + \frac{r^2 + a^2}{\Delta}\mathcal{T},\\
\diff{r}{\lambda} = {} & \pm \sqrt{V_r},\\
\diff{\theta}{\lambda} = {} & \pm \sqrt{V_\theta},\\
\diff{\phi}{\lambda} = {} & \frac{L_z}{\sin^2 \theta} - aE + \frac{a}{\Delta}\mathcal{T},
\end{align}
\end{subequations}
where $\Delta = r^2 - 2M r + a^2$; the signs of the $r$ and $\theta$ equations can be chosen independently, and we have introduced potentials
\begin{subequations}
\begin{align}
\mathcal{T} = {} & E\left(r^2 +a^2\right) - aL_z,\\
V_r = {} & \mathcal{T}^2 - \Delta\left[r^2 + \left(L_z -aE\right)^2 + Q\right],\\
V_\theta = {} & Q - \cos^2 \theta\left[a^2\left(1 - E^2\right) + {\displaystyle \frac{L_z^2}{\sin^2\theta}}\right].
\end{align}
\end{subequations}
As an affine parameter we have used Mino time which is related to the proper time $\tau$ by~\cite{Mino2003}
\begin{equation}
\tau = \intd{}{}{r^2 + a^2 \cos^2\theta}{\lambda}.
\end{equation}
Using Mino time allows us to decouple the $r$ and $\theta$ motions.

We only consider bound motion~\cite{Wilkins1972}: the radial motion covers a range $r\sub{p} \leq r \leq r\sub{a}$, where the turning points are the periapsis $r\sub{p}$ and apoapsis $r\sub{a}$. Drawing upon Keplerian orbits we parametrize the motion using
\begin{equation}
r = \frac{p M}{1+e\cos\psi},
\end{equation}
introducing eccentricity $e$, (dimensionless) semilatus rectum $p$ and relativistic anomaly $\psi$~\cite{Darwin1961,Drasco2004}. While $r$ oscillates between its maximum and minimum values, $\psi$ increases secularly, increasing by $2\pi$ across an orbit. The polar motion covers a range $\theta_- \leq \theta \leq \pi - \theta_-$. We also parametrize this motion in terms of an angular phase $\chi$, according to~\cite{Hughes2000}
\begin{equation}
\cos\theta = \cos\theta_-\cos\chi.
\end{equation}
While $\psi$ and $\chi$ are $2\pi$ periodic they are not the canonical action--angle variables~\cite{Schmidt2002}; they are, however, easy to work with.

The geodesic motion can equally be described by $\{E,L_z,Q\}$ or $\{p,e,\theta_-\}$~\cite{Schmidt2002}. Converting between them requires finding the solutions of $V_r = 0$ and $V_\theta = 0$. We employ a slightly different parameter set of $\{p,e,\iota\}$ where we have introduced the inclination~\cite{Ryan1996,Glampedakis2002}
\begin{equation}
\tan \iota = \frac{\sqrt{Q}}{L_z}.
\end{equation}
This is $0 \leq \iota < \pi/2$ for prograde orbits and $\pi/2 < \iota \leq \pi$ for retrograde orbits. Equatorial orbits ($\theta_- = \pi/2$) have $\iota = 0$ or $\pi$ and polar orbits ($\theta_- = 0$) have $\iota = \pi/2$. While formulae exist for conversion between the different parameters, these are complicated and uninsightful, so we do not reproduce them here.\footnote{In practice we find turning points numerically.}

\subsection{Orbital resonances}

The radial and polar orbital periods in Mino time are given by
\begin{subequations}
\begin{align}
\Lambda_r = {} & 2\intd{r\sub{p}}{r\sub{a}}{\recip{\sqrt{V_r}}}{r} = \intd{-\pi}{\pi}{\diff{\lambda}{\psi}}{\psi}, \\
\Lambda_\theta = {} & 4\intd{\theta_-}{\pi/2}{\recip{\sqrt{V_\theta}}}{\theta} = \intd{-\pi}{\pi}{\diff{\lambda}{\chi}}{\chi}.
\end{align}
\end{subequations}
The orbital frequencies are thus
\begin{equation}
\Upsilon_r = \frac{2\pi}{\Lambda_r}, \quad \Upsilon_\theta = \frac{2\pi}{\Lambda_\theta}.
\end{equation}
The geodesic equations for coordinate time $t$ and azimuthal angle $\phi$ are just functions of $r$ and $\theta$, hence their evolutions can be expressed as Fourier series~\cite{Drasco2004}
\begin{subequations}
\begin{align}
\diff{t}{\lambda} = {} & \sum_{k_r,\,k_\theta}T_{k_r,\, k_\theta}\exp\left[-i\left(k_r\Upsilon_r + k_\theta\Upsilon_\theta\right)\lambda\right], \\
\diff{\phi}{\lambda} = {} & \sum_{k_r,\,k_\theta}\Phi_{k_r,\, k_\theta}\exp\left[-i\left(k_r\Upsilon_r + k_\theta\Upsilon_\theta\right)\lambda\right].
\label{eq:Mino-Fourier}
\end{align}
\end{subequations}
The $(0,\,0)$ coefficients in these series give the average secular rate of increase of these quantities. We define
\begin{equation}
\Gamma = T_{0,\,0}, \quad \Upsilon_\phi = \Phi_{0,\,0}
\end{equation}
to act as Mino time frequencies. We can now convert to coordinate time frequencies with
\begin{equation}
\Omega_r = \frac{\Upsilon_r}{\Gamma}, \quad \Omega_\theta = \frac{\Upsilon_\theta}{\Gamma}, \quad \Omega_\phi = \frac{\Upsilon_\phi}{\Gamma}.
\end{equation}

Transient resonances occur when the radial and poloidal motions are commensurate, when
\begin{equation}
\nu = \frac{\Upsilon_r}{\Upsilon_\theta} = \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r}
\end{equation}
is the ratio of small integers. At this point any Fourier series like those in \eqnref{Mino-Fourier} goes from being an expansion in two frequencies to being an expansion in a single frequency~\cite{Bosley1992}.

For a general non-resonant orbit there is no fixed correlation between the radial and polar coordinates. After a sufficiently long time, the trajectory comes arbitrarily close to every point in the range of motion (with $r\sub{p} \leq r \leq r\sub{a}$ and $\theta_- \leq \theta \leq \pi - \theta_-$); on account of the orbital precession, the whole space is densely covered. This does not happen on resonance, as the radial and polar motions are locked together such that we can express one as a function of the other, and so the trajectory keeps cycling over the same path. The points visited are controlled by the relative phases of the $r$ and $\theta$ motions. To represent this, we use the $r$ phase at the $\theta$ turning point $\psi_{\theta_-} = \psi(\chi = 0)$. Varying $\psi_{\theta_-}$ across its full range allows every point in the range of motion to be reached. Hence averaging over all values of $\psi_{\theta_-}$ for a resonant orbit is equivalent to averaging over the $\psi$--$\chi$ $2$-torus for non-resonant orbits.

One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short time-scales, for small $\eta$ there is a small disparity. The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e., the case for motion in Kerr) is subject to a small perturbation the form of the orbits is preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}. % Chapter II, section 3.3 d)
This should ensure that, in general, there are only small shifts in the orbital frequencies. However, the KAM theory only applies for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}. % Chapter V section 1. c) 
This is a further reason why resonances merit an in-depth investigation.}
The dissipative piece causes the frequencies to evolve and, hence, the resonance cannot persist for multiple orbits (without some feedback coupling). In effect, we are really considering a period of time about the resonant crossing. The instantaneous orbital frequencies oscillate back and forth around their averaged values.

However, there shall be a time span when the frequencies are consistently close to being commensurate. During this time, the trajectory appears similar to a resonant trajectory, filling only a smaller region of the parameter space. It is this time period that is of interest for transient resonances~\cite{Bosley1992}.

%The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e., the case for motion in Kerr) is subject to a small perturbation the form of the orbits are preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}. % Chapter II, section 3.3 d)
%When we average the self-force and follow the adiabatic evolution of the trajectory we are exploiting this behaviour. However, the KAM theory only applies for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}. % Chapter V section 1. c)
% In these regions the adiabatic approximation breaks down and we need to follow the instantaneous evolution.

\subsection{Forced motion}
\label{sec:forced-motion}

For generic EMRIs, there are two timescales of interest: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change of fundamental frequencies $\sim\Omega/\dot{\Omega}$. These, along with the resonance timescale, are discussed more in \secref{res-time}. The two-timescale nature of the problem makes it ideally suited to the method of osculating elements~\cite{Gair2011a}: on short timescales, we analyse the unperturbed system resulting in geodesic motion, and then the long-term evolution is described by a sequence of instantaneous geodesics.

We require, at each instant in time, that the chosen geodesic matches the true position and velocity of the particle. This amounts to a specific choice of the orbital shape parameters (for example, the set $\{E,L_z,Q\}$ or the generalised action variables $J_\alpha$) and some initial phases at $t=t_0$ (for example, the set $\{\psi_0,\chi_0,\phi_0\}$). Collectively, these are referred to as \emph{osculating elements} and we denote them by $I^A(t)$, making explicit the variation with time.

Given geodesic motion, forced by some external acceleration $a^\alpha$, so that
\begin{equation}
\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = a^\alpha,
\end{equation}
we can calculate the evolution of the osculating elements $\dot{I}^A$. The specific equations for motion in Kerr are derived by Gair et al \cite{Gair2011a}.

\subsection{Self-force model}
\label{sec:self-force}

To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration. In this work we choose to work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}. For comparison, Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a} use a Teukolsky equation calculation of GW fluxes to account for the inspiral due to radiation reaction.

The self-force model uses the first-order PN terms of the dissipative self-force formulated in Flanagan and Hinderer~\cite{Flanagan2007} and the conservative force formulated in Iyer and Will~\cite{Iyer1993}, and Kidder~\cite{Kidder1995}. Since only the first PN terms are used, this prescription can only be of limited validity in strong fields. Both pieces of the self-force are computed assuming that the spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$. This is less than ideal for high spins. While this approximate self-force is not perfect, it should serve as a guide for the behaviour of the full self-force, allowing us to assess the qualitative impact of resonances on EMRI detection.

\subsection{Adiabatic evolution}

Beyond geodesic motion in the Kerr spacetime, a test particle will follow an accelerated trajectory determined by \ref{eq:Mino-E-o-M}. As mentioned previously, this may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}. This piece is purely dissipative and so signifies that flux is carried away from the system by gravitational waves.

In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform an appropriate average over the $2$-torus, given by
\begin{equation}
\label{eq:2torus-average}
\langle X\rangle_{\psi,\,\chi} = N \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\partialdiff{\psi}{t}\right)^{-1} \left(\partialdiff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} X }{\chi} }{\psi} .
\end{equation}
This form of averaging is successful at describing the leading order secular evolution of the trajectory, as is illustrated in \figref{good-traj}.

The effect of resonances must be completely contained within the pieces of the self-force left over after subtraction of the $2$-torus averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{\psi,\,\chi}$ for some quantity $X$ related to the self-force. The adiabatic evolution, which contains only a single averaged forcing term, will therefore not demonstrate any resonance behaviours\footnote{Resonances still occur as the fundamental frequencies become commensurate, but as the system passes through each resonance, there is no additional effect on the evolution of the system.}. One method of studying the effect of resonances is then to compare an adiabatic evolution with an instantaneous evolution, using the full knowledge of the self-force. This may be a useful exercise to calibrate and verify adiabatic models that can be generated cheaply, however it cannot be used to distinguish between dephasing effects arising from resonances and those arising from conservative corrections to the phase variables, as these are both explicitly eliminated from the adiabatic prescription. We therefore take a different approach, comparing the instantaneous evolution with a quasi-adiabatic evolution, generated by averaging the full available self-force over the $2$-torus, rather than just averaging the term $G_\alpha^{(1)}$.

To be specific, we write \ref{eq:Mino-E-o-M} as
\begin{subequations}
\begin{align}
\label{eq:Mino-E-o-M2}
\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{\psi,\,\chi}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
\diff{J_\alpha}{\lambda} = {} & \eta \langle G_\alpha\rangle_{\psi,\,\chi}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
\end{align}
\end{subequations}
where $g_\alpha$ and $G_\alpha$ encapsulate all our knowledge about the self-force and may contain contributions from sub-leading terms in the mass-ratio expansion. The quasi-adiabatic evolution is then obtained by dropping the terms proportional to $\hat{g}_\alpha$ and $\hat{G}_\alpha$. For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an "adiabatic evolution" in fact refer to a "quasi-adiabatic evolution".

\section{Properties of Transient Resonances}
\subsection{Location}

The first step in studying the effect of transient resonance is to locate orbital parameters for which the frequencies are commensurate. We can calculate the frequencies and so we are left with the problem of solving $\Omega = n_r \Omega_r - n_\theta \Omega_\theta = 0$ numerically. \Figref{res-plane-2-5-95} shows the semilatus rectum, eccentricity and (cosine of the) inclination angle of the $\nu = 2/5$ resonance surface for a BH of spin $a_\ast = 0.95$. 
\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_res-2-5-95-plane}
\caption{\label{fig:res-plane-2-5-95}Location of the $2/5$ resonance surface for an $a_\ast = 0.95$ BH in terms of orbital semilatus rectum $p$, eccentricity $e$ and inclination $\iota$.}
\end{figure}
The resonances occur at relatively small periapses, corresponding to regions of strong-field gravity.
Similar resonance surfaces can be found for other spin values and for other resonances. When considering the full parameter set of $\{p,e,\iota,a_\ast,\nu\}$ it is apparent that the search for resonances becomes expensive as a consequence of the dimensionality. It is therefore useful to have a guide of where to look. We shall attempt to build a model for the resonance plane.

Brink, Geyer and Hinderer~\cite{Brink2013} provide series expansions for the location of resonances in the limit of equatorial orbits for small spin and eccentricity. We do not follow this approach of trying to find analytic expressions for the resonance surface. The expressions become complicated when venturing away from limiting cases. Instead, we build an approximate phenomenological model and fit this to the resonance plane.%\footnote{A stopped clock is precisely correct twice a day, while a clock that is five minutes slow is never right but will often be close enough. Within this metaphor, our approximation would be a clock that varies between telling the correct time and being fifteen minutes off throughout the day: it is generally going to be good enough to give you a sense of the time but if you have an important event you should check with something better.}
This should be useful for designating the region in which resonance could be expected. To locate them precisely it is necessary to solve $\Omega = 0$ numerically; the approximate model should give a suitable starting point.

The resonant semilatus rectum for any particular spin and resonance ratio can be well approximated as
\begin{equation}
p(e,\iota;a_\ast,\nu) \simeq A\frac{1 + B e + D \cos\iota}{1 - C\exp(e)}.
\end{equation}
The coefficients $\{A,B,C,D\}$ depend upon the spin and the particular resonance; they can be approximated as
\begin{align} 
A(a_\ast,\nu) \simeq {} & a_0\frac{1 + a_1\nu - a_2 \nu^2 - a_3 \nu a_\ast^2}{1 + a_4\nu - (1 + a_4)\nu^2}, \\
B(a_\ast,\nu) \simeq {} & b_0(1 - b_1\nu)\exp(-b_2\nu)(1 - b_3 a_\ast), \\
C(a_\ast,\nu) \simeq {} & c_0, \\
D(a_\ast,\nu) \simeq {} & d_0\left[1 - \exp(a_\ast)\right]\left[1 - d_1\exp(\nu)\right].
\end{align}
This gives us a total of $12$ parameters for our fit. Whilst this may sound large, if we were fitting an expansion to quadratic order in any combination of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
\begin{equation}
\begin{array}{lll}
a_0 = 5.9854, & a_1 = 3.4116, & a_2 = 0.9253,\\
a_3 = 0.1959, & a_4 = 4.8846, & b_0 = 0.7692,\\
b_1 = 1.4752, & b_2 = 1.4861, & b_3 = 0.5974,\\
c_0 = 0.02332, & d_0 = 0.7968, & d_1 = 0.3115.
\end{array}
\end{equation}
These were fitted for all possible resonances with $n_r = 2$--$7$ as well as $\nu = 9/10$, $19/20$, $49/50$ and $99/100$; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$. 
%These were fitted for resonances with $\nu = 1/7, 1/6, 1/5, 1/4, 2/7, 1/3, 2/5, 3/7, 1/2, 4/7, 3/5, 2/3, 5/7, 3/4, 4/5, 5/6, 6/7, 9/10, 19/20, 49/50, 99/100$, with BH spins of $a_\ast = 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.93, 0.95, 0.97, 0.99, 0.999$.

Using this approximation, the maximum error in $p$ for a given $a_\ast$ and $\nu$ is typically $\sim10\%$ and less that $1$ in absolute terms. The relative error in the semilatus rectum is illustrated in \figref{p-error}. 
\begin{figure*}[htp]
\centering
\centerline{\includegraphics[width=0.47\textwidth]{Fig_fit-error-max-plane}\quad\includegraphics[width=0.47\textwidth]{Fig_fit-error-RMS-plane}}
\caption{\label{fig:p-error}Relative error in the approximate semilatus rectum compared to the accurate numerical result as a function of BH spin $a_\ast$ and resonance ratio $\nu$. The left panel shows the maximum relative error and the right shows the root-mean-square error; in both cases we are marginalising over eccentricity and inclination.}
\end{figure*}
The largest fractional error is $\sim50\%$, this is for $a\rightarrow 1$ and $\nu \rightarrow 0$, and corresponds to small $p$, such that the absolute error is still small. Taking the root-mean-square across $e$ and $\iota$, the fractional error for a given $a_\ast$ and $\nu$ never exceeds $9\%$ and is typically less than $4\%$.

\subsection{Time-scales}\label{sec:res-time}

When analysing resonances it is useful to refer to a number of reference time-scales.  We always use coordinate time $t$ for these, as this corresponds to what is measured by an observer at infinity. Translation to Mino time can be done with an appropriate factor of $\Gamma$. We use the orbital period $T$, the evolution time-scale $\tau\sub{ev}$, the precession time-scale $\tau\sub{pres}$ and the resonance time-scale $\tau\sub{res}$.

The simplest time-scales are the orbital periods $T_r = 2\pi/\Omega_r$, $T_\theta = 2\pi/\Omega_\theta$ and $T_\phi = 2\pi/\Omega_\phi$. These are the shortest in our set. We use $T$ to denote a time-scale of the same order as the orbital periods.

We define the evolution time-scale as
\begin{equation}
\tau\sub{ev} = \frac{\nu}{\dot{\nu}},
\end{equation}
where an overdot denotes a derivative with respect to $t$. In general, away from resonance, we take $\nu \equiv \Omega_r/\Omega_\theta$. This time-scale sets the period over which there is a significant change in the frequencies. It acts as an inspiral time-scale. It is long in all cases we study, $\tau\sub{ev} \sim \order{T/\eta}$. It is this property which makes extreme-mass-ratio inspirals interesting, as we can follow the waveform for many cycles, accruing high signal-to-noise ratios. This is also what allows us to use the adiabatic prescription, as it means the trajectory moves slowly through different orbital parameters.

We use the precession time-scale
\begin{equation}
\tau\sub{pres}(t) = \frac{2\pi}{|\Omega(t)|},
\label{eq:t-pres}
\end{equation}
with $\Omega(t) = n_r \Omega_r(t) - n_\theta \Omega_\theta(t)$, where the frequencies are calculated instantaneously and the integers are for the resonance of interest. This time-scale becomes infinite exactly on resonance, but decreases as we get further from resonance, eventually becoming $\order{T}$. It measures the relative precession rate of the radial and polar motions and hence gives an indication of how long it takes to fill the entire $\psi$--$\chi$ $2$-torus.

We also use the resonance time-scale
\begin{equation}
\tau\sub{res} = \left[\frac{2\pi}{\left|\langle\dot{\Omega}(0)\rangle_{q'}\right|}\right]^{1/2}.
\label{eq:t-res}
\end{equation}
Here $\dot{\Omega}(0)$ is the rate of change of $\Omega$ at resonance, which we take to be at $t = 0$. The instantaneous $\dot{\Omega}$ depends upon the orbital phase and oscillates about its mean trend over an orbit (see \figref{wibbles}). We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time-average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$. Close to resonance, $\Omega(t)$ is well approximated by a first-order Taylor expansion, decreasing linearly with time; hence we make the approximation
\begin{equation}
\left|\langle\dot{\Omega}\rangle_{q'}\right| \simeq \left|\frac{\Omega(t)}{t}\right|.
\end{equation}
The resonant time-scale should give an indication of the time over which we expect the effects of the resonance to be felt~\cite{Bosley1992}. Consider the phase of the Mino time Fourier expansion on resonance; neglecting the constant, the resonant Fourier component has form
\begin{equation}
\varphi_{n_r,\,-n_\theta} \simeq \left(n_r\Upsilon_r - n_\theta\Upsilon_\theta\right)\lambda + \left(n_r\dot{\Upsilon_r} - n_\theta\dot{\Upsilon}_\theta\right)\lambda^2 + \ldots
\end{equation}
Typically, the first term is non-zero and this gives the familiar oscillation. On resonance, it is zero, leaving the next order term to govern the behaviour~\cite{Flanagan2012}. Only once we have moved far enough away from resonance for the first term to dominate the second do we recapture the non-resonant behaviour. The first term (translating from Mino time to coordinate time) sets $\tau\sub{pres}$, the second sets $\tau\sub{res}$.

Since we have argued that the effect of resonance can be thought of as a consequence of not densely covering the $\psi$--$\chi$ $2$-torus, we might expect that $\tau\sub{pres}$, as well as $\tau\sub{res}$, could be used for setting the resonance duration: the resonance ends once sufficient time has elapsed that the $2$-torus could be filled. This is indeed the case. Let $t\sub{pres}$ be the time taken to fill the torus, then
\begin{align}
t\sub{pres} = {} & \tau\sub{pres}(t\sub{pres}) \\
 \simeq {} & \dfrac{2\pi}{\left|\langle\dot{\Omega}(0)\rangle_{q'} t\sub{pres}\right|}, \nonumber 
\end{align}
using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \langle\dot{\Omega}\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
\begin{equation}
t\sub{pres} \simeq \tau\sub{res}.
\end{equation}
The two time-scales are equivalent. We preferentially use $\tau\sub{res}$ to denote the resonance width. It is shorter than the inspiral time-scale, but longer than an orbital period, $\tau\sub{res} \sim \order{\eta^{1/2}\tau\sub{ev}} \sim \order{\eta^{-1/2}T}$~\cite{Flanagan2012,Gair2011a}; it therefore acts as a bridge between the two time-scales~\cite{Hinderer2008}.

Since we shall be considering Fourier decompositions, in anticipation of future results, we also define a time-scale for the $s$-th resonant frequency harmonic
\begin{align}
%\tau_{\mathrm{pres},\,s} = {} & \frac{2\pi}{|s\Omega(t)|};\\
\tau_{\mathrm{res},\,s} = {} & \left[\frac{2\pi}{\left|s\langle \dot{\Omega}(0)\rangle_{q'}\right|}\right]^{1/2}.
\label{eq:T-res-s}
\end{align}
This assumes that $s$ is a non-zero integer.

\subsection{Resonant flux enhancement}

Evolving through a resonance can lead to an enhancement (or decrement) of fluxes relative to the adiabatic prescription. After crossing the resonance region, the orbital parameters are different from those calculated from an adiabatic evolution. Flanagan and Hinderer~\cite{Flanagan2012} gave an expression for this deviation. If we denote the orbital parameters by $\mathcal{I}_a = \{E,L_z,Q\}$, then the change across resonance is
\begin{equation}
\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\sqrt{\dfrac{2\pi}{\left|s \langle\dot{\Omega}\rangle_{q'}\right|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
\end{equation}
Here $\widehat{\kappa}_0$ is the orbital phase on resonance and $F_{a,\,s}^{(1)}$ is the $s$-th harmonic of the first-order self-force on resonance, defined such that\footnote{Since the geodesic equations decouple in Mino time rather than coordinate time, this is true only in an average sense.}
\begin{equation}
\diff{\mathcal{I}_a}{t} = \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
\end{equation}
A derivation is presented in \apref{res-asymptotic}, which contains a more comprehensive explanation of the various terms. This employs matched asymptotic expansions to track the evolution through resonance, following the approach of Kevorkian~\cite{Kevorkian1987}.

The expression for the change can be understood using our expression for the resonance width, substituting in \eqnref{T-res-s},
\begin{equation}
\label{eq:delta-I-a}
\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
\end{equation}
Schematically, this can be understood as the magnitude of the forcing function on resonance $\sim \eta F_{a,\,s}$ multiplied by the time on resonance $\sim \tau_{\mathrm{res},\,s}$ and a function that varies with the phase $\widehat{\kappa}_0$. Averaging over all values of $\widehat{\kappa}_0$ is equivalent to averaging over all values of $\psi_{\theta_-}$, and has the same effect as averaging over the $\psi$--$\chi$ $2$-torus; this gives an average discrepancy relative to the adiabatic evolution of
\begin{equation}
\left\langle \Delta \mathcal{I}_a \right\rangle_{\hat{\kappa}_0} = 0,
\end{equation}
exactly as expected.

\section{Astrophysical implications}

To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs\footnote{The actual distribution in parameter space of observable EMRI systems is unknown, but we can make reasonable estimates.} and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.

We shall now discuss our chosen waveform generation scheme, giving demonstrations of its accuracy for our purposes. We will then detail our procedure for generating a realistic distribution of detectable EMRIs, before turning to the statistical effect of resonances on the population at the end of the section.

\subsection{Waveform model}

To generate gravitational waveforms, we choose to employ the 'Numerical Kludge' (NK) method of Babak \emph{et al}~\cite{Babak2007}. The principle behind the class of so-called kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory. This can be much quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives very similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher~\cite{Babak2007}.

Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the post-Newtonian self-force from \secref{self-force}. We then interpret the Boyer-Lindquist coordinates as spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique. Specifically, we apply the standard quadrupole formula~\cite{Misner1973}.

We expect these waveforms to be sufficiently accurate for our purposes, given the inaccuracy of the self-force model being employed. It would, however, be straight-forward to adopt a different waveform generation scheme, for example by including higher order multipole moments, should advances be made in determining the self-force.

Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s_A(t)$ and $h_A(t)$, according to
\begin{equation}
\label{eq:innerprod}
\overlap{s}{h} = 2 \sum_{A=+,\times}\intd{0}{\infty}{\frac{\tilde{s}_A(f)\tilde{h}_A^*(f)+\tilde{s}_A^*(f)\tilde{h}_A(f)}{S_n(f)}}{f},
\end{equation}
where $\tilde{s}_A(f)$ represents the Fourier transform of $s_A(t)$ and similarly for $\tilde{h}_A(f)$, $A = \{+,\times\}$ denotes the plus- or cross-polarisation state of the GW and $S_n(f)$ is the one-sided noise power spectral density (PSD). We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept. The analytic approximation we use is~\cite{Amaro-Seoane2012a}
\begin{equation}
S_n(f) = \frac{20}{3}\frac{4 S_{acc}(f)/(2\pi f)^4 + S_{sn}(f) + S_{omn}(f)}{L^2} \left(1+\left(\frac{f}{0.41 c/2L}\right)\right)^2,
\end{equation}
where $L = 1\;\mathrm{Mkm}$ is the length of the interferometer arms, $c$ is the speed of light and the components of noise separate into: the noise due to residual test mass acceleration $S_{acc}(f) = 2.13\times10^{-29}(1+10^{-4}\mathrm{Hz}/f)\;\mathrm{m^2\,s^{-4}\,Hz^{-1}}$; the shot noise $S_{sn} = 5.25\times10^{-23}\;\mathrm{m^2\,Hz^{-1}}$; and all other measurement noises $S_{omn} = 6.28\times10^{-23}\;\mathrm{m^2\,Hz^{-1}}$. We show a plot of this PSD against frequency in \figref{noise-curves}, along with the corresponding curve for the LISA detector for comparison. The effect of white dwarf (WD) confusion noise, which is important for LISA (as shown in the figure), doesn't significantly impact the eLISA mission and so we neglect it.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_noise_curves}
\caption{\label{fig:noise-curves}One-sided noise power spectral densities for space-based GW detectors, eLISA and LISA. The effect of WD confusion noise is shown for the LISA curve.}
\end{figure}

In practice, the functions $s_A(t)$ and $h_A(t)$ will not be available, instead replaced by the time series $s_A(t_i)$ and $h_A(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values but which do not necessarily have the same length or spacing. To take this into account, when evaluating overlaps, we first project the $h_A(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme, setting any values outside the domain of $\{t_j\}$ to zero. A Planck-taper window function~\cite{McKechan2010} is applied to the resulting time series. This can be defined in a continuous form by
\begin{equation}
w(t) = \left\{
\begin{array}{ll}
	0 & t\leq t_1 \\
	1\big/\left(1+\exp\left(\frac{t_2-t_1}{t-t_1}+\frac{t_2-t_1}{t-t_2}\right)\right) & t_1 < t < t_2 \\
	1 & t_2\leq t\leq t_3 \\
	1\big/\left(1+\exp\left(\frac{t_3-t_4}{t-t_3}+\frac{t_3-t_4}{t-t_4}\right)\right) & t_3 < t < t_4 \\
	0 & t_4\leq t,
\end{array}\right.
\end{equation}
where $t_1$ and $t_4$ are chosen to encompass the whole waveform and $t_2$ and $t_3$ are the times of the second closest stationary points to each end. We demonstrate the application of this window function to a simple sinusoid in \figref{window-func}, with a comparison to the uniform window, which is unity for $t_1\leq t\leq t_4$ and zero elsewhere. The use of the second stationary points from each end in the Planck-taper ensures that the transition is not too sharp; the smoothness of the waveform at each end reduces unwanted spectral leakage in the Fourier transforms.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_window_func}
\caption{\label{fig:window-func}Two window functions applied to a simple sinusoid: the uniform (dashed) and the Planck-taper (solid). The Planck-tapered sinusoid has no discontinuities at the endpoints.}
\end{figure}

After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes. In addition, we maximise every overlap calculation with respect to some time-offset between the models by computing the correlation between the waveforms. This can be performed cheaply using Fourier transforms and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.

%This whole procedure gives more accurate results than interpolating in the Fourier domain since the waveforms are smoother in the time domain.

One of the targets of pre-eLISA research is to generate a bank of waveform templates, $h_A(t;\lambda^a)$ described by some set of parameters $\lambda^a$. Once real-world data, $s_A(t)$, has been collected, we can then search over the template bank to find the model that most closely matches the data. At the current time, with a lack of experimental data, we are interested in ensuring that the waveform templates accurately reproduce what we expect to observe in nature, without being too computationally expensive.

Ideally, we would like to use waveform templates from EMRI systems that are evolved under the entire self-force model, but these are too expensive. In practice, we are therefore forced to use cheaper adiabatic waveforms, but these do not include the effect of resonances. To assess the impact of this, we compare data $s_A(t;\bar{\lambda}^a)$ generated using the full self-force model to templates $h_A(t;\lambda^a)$ generated using the expensive 2-torus averaged self-force model. As discussed before, this eliminates any other effects and allows us to focus on the resonance problem.

We wish to test whether there exists some set of parameters $\lambda^a$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform. We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template, defined by
\begin{equation}
\frac{S}{N}\left[h\right] = \frac{\langle\overlap{s}{h}\rangle_n}{\sqrt{\left\langle\overlap{n}{h}^2\right\rangle_n}} = \frac{\overlap{s}{h}}{\sqrt{\overlap{h}{h}}},
\end{equation}
where $\langle\rangle_n$ represents an average over all possible noise realisations. If a template were to exactly match the data, it would produce an SNR of $\sqrt{\overlap{s}{s}}$ and so the overlap
\begin{equation}
\label{eq:overlap}
\mathcal{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}}
\end{equation}
provides an indication of how well-matched the template is to the data. We leave an investigation into parameter estimation biases (where the parameters of the best-matching adiabatic, $\lambda^a$, do not agree with the parameters of the system, $\bar{\lambda}^a$) to future work.

To demonstrate the approach, we calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$ years for a set of initial parameters chosen to avoid any significant resonances. Specifically, we choose: the mass of the CO $\mu = 10 M_{\odot}$; the mass ratio $\mu / M = 3\times 10^{-6}$ where $M$ is the mass of the central BH; the spin of the central BH $a=0.95$; the initial dimensionless semi-latus rectum $p=7.5$; the initial eccentricity $e=0.7$; the initial inclination angle $\cos \iota = 0.5$; and the redshift of the source $z=0.204$. The remaining intrinsic (the initial phases $\psi$ and $\chi$) and extrinsic (the position of the source on the sky, the direction of the spin of the central BH and the azimuthal angle) parameters are set to arbitrary values. The resulting evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, where it can be seen that the adiabatic evolution closely matches the full evolution on longer inspiral timescales. The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the 2-torus averaging explicitly smoothes out the visible structure on this scale.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{Fig_good_traj}
\caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) under the full (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances. The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.}
\end{figure}

Once the trajectories have been computed using both methods of evolution, we can calculate the corresponding waveforms. The plus-polarised GW at the start and end of the evolution is shown in \figref{good-waveform}. The adiabatic waveform shows good agreement with the full waveform in both amplitude and phase. Calculating the SNR for each of the models, we find an overlap $\mathcal{O} = 0.984$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{Fig_good_waveform}
\caption{\label{fig:good-waveform}The plus-polarised waveform for an illustrative EMRI system with $p=7.5$, shown for two short segments at the start and end of a $2$ year evolution. The full and adiabatic models are both shown, but are indistinguishable by eye.}
\end{figure}

\subsection{The effect of resonances: Dephasing}

The illustrative EMRI system of the previous section does not encounter a significant resonance and the adiabatic evolution provides a good match to the full evolution. We now study a system that does pass through a resonance during its $2$ year evolution. Specifically, we choose the initial conditions to be the same as before but with an initial semi-latus rectum $p=7.85$. This system passes through the 2:3 resonance, the effect of which is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms. Plotting the plus-polarised GWs at the start and end of the evolution, as before, demonstrates this dephasing, as shown in \figref{dephased-waveform}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{Fig_dephased_waveform}
\caption{\label{fig:dephased-waveform}The plus-polarised waveform for an illustrative EMRI system with $p=7.85$, shown for two short segments at the start and end of a $2$ year evolution. The full and adiabatic models are both shown; there is a significant dephasing between them by the end of the evolution.}
\end{figure}

To highlight the dephasing more quantitatively, we calculate the shortened overlap between the two models as a function of time $t$, defined as the overlap obtained by including only the part of the waveform within some $\Delta t$ of $t$. For our illustrative system, we choose $\Delta t$ such that we can calculate 25 non-overlapping values of this shortened overlap. Before the resonance occurs, the adiabatic model provides a good match to the full evolution but the overlap is reduced near to the resonance and never fully recovers afterwards. This is shown in \figref{overlap-dephasing}, which is centered on the time at which the full evolution crosses the 2:3 resonance.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, chosen to give 25 independent (non-overlapping) calculations. The time axis is centred on the value at which the full evolution crosses the 2:3 resonance.}
\end{figure}

Also shown is the shortened overlap computed between the full evolution and a different adiabatic evolution that is chosen to match the full evolution at the end of the integration. In this case, we see similar behaviour as before: the adiabatic waveform has a high overlap in the region where it is constructed to match the full evolution and the boundary of that region is set by the presence of an important resonance.

\subsection{The effect of resonances: Loss of SNR}
\label{sec:SNRloss}

To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral. The dephasing means that an adiabatic evolution with the same initial conditions as a full evolution is likely to have a low overlap if a significant resonance is encountered. However, this does not necessarily mean that all adiabatic evolutions will have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.

The extremely large parameter space of adiabatic waveforms renders a brute-force approach computationally impossible; it is therefore necessary to restrict the search in some way. For this preliminary investigation into the importance of resonances, we do this by focussing on a small subset of parameters that we suspect will produce a large overlap. We make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t_{\mathrm{match}}$. This reduces the search to a one parameter family of waveforms that can easily be computed concurrently with the full evolution.

The most restrictive aspect of this approach is that all extrinsic parameters and any intrinsic ones that remain constant throughout the evolution, such as the masses of the BHs and the central BH spin, do not get explored by our family of adiabatic models. It is possible that small deviations from these true parameters will give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects. Indeed, systematic biases are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.

The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$. Anticipating that resonances will play a key role in determining the degree of waveform overlap, we choose matching times situated at $5$ resonance-timescales after each resonance of interest, namely the low-order 2:3, 1:2, 2:5 and 1:3. These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral. We also include templates that match at the start of the evolution, so that we have a model that should match the portion of the evolution before the first resonance, and at the end of the evolution, as this is often where most SNR is accumulated. In addition, we include models that match exactly on each of the resonances as these may achieve partial matching both before and after the resonance and hence perform well.

We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.

In our experience, we found that these choices were not sufficient to produce a good family of adiabatic waveforms. In particular, if an adiabatic model happened to match the full model close to a turning point in the fast orbital evolution, then it eventually diverged from the full model and the resulting waveform provided a poor match. We therefore adjust the values of $t_{\mathrm{match}}$ so that they correspond to times in the inspiral when the time-derivative of a particular orbital parameter (we choose the semi-latus rectum $p$) is maximally different from the adiabatic time-derivative of that orbital parameter. This ensures that the adiabatic model intersects with the full model close to the centre of the envelope created by the fast oscillations, as demonstrated by the inset plots in \figref{good-traj}, and as a result gives much higher overlaps in general.

For the illustrative EMRI system detailed in the previous section, we have computed the family of adiabatic models set by $t_{\mathrm{match}}$ and the resulting trajectories are consistent for the length of the evolution. As they cannot be easily distinguished by eye, we plot the difference (in $E$, $L_z$ and $Q$) between the models and the adiabatic evolution that matches at the start, shown in \figref{res-diff-traj}. The jumps in the orbital parameters due to the 2:3 resonance can be clearly seen, as can the fast orbital oscillations present in the full evolution but absent in the adiabatics.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{Fig_res_diff_traj}
\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start. The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.202), at the 2:3 resonance (0.458), after the 2:3 resonance (0.259) and at the end (0.677).}
\end{figure}

None of the adiabatic models presented here give a particularly high overlap with the full evolution, because of the effects of the resonance. The best-performing adiabatic model was that which matched at the end, giving $\mathcal{O} = 0.677$, while the model that matches at the start gives only $\mathcal{O} = 0.202$. These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.

\subsection{The effect of resonances: Jump sizes}

As demonstrated in \figref{res-diff-traj} and computed analytically in \apref{res-asymptotic}, the full evolution undergoes a rapid change in the orbital parameters with respect to the adiabatic evolutions. To extract the magnitude of this jump from the trajectory data, we first must account for the fast orbital oscillations as well as ensuring that we include the full effect of the resonance, while minimising any changes resulting from other sources.

To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away. For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}_\mathrm{ad}[t_{\mathrm{match}} = t_\mathrm{res}]$ for each of the orbital parameters. We then identify the most horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations. The resulting data and bounding lines are shown for our illustrative system in \figref{res-jump-calc}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_res_jump_calc}
\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the 2:3 resonance, scaled by the magnitude of the 2:3 resonance jump. The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump. The dotted red line indicates the computed size of the jump. The time axis is centred on the 2:3 resonance and is scaled by the appropriate resonance timescale.}
\end{figure}

To estimate the size of the jump $\Delta \mathcal{I}_{\mathrm{jump}}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits. This allows us to take into account any linear drift in $\Delta \mathcal{I}$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation. In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down. For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.

The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems. Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance. For the orbital parameter $\mathcal{I}$,
\begin{equation}
\label{eq:res-jump-ratio}
\frac{\Delta \mathcal{I}_\mathrm{jump}}{\mathcal{I}_\mathrm{ad}(t>t_\mathrm{res})-\mathcal{I}_\mathrm{ad}(t<t_\mathrm{res})} = \frac{\Delta \mathcal{I}_\mathrm{jump}}{\dot{\mathcal{I}}_\mathrm{ad}(t_\mathrm{res})\tau_\mathrm{res}}.
\end{equation}

The resonance jump depends sensitively on the relative phase of radial and poloidal motions. It is therefore instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi$, which we do for our illustrative system parameters. Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqref{eq:delta-I-a}.

Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump. We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$. The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqref{eq:delta-I-a} and giving us confidence in our numerical jump calculations. We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}); they are highly correlated to the magnitude of the jump in $E$, due to the dependence on $u_t$.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$. The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqref{eq:res-jump-ratio}. The individual jumps as well as a sinusoidal fit are plotted for the three orbital shape parameters (solid lines) and the three orbital frequencies (dotted).}
\end{figure}

For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t_\mathrm{match}$ described in the previous section. The largest overlap in each case is plotted in \figref{overlap-vs-resjump-ill}, as a function of the percentage resonance jumps in the orbital parameters. We observe no strong trend with the size of the jump, although systems with small jumps in $E$ (less than $1\%$) have higher than average overlaps. These systems also have small jumps in the fundamental frequencies and so the adiabatic waveform dephases more slowly from the full waveform, leading to a larger overlap.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{Fig_overlap_vs_resjump_ill}
\caption{\label{fig:overlap-vs-resjump-ill}The overlap \eqref{eq:overlap} between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the resonance jumps in $E$ (left), $L_z$ (middle), and $Q$ (right). Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase. The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
\end{figure}

The best performing adiabatic for the majority of our systems matched at the end of the inspiral. We can interpret the low overlaps by assuming that these adiabatic waveforms recover the entire SNR in the region post-resonance, but none pre-resonance. As a result, we would expect the overlap to roughly equal the fraction of time spent post-resonance, which in this case is $0.7$, demonstrated by the red line in \figref{overlap-vs-resjump-ill}. There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the base of the eLISA noise curve.

A small fraction of systems have a much lower overlap than expected, at around $0.5$. This is due to a sub-optimal choice of the matching times; making small adjustments to $t_\mathrm{match}$ can lead to higher overlaps around $0.7$, but the exact adjustments required are difficult to predict in advance.

\subsection{Sample EMRI population}

Strong resonances can severely limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic modes, which may be individually undetectable. To assess the impact of this on future gravitational wave missions, we need to create a sample population of detectable EMRIs.

We first identify 14 parameter distributions that describe an astrophysically typical EMRI. We choose the mass of the compact object $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs. We take the central object to be a typical supermassive BH at the centre of a galaxy, with mass $M$ uniformly distributed in log-space between $10^5$ and $10^7 M_\odot$, and dimensionless spin $a$ uniformly between its limiting values, $0$ and $1$. The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere. The angles specifying the location of pericenter and the compact object around the orbit are distributed uniformly between $0$ and $2\pi$.

We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\,\mathrm{km s^{-1}\, Mpc^{-1}}$, and sample the redshift of the source uniformly from comoving volume, up to a maximum of $z_\mathrm{max} = 0.7$. We don't expect the exact details of the cosmology to significantly alter our results and our choice of $z_\mathrm{max}$ is justified \textit{post hoc}.

The astrophysics in dense stellar clusters at the center of galaxies leading to EMRI production is highly uncertain. The currently favoured formation mechanism scatters compact objects onto highly eccentric orbits, which lose a burst of energy to GW emission with each periapse passage. This tends to circularise the orbit, as the compact object spirals inwards. Hopman and Alexander \cite{Hopman2005} model this scattering and give an eccentricity distribution for compact objects spiralling into a $3 \times 10^6 M_\odot$ Schwarzschild BH at the point when the orbital period enters the LISA band at $10^4\,\mathrm{s}$. The overall results of their work are still valid for spinning BHs and so assuming that this distribution holds for other BH masses, we can sample the orbit eccentricity and calculate the semi-latus rectum from the orbital period.

These EMRIs are astrophysically motivated, but may not be detectable by eLISA. To assess this, we evolve the systems forward in time using the analytic kludge method of Barack and Cutler \cite{Barack2004} until they reach the last stable orbit. We then calculate the Peters-Matthews waveform \cite{Peters1963} for the last $t_\mathrm{insp}$ years of the inspiral, where $t_\mathrm{insp}$ is chosen uniformly between $0.5$ and $2$. We focus on the late stages of inspiral as these are most likely to be detected by eLISA. Given the waveform, we can estimate the expected SNR by approximating the overlap with a time-domain integral
\begin{equation}
\overlap{s}{h} \approx 4 \sum_{A=+,\times}\intd{0}{T}{\frac{s_A(t)h_A(t)}{S_n(f(t))}}{t},
\end{equation}
where $f(t)$ is some characteristic frequency of the orbit at time $t$. We find that taking $f$ to be twice the (radial) orbital frequency gives a good approximation to the SNR.

To obtain our final distribution of detectable EMRIs, shown as dark blue in \figref{EMRIpar-dists}, we quote the parameter values $t_\mathrm{insp}$ years before plunge, including only those systems with an estimated SNR above 10. Also shown in the figure, in light pink, are the systems with an estimated SNR below 10, to illustrate which systems tend to have lower SNRs. The acceptance rate for generating a detectable EMRI from the entire distribution is $32\%$.

We note that prograde orbits (those with $\cos\iota > 0$) and systems with higher BH spins tend to be favoured; this is likely due to the fact that such systems can inspiral much closer before finally plunging and so experience the strongest gravitational fields. This effect is amplified by the fact that the majority of the SNR is recovered from the final part of the inspiral, which also explains why larger values of $t_\mathrm{insp}$ are only marginally preferred over smaller values.

Systems at higher redshifts start to become undetectable as the amplitude of the signal scales inversely proportional to the luminosity distance to the source. At $z=0.7$, the fraction of detectable systems drops to $5\%$; this is chosen as the upper cutoff in the redshift distribution as a compromise between including as many detectable systems as possible, but without having too low an overall acceptance rate. We quote the masses in the system as "redshifted masses" to account for the FRW background metric.

The circularising effect of GW emission can be observed in the eccentricity distribution: even though the input systems from Hopman and Alexander have high eccentricites, with a peak in the distribution at $e\sim0.7$, the final distribution contains only low eccentricity systems, peaked at $e\sim0.05$ with a maximum at $e=0.31$.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{Fig_EMRIpar_dists}
\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $50,000$ realisations. The (darker) blue bars represent the final distribution, including an SNR cut of 10. The (lighter) pink bars stacked on top indicate the full distribution before the SNR cut.}
\end{figure}

\subsection{Population results}

We take draws from the EMRI parameter distributions, using them as initial conditions for the full instantaneous evolution, from which we generate a family of adiabatic evolutions, as discussed in \secref{SNRloss}. There are broadly three categories of system regarding the effect of resonances on the evolution:

\begin{enumerate}
\item No significant resonance jumps are observed, and all the adiabatic models agree with the full evolution reasonably well. The overlaps tend to be noticeably lower than $1$ since the adiabatic approach breaks down near the end of the inspiral.

\item At least one resonance jump can be observed (typically the 2:3 resonance), but the adiabatic models still produce large overlaps. Most of the SNR from a signal is accumulated near to the end of the inspiral, at which point there are few significant resonances.

\item At least one resonance jump can be observed, and all of our chosen adiabatic models have low overlaps with the full evolution. This does not rule out the possibility of adiabatic models existing that would produce a high overlap, but it is difficult to predict the combination of parameters that would achieve this.
\end{enumerate}

To ensure that our chosen adiabatics are reasonable estimates to maximise the overlap, we computed many adiabatic waveforms on a rigid grid for one particular system; there can be large fluctuations in the overlap, but our chosen parameters perform at least averagely well.

Across the full population, of $140$ different realisations, we find that $77\%$ of systems pass through the 2:3 resonance within the later stages of inspiral. In addition, $99\%$ of the systems we studied encountered the 1:2 resonance. This suggests that transient resonances are an important consideration in the evolution of EMRIs.

While resonance jumps may in principle be as large as a few percent (see the example in \figref{resjump-vs-q}), this tends to only occur in high eccentricity systems. The typical size of any resonance jump within our astrophysical population is less than $0.01\%$, with many systems showing no visible changes through resonance. \Figref{pop-SNR-vs-jump} shows the distribution of computed resonance jumps in $E$ for our population, along with the corresponding loss of overlap. There is no observed trend with the size of the resonance jump, although this is not surprising given the small size of the jumps. The large scatter on the plot is due to our choice of adiabatic family; in each case, the quoted loss of SNR is a lower bound for the actual loss of SNR across the entire adiabatic model space.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{Fig_pop_SNR_vs_jump}
\caption{\label{fig:pop-SNR-vs-jump}The fraction of the full SNR recovered using our family of adiabatic models, as a function of the percentage resonance jump in $E$. The histograms show the marginalised distributions over each parameter.}
\end{figure}

It is possible for EMRIs to form through alternative channels, which may result in a population of high eccentricity systems. For example, 'plunging' compact objects can be mistaken for slow inspirals since they still undergo many hundreds of thousands of cycles in the bandwidth of a detector like eLISA \cite{Amaro-Seoane2013}. This type of system is expected to be orders of magnitude more common than classic EMRIs. It is expected that resonance behaviours would be much more important in this case.

Even within the low eccentricity systems considered here, it is possible for some EMRIs to be undetectable using adiabatic models, given the resonances they encounter. These are any systems for which the full SNR is larger than some cutoff value, but the largest adiabatic SNR is smaller than the cutoff. The estimated fraction of systems that would be lost in this manner, as a function of cutoff SNR is shown in \figref{SNRcutoff}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46\textwidth]{Fig_SNRcutoff}
\caption{\label{fig:SNRcutoff}The fraction of EMRI systems that would be lost using adiabatic models to analyse data that contain the effects of transient resonances. The error bars are estimated using the beta distribution.}
\end{figure}

\section{Concluding remarks}

Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral. However, due to the low eccentricity of detectable sources, the overall effect on SNR recovery is small; it is therefore likely to be sufficient to use adiabatic models to detect EMRIs. An unresolved question here is how these adiabatic models affect parameter estimation: systematic biases may be introduced if we neglect the presence of important resonances.

High eccentricity EMRIs may also be possible, in which case transient resonances will play an important role in their detectability; systems that undergo a jump of a few percent dephase very rapidly from the adiabatic approximation. In such scenarios, adiabatic models can still be used, but there must be sufficient SNR in each non-resonant region that they can be identified as individual sources and then joined together as one long inspiral. Otherwise, if the phase difference on resonance can be predicted, \eqref{eq:delta-I-a} can be used to join together different adiabatic models.

This work is highly dependent on the chosen self-force model, and so the results should be taken as qualitative estimates, given the unsuitability of a PN self-force to the problem of fast motion near to highly-spinning BHs. In the future, with a more accurate self-force model, more quantitative results can be derived.

\begin{acknowledgments}
The authors extend our sincere gratitude to \'{E}anna Flanagan and Tanja Hinderer for providing their PN self-force code, without which this work would not be possible. We also wish to thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent and Leor Barack for useful conversations. RHC is supported by STFC; CPLB was supported by STFC and the Cambridge Philosophical Society; PC is supported by a Marie Curie Fellowship, and JRG is supported by the Royal Society.
\end{acknowledgments}

\appendix

\section{Asymptotic solution for passage through resonance}\label{sec:res-asymptotic}

\input{./asymptotic.tex}

\bibliography{Resonances}

\end{document}

### 2014-08-04 res_char.tex
two column layout

@@
-\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,reprint,showpacs]{revtex4}
+\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,reprint,showpacs,superscriptaddress,twocolumn]{revtex4}
 \usepackage{graphicx}
@@
 \author{Robert H. Cole}
-\email[]{rhc2@ast.cam.ac.uk}
+\email[]{rhc26@ast.cam.ac.uk}
+\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
 \author{Christopher P.L. Berry}
-\email[]{cplb2@ast.cam.ac.uk}
+\email[]{cplb@star.sr.bham.ac.uk}
+\affiliation{School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
 \author{Priscilla Ca\~{n}izares}
@@
 \begin{abstract}
-The inspiral of stellar mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational field regime, via the emission of gravitational waves.
+The inspiral of stellar-mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
 In order to detect such signals and extract the desired information, it is necessary to possess a bank of accurate waveform templates.
 For computational efficiency, adiabatic templates are often used, but these neglect the effect of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
-During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object trajectory and ultimately the phase of the emitted radiation.
-An astrophysical population of detectable extreme-mass-ratio-inspirals is studied and it is found that a large proportion will encounter a low-order resonance in the later stages of inspiral.
-The resulting effect on SNR recovery is small due to the low eccentricity of the systems, but nevertheless at least 20\% of EMRIs may become undetectable if transient resonances aren't accounted for.
+During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact-object trajectory and, ultimately, the phase of the emitted radiation.
+In this work, an astrophysical population of detectable extreme-mass-ratio-inspirals is studied and it is found that a large proportion shall encounter a low-order resonance in the later stages of inspiral.
+The resulting effect on signal-to-noise recovery is small due to the low eccentricity of the studied population, but nevertheless at least one fifth of the systems may become undetectable if transient resonances aren't accounted for.
 \end{abstract}
@@
 The two-body problem in general relativity (GR) is well studied.
-It is of paramount importance for gravitational wave (GW) astronomy, where binary systems are the dominant source of radiation.
+It is of paramount importance for gravitational-wave (GW) astronomy, where binary systems are the dominant source of radiation.
 Correctly modelling the dynamics of these systems is necessary to interpret and extract information from gravitational waveforms.
@@
 These simulations shall allow us to understand BH--BH mergers.
-Stellar-mass BH mergers are targets for ground-based GW detectors, such as the currently mid-upgrade LIGO~\cite{Harry2010} and Virgo~\cite{Accadia2011}, and the in-construction Kagra~\cite{Kuroda2010}.
+Stellar-mass BH mergers are targets for ground-based GW detectors, such as the currently mid-upgrade LIGO~\cite{Harry2010} and Virgo~\cite{Accadia2011}, and the in-construction KAGRA~\cite{Kuroda2010}.
 Massive BH mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
@@
 Calculations can instead be performed perturbatively.
-The paradigm unequal mass system has a stellar-mass BH orbiting a massive black hole, such as those expected to be found at the centre of most galaxies.
+The paradigm unequal-mass system has a stellar-mass BH orbiting a massive black hole, such as those expected to be found at the centre of most galaxies.
 These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors~\cite{Amaro-Seoane2007}.
@@
 In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}.
-In the test particle limit the smaller body follows an exact geodesic of the BH's spacetime.
+In the test particle limit, the smaller body follows an exact geodesic of the BH's spacetime.
 Including the effects of the smaller body's finite mass, the background spacetime is perturbed.
@@
 The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase although this can accumulate over many orbits.
-Being able to accurately model the influence of the self-force shall allow us to create accurate waveform models.
+Being able to accurately model the influence of the self-force shall allow us to create reliable waveform models.
 
@@
 Here $\dot{\Omega}(0)$ is the rate of change of $\Omega$ at resonance, which we take to be at $t = 0$.
-The instantaneous $\dot{\Omega}$ depends upon the orbital phase and oscillates about its mean trend over an orbit (see \figref{wibbles}).
+The instantaneous $\dot{\Omega}$ depends upon the orbital phase and oscillates about its mean trend over an orbit.
 We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time-average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$.
@@
 The analytic approximation we use is~\cite{Amaro-Seoane2012a}
-\begin{equation}
+\begin{widetext}\begin{equation}
 S_n(f) = \frac{20}{3}\frac{4 S_{acc}(f)/(2\pi f)^4 + S_{sn}(f) + S_{omn}(f)}{L^2} \left(1+\left(\frac{f}{0.41 c/2L}\right)\right)^2,
-\end{equation}
+\end{equation}\end{widetext}
 where $L = 1\;\mathrm{Mkm}$ is the length of the interferometer arms, $c$ is the speed of light and the components of noise separate into: the noise due to residual test mass acceleration $S_{acc}(f) = 2.13\times10^{-29}(1+10^{-4}\mathrm{Hz}/f)\;\mathrm{m^2\,s^{-4}\,Hz^{-1}}$; the shot noise $S_{sn} = 5.25\times10^{-23}\;\mathrm{m^2\,Hz^{-1}}$; and all other measurement noises $S_{omn} = 6.28\times10^{-23}\;\mathrm{m^2\,Hz^{-1}}$.
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \centering
@@
 The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.}
-\end{figure}
+\end{figure*}
 
@@
 The full and adiabatic models are both shown, but are indistinguishable by eye.}
-\end{figure}
+\end{figure*}
 
@@
 The full and adiabatic models are both shown; there is a significant dephasing between them by the end of the evolution.}
-\end{figure}
+\end{figure*}
 
@@
 The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.202), at the 2:3 resonance (0.458), after the 2:3 resonance (0.259) and at the end (0.677).}
-\end{figure}
+\end{figure*}
 
@@
 The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
-\end{figure}
+\end{figure*}
 
@@
 The (lighter) pink bars stacked on top indicate the full distribution before the SNR cut.}
-\end{figure}
+\end{figure*}
 
@@
 The histograms show the marginalised distributions over each parameter.}
-\end{figure}
+\end{figure*}
 

### 2014-08-04 res_char.tex
small changes

@@
 
-\title{Characterization of transient resonances in extreme-mass-ratio-inspirals}
+\title{Characterisation of transient resonances in extreme-mass-ratio-inspirals}
 
@@
 In order to detect such signals and extract the desired information, it is necessary to possess a bank of accurate waveform templates.
-For computational efficiency, adiabatic templates are often used, but these neglect the effect of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first order dissipative gravitational self-force.
+Other effects are however neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
 During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact-object trajectory and, ultimately, the phase of the emitted radiation.
@@
 % Section 10.6
-In the weak-field limit these all tend towards the Keplerian frequency; in the strong-field $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
+In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field, $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
 For extreme-mass-ratio systems, the evolution time-scale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital time-scales.
@@
 Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We shall consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$, and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
+We shall consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.}, and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
 We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
@@
 where $\lambda$ is Mino time~\cite{Mino2003}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force.
-By working with $lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
+By working with $\lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
 At zeroth order in the mass ratio we recover the limit of purely geodesic motion: the integrals of the motion are actually constants and the angle variables evolve according to their associated frequencies $\omega_\alpha$.
@@
 However, this averaging fails when the ratio of frequencies is the ratio of integers.
-In this case the trajectory does not ergodically fill the $2$-torus but instead traces out a $1$-dimensional subspace.\footnote{For illustrations, see Grossman, Levin and Perez-Giz~\cite{Grossman2012}.} There are then contributions to the self-force that no longer average out beyond $\langle G_A^{(1)}\rangle_{q_r,\,q_\theta}$.
+In this case the trajectory does not ergodically fill the $2$-torus, but instead traces out a $1$-dimensional subspace.\footnote{For illustrations, see Grossman, Levin and Perez-Giz~\cite{Grossman2012}.} There are then contributions to the self-force that no longer average out beyond $\langle G_A^{(1)}\rangle_{q_r,\,q_\theta}$.
 Intuitively, we expect that this effect is more important for ratios of small integers as when the integers are large, the orbit comes close to all points on the $2$-torus.
@@
 In this work we seek to characterise the importance of these resonances for the purposes of modelling EMRIs.
-The amplitude of expected signals will be below the level of noise in a space-based GW detector.
+The amplitude of expected signals shall be below the level of noise in a space-based GW detector.
 However, systems remain in band for many hundreds of thousands of cycles and so may be detected using a matched filter, provided we have sufficiently accurate waveform templates.

### 2014-08-04 res_char.tex
paper summary

@@
 
+In \secref{problem}, we formulate the specific problem at hand: that of geodesic motion in Kerr spacetime, perturbed by the gravitational self-force.
+We then study generic properties of transient resonances in \secref{properties}, detailing their location in parameter space, the time-scales over which they affect the motion and the resulting GW flux enhancements.
+Specific examples are considered to illustrate the effects of resonances at the start of \secref{astrophysics}, before finally turning to an astrophysical population in \secref{population}.
+Our conclusions can be found in \secref{conclusion}.
+
 We use geometric units with $G = c = 1$ throughout.
@@
 \section{Problem formulation}
-
+\label{sec:problem}
 \subsection{Kerr geodesics}
@@
 \section{Properties of Transient Resonances}
+\label{sec:properties}
 \subsection{Location}
@@
 \section{Astrophysical implications}
+\label{sec:astrophysics}
 
@@
 \subsection{Population results}
+\label{sec:population}
 
@@
 \section{Concluding remarks}
+\label{sec:conclusion}
 

### 2014-08-05 res_char.tex
adiabatic averaging

@@
 \end{equation}
-to act as Mino time frequencies. We can now convert to coordinate time frequencies with
+to act as Mino-time frequencies. We can now convert to coordinate-time frequencies with
 \begin{equation}
@@
 is the ratio of small integers.
-At this point any Fourier series like those in \eqnref{Mino-Fourier} goes from being an expansion in two frequencies to being an expansion in a single frequency~\cite{Bosley1992}.
+At this point, any Fourier series like those in \eqnref{Mino-Fourier} goes from being an expansion in two frequencies to being an expansion in a single frequency~\cite{Bosley1992}.
 
@@
 Varying $\psi_{\theta_-}$ across its full range allows every point in the range of motion to be reached.
-Hence averaging over all values of $\psi_{\theta_-}$ for a resonant orbit is equivalent to averaging over the $\psi$--$\chi$ $2$-torus for non-resonant orbits.
+Hence averaging over all values of $\psi_{\theta_-}$ for resonant orbits is equivalent to averaging over the $\psi$--$\chi$ $2$-torus for non-resonant orbits.
 
@@
 
-For generic EMRIs, there are two timescales of interest: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change of fundamental frequencies $\sim\Omega/\dot{\Omega}$.
+For generic EMRIs, there are two timescales of interest: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$.
 These, along with the resonance timescale, are discussed more in \secref{res-time}.
@@
 
-Given geodesic motion, forced by some external acceleration $a^\alpha$, so that
+Given geodesic motion in some background spacetime with Christoffel symbols $\Gamma^\alpha_{\beta\gamma}$, forced by some external acceleration $f^\alpha$, so that
 \begin{equation}
-\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = a^\alpha,
+\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = f^\alpha,
 \end{equation}
@@
 
-Beyond geodesic motion in the Kerr spacetime, a test particle will follow an accelerated trajectory determined by \ref{eq:Mino-E-o-M}.
-As mentioned previously, this may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
-This piece is purely dissipative and so signifies that flux is carried away from the system by gravitational waves.
+Beyond geodesic motion in the Kerr spacetime, a test particle follows an accelerated trajectory determined by \eqnref{Mino-E-o-M}.
+This may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
+This piece is purely dissipative and so signifies that flux is carried away from the system by GWs.
 
-In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform an appropriate average over the $2$-torus, given by
+In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform appropriate averages over this new $2$-torus.
+The osculating elements integration scheme involves computing derivatives with respect to coordinate time, and we shall often be interested in the averaged form of such derivatives.
+Given a function $X(\psi,\chi)$, we first transform to Mino time where the motion is biperiodic
 \begin{equation}
+\langle \dot{X}\rangle_{\psi,\,\chi} \simeq \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right..
+\end{equation}
+It is then possible to calculate
+\begin{widetext}\begin{equation}
 \label{eq:2torus-average}
-\langle X\rangle_{\psi,\,\chi} = N \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\partialdiff{\psi}{t}\right)^{-1} \left(\partialdiff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} X }{\chi} }{\psi} .
-\end{equation}
-This form of averaging is successful at describing the leading order secular evolution of the trajectory, as is illustrated in \figref{good-traj}.
+\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi},
+\end{equation}\end{widetext}
+in terms of precomputed quantities. This form of averaging is successful at describing the leading order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
 
-The effect of resonances must be completely contained within the pieces of the self-force left over after subtraction of the $2$-torus averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{\psi,\,\chi}$ for some quantity $X$ related to the self-force.
-The adiabatic evolution, which contains only a single averaged forcing term, will therefore not demonstrate any resonance behaviours\footnote{Resonances still occur as the fundamental frequencies become commensurate, but as the system passes through each resonance, there is no additional effect on the evolution of the system.}.
+The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{\psi,\,\chi}$.
+The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurence of any transient resonances.
+Resonances still occur as the fundamental frequencies become commensurate, but as the system passes through each resonance, there is no additional effect on the evolution of the system.
+
 One method of studying the effect of resonances is then to compare an adiabatic evolution with an instantaneous evolution, using the full knowledge of the self-force.
-This may be a useful exercise to calibrate and verify adiabatic models that can be generated cheaply, however it cannot be used to distinguish between dephasing effects arising from resonances and those arising from conservative corrections to the phase variables, as these are both explicitly eliminated from the adiabatic prescription.
+This may be a useful exercise to calibrate and verify adiabatic models that can be generated cheaply; however it cannot be used to distinguish between dephasing effects arising from resonances and those arising from conservative corrections to the phase variables, as both are explicitly eliminated from the adiabatic prescription.
 We therefore take a different approach, comparing the instantaneous evolution with a quasi-adiabatic evolution, generated by averaging the full available self-force over the $2$-torus, rather than just averaging the term $G_\alpha^{(1)}$.
 
-To be specific, we write \ref{eq:Mino-E-o-M} as
+To be specific, we write \eqnref{Mino-E-o-M} as
 \begin{subequations}
@@
 where $g_\alpha$ and $G_\alpha$ encapsulate all our knowledge about the self-force and may contain contributions from sub-leading terms in the mass-ratio expansion.
-The quasi-adiabatic evolution is then obtained by dropping the terms proportional to $\hat{g}_\alpha$ and $\hat{G}_\alpha$.
-For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an "adiabatic evolution" in fact refer to a "quasi-adiabatic evolution".
+The quasi-adiabatic evolution is then obtained by dropping the terms containing $\hat{g}_\alpha$ and $\hat{G}_\alpha$.
+For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an ``adiabatic evolution'' in fact refer to a ``quasi-adiabatic evolution''.
 
@@
 It is therefore instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi$, which we do for our illustrative system parameters.
-Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqref{eq:delta-I-a}.
+Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqnref{delta-I-a}.
 
@@
 We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$.
-The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqref{eq:delta-I-a} and giving us confidence in our numerical jump calculations.
+The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
 We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}); they are highly correlated to the magnitude of the jump in $E$, due to the dependence on $u_t$.
@@
 \caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$.
-The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqref{eq:res-jump-ratio}.
+The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqnref{res-jump-ratio}.
 The individual jumps as well as a sinusoidal fit are plotted for the three orbital shape parameters (solid lines) and the three orbital frequencies (dotted).}
@@
 \includegraphics[width=\textwidth]{Fig_overlap_vs_resjump_ill}
-\caption{\label{fig:overlap-vs-resjump-ill}The overlap \eqref{eq:overlap} between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the resonance jumps in $E$ (left), $L_z$ (middle), and $Q$ (right).
+\caption{\label{fig:overlap-vs-resjump-ill}The overlap \eqnref{overlap} between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the resonance jumps in $E$ (left), $L_z$ (middle), and $Q$ (right).
 Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
@@
 In such scenarios, adiabatic models can still be used, but there must be sufficient SNR in each non-resonant region that they can be identified as individual sources and then joined together as one long inspiral.
-Otherwise, if the phase difference on resonance can be predicted, \eqref{eq:delta-I-a} can be used to join together different adiabatic models.
+Otherwise, if the phase difference on resonance can be predicted, \eqnref{delta-I-a} can be used to join together different adiabatic models.
 

### 2014-08-06 res_char.tex
adiabatic averaging extra eq
shortened waveform model section

@@
 \subsection{Adiabatic evolution}
+\label{sec:adiabatic}
 
@@
 \begin{equation}
-\langle \dot{X}\rangle_{\psi,\,\chi} \simeq \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right..
+\left\langle \dot{X}\right\rangle_{\psi,\,\chi} \simeq \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right..
 \end{equation}
@@
 \left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi},
+\end{equation}
+\begin{equation}
+\left\langle \dot{X}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-2} \dot{X} }{\chi} }{\psi},
 \end{equation}\end{widetext}
-in terms of precomputed quantities. This form of averaging is successful at describing the leading order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
+in terms of precomputed quantities \cite{Drasco2005}. This form of averaging is successful at describing the leading order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
 
@@
 Similar resonance surfaces can be found for other spin values and for other resonances.
-When considering the full parameter set of $\{p,e,\iota,a_\ast,\nu\}$ it is apparent that the search for resonances becomes expensive as a consequence of the dimensionality.
+When considering the full parameter set of $\{p,e,\iota,a_\ast,\nu\}$, it is apparent that the search for resonances becomes expensive as a consequence of the dimensionality.
 It is therefore useful to have a guide of where to look.
@@
 The expressions become complicated when venturing away from limiting cases.
-Instead, we build an approximate phenomenological model and fit this to the resonance plane.%\footnote{A stopped clock is precisely correct twice a day, while a clock that is five minutes slow is never right but will often be close enough.
+Instead, we build an approximate phenomenological model and fit this to the resonance plane.
+%\footnote{A stopped clock is precisely correct twice a day, while a clock that is five minutes slow is never right but will often be close enough.
 Within this metaphor, our approximation would be a clock that varies between telling the correct time and being fifteen minutes off throughout the day: it is generally going to be good enough to give you a sense of the time but if you have an important event you should check with something better.}
 This should be useful for designating the region in which resonance could be expected.
-To locate them precisely it is necessary to solve $\Omega = 0$ numerically; the approximate model should give a suitable starting point.
+To locate them precisely, it is necessary to solve $\Omega = 0$ numerically; the approximate model should give a suitable starting point.
 
@@
 \begin{equation}
-\tau\sub{res} = \left[\frac{2\pi}{\left|\langle\dot{\Omega}(0)\rangle_{q'}\right|}\right]^{1/2}.
+\tau\sub{res} = \left[\frac{2\pi}{\left|\left\langle\dot{\Omega}(0)\right\rangle_{q'}\right|}\right]^{1/2}.
 \label{eq:t-res}
@@
 \begin{equation}
-\left|\langle\dot{\Omega}\rangle_{q'}\right| \simeq \left|\frac{\Omega(t)}{t}\right|.
+\left|\left\langle\dot{\Omega}\right\rangle_{q'}\right| \simeq \left|\frac{\Omega(t)}{t}\right|.
 \end{equation}
@@
 \begin{equation}
-\varphi_{n_r,\,-n_\theta} \simeq \left(n_r\Upsilon_r - n_\theta\Upsilon_\theta\right)\lambda + \left(n_r\dot{\Upsilon_r} - n_\theta\dot{\Upsilon}_\theta\right)\lambda^2 + \ldots
+\varphi_{n_r,\,-n_\theta} \simeq \left(n_r\Upsilon_r - n_\theta\Upsilon_\theta\right)\lambda + \left(n_r\dot{\Upsilon}_r - n_\theta\dot{\Upsilon}_\theta\right)\lambda^2 + \ldots
 \end{equation}
@@
 t\sub{pres} = {} & \tau\sub{pres}(t\sub{pres}) \\
- \simeq {} & \dfrac{2\pi}{\left|\langle\dot{\Omega}(0)\rangle_{q'} t\sub{pres}\right|}, \nonumber 
+ \simeq {} & \dfrac{2\pi}{\left|\left\langle\dot{\Omega}(0)\right\rangle_{q'} t\sub{pres}\right|}, \nonumber 
 \end{align}
@@
 \end{equation}
-The two time-scales are equivalent.
-We preferentially use $\tau\sub{res}$ to denote the resonance width.
+The two time-scales are equivalent: we preferentially use $\tau\sub{res}$ to denote the resonance width.
 It is shorter than the inspiral time-scale, but longer than an orbital period, $\tau\sub{res} \sim \order{\eta^{1/2}\tau\sub{ev}} \sim \order{\eta^{-1/2}T}$~\cite{Flanagan2012,Gair2011a}; it therefore acts as a bridge between the two time-scales~\cite{Hinderer2008}.
@@
 %\tau_{\mathrm{pres},\,s} = {} & \frac{2\pi}{|s\Omega(t)|};\\
-\tau_{\mathrm{res},\,s} = {} & \left[\frac{2\pi}{\left|s\langle \dot{\Omega}(0)\rangle_{q'}\right|}\right]^{1/2}.
+\tau_{\mathrm{res},\,s} = {} & \left[\frac{2\pi}{\left|s\left\langle \dot{\Omega}(0)\right\rangle_{q'}\right|}\right]^{1/2}.
 \label{eq:T-res-s}
@@
 \begin{equation}
-\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\sqrt{\dfrac{2\pi}{\left|s \langle\dot{\Omega}\rangle_{q'}\right|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right].
+\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\sqrt{\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right].
 \end{equation}
@@
 
-We shall now discuss our chosen waveform generation scheme, giving demonstrations of its accuracy for our purposes.
-We will then detail our procedure for generating a realistic distribution of detectable EMRIs, before turning to the statistical effect of resonances on the population at the end of the section.
+We discuss our chosen waveform generation scheme, giving demonstrations of its accuracy for our purposes.
+We then detail our procedure for generating a realistic distribution of detectable EMRIs, before turning to the statistical effect of resonances on the population at the end of the section.
 
@@
 
-To generate gravitational waveforms, we choose to employ the 'Numerical Kludge' (NK) method of Babak \emph{et al}~\cite{Babak2007}.
-The principle behind the class of so-called kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
-This can be much quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives very similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher~\cite{Babak2007}.
+To generate gravitational waveforms, we choose to employ the numerical kludge (NK) method of Babak \emph{et al}~\cite{Babak2007}.
+The principle behind the class of kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
+This can be much quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
 
 Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the post-Newtonian self-force from \secref{self-force}.
-We then interpret the Boyer-Lindquist coordinates as spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique.
-Specifically, we apply the standard quadrupole formula~\cite{Misner1973}.
+We then interpret the Boyer--Lindquist coordinates as spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique; specifically, we apply the standard quadrupole formula~\cite{Misner1973}.
 
@@
 
-Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s_A(t)$ and $h_A(t)$, according to
+Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s(t)$ and $h(t)$, according to
 \begin{equation}
 \label{eq:innerprod}
-\overlap{s}{h} = 2 \sum_{A=+,\times}\intd{0}{\infty}{\frac{\tilde{s}_A(f)\tilde{h}_A^*(f)+\tilde{s}_A^*(f)\tilde{h}_A(f)}{S_n(f)}}{f},
+\overlap{s}{h} = 2 \intd{0}{\infty}{\frac{\tilde{s}(f)\tilde{h}^*(f)+\tilde{s}^*(f)\tilde{h}(f)}{S_n(f)}}{f},
 \end{equation}
-where $\tilde{s}_A(f)$ represents the Fourier transform of $s_A(t)$ and similarly for $\tilde{h}_A(f)$, $A = \{+,\times\}$ denotes the plus- or cross-polarisation state of the GW and $S_n(f)$ is the one-sided noise power spectral density (PSD).
+where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD).
 We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept.
-The analytic approximation we use is~\cite{Amaro-Seoane2012a}
-\begin{widetext}\begin{equation}
-S_n(f) = \frac{20}{3}\frac{4 S_{acc}(f)/(2\pi f)^4 + S_{sn}(f) + S_{omn}(f)}{L^2} \left(1+\left(\frac{f}{0.41 c/2L}\right)\right)^2,
-\end{equation}\end{widetext}
-where $L = 1\;\mathrm{Mkm}$ is the length of the interferometer arms, $c$ is the speed of light and the components of noise separate into: the noise due to residual test mass acceleration $S_{acc}(f) = 2.13\times10^{-29}(1+10^{-4}\mathrm{Hz}/f)\;\mathrm{m^2\,s^{-4}\,Hz^{-1}}$; the shot noise $S_{sn} = 5.25\times10^{-23}\;\mathrm{m^2\,Hz^{-1}}$; and all other measurement noises $S_{omn} = 6.28\times10^{-23}\;\mathrm{m^2\,Hz^{-1}}$.
-We show a plot of this PSD against frequency in \figref{noise-curves}, along with the corresponding curve for the LISA detector for comparison.
-The effect of white dwarf (WD) confusion noise, which is important for LISA (as shown in the figure), doesn't significantly impact the eLISA mission and so we neglect it.
+We use the analytic approximation of Amaro-Seoane \emph{et al}~\cite{Amaro-Seoane2012a}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_noise_curves}
-\caption{\label{fig:noise-curves}One-sided noise power spectral densities for space-based GW detectors, eLISA and LISA. The effect of WD confusion noise is shown for the LISA curve.}
-\end{figure}
-
-In practice, the functions $s_A(t)$ and $h_A(t)$ will not be available, instead replaced by the time series $s_A(t_i)$ and $h_A(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values but which do not necessarily have the same length or spacing.
-To take this into account, when evaluating overlaps, we first project the $h_A(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme, setting any values outside the domain of $\{t_j\}$ to zero.
-A Planck-taper window function~\cite{McKechan2010} is applied to the resulting time series.
-This can be defined in a continuous form by
-\begin{equation}
-w(t) = \left\{
-\begin{array}{ll}
-	0 & t\leq t_1 \\
-	1\big/\left(1+\exp\left(\frac{t_2-t_1}{t-t_1}+\frac{t_2-t_1}{t-t_2}\right)\right) & t_1 < t < t_2 \\
-	1 & t_2\leq t\leq t_3 \\
-	1\big/\left(1+\exp\left(\frac{t_3-t_4}{t-t_3}+\frac{t_3-t_4}{t-t_4}\right)\right) & t_3 < t < t_4 \\
-	0 & t_4\leq t,
-\end{array}\right.
-\end{equation}
-where $t_1$ and $t_4$ are chosen to encompass the whole waveform and $t_2$ and $t_3$ are the times of the second closest stationary points to each end.
-We demonstrate the application of this window function to a simple sinusoid in \figref{window-func}, with a comparison to the uniform window, which is unity for $t_1\leq t\leq t_4$ and zero elsewhere.
-The use of the second stationary points from each end in the Planck-taper ensures that the transition is not too sharp; the smoothness of the waveform at each end reduces unwanted spectral leakage in the Fourier transforms.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_window_func}
-\caption{\label{fig:window-func}Two window functions applied to a simple sinusoid: the uniform (dashed) and the Planck-taper (solid).
-The Planck-tapered sinusoid has no discontinuities at the endpoints.}
-\end{figure}
+In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
+To take this into account, when evaluating overlaps, we first project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme, setting any values outside the domain of $\{t_j\}$ to zero.
+A Planck-taper window function~\cite{McKechan2010} is applied to the resulting time series to reduce unwanted spectral leakage in the Fourier transforms.
 
@@
 
-One of the targets of pre-eLISA research is to generate a bank of waveform templates, $h_A(t;\lambda^a)$ described by some set of parameters $\lambda^a$.
-Once real-world data, $s_A(t)$, has been collected, we can then search over the template bank to find the model that most closely matches the data.
+One of the targets of pre-eLISA research is to generate a bank of waveform templates, $h(t;\Theta^a)$ described by some set of parameters $\Theta^a$.
+Once real-world data, $s(t)$, has been collected, we can then search over the template bank to find the model that most closely matches the data.
 At the current time, with a lack of experimental data, we are interested in ensuring that the waveform templates accurately reproduce what we expect to observe in nature, without being too computationally expensive.
@@
 Ideally, we would like to use waveform templates from EMRI systems that are evolved under the entire self-force model, but these are too expensive.
-In practice, we are therefore forced to use cheaper adiabatic waveforms, but these do not include the effect of resonances.
-To assess the impact of this, we compare data $s_A(t;\bar{\lambda}^a)$ generated using the full self-force model to templates $h_A(t;\lambda^a)$ generated using the expensive 2-torus averaged self-force model.
-As discussed before, this eliminates any other effects and allows us to focus on the resonance problem.
+In practice, we are forced to use cheaper adiabatic waveforms, but these do not include the effect of resonances.
+To assess the impact of this, we compare data $s(t;\bar{\Theta}^a)$ generated using the full self-force model to templates $h(t;\Theta^a)$ generated using the expensive 2-torus averaged self-force model (see \secref{adiabatic}).
 
-We wish to test whether there exists some set of parameters $\lambda^a$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
+We wish to test whether there exists some set of parameters $\Theta^a$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
 We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template, defined by
 \begin{equation}
-\frac{S}{N}\left[h\right] = \frac{\langle\overlap{s}{h}\rangle_n}{\sqrt{\left\langle\overlap{n}{h}^2\right\rangle_n}} = \frac{\overlap{s}{h}}{\sqrt{\overlap{h}{h}}},
+\rho\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{h}{h}}}.
 \end{equation}
-where $\langle\rangle_n$ represents an average over all possible noise realisations. If a template were to exactly match the data, it would produce an SNR of $\sqrt{\overlap{s}{s}}$ and so the overlap
+If a template were to exactly match the data, it would produce an SNR of $\sqrt{\overlap{s}{s}}$ and so the overlap
 \begin{equation}
@@
 provides an indication of how well-matched the template is to the data.
-We leave an investigation into parameter estimation biases (where the parameters of the best-matching adiabatic, $\lambda^a$, do not agree with the parameters of the system, $\bar{\lambda}^a$) to future work.
+We leave an investigation into parameter estimation biases (where the parameters of the best-matching adiabatic $\Theta^a$ do not agree with the parameters of the system $\bar{\Theta}^a$) to future work.
 

### 2014-08-06 asymptotic.tex
more small changes

@@
 
-The following derivation is based upon the analysis of Kevorkian~\cite{Kevorkian1987}.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} Small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.
+The following derivation is reproduced from Berry~\cite{BerryThesis2013} and is based upon the analysis of Kevorkian~\cite{Kevorkian1987}.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} Small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.
+A similar derivation can be found by van de Meent~\cite{VanDeMeent2014}.
 
@@
 We are only concerned with the $r$ and $\theta$ motions, so we have a $2$-dimensional system.
-We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992}.
+We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992,VanDeMeent2014}.
 This becomes one of the new angle variables, the other variable $q'$ can be either $q_r$ or $q_\theta$ (as, on resonance, varying one necessarily changes the other).
@@
 In the limit of $\eta \rightarrow 0$ we are left with a constant frequency $\omega_0(0)$.
-The higher order terms are identified below by matching terms in the series expansions of the equations of motion.
+The higher-order terms are identified below by matching terms in the series expansions of the equations of motion.
 Taking the two time-scales as independent, we may write the time derivative to $\order{\eta}$ as
@@
 This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest order piece of) the adiabatic solution~\cite{Hinderer2008}.
-If we similarly average \eqnref{q-1} we find
+If we similarly average \eqnref{q-1}, we find
 \begin{equation}

### 2014-08-06 res_char.tex
more small changes

@@
 
-To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs\footnote{The actual distribution in parameter space of observable EMRI systems is unknown, but we can make reasonable estimates.} and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.
+To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.
 
@@
 \includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
-\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, chosen to give 25 independent (non-overlapping) calculations.
+\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give 25 independent (non-overlapping) calculations.
 The time axis is centred on the value at which the full evolution crosses the 2:3 resonance.}
@@
 The dephasing means that an adiabatic evolution with the same initial conditions as a full evolution is likely to have a low overlap if a significant resonance is encountered.
-However, this does not necessarily mean that all adiabatic evolutions will have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.
+However, this does not necessarily mean that all adiabatic evolutions have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.
 
@@
 The most restrictive aspect of this approach is that all extrinsic parameters and any intrinsic ones that remain constant throughout the evolution, such as the masses of the BHs and the central BH spin, do not get explored by our family of adiabatic models.
-It is possible that small deviations from these true parameters will give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
+It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
 Indeed, systematic biases are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.
@@
 The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
-Anticipating that resonances will play a key role in determining the degree of waveform overlap, we choose matching times situated at $5$ resonance-timescales after each resonance of interest, namely the low-order 2:3, 1:2, 2:5 and 1:3.
+Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau_\mathrm{res}$ after each resonance of interest, namely the low-order 2:3, 1:2, 2:5 and 1:3.
 These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral.
@@
 The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
-We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}); they are highly correlated to the magnitude of the jump in $E$, due to the dependence on $u_t$.
+We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).
 
@@
 Strong resonances can severely limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic modes, which may be individually undetectable.
-To assess the impact of this on future gravitational wave missions, we need to create a sample population of detectable EMRIs.
+To assess the impact of this on future GW missions, we need to create a sample population of detectable EMRIs.
 
-We first identify 14 parameter distributions that describe an astrophysically typical EMRI.
+We first identify 14 parameter distributions that describe an astrophysically typical EMRI\footnote{The actual distribution in parameter space of observable EMRI systems is unknown, but we can make reasonable estimates.}.
 We choose the mass of the compact object $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs.
-We take the central object to be a typical supermassive BH at the centre of a galaxy, with mass $M$ uniformly distributed in log-space between $10^5$ and $10^7 M_\odot$, and dimensionless spin $a$ uniformly between its limiting values, $0$ and $1$.
+We take the central object to be a typical supermassive BH at the centre of a galaxy, with mass $M$ uniformly distributed in log-space between $10^5 M_\odot$ and $10^7 M_\odot$, and dimensionless spin $a_\ast$ uniformly between its limiting values, $0$ and $1$.
 The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere.
@@
 To assess this, we evolve the systems forward in time using the analytic kludge method of Barack and Cutler \cite{Barack2004} until they reach the last stable orbit.
-We then calculate the Peters-Matthews waveform \cite{Peters1963} for the last $t_\mathrm{insp}$ years of the inspiral, where $t_\mathrm{insp}$ is chosen uniformly between $0.5$ and $2$.
+We then calculate the Peters-Matthews waveform \cite{Peters1963} for a time $t_\mathrm{insp}$ at the end of the inspiral, where $t_\mathrm{insp}$ is chosen uniformly between $0.5$ years and $2$ years.
 We focus on the late stages of inspiral as these are most likely to be detected by eLISA.
@@
 \begin{equation}
-\overlap{s}{h} \approx 4 \sum_{A=+,\times}\intd{0}{T}{\frac{s_A(t)h_A(t)}{S_n(f(t))}}{t},
+\overlap{s}{h} \approx 4 \intd{0}{t_\mathrm{insp}}{\frac{s(t)h(t)}{S_n(f(t))}}{t},
 \end{equation}
@@
 
-To obtain our final distribution of detectable EMRIs, shown as dark blue in \figref{EMRIpar-dists}, we quote the parameter values $t_\mathrm{insp}$ years before plunge, including only those systems with an estimated SNR above 10.
+To obtain our final distribution of detectable EMRIs, shown as dark blue in \figref{EMRIpar-dists}, we quote the parameter values $t_\mathrm{insp}$ before plunge.
+Only systems with $\rho > 10$ are treated as detectable.
 Also shown in the figure, in light pink, are the systems with an estimated SNR below 10, to illustrate which systems tend to have lower SNRs.
@@
 
-We note that prograde orbits (those with $\cos\iota > 0$) and systems with higher BH spins tend to be favoured; this is likely due to the fact that such systems can inspiral much closer before finally plunging and so experience the strongest gravitational fields.
+We note that prograde orbits (those with $\cos\iota > 0$) and systems with higher BH spins tend to be favoured; this is likely due to the fact that such systems can inspiral much closer before finally plunging, and so experience the strongest gravitational fields.
 This effect is amplified by the fact that the majority of the SNR is recovered from the final part of the inspiral, which also explains why larger values of $t_\mathrm{insp}$ are only marginally preferred over smaller values.
@@
 Systems at higher redshifts start to become undetectable as the amplitude of the signal scales inversely proportional to the luminosity distance to the source.
-At $z=0.7$, the fraction of detectable systems drops to $5\%$; this is chosen as the upper cutoff in the redshift distribution as a compromise between including as many detectable systems as possible, but without having too low an overall acceptance rate.
-We quote the masses in the system as "redshifted masses" to account for the FRW background metric.
+At $z=0.7$, the fraction of detectable systems drops to $5\%$; this is chosen as the upper limit in the redshift distribution to minimise unneccessary computations.
+We quote the masses in the system as redshifted masses, which accounts for the frequency shift due to the expansion of the Universe.
 
@@
 \begin{enumerate}
-\item No significant resonance jumps are observed, and all the adiabatic models agree with the full evolution reasonably well.
-The overlaps tend to be noticeably lower than $1$ since the adiabatic approach breaks down near the end of the inspiral.
+\item No significant resonance jumps are observed, and all the adiabatic models agree with the full evolution reasonably well: the overlaps are typically larger than $0.9$.
+The models have overlaps less than $1$ since the adiabatic approach breaks down near the end of the inspiral.
 
@@
 
-To ensure that our chosen adiabatics are reasonable estimates to maximise the overlap, we computed many adiabatic waveforms on a rigid grid for one particular system; there can be large fluctuations in the overlap, but our chosen parameters perform at least averagely well.
+To ensure that our chosen adiabatics are reasonable estimates to maximise the overlap, we computed many adiabatic waveforms with regular values of $t_\mathrm{match}$ for one particular system; there can be large fluctuations in the overlap, but our chosen parameters perform at least averagely well.
 
@@
 There is no observed trend with the size of the resonance jump, although this is not surprising given the small size of the jumps.
-The large scatter on the plot is due to our choice of adiabatic family; in each case, the quoted loss of SNR is a lower bound for the actual loss of SNR across the entire adiabatic model space.
+The large scatter on the plot is due to our choice of adiabatic family; in each case, the quoted fractional SNR is a lower bound for the actual fractional SNR across the entire adiabatic model space.
 
@@
 It is possible for EMRIs to form through alternative channels, which may result in a population of high eccentricity systems.
-For example, 'plunging' compact objects can be mistaken for slow inspirals since they still undergo many hundreds of thousands of cycles in the bandwidth of a detector like eLISA \cite{Amaro-Seoane2013}.
-This type of system is expected to be orders of magnitude more common than classic EMRIs.
+As an example, compact objects can be scattered onto orbits that plunge directly into the central BH; these systems can be mistaken for slow inspirals since they still undergo many hundreds of thousands of cycles in the bandwidth of a detector like eLISA \cite{Amaro-Seoane2013}.
+This kind of event is expected to be orders of magnitude more common than classic EMRIs.
 It is expected that resonance behaviours would be much more important in this case.
@@
 \caption{\label{fig:SNRcutoff}The fraction of EMRI systems that would be lost using adiabatic models to analyse data that contain the effects of transient resonances.
-The error bars are estimated using the beta distribution.}
+The error bars are estimated using the beta distribution \cite{Cameron2011}.}
 \end{figure}

### 2014-08-06 asymptotic.tex
Small tweaks

@@
 
-The following derivation is reproduced from Berry~\cite{BerryThesis2013} and is based upon the analysis of Kevorkian~\cite{Kevorkian1987}.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} Small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.
-A similar derivation can be found by van de Meent~\cite{VanDeMeent2014}.
+The following derivation is reproduced from Berry~\cite{BerryThesis2013}.
+It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found by van de Meent~\cite{VanDeMeent2014}.
 
@@
 \begin{align}
-\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\sqrt{\dfrac{2\pi}{\left|s \langle\dot{\Omega}\rangle_{q'}\right|}} \nonumber \\*
+\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
  {} & \times \left.\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\dot{\Omega}\right)\right]\right. \\

### 2014-08-07 res_char.tex
larger plot labels
Small tweaks

@@
-\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,reprint,showpacs,superscriptaddress,twocolumn]{revtex4}
+\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,showpacs,superscriptaddress,twocolumn]{revtex4}
 \usepackage{graphicx}
@@
 \email[]{cplb@star.sr.bham.ac.uk}
+\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
 \affiliation{School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
@@
 \begin{equation}
-\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\sqrt{\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right].
+\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right].
 \end{equation}
@@
 \centering
-\includegraphics[width=0.92\textwidth]{Fig_good_traj}
+\includegraphics[width=\textwidth]{Fig_good_traj}
 \caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) under the full (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances.
@@
 \includegraphics[width=\textwidth]{Fig_EMRIpar_dists}
-\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $50,000$ realisations.
+\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $\mathcal{O}(10^4)$ realisations.
 The (darker) blue bars represent the final distribution, including an SNR cut of 10.

### 2014-08-07 res_char.tex
adiabatic evolution clean-up

@@
 The osculating elements integration scheme involves computing derivatives with respect to coordinate time, and we shall often be interested in the averaged form of such derivatives.
-Given a function $X(\psi,\chi)$, we first transform to Mino time where the motion is biperiodic
-\begin{equation}
-\left\langle \dot{X}\right\rangle_{\psi,\,\chi} \simeq \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right..
-\end{equation}
-It is then possible to calculate
+Given a function $X(\psi,\chi)$, we can calculate the 2-torus average \cite{Drasco2005} of the derivative with respect to Mino time in terms of precomputed quantities
 \begin{widetext}\begin{equation}
 \label{eq:2torus-average}
-\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi},
+\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi}.
+\end{equation}\end{widetext}
+We then obtain a coordinate time derivative
+\begin{equation}
+\dot{\left\langle X\right\rangle}_{\psi,\,\chi} = \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right.,
 \end{equation}
-\begin{equation}
-\left\langle \dot{X}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-2} \dot{X} }{\chi} }{\psi},
-\end{equation}\end{widetext}
-in terms of precomputed quantities \cite{Drasco2005}. This form of averaging is successful at describing the leading order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
+which is successful at describing the leading order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
 

### 2014-08-07 res_char.tex
marginally improved summaries and conclusions

@@
 
+The combination of a full instantaneous evolution and an adiabatic evolution allows us to systematically study the effect of transient resonances on EMRIs over the course of an inspiral.
+Before approaching this problem, we first investigate the properties of the resonances themselves.
+
 \section{Properties of Transient Resonances}
@@
 
-The first step in studying the effect of transient resonance is to locate orbital parameters for which the frequencies are commensurate.
+The first step in studying the effect of transient resonances is to locate orbital parameters for which the frequencies are commensurate.
 We can calculate the frequencies and so we are left with the problem of solving $\Omega = n_r \Omega_r - n_\theta \Omega_\theta = 0$ numerically.
@@
 exactly as expected.
+
+Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on a population of inspirals.
 
@@
 Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral.
+We have presented an evolution scheme appropriate for studying the effects of resonances on inspirals, which has been applied to an astrophysically realistic population.
+The location and duration of resonances, as well as their effect on orbital parameters, can all be calculated.
+
+Amongst a population of sources, unmodelled resonances could diminish detection prospects.
 However, due to the low eccentricity of detectable sources, the overall effect on SNR recovery is small; it is therefore likely to be sufficient to use adiabatic models to detect EMRIs.
-An unresolved question here is how these adiabatic models affect parameter estimation: systematic biases may be introduced if we neglect the presence of important resonances.
+An unresolved question here is how using these adiabatic models would affect parameter estimation: systematic biases may be introduced if we neglect the presence of important resonances.
+One method of providing an answer to this problem is to perform an MCMC parameter estimation study using our adiabatic waveforms as templates; we leave this to future work.
 
-High eccentricity EMRIs may also be possible, in which case transient resonances will play an important role in their detectability; systems that undergo a jump of a few percent dephase very rapidly from the adiabatic approximation.
-In such scenarios, adiabatic models can still be used, but there must be sufficient SNR in each non-resonant region that they can be identified as individual sources and then joined together as one long inspiral.
-Otherwise, if the phase difference on resonance can be predicted, \eqnref{delta-I-a} can be used to join together different adiabatic models.
+High eccentricity EMRIs may be possible, in which case transient resonances will play an important role in their detectability; systems that undergo a jump of a few percent dephase very rapidly from the adiabatic approximation.
+In such scenarios, adiabatic models can still be used, but there must be sufficient SNR in each non-resonant region that they can be detected as individual sources and then identified as arising from the same inspiral.
+If this is not possible, \eqnref{delta-I-a} can be used to join together different adiabatic models, if the phase on resonance and the magnitude of the self-force can be predicted sufficiently accurately.
+Alternatively, the jump $\Delta\mathcal{I}_\mathrm{jump}$ for each orbital parameter could be included as a free parameter in the template space, and then searched over.
+Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.
 
 This work is highly dependent on the chosen self-force model, and so the results should be taken as qualitative estimates, given the unsuitability of a PN self-force to the problem of fast motion near to highly-spinning BHs.
-In the future, with a more accurate self-force model, more quantitative results can be derived.
+In the future, with a more accurate self-force model, more exact quantitative results can be derived.
 

### 2014-08-08 res_char.tex
Premerge

--- 2014-08-07 res_char.tex
@@
 
-\title{Characterisation of transient resonances in extreme-mass-ratio-inspirals}
+\title{Characterisation of transient resonances in extreme-mass-ratio inspirals}
 
@@
 During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact-object trajectory and, ultimately, the phase of the emitted radiation.
-In this work, an astrophysical population of detectable extreme-mass-ratio-inspirals is studied and it is found that a large proportion shall encounter a low-order resonance in the later stages of inspiral.
+In this work, an astrophysical population of detectable extreme-mass-ratio inspirals is studied and it is found that a large proportion shall encounter a low-order resonance in the later stages of inspiral.
 The resulting effect on signal-to-noise recovery is small due to the low eccentricity of the studied population, but nevertheless at least one fifth of the systems may become undetectable if transient resonances aren't accounted for.
@@
 Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We shall consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.}, and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
+We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.}, and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
 We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
@@
 \end{subequations}
-As an affine parameter we have used Mino time which is related to the proper time $\tau$ by~\cite{Mino2003}
+As an affine parameter, we have used Mino time which is related to the proper time $\tau$ by~\cite{Mino2003}
 \begin{equation}
@@
 \end{equation}
-we can calculate the evolution of the osculating elements $\dot{I}^A$. The specific equations for motion in Kerr are derived by Gair et al \cite{Gair2011a}.
+we can calculate the evolution of the osculating elements $\dot{I}^A$. The specific equations for motion in Kerr are derived by Gair {\it{et al}}.~\cite{Gair2011a}.
 
@@
 
-To generate gravitational waveforms, we choose to employ the numerical kludge (NK) method of Babak \emph{et al}~\cite{Babak2007}.
+To generate gravitational waveforms, we choose to employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}.
 The principle behind the class of kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
@@
 We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept.
-We use the analytic approximation of Amaro-Seoane \emph{et al}~\cite{Amaro-Seoane2012a}
+We use the analytic approximation of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}
 
@@
 \includegraphics[width=\textwidth]{Fig_EMRIpar_dists}
-\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $\mathcal{O}(10^4)$ realisations.
+\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $50,000$ realisations.
 The (darker) blue bars represent the final distribution, including an SNR cut of 10.

### 2014-08-08 res_char.tex
Attmept to clear up the averaging

@@
 Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.}, and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
+We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
 We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
@@
 This may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
-This piece is purely dissipative and so signifies that flux is carried away from the system by GWs.
+This piece is purely dissipative and determines how the inspiral evolves due to the radiation of GWs.
 
-In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform appropriate averages over this new $2$-torus.
+To construct an adiabatic trajectory we need the $2$-torus-averaged fluxes of our osculating elements.
+Computing an average of a quantity over the $\{q_r, q_\theta\}$ is trivial if it is parametrized in terms of these variables,
+\begin{equation}
+\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta} = \frac{1}{(2\pi)^2}\intd{0}{2\pi}{\intd{0}{2\pi}{\diff{X}{\lambda}}{q_r}}{q_\theta}.
+\end{equation}
+However, we are using $\psi$ and $\chi$, as these are simpler to evolve; furthermore, we compute instantaneous co-ordinate time fluxes $\dot{X}$, not Mino-time fluxes.
+Changing variables gives an average of \cite{Drasco2005}
+\begin{widetext}\begin{align}
+\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta} = {} & \frac{1}{\Lambda_r \Lambda_\theta}\intd{0}{2\pi}{\intd{0}{2\pi}{\left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-2} \diff{X}{\lambda}}{\psi}}{\chi} \\
+ = {} & \frac{1}{\Lambda_r \Lambda_\theta}\intd{0}{2\pi}{\intd{0}{2\pi}{\left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X}}{\psi}}{\chi}.
+\label{eq:2torus-average}
+\end{align}\end{widetext}
+This average describes the Mino-time rate of change of the quantity $X$ over an orbit. To convert to a co-ordinate flux of the averaged quantity, we simply divide by the period $\Gamma$, defining
+\begin{equation}
+\dot{\left\langle X\right\rangle}_{q_r,\,q_\theta} = \frac{1}{\Gamma}\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta}.
+\end{equation}
+Is in convenient to calculate $\Gamma$ as
+\begin{equation}
+\Gamma = \left\langle \diff{t}{\lambda}\right\rangle_{q_r,\,q_\theta},
+\end{equation}
+using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.
+The averaged fluxes successfully describe the leading-order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
+
+%In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform appropriate averages over this new $2$-torus.
 The osculating elements integration scheme involves computing derivatives with respect to coordinate time, and we shall often be interested in the averaged form of such derivatives.
 Given a function $X(\psi,\chi)$, we can calculate the 2-torus average \cite{Drasco2005} of the derivative with respect to Mino time in terms of precomputed quantities
-\begin{widetext}\begin{equation}
-\label{eq:2torus-average}
-\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi}.
-\end{equation}\end{widetext}
-We then obtain a coordinate time derivative
-\begin{equation}
-\dot{\left\langle X\right\rangle}_{\psi,\,\chi} = \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right.,
-\end{equation}
-which is successful at describing the leading order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
+%\begin{widetext}\begin{equation}
+%\label{eq:2torus-average}
+%\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi}.
+%\end{equation}\end{widetext}
+%We then obtain a coordinate time derivative
+%\begin{equation}
+%\dot{\left\langle X\right\rangle}_{\psi,\,\chi} = \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right.,
+%\end{equation}
+%which is successful at describing the leading-order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
 
-The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{\psi,\,\chi}$.
+The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$.
 The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurence of any transient resonances.
@@
 \label{eq:Mino-E-o-M2}
-\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{\psi,\,\chi}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
-\diff{J_\alpha}{\lambda} = {} & \eta \langle G_\alpha\rangle_{\psi,\,\chi}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
+\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
+\diff{J_\alpha}{\lambda} = {} & \eta \langle G_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
 \end{align}
@@
 \end{align}
-using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \langle\dot{\Omega}\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
+using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \left\langle\dot{\Omega}\right\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
 \begin{equation}
@@
 If we denote the orbital parameters by $\mathcal{I}_a = \{E,L_z,Q\}$, then the change across resonance is
-\begin{equation}
-\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right].
-\end{equation}
+\begin{align}
+\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
+ {} & \times {} \exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
+\end{align}
 Here $\widehat{\kappa}_0$ is the orbital phase on resonance and $F_{a,\,s}^{(1)}$ is the $s$-th harmonic of the first-order self-force on resonance, defined such that\footnote{Since the geodesic equations decouple in Mino time rather than coordinate time, this is true only in an average sense.}

### 2015-05-29 res_char.tex
post-viva, post chop house thoughts

@@
 
-\title{Characterisation of transient resonances in extreme-mass-ratio inspirals}
+\title{The importance of transient resonances in extreme-mass-ratio inspirals (new title?)}
 
@@
 \begin{abstract}
+\emph{Update abstract}
 The inspiral of stellar-mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
@@
 
+\emph{Add detail about why self-force model is rubbish. Comment on switching conservative piece off for population results.}
+
 \subsection{Adiabatic evolution}
@@
 
+\emph{Remove from here ...}
 The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$.
@@
 For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an ``adiabatic evolution'' in fact refer to a ``quasi-adiabatic evolution''.
+\emph{... to here.}
 
@@
 \label{sec:properties}
-\subsection{Location}
+
+\emph{Move location stuff to appendix ... the bits here need tweaking to reflect that.}
 
@@
 We shall attempt to build a model for the resonance plane.
-
-Brink, Geyer and Hinderer~\cite{Brink2013} provide series expansions for the location of resonances in the limit of equatorial orbits for small spin and eccentricity.
-We do not follow this approach of trying to find analytic expressions for the resonance surface.
-The expressions become complicated when venturing away from limiting cases.
-Instead, we build an approximate phenomenological model and fit this to the resonance plane.
-%\footnote{A stopped clock is precisely correct twice a day, while a clock that is five minutes slow is never right but will often be close enough.
-Within this metaphor, our approximation would be a clock that varies between telling the correct time and being fifteen minutes off throughout the day: it is generally going to be good enough to give you a sense of the time but if you have an important event you should check with something better.}
-This should be useful for designating the region in which resonance could be expected.
-To locate them precisely, it is necessary to solve $\Omega = 0$ numerically; the approximate model should give a suitable starting point.
-
-The resonant semilatus rectum for any particular spin and resonance ratio can be well approximated as
-\begin{equation}
-p(e,\iota;a_\ast,\nu) \simeq A\frac{1 + B e + D \cos\iota}{1 - C\exp(e)}.
-\end{equation}
-The coefficients $\{A,B,C,D\}$ depend upon the spin and the particular resonance; they can be approximated as
-\begin{align} 
-A(a_\ast,\nu) \simeq {} & a_0\frac{1 + a_1\nu - a_2 \nu^2 - a_3 \nu a_\ast^2}{1 + a_4\nu - (1 + a_4)\nu^2}, \\
-B(a_\ast,\nu) \simeq {} & b_0(1 - b_1\nu)\exp(-b_2\nu)(1 - b_3 a_\ast), \\
-C(a_\ast,\nu) \simeq {} & c_0, \\
-D(a_\ast,\nu) \simeq {} & d_0\left[1 - \exp(a_\ast)\right]\left[1 - d_1\exp(\nu)\right].
-\end{align}
-This gives us a total of $12$ parameters for our fit.
-Whilst this may sound large, if we were fitting an expansion to quadratic order in any combination of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
-\begin{equation}
-\begin{array}{lll}
-a_0 = 5.9854, & a_1 = 3.4116, & a_2 = 0.9253,\\
-a_3 = 0.1959, & a_4 = 4.8846, & b_0 = 0.7692,\\
-b_1 = 1.4752, & b_2 = 1.4861, & b_3 = 0.5974,\\
-c_0 = 0.02332, & d_0 = 0.7968, & d_1 = 0.3115.
-\end{array}
-\end{equation}
-These were fitted for all possible resonances with $n_r = 2$--$7$ as well as $\nu = 9/10$, $19/20$, $49/50$ and $99/100$; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
-%These were fitted for resonances with $\nu = 1/7, 1/6, 1/5, 1/4, 2/7, 1/3, 2/5, 3/7, 1/2, 4/7, 3/5, 2/3, 5/7, 3/4, 4/5, 5/6, 6/7, 9/10, 19/20, 49/50, 99/100$, with BH spins of $a_\ast = 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.93, 0.95, 0.97, 0.99, 0.999$.
-
-Using this approximation, the maximum error in $p$ for a given $a_\ast$ and $\nu$ is typically $\sim10\%$ and less that $1$ in absolute terms.
-The relative error in the semilatus rectum is illustrated in \figref{p-error}.
-\begin{figure*}[htp]
-\centering
-\centerline{\includegraphics[width=0.47\textwidth]{Fig_fit-error-max-plane}\quad\includegraphics[width=0.47\textwidth]{Fig_fit-error-RMS-plane}}
-\caption{\label{fig:p-error}Relative error in the approximate semilatus rectum compared to the accurate numerical result as a function of BH spin $a_\ast$ and resonance ratio $\nu$.
-The left panel shows the maximum relative error and the right shows the root-mean-square error; in both cases we are marginalising over eccentricity and inclination.}
-\end{figure*}
-The largest fractional error is $\sim50\%$, this is for $a\rightarrow 1$ and $\nu \rightarrow 0$, and corresponds to small $p$, such that the absolute error is still small.
-Taking the root-mean-square across $e$ and $\iota$, the fractional error for a given $a_\ast$ and $\nu$ never exceeds $9\%$ and is typically less than $4\%$.
 
@@
 
+\emph{Add in here some numerical results. Definitely plot showing resonant flux enhancement vs eccentricity. Probably not much else.}
+
 Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on a population of inspirals.
@@
 \subsection{Waveform model}
+\emph{Can probably shorten this section a bit.}
 
@@
 
+\emph{Update this to reflect thesis.}
+
 Once the trajectories have been computed using both methods of evolution, we can calculate the corresponding waveforms.
@@
 
+\emph{Update this to reflect thesis.}
+
 To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral.
@@
 \subsection{The effect of resonances: Jump sizes}
+
+\emph{Update this to reflect thesis.}
 
@@
 
+\emph{COMPLETELY REWRITE}
+
+\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix??}
+
 Strong resonances can severely limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic modes, which may be individually undetectable.
@@
 
+\emph{THIS IS ALL GARBAGE (oh :()}
+
 We take draws from the EMRI parameter distributions, using them as initial conditions for the full instantaneous evolution, from which we generate a family of adiabatic evolutions, as discussed in \secref{SNRloss}.
@@
 
+\emph{Need to reconclude}
+
 Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral.
@@
 
+\section{Location of resonances}
+Brink, Geyer and Hinderer~\cite{Brink2013} provide series expansions for the location of resonances in the limit of equatorial orbits for small spin and eccentricity.
+We do not follow this approach of trying to find analytic expressions for the resonance surface.
+The expressions become complicated when venturing away from limiting cases.
+Instead, we build an approximate phenomenological model and fit this to the resonance plane.
+%\footnote{A stopped clock is precisely correct twice a day, while a clock that is five minutes slow is never right but will often be close enough.
+Within this metaphor, our approximation would be a clock that varies between telling the correct time and being fifteen minutes off throughout the day: it is generally going to be good enough to give you a sense of the time but if you have an important event you should check with something better.}
+This should be useful for designating the region in which resonance could be expected.
+To locate them precisely, it is necessary to solve $\Omega = 0$ numerically; the approximate model should give a suitable starting point.
+
+The resonant semilatus rectum for any particular spin and resonance ratio can be well approximated as
+\begin{equation}
+p(e,\iota;a_\ast,\nu) \simeq A\frac{1 + B e + D \cos\iota}{1 - C\exp(e)}.
+\end{equation}
+The coefficients $\{A,B,C,D\}$ depend upon the spin and the particular resonance; they can be approximated as
+\begin{align} 
+A(a_\ast,\nu) \simeq {} & a_0\frac{1 + a_1\nu - a_2 \nu^2 - a_3 \nu a_\ast^2}{1 + a_4\nu - (1 + a_4)\nu^2}, \\
+B(a_\ast,\nu) \simeq {} & b_0(1 - b_1\nu)\exp(-b_2\nu)(1 - b_3 a_\ast), \\
+C(a_\ast,\nu) \simeq {} & c_0, \\
+D(a_\ast,\nu) \simeq {} & d_0\left[1 - \exp(a_\ast)\right]\left[1 - d_1\exp(\nu)\right].
+\end{align}
+This gives us a total of $12$ parameters for our fit.
+Whilst this may sound large, if we were fitting an expansion to quadratic order in any combination of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
+\begin{equation}
+\begin{array}{lll}
+a_0 = 5.9854, & a_1 = 3.4116, & a_2 = 0.9253,\\
+a_3 = 0.1959, & a_4 = 4.8846, & b_0 = 0.7692,\\
+b_1 = 1.4752, & b_2 = 1.4861, & b_3 = 0.5974,\\
+c_0 = 0.02332, & d_0 = 0.7968, & d_1 = 0.3115.
+\end{array}
+\end{equation}
+These were fitted for all possible resonances with $n_r = 2$--$7$ as well as $\nu = 9/10$, $19/20$, $49/50$ and $99/100$; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
+%These were fitted for resonances with $\nu = 1/7, 1/6, 1/5, 1/4, 2/7, 1/3, 2/5, 3/7, 1/2, 4/7, 3/5, 2/3, 5/7, 3/4, 4/5, 5/6, 6/7, 9/10, 19/20, 49/50, 99/100$, with BH spins of $a_\ast = 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.93, 0.95, 0.97, 0.99, 0.999$.
+
+Using this approximation, the maximum error in $p$ for a given $a_\ast$ and $\nu$ is typically $\sim10\%$ and less that $1$ in absolute terms.
+The relative error in the semilatus rectum is illustrated in \figref{p-error}.
+\begin{figure*}[htp]
+\centering
+\centerline{\includegraphics[width=0.47\textwidth]{Fig_fit-error-max-plane}\quad\includegraphics[width=0.47\textwidth]{Fig_fit-error-RMS-plane}}
+\caption{\label{fig:p-error}Relative error in the approximate semilatus rectum compared to the accurate numerical result as a function of BH spin $a_\ast$ and resonance ratio $\nu$.
+The left panel shows the maximum relative error and the right shows the root-mean-square error; in both cases we are marginalising over eccentricity and inclination.}
+\end{figure*}
+The largest fractional error is $\sim50\%$, this is for $a\rightarrow 1$ and $\nu \rightarrow 0$, and corresponds to small $p$, such that the absolute error is still small.
+Taking the root-mean-square across $e$ and $\iota$, the fractional error for a given $a_\ast$ and $\nu$ never exceeds $9\%$ and is typically less than $4\%$.
+
 \section{Asymptotic solution for passage through resonance}\label{sec:res-asymptotic}

### 2015-05-29 thesis_chapters/EMRIs.tex
post-viva, post chop house thoughts

\chapter[Extreme-mass-ratio inspirals (EMRIs)]{Extreme-mass-ratio inspirals}\label{ch:EMRIs}

Extreme-mass-ratio inspirals (EMRIs) are one of the main sources of GWs for space-based interferometric detectors \citep{amaro-seoane_intermediate_2007}. They are formed from binary systems where one component is significantly more massive than the other; typically this occurs at the centre of galaxies where SMBHs with masses in the range $10^4 - 10^7 M_\odot$ capture compact objects with masses of $1-10 M_\odot$ from the surrounding galactic nucleus. Due to the effects of mass segregation, heavier objects sink to lower values of the galactic potential closer to the SMBH and so are more likely to be captured \citep{freitag_stellar_2006}. This tends to favour EMRIs involving heavier stellar mass BHs, although it is possible for EMRIs to also form from NSs and WDs. Non-compact objects are tidally disrupted by the SMBH early in their evolution and so do not generate detectable emission, other than from the centre of our own galaxy \citep{freitag_gravitational_2003}.

The extreme mass ratio between the two components in the binary and the compact nature of the smaller object allow for EMRIs to be accurately modelled as a point particle moving in the Kerr spacetime, which we now discuss.

\section{Kerr spacetime}
The Kerr solution, given by \eqnref{kerr-metric}, is extensively used to describe the spacetime around astrophysical BHs, which are generically expected to possess angular momentum due to accretion from their surroundings.

\subsection{Geodesic motion}
\label{sec:kerr-geodesic}
It is well-known that test particles move along geodesics of the background spacetime. As a zeroth-order attempt at describing the behaviour of an EMRI, it is instructive to consider solutions to the geodesic equations
\begin{equation}
\label{eq:geodesic-equation}
\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = 0,
\end{equation}
where $x^\alpha$ denotes the worldline of a particle, $\tau$ is its proper time, and $\Gamma^\alpha_{\beta\gamma}$ are the connection coefficients, given in terms of the metric by \eqnref{christoffel-symbols}.

It is desirable to find first integrals of \eqnref{geodesic-equation} to obtain expressions for the components of the 4-velocity $u^\alpha$ in terms of conserved quantities. For the Kerr metric, there exist a complete set of constants of the motion: the mass of the test particle, which we henceforth take to be unity; the energy $E = -u_t$ and angular momentum $L_z = u_\phi$, which are conserved due to the isometries of the metric in the coordinates $t$ and $\phi$; and the Carter constant $Q$, which is related to the separability of the motion in $r$ and $\theta$. The geodesic equations can then be written in Boyer--Lindquist coordinates as \citep{schmidt_celestial_2002}
\begin{subequations}\label{eq:kerr-geodesics}\begin{align}
\label{eq:kerr-geodesics-t}
\Sigma \diff{t}{\tau} &= \frac{r^2 + a^2}{\Delta}\mathcal{T} - a\left(aE\sin^2\theta-L_z\right),\\
\left(\Sigma \diff{r}{\tau}\right)^2 &= R(r),\\
\left(\Sigma \diff{\theta}{\tau}\right)^2 &= \Theta(\theta),\\
\Sigma \diff{\phi}{\tau} &= \frac{a}{\Delta}\mathcal{T} - aE + \frac{L_z}{\sin^2\theta},
\end{align}\end{subequations}
where the coordinates have been appropriately normalised by the mass of the BH,\footnote{For example, we can define a dimensionless radial Boyer--Lindquist coordinate $\tilde{r}=r/M$ where $M$ is the mass of the BH, and then relabel $\tilde{r} \rightarrow r$. Following this process, the equations are found to be independent of $M$.} $\Delta = r^2 - 2 r + a^2$, $\Sigma = r^2 + a^2 \cos^2\theta$, and we have introduced the potentials
\begin{align}
\mathcal{T} &= E(r^2+a^2)-aL_z,\\
R(r) &= \mathcal{T}^2 - \Delta\left(r^2+(L_z-aE)^2+Q\right),\\
\Theta(\theta) &= Q - \left((1-E^2)a^2 + \frac{L_z^2}{\sin^2\theta}\right)\cos^2\theta.
\end{align}

Bound orbits in the Kerr spacetime occur between the two largest roots of $R(r)$, the periapsis $r_p$ and apoapsis $r_a$, such that $r_p \leq r \leq r_a$. By analogy with Keplerian orbits, it is useful to parameterise the radial motion using
\begin{equation}
r = \frac{p}{1+e \cos\psi},
\end{equation}
where $p$ is the (dimensionless) semi-latus rectum, $e$ is the eccentricity and $\psi$ is the relativistic anomaly, which increases secularly by $2\pi$ as $r$ goes through a complete cycle. Similarly, the polar coordinate oscillates about the equatorial plane, in the range $\theta_\mathrm{inc} \leq \theta \leq \pi - \theta_\mathrm{inc}$, where the inclination angle $\theta_\mathrm{inc}$ is the smallest real root of $\Theta(\theta)$. We parameterise this using
\begin{equation}
\cos\theta = \cos\theta_\mathrm{inc} \cos\chi,
\end{equation}
where $\chi$ is an auxilliary phase variable. By considering the roots of $R(r)$ and $\Theta(\theta)$, it is possible to convert between the two parameterisations $\{E,\,L_z,\,Q\}$ and $\{p,\,e,\,\theta_\mathrm{inc}\}$, which can both be used to describe the shape of the geodesic orbit; it is easiest to calculate the transformations numerically. We shall also make use of an alternative parameterisation, introducing the inclination $\iota$ to replace $\theta_\mathrm{inc}$. It is defined according to
\begin{equation}
\tan \iota = \frac{\sqrt{Q}}{L_z}.
\end{equation}
Prograde (retrograde) equatorial orbits have $\cos\iota = (-)1$, while polar orbits have $\cos\iota = 0$.

%Discussion of last stable orbit here? Perhaps show a plot demonstrating the separatrix in Kerr?

The form of the geodesic equations (\ref{eq:kerr-geodesics}), as a function of proper time, couples the radial and polar motions through $\Sigma$. However, there exists a time coordinate $\lambda$, named after \citet{mino_perturbative_2003}, such that
\begin{equation}
\dd \lambda = \frac1{\Sigma} \dd \tau,
\end{equation}
and which decouples the motions. It is straightforward to calculate the Mino-time periods of the motion
\begin{subequations}\begin{align}
\Lambda_r &= 2 \intd{r_p}{r_a}{\frac1{\sqrt{R}}}{r} = \intd{-\pi}{\pi}{\diff{\lambda}{\psi}}{\psi},\\
\Lambda_\theta &= 4 \intd{\theta_\mathrm{inc}}{\pi/2}{\frac1{\sqrt{\Theta}}}{\theta} = \intd{-\pi}{\pi}{\diff{\lambda}{\chi}}{\chi},
\end{align}\end{subequations}
with corresponding frequencies
\begin{equation}
\Upsilon_{r,\theta} = \frac{2\pi}{\Lambda_{r,\theta}}.
\end{equation}
Any dynamical quantity $q$ that depends only on the values of $r$ and $\theta$ along the worldline may be expanded as a Fourier series
\begin{equation}
q[r(\lambda),\theta(\lambda)] = \sum_{n,m} q_{nm} \exp[-\iu(n\Upsilon_r+m\Upsilon_\theta)\lambda].
\end{equation}
In particular, the worldline coordinates $t$ and $\phi$ are such quantities, and so we may write down expansions for their Mino-time rates of change $\dd t/\dd\lambda$ and $\dd \phi / \dd\lambda$. Their average secular rates of increase, which we shall denote by $\Gamma$ and $\Upsilon_\phi$ respectively, may then be computed by taking the $n=m=0$ component of these expansions \citep{drasco_rotating_2004}.

It is not immediately obvious how to calculate coordinate-time frequencies,\footnote{These are desirable quantities to compute because they correspond to the frequencies as measured by an observer at infinity.} $\Omega_k$ ($k=\{r,\,\theta,\,\phi\}$), in an analagous manner to $\Upsilon_k$. This is because the elapsed coordinate time is different for different cycles of the radial or polar motion. However, with the notion of the average rate of coordinate time increase $\Gamma$, it is possible to define suitable quantities according to
\begin{equation}
\Omega_k = \frac{\Upsilon_k}{\Gamma}.
\end{equation}
These are consistent with results derived by \citet{schmidt_celestial_2002}, who showed that it is possible to calculate the fundamental\footnote{The term fundamental refers to the fact that these quantities demonstrate properties of the orbital motion that are independent of the coordinate representation.} proper-time frequencies, $\omega_k$, of the system using action-angle variables \citep{goldstein_classical_2001}. These are a particular set of generalised angle coordinates $q_\alpha = \{q_t,\,q_r,\,q_\theta,\,q_\phi\}$ and corresponding actions $J_\alpha$ such that the Hamiltonian of the system $H(q,J) = H(J)$ does not depend on the angle variables. According to Hamilton's equations, the system then evolves following
\begin{equation}
\label{eq:EMRI-AAgeo}
\diff{q_\alpha}{\tau} = \omega_\alpha(\boldsymbol{J}), \quad \diff{J_\alpha}{\tau} = 0.
\end{equation}
We can obtain suitable coordinate-time frequencies using $\omega_t$ according to
\begin{equation}
\Omega_k = \frac{\omega_k}{\omega_t}.
\end{equation}
In both formulations of the problem, it should be noted that ratios of frequencies are independent of the time coordinate that is used.

\subsection{Beyond geodesic motion}
\label{sec:EMRI-beyond-geodesics}
Realistic compact objects do not follow geodesics as, by virtue of their own mass, they generate additional spacetime curvature that interacts with the background geometry. We previously introduced the concept of the gravitational self-force in \secref{intro-self-force}, but we now seek to establish its mathematical foundation.

We denote the physical, retarded metric perturbation by $h_{\alpha\beta}^{\mathrm{ret}}$, which is obtained by solving the Einstein field equations subject to causal boundary conditions, and sourced by a suitable energy-momentum tensor representing the motion of a compact object along a geodesic of some background spacetime $g_{\alpha\beta}$. This quantity is singular along the worldline of the particle and so it is physically meaningless to consider the particle motion as a geodesic in the ``full" spacetime $g + h^\mathrm{ret}$. Instead, the motion is better described by a trajectory in $g$, perturbed by the action of the gravitational self-force $F^\alpha_\mathrm{self}$.

Considering a particle of mass $\mu$, the self-force can be calculated at first order in the mass ratio $\eta\equiv\mu/M\ll 1$, by splitting the retarded field into two components at each spacetime point (P),
\begin{equation}
h_{\alpha\beta}^{\mathrm{ret}} = h_{\alpha\beta}^{\mathrm{dir}} + h_{\alpha\beta}^{\mathrm{tail}},
\end{equation}
corresponding to a direct contribution from the intersection of the past lightcone of P with the worldline, and a tail contribution from the worldline \textit{within} the past lightcone of P. The MiSaTaQuWa formula \citep{mino_gravitational_1997, quinn_axiomatic_1997} then states that in the Lorenz gauge, the self-force is given by
\begin{equation}
F^\alpha_\mathrm{self} = \mu \bar{\nabla}^{\alpha\beta\gamma} \bar{h}_{\beta\gamma}^{\mathrm{tail}},
\end{equation}
where an overbar represents trace-reversal using \eqnref{GW-trace-reversal}, and $\bar{\nabla}^{\alpha\beta\gamma}$ is the GR operator that computes forces from metric perturbations \citep{barack_gravitational_2009}.

The tail field is not smooth along the worldline. However, an alternative formulation by \citet{detweiler_self-force_2003} splits the retarded field instead into a singular (S) piece and a regular (R) piece according to
\begin{equation}
h_{\alpha\beta}^{\mathrm{ret}} = h_{\alpha\beta}^{\mathrm{S}} + h_{\alpha\beta}^{\mathrm{R}},
\end{equation}
with a corresponding force $F^\alpha_\mathrm{self} = \mu \bar{\nabla}^{\alpha\beta\gamma} \bar{h}_{\beta\gamma}^{\mathrm{R}}$. In this case, the R field is smooth everywhere and so the particle can be said to travel along geodesics in the effective metric $g+h^\mathrm{R}$.

It is conceptually useful to split the retarded field into yet another two components: the conservative and dissipative pieces are defined via
\begin{subequations}\begin{align}
F^\alpha_\mathrm{cons} = \frac1{2}\left(F^\alpha_\mathrm{self} + F^\alpha_\mathrm{self(adv)}\right),\\
F^\alpha_\mathrm{diss} = \frac1{2}\left(F^\alpha_\mathrm{self} - F^\alpha_\mathrm{self(adv)}\right),
\end{align}\end{subequations}
where $F^\alpha_\mathrm{self(adv)} = \mu \bar{\nabla}^{\alpha\beta\gamma}( \bar{h}_{\beta\gamma}^{\mathrm{adv}} - \bar{h}_{\beta\gamma}^{\mathrm{S}})$ is the (unphysical) equivalent self-force arising from the advanced metric perturbation. The dissipative part is responsible for secular changes in the orbital shape parameters (for example, the set $\{E,L_z,Q\}$), and so drives the inspiral of a particle via GW emission. The effects of the conservative part are more subtle, changing how the particle traverses an orbit rather than the orbit itself; it causes a shift in the fundamental frequencies along with secular changes to the positional elements (for example, the initial phases at $t=t_0$, given by the set $\{\psi_0,\chi_0,\phi_0\}$).

We can incorporate the self-force into the equations of motion using the action-angle formalism; employing an expansion in the mass ratio $\eta$, \eqnref{EMRI-AAgeo} becomes
\begin{subequations}
\label{eq:EMRI-AAsf}
\begin{align}
\diff{q_\alpha}{\tau} &= \omega_\alpha(\boldsymbol{J}) + \eta g_\alpha^{(1)}(q_r,q_\theta,\boldsymbol{J}) + \mathcal{O}(\eta^2),\\
\diff{J_\alpha}{\tau} &= \eta G_\alpha^{(1)}(q_r,q_\theta,\boldsymbol{J}) + \mathcal{O}(\eta^2),
\end{align}
\end{subequations}
where the forcing functions $g_\alpha^{(1)}$ and $G_\alpha^{(1)}$, which are $2\pi$-periodic in $q_r$ and $q_\theta$, arise from the first order self-force discussed above. The dissipative part is represented by the piece of $G_\alpha^{(1)}$ that is even under the transformation $q_{r,\,\theta}\rightarrow 2\pi-q_{r,\,\theta}$ and the piece of $g_\alpha^{(1)}$ that is odd under the same transformation \citep{hinderer_two-timescale_2008}. The remaining pieces constitute the conservative part.

Since the inspiral proceeds much slower than the orbital motion, these equations may be approximated by the adiabatic prescription by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\citep{drasco_computing_2005}. This piece is then purely dissipative and determines how the inspiral evolves due to the leading order radiation of GWs.

\subsubsection{The adiabatic approximation}
%To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration; a calculation of the full self-force at this stage is not feasible and so various approximation schemes have to be adopted. An initial simplification arises from considering only the dissipative part of the first-order self-force, i.e. the part that predominantly drives the inspiral, and neglecting behaviour on the orbital timescale.

The adiabatic prescription allows us to bypass many of the complexities of the full problem by using balance arguments \citep{hughes_gravitational_2005}. \Citet{teukolsky_perturbations_1973} derived a now eponymous equation for the Weyl scalar, $\psi_4$, that takes the form of a linearised wave equation for the perturbation due to a source. Solving this for a mass moving along a Kerr geodesic, the GWs at infinity can be extracted using the result that
\begin{equation}
\label{eq:teukolsky-waveform}
\psi_4 (r\rightarrow\infty) = \frac1{2}\left(\ddot{h}_+ - \iu \ddot{h}_\times\right).
\end{equation}
It is then possible to compute the orbital-averaged fluxes of energy and angular momentum that escape to infinity, $\langle\dot{E}^\mathrm{GW}\rangle_{\infty}$ and $\langle\dot{L}_z^\mathrm{GW}\rangle_{\infty}$. A similar procedure can be followed to calculate the fluxes lost to the central BH, $\langle\dot{E}^\mathrm{GW}\rangle_{\mathrm{H}}$ and $\langle\dot{L}_z^\mathrm{GW}\rangle_{\mathrm{H}}$. Taking these together, flux-balance implies that the orbital-averaged evolution of the energy must be given by
\begin{equation}
\left\langle\dot{E}\right\rangle = - \left\langle\dot{E}^\mathrm{GW}\right\rangle_{\infty} - \left\langle\dot{E}^\mathrm{GW}\right\rangle_{\mathrm{H}},
\end{equation}
and similarly for $\langle\dot{L_z}\rangle$. With these results, it is possible to construct adiabatic trajectories through phase space. For certain special cases, this is sufficient to fully describe the orbit; adiabatic evolutions have been generated for non-spinning systems \citep{cutler_gravitational_1994}, equatorial orbits \citep{glampedakis_zoom_2002}, which remain in the plane and so can be described using only $E$ and $L_z$, and circular inclined orbits \citep{hughes_evolution_2000}, which remain circular \citep{ryan_effect_1996} and so again can be represented by just two constants.

Generic orbits also depend on the value of the Carter constant $Q$, which must be evolved in a similar manner. The same balance argument cannot be used for $Q$ since there is no associated flux carried by the GWs. However, an equivalent orbital-averaged quantity can be derived, due to an argument by \citet{mino_perturbative_2003}, by replacing a long-time average with an average over the $2$-torus describing bound motion in $r$ and $\theta$
\begin{equation}
\label{eq:emri-longtime-av}
\lim_{T\rightarrow\infty} \frac1{2T}\intd{-T}{T}{}{\lambda} \rightarrow \frac1{(2\pi)^2} \intd{0}{2\pi}{}{q_r} \intd{0}{2\pi}{}{q_\theta},
\end{equation}
where $q_r$ and $q_\theta$ are the generalised angle variables for the $r$ and $\theta$ oscillations \citep{sago_adiabatic_2006}. This is valid for generic orbits as there is no strict phase relation between the radial and polar motions, and so all possible combinations of $\{q_r,\,q_\theta\}$ (modulo $2\pi$) occur over an infinite timespan; this is not true for resonant orbits, see \chapref{resonances}. Making use of this result, adiabatic trajectories and waveforms can be computed for generic orbits in Kerr \citep{drasco_gravitational_2006,sundararajan_towards_2008}, although this is limited by the ability to solve the Teukolsky equation efficiently. Approximate adiabatic inspirals can be generated more cheaply using PN expansions for the fluxes, augmented to give better agreement with Teukolsky results \citep{gair_improved_2006}.

These adiabatic approximations are simple to work with, but the neglected conservative piece of the self-force, which shifts the fundamental frequencies and produces a secular change in the positional elements, is important \citep{pound_limitations_2005} and may lead to observable consequences for parameter estimation \citep{huerta_influence_2009}.

\subsection{Gravitational waveforms}
\label{sec:EMRI-GWs}
An intrinsic part of the evolution of a two-body system in GR is the emission of GWs; a fully self-consistent solution to the problem must specify the metric coefficients at all points in spacetime, and thus completely encode information about both the motion of the bodies and the radiative fields at infinity. In practice, we also wish to calculate the GWs as they provide a handle on methods of observing the systems in question. However, in general this is a difficult task due to the nonlinear interaction between matter and spacetime curvature.

Ignoring any conservative corrections, Teukolsky-based waveforms, i.e. those obtained through \eqnref{teukolsky-waveform}, are the most accurate, but rely on solving the Teukolsky equation, which is computationally intensive. For data analysis purposes, it is useful to have EMRI waveforms that are cheap to compute but still contain many of the same features. `Kludge' waveforms provide for this possibility by separating the trajectory calculation from the waveform generation; there are two main approaches to this technique, the analytic kludge (AK) and the numerical kludge (NK).

AK waveforms, introduced by \citet{barack_lisa_2004}, are based on Keplerian orbits, and are parameterised by the quantities listed in \tabref{AK-parameters}. The trajectory is instantaneously described by an ellipse, but this is evolved in time using PN expressions to account for inspiral as well as relativistic perihelion and Lense-Thirring precessions. Notably, the inclination of the orbit is assumed to be constant. Once the trajectory has been determined, a corresponding waveform is generated using the Peters--Matthews formalism \citep{peters_gravitational_1963, peters_gravitational_1964}. This is the lowest-order consideration leading to GWs using just the quadrupole moment, but provides analytic expressions for the waveforms, which make them very quick to generate.

\begin{table}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{symbol}    & \textbf{description}                                 & \textbf{distribution in EMRI population}                      \\ \midrule
$\mu$              & mass of compact object                               & $10 M_\odot$                                                  \\
$M$                & mass of SMBH                                         & described in text, $10^4 \leq M/M_\odot \leq 10^7$            \\
$a$                & dimensionless spin of SMBH                           & uniform in $[0,0.95]$                                       \\
$\nu_0$            & orbital frequency at some fiducial time $t_0$        & described in text                                             \\
$e_0$              & eccentricity at $t_0$                                & described in text                                             \\
$\cos\iota$        & cosine of inclination angle                          & uniform in $[-1,1]$                                         \\
$\cos\theta_S$     & cosine of polar angle of direction to source         & uniform in $[-1,1]$                                         \\
$\phi_S$           & azimuthal angle of direction to source               & uniform in $[0,2\pi)$                                        \\
$\cos\theta_k$     & cosine of polar angle of SMBH spin direction         & uniform in $[-1,1]$                                         \\
$\phi_k$           & azimuthal angle of SMBH spin direction               & uniform in $[0,2\pi)$                                        \\
$\alpha_0$         & direction of orbital plane around SMBH spin at $t_0$ & uniform in $[0,2\pi)$                                        \\
$\tilde{\gamma}_0$ & direction of pericenter in orbital plane at $t_0$    & uniform in $[0,2\pi)$                                        \\
$\Phi_0$           & orbital mean anomaly at $t_0$                        & uniform in $[0,2\pi)$                                        \\
$z$                & redshift of source                                   & uniform in comoving volume, $z \leq z_\mathrm{max} = 1.5$ \\ \bottomrule
\end{tabular}
}
\caption{Parameters of the analytic kludge GW model of \citet{barack_lisa_2004}. Also given are the parameter distributions chosen for the sample EMRI population discussed in \secref{EMRI-detectability}.}
\label{tab:AK-parameters}
\end{table}

NK waveforms seek to improve upon the AK approach by using instantaneous Kerr geodesics instead of Keplerian ellipses. Trajectories are calculated by evolving the Kerr geodesic orbital parameters according to some forcing scheme; originally, PN fluxes supplemented with Teukolsky-based fits were used, but this is not a requirement. With a realistic trajectory in hand, the Boyer--Lindquist coordinates are then identified with flat-space polar coordinates and the gravitational waveform is computed using a flat-space emission formula. It is found that, even in the strong-field, NK waveforms have high overlaps (over $95\%$ in many cases) with Teukolsky-based waveforms, but are considerably quicker to produce \citep{babak_kludge_2007}.

\section{Astrophysics of extreme-mass-ratio inspirals}
The evolution of EMRIs has been studied extensively as a mathematical application for perturbation theory within GR. The resulting theoretical understanding of the gravitational self-force may also be of interest astrophysically if detectable systems occur regularly in the local Universe. Despite many areas of significant uncertainty, it is believed that a space-based GW detector should detect at least a handful of EMRI events. We now discuss the considerations that go into event rate estimates, before computing specific predictions.

\subsection{Formation of EMRIs}
The standard EMRI formation scenario considers the effect of two-body relaxation in the galactic nucleus, where the stellar population is dense enough for the process to be efficient. Components interact with each other via gravitational forces or direct contact collisions \citep{freitag_dynamics_2007}, causing deflections in their motion. The relaxation time $t_\mathrm{rlx}$ is the timescale on which an object's velocity is changed by an amount of order itself. Low values of $t_\mathrm{rlx}$ make it more likely for objects to be deflected into the loss cone, which is the volume of phase space such that the orbit passes close to the SMBH. Loss cone trajectories that cross the tidal disruption radius or the event horizon result in the prompt destruction of the compact object, but if this does not occur then it is possible for an inspiral to start \citep{alexander_orbital_2003}. It has recently been argued by \citet{amaro-seoane_role_2013} that the plunging orbits, which vastly outnumber traditional inspirals, are in fact indistinguishable as they can undergo hundreds of thousands of GW cycles before actually merging. This would enhance the EMRI event rate, producing a population of systems with very high eccentricities.

For non-plunging orbits, the system emits a burst of GWs with each periapse passage. These GW bursts are potentially detectable by space-based GW detectors, which might see $\mathcal{O}(1)$ per year \citep{berry_expectations_2013}. The compact object therefore loses energy and angular momentum, causing the orbit to shrink.  If the energy loss due to GWs occurs rapidly enough such that the radiation timescale $t_\mathrm{GW}$ is sufficiently shorter than the periapse relaxation timescale $(1-e)t_\mathrm{rlx}$, the compact object becomes decoupled from the stellar population and thus slowly spirals in towards the SMBH, producing an EMRI \citep{merritt_stellar_2011}.

This simple model of EMRI formation relies on two-body relaxation being the dominant mechanism through which a compact object can change its angular momentum. In fact, it is found that resonant relaxation \citep{rauch_resonant_1996} plays an important role \citep{hopman_resonant_2006}. On timescales longer than the orbital timescale, but shorter than the precessional timescale, the stellar trajectories around the SMBH can be modelled as massive wires with a local mass density proportional to the time that the compact object spends at each point. These wires exert mutual torques on each other that are roughly constant on the orbital timescale, leading to enhanced relaxation of the angular momenta. Objects that are scattered onto high eccentricity orbits come close to the SMBH, where GR precessional effects are significant, negating the effect of resonant relaxation and ensuring the object does not get scattered back. This process enhances the EMRI formation rate by a factor of a few \citep{hopman_resonant_2006}.

In addition to two-body capture, as described above, it is possible to form an EMRI through other channels. If a binary system approaches the SMBH it can be tidally separated, ejecting one component at high velocity while capturing the other \citep{miller_binary_2005}. This process occurs at much larger radii than the two-body capture events, vastly increasing the cross-section for interaction, although the event rate is restricted by the limited abundance of systems in binary configurations. These EMRIs are expected to have very low eccentricities, since they evolve under GW emission (and hence circularise) from large radii without interacting significantly with other objects. Alternatively, EMRIs may be formed from accretion discs that are unstable to star formation \citep{levin_starbursts_2007}. This would lead to a population of systems in the equatorial plane of the SMBH in contrast to other formation channels, which lead to spherically symmetric distributions.

\subsection{EMRI event rate}
\label{sec:EMRI-event-rate}
Estimates for the event rate of EMRIs are split into separate parts to help illuminate the different approximations that have to be made, and apply any available observational constraints. Perhaps the most uncertain component is that of the ``intrinsic" event rate $\mathcal{R}$, i.e. the number of inspiral events per unit time for any given galaxy. We are interested in galaxies that possess a SMBH with a mass in the range $(10^{4} - 10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors. In principle, $\mathcal{R}$ will depend on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each component, and the mass and spin of each SMBH. All of these properties are highly uncertain, even for our own galaxy.% In addition, there is no \emph{a priori} reason to expect that different galaxies have similar predictable properties very close to the SMBH.

Despite the difficulties, simple estimates of $\mathcal{R}$ have been carried out using Monte Carlo methods to count the number of stars from isothermal distributions that spiral in to a SMBH without plunging. The result is a scaling law for each species of compact object of the form
\begin{equation}
\label{eq:EMRI-intrinsic-rate}
\mathcal{R}(M) = \mathcal{R}_0 \left(\frac{M}{M_0}\right)^\alpha,
\end{equation}
where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is some fiducial mass. \Citet{hopman_extreme_2009} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, NSs and WDs, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}\; \mathrm{Gyr}^{-1}$ for each component. This neglects the effects of resonant relaxation, ignores the spin of the SMBH, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses. Each of these is likely to significantly impact the event rate, but \eqnref{EMRI-intrinsic-rate} can still be used as a rough guide to the expected number of events.

The different compact objects emit GWs with different amplitudes, scaling with mass $\mu$ and distance $D$ as $\mu/D$. Assuming that a detector is sensitive to GWs above a certain amplitude, we can then observe heavier systems at much greater distances. The observational volume is a factor of $\mathcal{O}(10^3)$ larger for stellar mass BHs of typical mass $10 M_\odot$, in comparison to that for WDs and NSs. The total event rate will therefore be dominated by BHs \citep{gair_event_2004}. Henceforth, we neglect the contribution from NSs and WDs and set the mass of the compact object to be $10 M_\odot$.

%Henceforth, we neglect the contribution from NSs and WDs as their rates are significantly lower than that of BHs.\footnote{The different compact objects emit GWs with different amplitudes and frequency evolutions. Despite the lower intrinsic rates, it is therefore possible for NSs and WDs to contribute more significantly to \emph{detectable} systems, if they emit more strongly at the most sensitive frequencies of a detector. However, this is found not to be the case \citep{gair_event_2004}.} We therefore set the mass of the compact object to be $10 M_\odot$, a typical size for stellar mass BHs.

\Citet{amaro-seoane_impact_2011} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker-Planck description for the bulk properties of the stellar distribution. They found a better fit for the power law spectral index for BHs of $\alpha = -0.19$.

Once the intrinsic EMRI rate has been established, it is necessary to consider the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH. This is challenging to estimate because the number of SMBHs will depend on many parameters, including local structure in the Universe, the evolution of that structure, and properties of the SMBHs themselves. We simplify the problem by assuming a homogeneous distribution that does not evolve with redshift, which is reasonable for the typical scales considered by GW detectors. We also neglect correlations between the SMBH mass and spin, and impose a power law scaling relation for the comoving number density
\begin{equation}
\diff{n}{\ln M}(M) = n_0 \left(\frac{M}{M_0}\right)^\beta.
\end{equation}
There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002 \; \mathrm{Mpc}^{-3}$ for SMBHs with $M < \mathcal{O}(10^7 M_\odot)$ \citep{greene_mass_2007, gair_lisa_2010}. \Citet{gair_lisa_2010} show that it is possible to reverse this calculation and constrain the slope $\beta$ using the observed number of EMRI events.

%A further complication arises from the fact that EMRIs are long-lived sources of GWs. Each potential source can be characterised by the remaining time before plunge, $t_\mathrm{plunge}$, when a GW detector starts observing. If the mission lifetime $T$ is longer than $t_\mathrm{plunge}$ then the entire remainder of the evolution can be seen, and the SNR increases with increasing $t_\mathrm{plunge}$. However, for $T < t_\mathrm{plunge}$, only a fraction of the remaining inspiral can be observed and this leads to a decreasing SNR with $t_\mathrm{plunge}$ as the GW emission moves out of band. We can define an observable lifetime for EMRIs, $\tau$, as the range of $t_\mathrm{plunge}$ such that the SNR is above some threshold level; the number of events seen from a particular galaxy is then $\tau \mathcal{R}$. Making the approximation that all of the SNR is accumulated at plunge and that this is sufficient for detection, we can set $\tau = T$ for all systems.

\subsection{Detectability of EMRIs}
\label{sec:EMRI-detectability}
To evaluate the number of \emph{observable} EMRIs for a given detector configuration requires a sample population of potentially observable systems to be generated. A complication arises from the fact that EMRIs are long-lived sources of GWs, and their detectability depends on the part of their evolution history that overlaps with the observation window. A typical source may be undetectable if it plunges close to the start of an observation, become detectable once the plunge occurs closer to the end of the observation, remain detectable even if it plunges after the observation, before becoming undetectable again if it plunges much later. It is not possible to model this behaviour in a detector-independent manner. We therefore consider only those systems that plunge within some time $T$, producing a population with $\mathcal{R}T$ potentially observable events from each galaxy.

A given EMRI is classified as observable if its SNR exceeds some threshold value $\mathrm{SNR}_\mathrm{thres}$. The size of the required population can be estimated by combining the different components discussed in the previous section, to give an integral of the form \citep{gair_probing_2009}
\begin{align}
\label{eq:EMRI-number}
N_\mathrm{EMRI} &= \intd{z=0}{z_\mathrm{max}}{ \intd{M=M_\mathrm{min}}{M_\mathrm{max}}{   \mathcal{R}T\diff{n}{\ln M} \diff{V_\mathrm{c}}{z}   }{\ln M}  }{z},\\
&= \mathcal{R}_0 T n_0 V_\mathrm{c}(z_\mathrm{max}) \intd{M=M_\mathrm{min}}{M_\mathrm{max}}{   \left(\frac{M}{M_0}\right)^{\alpha+\beta}   }{\ln M}, \label{eq:EMRI-num-Mdist}
\end{align}
where $V_\mathrm{c}(z)$ is the comoving volume at redshift $z$. For a mission lifetime of 2 years and a maximum redshift of $1.5$, along with values of the constants quoted in \secref{EMRI-event-rate}, the integral in \eqnref{EMRI-number} gives a total of $6333$ EMRI events.\footnote{We use a $\Lambda$CDM cosmology to compute the comoving volume, with $\Omega_m = 0.3$, $\Omega_\Lambda = 0.7$ and $H_0 = 70\;\mathrm{km}\,\mathrm{s}^{-1}\,\mathrm{Mpc}^{-1}$.} This is a lower bound for $N_\mathrm{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.

We generate a representative EMRI population of this size, using the AK parameters to describe each system. The mass of the SMBH is sampled from the power-law distribution indicated by \eqnref{EMRI-num-Mdist}, i.e. with a probability distribution function (pdf) $f(M) \propto M^{\alpha+\beta-1}$. Most of the other parameters are sampled from uniform distributions, detailed in \tabref{AK-parameters}, but the eccentricity distribution is more complicated.

Eccentricities for EMRIs are uncertain, and depend strongly on the formation scenario being considered. Here, we adopt a fit to a distribution computed using Monte Carlo simulations by \citet{hopman_orbital_2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH. We assume this can be extended to provide a rough estimate of the distribution around BHs of other masses and spins. At the point in the inspiral when the orbital period takes a fiducial value $\nu_0 = 10^4\,\mathrm{s}$, we find that the Monte Carlo eccentricity pdf is well-described by a power-law with an exponential cutoff
\begin{equation}
\label{eq:EMRI-e-distribution}
f(e;\,e_\mathrm{m},\,e_\mathrm{p},\,b) = 
	\begin{cases}
		\displaystyle\frac{b^{1+b(e_\mathrm{m}-e_\mathrm{p})} \left(e_\mathrm{m}-e\right)^{b(e_\mathrm{m}-e_\mathrm{p})} \exp\left\{b(e-e_\mathrm{m})\right\} }{\Gamma\left(1+b(e_\mathrm{m}-e_\mathrm{p})\right) - \Gamma\left(1+b(e_\mathrm{m}-e_\mathrm{p}), b e_\mathrm{m}\right)}, &\quad 0 \leq e \leq e_\mathrm{m}\\
		0, &\quad\text{otherwise}
	\end{cases}
\end{equation}
where $\Gamma(z)$ is the Euler gamma function, $\Gamma(z,a)$ is the incomplete gamma function, $e_\mathrm{m} = 0.81$ is the maximum observed eccentricity, $e_\mathrm{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index. This distribution, as shown in \figref{EMRI-e-distribution}, is peaked at higher eccentricities.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{e_distribution}
\caption{\label{fig:EMRI-e-distribution}Probability distribution function (pdf) for the eccentricity of an inspiral when the orbital period is $10^4\,\mathrm{s}$ (black line), fit to data by \citet{hopman_orbital_2005}. Also shown are histograms of the eccentricites of the EMRI population as originally sampled (blue), and at the last stable orbit (red).}
\end{figure}

We are interested in EMRI systems close to plunge, as it is here where the GW amplitude is largest. To create our EMRI population, we therefore sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value of $10^4\,\mathrm{s}$, and evolve the system using the AK prescription up until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $R(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.} The resulting eccentricity distribution at plunge is shown in red in \figref{EMRI-e-distribution}, which demonstrates the circularising effect of GW emission. The corresponding orbital frequency distribution is shown in \figref{EMRI-nu-distribution}; the turn-off at low frequencies is an unphysical artifact due to the sharp cut-off in our mass distribution.

Each of these systems is then evolved backwards for some time $t_\mathrm{insp}$, chosen uniformly from the range $[0,2]$ years, and the expected gravitational waveforms $h_{+,\times}$ are calculated using the AK formalism. The evolution is carried out using the Euler method, including an additional approximation for the second time derivative.\footnote{To integrate the differential equation $\dd y/\dd x = f(y)$ from an initial value $y_0$, we set $y_{i+1} = y_i + h f(y_i) + h(f(y_i) - f(y_{i-1}))/2$, where $y_i = y(x_i)$, $h$ is the step size, and we compute $y_1$ using only the first two terms.} This is a crude evolution scheme but should be sufficiently accurate if we choose a step size on the orbital timescale, which is small with respect to the evolution timescale; we use $\dd t = \mathrm{min}(10\,\mathrm{s}, 1/(8\nu_\mathrm{LSO}))$ to ensure a minimum number of eight evaluation points within any GW period.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{nu_distribution}
\caption{\label{fig:EMRI-nu-distribution}Probability distribution function of the orbital frequencies of the EMRI population at the last stable orbit.}
\end{figure}

In order to assess whether or not EMRIs are observable, we must specify the noise properties of the detector being used. If we assume stationary, Gaussian noise in the detector, it is sufficient to specify the (one-sided) noise PSD $S_n(f)$, defined by \eqnref{noise-PSD}. We consider seven different PSDs corresponding to different detector configurations: a standard fit used to represent the eLISA detector \citep{amaro-seoane_elisa:_2013}; and six similar detectors, labelled A--F, formed by varying the arm lengths (using either $1$, $2$ or $5\times 10^9\;\mathrm{m}$) and the acceleration noise model (taken to match either the expected LISA Pathfinder performance or the requirements of the LISA mission). The different detectors are detailed in \tabref{EMRI-numEMRIs} and the corresponding PSDs are plotted in \figref{EMRI-detector-PSDs}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{GOAT_detectors}
\caption{\label{fig:EMRI-detector-PSDs}The (one-sided) noise power spectral densities for the detectors used to evaluate the number of observable EMRIs.}
\end{figure}

Each detector configuration consists of three satellites with six laser links between them, from which it is possible to construct two independent GW responses $h_{I,II}$, which are linear combinations of $h_{+,\times}$. Descoped missions often remove two of the laser links meaning that only a single response $h_I$ can be constructed. Combining equations \eqref{eq:noise-overlap} and \eqref{eq:general-SNR}, we compute the SNR in the frequency domain for each of our EMRI systems according to
\begin{equation}
\label{eq:EMRI-SNR}
\mathrm{SNR}^2 = 4\sum_{\alpha}\intd{0}{\infty}{\frac{\tilde{h}^*_\alpha(f) \tilde{h}_\alpha(f)}{S_n(f)}}{f},
\end{equation}
where the sum is over $\alpha = \{I\}$ for four laser link systems, and $\alpha = \{I,\,II\}$ for six laser links. The resulting survival distribution functions for four and six laser links are shown in \figref{EMRI-SNRs}, that is for each value of SNR, we plot the probability of an observation exceeding that SNR. The inclusion of six laser links instead of four only marginally improves the detection prospects for EMRIs,\footnote{The main scientific advantage of additional laser links is that the degeneracy between sky position and polarisation can be broken for monochromatic sources \citep{vallisneri_sensitivity_2008}.} increasing the SNR by a factor of $\sqrt{2}$. The numbers of EMRIs with SNRs above threshold values of $15$, $20$ and $25$ are listed in \tabref{EMRI-numEMRIs}. Given the many uncertainties involved, particularly in regards to the astrophysical event rate, these numbers should be treated as a rough guide to the number of observable events for a two year mission lifetime.

\begin{table}[htbp]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}cclc>{\centering}p{3cm}<{\centering}>{\centering}p{3cm}<{\centering}>{\centering}p{3cm}<{\centering}@{}}
\toprule
\multicolumn{4}{c}{\textbf{detector specification}}                                                                                                          & \multicolumn{3}{c}{\textbf{number of observable events in 2 years}} \\
\multirow{2}{*}{\textit{label}} & \multirow{2}{*}{\textit{arm length}}                 & \multicolumn{1}{c}{\multirow{2}{*}{\textit{acceleration noise}}} & \multirow{2}{*}{\textit{laser links}} & \multicolumn{3}{c}{\textit{SNR threshold}}                                   \\
                       &                                             & \multicolumn{1}{c}{}                                    &                              & 15                    & 20                    & 25                   \tabularnewline \midrule
\multirow{2}{*}{A}     & \multirow{2}{*}{$1\times 10^9\;\mathrm{m}$} & \multirow{2}{*}{LISA Pathfinder}                        & 4                            & 95                    & 45                    & 25                   \tabularnewline %\cmidrule{4-7} 
                       &                                             &                                                         & 6                            & 215                   & 110                   & 70                   \tabularnewline \midrule%\cmidrule{4-7} 
\multirow{2}{*}{B}     & \multirow{2}{*}{$1\times 10^9\;\mathrm{m}$} & \multirow{2}{*}{LISA}                                   & 4                            & 675                   & 355                   & 200                  \tabularnewline %\cmidrule{4-7} 
                       &                                             &                                                         & 6                            & 1355                  & 760                   & 460                  \tabularnewline \midrule
\multirow{2}{*}{C}     & \multirow{2}{*}{$2\times 10^9\;\mathrm{m}$} & \multirow{2}{*}{LISA Pathfinder}                        & 4                            & 375                   & 205                   & 115                  \tabularnewline %\cmidrule{4-7} 
                       &                                             &                                                         & 6                            & 840                   & 445                   & 250                  \tabularnewline \midrule%\cmidrule{4-7} 
\multirow{2}{*}{D}     & \multirow{2}{*}{$2\times 10^9\;\mathrm{m}$} & \multirow{2}{*}{LISA}                                   & 4                            & 2245                  & 1430                  & 915                  \tabularnewline %\cmidrule{4-7} 
                       &                                             &                                                         & 6                            & 3405                  & 2445                  & 1790                 \tabularnewline \midrule
\multirow{2}{*}{E}     & \multirow{2}{*}{$5\times 10^9\;\mathrm{m}$} & \multirow{2}{*}{LISA Pathfinder}                        & 4                            & 1285                  & 735                   & 440                  \tabularnewline %\cmidrule{4-7} 
                       &                                             &                                                         & 6                            & 2280                  & 1450                  & 930                  \tabularnewline \midrule%\cmidrule{4-7} 
\multirow{2}{*}{F}     & \multirow{2}{*}{$5\times 10^9\;\mathrm{m}$} & \multirow{2}{*}{LISA}                                   & 4                            & 4375                  & 3550                  & 2860                 \tabularnewline %\cmidrule{4-7} 
                       &                                             &                                                         & 6                            & 5180                  & 4570                  & 3940                 \tabularnewline \midrule
\multicolumn{3}{c}{\multirow{2}{*}{eLISA}}                                                                                     & 4                            & 215                   & 110                   & 60                   \tabularnewline %\cmidrule{4-7} 
\multicolumn{3}{c}{}                                                                                                           & 6                            & 515                   & 255                   & 145                  \tabularnewline \bottomrule
\end{tabular}
}
\caption{Detector configurations used to study the number of EMRIs that will be observable over a two year mission lifetime. The numbers for varying threshold SNRs are shown rounded to the nearest $5$ to the right-hand side of the table. The total number of possible EMRI events in the population was 6333. Detector F corresponds to the original LISA mission design.}
\label{tab:EMRI-numEMRIs}
\end{table}

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.8\textwidth]{GOAT_SNRs_4links}}\\
\subfloat{\includegraphics[width=0.8\textwidth]{GOAT_SNRs_6links}}
\caption{\label{fig:EMRI-SNRs}The survival function for the SNR distributions of our EMRI population, using the given detector configurations with four (upper figure) and six (lower) laser links.}
\end{figure}

As well as the raw number of observable events, it is useful to consider the full distribution of SNRs as a function of the population parameters. For a given parameter, we bin the population data and then count the number of events $\mathcal{N}_s$ in each bin that exceed a certain threshold SNR. Given these successes and the size of the bin $\mathcal{N}$, the probability of a new observation within the bin exceeding the threshold follows the beta distribution, $p \sim \mathrm{Beta}\left(\mathcal{N}_s + 1/2, \, \mathcal{N} - \mathcal{N}_s + 1/2\right)$.\footnote{We use the uninformative Jeffreys prior for the probability, $p \sim \mathrm{Beta}\left(1/2,\,1/2\right)$. This results in the expected value after $\mathcal{N}$ trials given by \eqnref{EMRI-beta-mean} rather than the obvious alternative $\mu = \mathcal{N}_s/\mathcal{N}$, which is the result of taking the Haldane prior, $p \sim \mathrm{Beta}\left(0,\,0\right)$.} This has an expected value
\begin{equation}
\label{eq:EMRI-beta-mean}
\mu = \frac{\mathcal{N}_s + 1/2}{\mathcal{N} + 1},
\end{equation}
and variance
\begin{equation}
\sigma^2 = \frac{\mu\left(1-\mu\right)}{\mathcal{N} + 2}.
\end{equation}
Using the eLISA detector configuration with $6$ links, we plot the probability of detection in \figref{EMRI-detrate}, as a function of $a$, $\cos\iota$ and $z$. We find mild selection biases in spin and inclination; it is easier to detect prograde equatorial EMRIs around highly spinning BHs. This makes sense intuitively as these orbits penetrate to much lower values of $r$ before becoming unstable. The detection probability decreases with $z$, as the source gets further away from the detector. The maximum redshift that eLISA could observe an EMRI with an SNR greater than $25$ is approximately $0.7$, with most detections occuring below $0.2$.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.7\textwidth]{eLISA_detrate_a}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{eLISA_detrate_cosi}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{eLISA_detrate_z}}
\caption{\label{fig:EMRI-detrate}The probability of an EMRI signal observed with the eLISA $6$-link detector configuration having an SNR greater than some threshold, as a function of the spin of the SMBH (upper figure), the cosine of the orbital inclination (middle), and the redshift (lower). Three different thresholds are considered, $15$ (blue), $20$ (green), and $25$ (red), with larger values being represented by thicker lines.}
\end{figure}

The noise properties of the detectors are frequency dependent, and so we expect sources with characteristic frequencies in the most sensitive region to have higher SNRs; this fact motivates the restricted mass range for the SMBH in our sample population to those systems that produce EMRIs of the ``correct" frequency. The calculation of a characteristic frequency is, however, rather arbitrary. Here, we follow \citet{barack_lisa_2004} and define
\begin{equation}
f_\mathrm{GW} = n \nu + \frac{\dot{\tilde{\gamma}}}{\pi},
\end{equation}
as a combination of the radial orbital frequency $\nu$ and the azimuthal orbital frequency $\nu + \dot{\tilde{\gamma}}/(2\pi)$, and where $n$ is an integer specifying different harmonics. In \figref{EMRI-eLISA-SNRs}, we plot the SNR of each EMRI system as a function of $f_\mathrm{GW}$ evaluated at the LSO, along with an overlay of the eLISA PSD. As expected, the SNRs peak near the minimum of the PSD, with scatter due to the specific combination of other parameters. We can estimate the SNR using a simple prescription, as follows. We assume a constant noise PSD across the inspiral, so that \eqnref{EMRI-SNR} becomes
\begin{equation}
\mathrm{SNR}^2 = 4\sum_{\alpha}\intd{0}{\infty}{\frac{\tilde{h}_{\alpha}^*(f) \tilde{h}_{\alpha}(f)}{S_n(f)}}{f} = \frac{4}{S_n(f_\star)}\sum_{\alpha}\intd{0}{\infty}{\tilde{h}_{\alpha}^*(f) \tilde{h}_{\alpha}(f)}{f},
\end{equation}
which should be taken as a definition of $f_\star$ for any given system. We can evaluate the integral using Plancherel's theorem, according to
\begin{align}
\intd{0}{\infty}{\tilde{h}_{\alpha}^*(f) \tilde{h}_{\alpha}(f)}{f} &= \frac1{2} \intd{-\infty}{\infty}{\tilde{h}_{\alpha}^*(f) \tilde{h}_{\alpha}(f)}{f},\nonumber\\
&= \frac1{2}\intd{-\infty}{\infty}{\!}{f} \intd{-\infty}{\infty}{\!}{t} \intd{-\infty}{\infty}{\!}{t'} h_{\alpha}^*(t') h_{\alpha}(t) \exp\left(2\pi\iu f(t-t')\right),\nonumber\\
&= \frac1{2}\intd{-\infty}{\infty}{\!}{t} \intd{-\infty}{\infty}{\!}{t'} h_{\alpha}^*(t') h_{\alpha}(t) \delta\left(t-t'\right),\nonumber\\
&= \frac1{2}\intd{-\infty}{\infty}{h_{\alpha}^*(t) h_{\alpha}(t)}{t},\nonumber\\
&\approx \frac{|h_{\alpha}|^2 T}{2},
\end{align}
where we have used the definition of the Fourier transform, evaluated the time integral over the finite length $T$, and approximated that the signal amplitude remains roughly constant in time $|h_{\alpha}(t)| \approx |h_{\alpha}|$. The SNR can then be estimated using
\begin{equation}
\label{eq:EMRI-SNR-estimate}
\mathrm{SNR}^2 \approx \frac{2T}{S_n(f_\star)} \sum_{\alpha}|h_\alpha|^2.
\end{equation}
This is only useful if we have a practical means of estimating $f_\star$. For a given system, if the frequency at plunge, $f_\mathrm{GW}$, is lower than the location of the minimum of the detector PSD at $f_\mathrm{min}$, then the dominant contribution to the SNR will occur at plunge; we might therefore estimate $f_\star \approx f_\mathrm{GW}$. In contrast, if the frequency is greater than $f_\mathrm{min}$, a significant part of the SNR will be generated at $f_\mathrm{min}$ and so a reasonable estimate might be $f_\star \approx f_\mathrm{min}$. Smoothly transitioning between these regimes, we take
\begin{equation}
\frac1{f_\star} = \frac1{2} \left( \frac1{f_\mathrm{GW}} + \frac1{f_\mathrm{min}} \right).
\end{equation}
The resulting estimate of the SNR well-approximates the data across a wide range of frequencies close to where the detector is most sensitive. It is shown for eLISA in \figref{EMRI-eLISA-SNRs} alongside the raw data points.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{eLISA_SNRs}
\caption{\label{fig:EMRI-eLISA-SNRs}The SNRs for the sample EMRI population, computed using the eLISA detector configuration with 6 laser links. Each SNR is plotted against the characteristic GW frequency at plunge. An estimate of the SNR, as computed using \eqnref{EMRI-SNR-estimate}, is shown by the solid golden line; the amplitude of this estimate is chosen to match the data. Also shown, for comparison, is the eLISA noise PSD (black, dashed line), although this is on a different scale (not given).}
\end{figure}

\section{Summary}
It should be possible to observe extreme-mass-ratio inspirals using a future space-based GW detector, currently in the early stages of planning, and due for launch in the 2030s. A typical mission design may be able to see hundreds of events over a two-year lifetime, although any estimates are highly uncertain due to the complicated astrophysics within galactic nuclei. The number of observations, along with the physical parameter distributions, will be able to inform us about different EMRI formation mechanisms, as well as processes such as mass segregation and resonant relaxation, and even provide information on cosmological scales, constraining the SMBH mass function.

We have considered a variety of potential detector configurations, as part of an investigation into the design of mission concepts to address the European Space Agency's L3 science theme, \inlinequote{The Gravitational Universe} \citep{the_elisa_consortium_gravitational_2013}. We find that all of the proposed detectors are capable of observing some EMRI events, although this is not guaranteed due to the uncertain intrinsic event rates. The expected number of observable events increases by a factor of roughly $2$ if we use six laser links instead of four, and by a factor of around $7$--$8$ if we implement the more stringent LISA acceleration noise requirements rather than those for LISA Pathfinder. Increased arm lengths also result in larger numbers of observable events, which increase by a factor of roughly $4$ going from $1$ to $2\times  10^9\;\mathrm{m}$, and approximately $10$ if we use $5\times  10^9\;\mathrm{m}$. However, these improvements are all diminished for the more sensitive configurations, as we begin to become limited by the source abundance.

GW signals from EMRIs contain a wealth of scientific information, not just from population statistics but on an individual level as well. The modelling process depends on details of the Kerr metric in GR and so any deviations from this can influence the emitted GWs. This is aided by the fact that the compact object can spend hundreds of thousands of cycles in the strong field regime close to the SMBH, where deviations are likely to be strongest. In order to extract such information from the signals, we need accurate waveform models, which can be approximated using the adiabatic prescription. However, this neglects certain effects, one of which is the presence of transient resonances, which we discuss further in the following chapter.

### 2015-05-29 thesis_chapters/effect_of_resonances.tex
post-viva, post chop house thoughts

\chapter[Effect of resonances on the detectability of EMRIs]{Effect of resonances on the detectability of extreme-mass-ratio inspirals}\label{ch:effect-resonances}

One of the targets of pre-eLISA research is to generate a bank of waveform templates, $h(t;\boldsymbol{\lambda})$ described by some set of parameters $\boldsymbol{\lambda}$. Once observational data, $s(t)$, has been collected, we can then search over the template bank to find the model that most closely matches the data. In order for the best-fit model to have any physical significance, we require that the waveform templates accurately reproduce what we expect to observe in nature. However, they cannot be too computationally expensive, as this would prohibit an effective search over the parameter space.

Ideally, we would like to use waveform templates from EMRI systems that are evolved under the action of the full gravitational self-force. Such a computational scheme is not yet available and so we are forced to use waveforms computed under the adiabatic approximation. We know that this explicitly eliminates any effects of transient resonances but it is unclear how that impacts the detectability of EMRIs using adiabatic templates. To assess this, we compare simulated data $s(t;\bar{\boldsymbol{\lambda}})$ generated using the instantaneous self-force model, with templates $h(t;\boldsymbol{\lambda})$ generated using the 2-torus-averaged adiabatic model, detailed in \secref{res-ad-averaging}.

\section{Gravitational wave generation}
To generate gravitational waveforms, we choose to employ the NK method of \citet{babak_kludge_2007}. As discussed in \secref{EMRI-GWs}, the principle behind the class of kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory. This can be much quicker and easier to calculate than waveforms using Teukolsky-based methods and yet gives similar results.

Following the NK method, we first compute the inspiral trajectory of the compact object using the method of osculating elements detailed in \secref{EMRI-evolution-scheme}. We then interpret the Boyer--Lindquist coordinates as spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique. Specifically, we apply the standard quadrupole formula. We expect these waveforms to be sufficiently accurate for our purposes, given the inaccuracy of the self-force model being employed. It would, however, be straightforward to adopt a different waveform generation scheme, for example by including higher order multipole moments, should advances be made in determining the self-force.

Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s_\alpha(t)$ and $h_\alpha(t)$, where the subscript $\alpha$ refers to different detector outputs. Following \eqnref{noise-overlap}, we write
\begin{equation}
\label{eq:effres-innerprod}
\innerii{s}{h} = 2 \sum_\alpha \intd{0}{\infty}{\frac{\tilde{s}_\alpha(f)\tilde{h}_\alpha^*(f)+\tilde{s}_\alpha^*(f)\tilde{h}_\alpha(f)}{S_{n,\alpha}(f)}}{f},
\end{equation}
where $\tilde{s}_\alpha(f)$ represents the Fourier transform of $s_\alpha(t)$, similarly for $\tilde{h}_\alpha(f)$, and $S_{n,\alpha}(f)$ is the one-sided noise PSD for detector $\alpha$. Given a specific data realisation, we can evaluate the signal-to-noise ratio (SNR) for each waveform template, defined by
\begin{equation}
\rho\left[h\right] = \frac{\innerii{s}{h}}{\sqrt{\innerii{h}{h}}}.
\end{equation}
If a template were to exactly match the data, it would produce an SNR of $\sqrt{\innerii{s}{s}}$ and so the overlap
\begin{equation}
\label{eq:effres-overlap}
\mathcal{O}\left[h\right] = \frac{\innerii{s}{h}}{\sqrt{\innerii{s}{s}}\sqrt{\innerii{h}{h}}}
\end{equation}
provides an indication of how well-matched the template is to the data.

We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept, specifically choosing the analytic approximation of \citet{amaro-seoane_low-frequency_2012}. We compute two orthogonal detector responses corresponding to the two GW-sensitive datastreams of a 6 laser link eLISA mission, and carry out the appropriate sum over these outputs according to \eqnref{effres-innerprod}. For brevity, we shall henceforth drop the subscript $\alpha$.

In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing. To account for this, when evaluating overlaps, we first project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme, setting any values outside the domain of $\{t_j\}$ to zero. A Planck-taper window function \citep{mckechan_tapering_2010} is applied to the resulting time series. This can be defined in a continuous form by
\begin{equation}
w(t) = \left\{
\begin{array}{ll}
	0 & t\leq t_1 \\
	\left(1+\exp\left(\frac{t_2-t_1}{t-t_1}+\frac{t_2-t_1}{t-t_2}\right)\right)^{-1} & t_1 < t < t_2 \\
	1 & t_2\leq t\leq t_3 \\
	\left(1+\exp\left(\frac{t_3-t_4}{t-t_3}+\frac{t_3-t_4}{t-t_4}\right)\right)^{-1} & t_3 < t < t_4 \\
	0 & t_4\leq t,
\end{array}\right.
\end{equation}
where $t_1$ and $t_4$ are chosen to encompass the whole waveform and $t_2$ and $t_3$ are the times of the second closest stationary points to each end, ensuring that the transition is not too sharp; the smoothness of the waveform at each end reduces unwanted spectral leakage in the Fourier transforms.

After tapering the projected waveforms, we then calculate inner products according to \eqnref{effres-innerprod}, using discrete Fourier transforms and replacing the integral over frequency with an appropriate sum over Fourier modes. In addition, we maximise every overlap calculation with respect to some time-offset between the models by finding the maximum value of the cross-correlation between the waveforms. This process can be performed cheaply using Fourier transforms and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.

We demonstrate our waveform generation using the illustrative system introduced at the end of \secref{res-ad-averaging}. Once the trajectories have been computed using both the instantaneous and adiabatic evolutions, we can calculate the corresponding waveforms. The plus-polarised GW at the start and end of the evolution is shown in \figref{good-waveform}. The waveforms show good agreement in both amplitude and phase; we find an overlap of $0.993$, illustrating that adiabatic models can be used in place of the instantaneous evolution in cases where resonances are not encountered.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{good_waveform}
\caption{\label{fig:good-waveform}The plus-polarised waveform for an illustrative EMRI system with $p=7.5$, shown for two short segments at the start (left plot) and end (right) of a $2$ year evolution. The instantaneous (thick blue line) and adiabatic (thin golden line) models are both shown, but are indistinguishable by eye.}
\end{figure}

\section{The effect of transient resonances on a single system}
\label{sec:effres-single-system}
The illustrative EMRI system of the previous section does not encounter a significant resonance and the adiabatic evolution provides a good match to the instantaneous evolution. We now study a system that does pass through a resonance during its $2$ year evolution.

\subsection{Dephasing}
\label{sec:effres-dephasing}
We choose the initial conditions to be the same as before but with an initial semi-latus rectum $p=7.85$. This system passes through the 2:3 resonance, the effect of which is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms. Plotting the plus-polarised GWs at the start and end of the evolution, as before, demonstrates this dephasing; the waveforms are shown in \figref{dephased-waveform}.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{dephased_waveform}
\caption{\label{fig:dephased-waveform}The plus-polarised waveform for an illustrative EMRI system with $p=7.85$, shown for two short segments at the start (left plot) and end (right) of a $2$ year evolution. The instantaneous (thick blue line) and adiabatic (thin golden line) models are both shown; there is a significant dephasing between them by the end of the evolution.}
\end{figure}

To highlight the dephasing more quantitatively, we calculate the shortened overlap between the two models as a function of time, $t$. We define this as the overlap obtained by including only the part of the waveform within some $\Delta t$ of $t$. For our illustrative system, we choose $\Delta t$ such that we can calculate 25 non-overlapping values of this shortened overlap. Before the resonance occurs, the adiabatic model provides a good match to the instantaneous evolution but the overlap is reduced near to the resonance and never fully recovers afterwards. This is shown in \figref{overlap-dephasing}, which is centred on the time at which the instantaneous evolution crosses the 2:3 resonance.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{overlap_vs_time}
\caption{\label{fig:overlap-dephasing}The overlap computed between the instantaneous evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give 25 independent (non-overlapping) calculations. The time axis is centred on the value at which the instantaneous evolution crosses the 2:3 resonance.}
\end{figure}

Also shown is the shortened overlap computed using a different adiabatic evolution that is chosen to match the instantaneous evolution at the end of the observation. In this case, we see similar behaviour as before: the adiabatic waveform has a high overlap in the region where it is constructed to match the instantaneous evolution; and the boundary of that region is set by the location of the 2:3 resonance.

\subsection{A one-parameter family of adiabatic evolutions}
\label{sec:effres-1p-family}
To be able to quantify the effect that resonances have on EMRIs, we wish to use the full overlap between the instantaneous and adiabatic models computed for the entire length of the observed inspiral. The dephasing means that an adiabatic evolution with the same initial conditions is likely to have a low overlap if a significant resonance is encountered. However, this does not necessarily mean that all adiabatic evolutions have low overlaps; we need to search for optimal adiabatic parameters that maximise the overlap.

The large parameter space of waveforms, coupled with the expensive nature of our 2-torus averaging routine, renders a brute force approach computationally unfeasible. It is therefore necessary to restrict the search in some way. As a preliminary investigation into the importance of resonances, we do this by focussing on a small subset of parameters that we suspect will produce a large overlap. We make the assumption that a good adiabatic model will \emph{exactly} match the instantaneous model at some time $t_{\mathrm{match}}$. This reduces the search to a one-parameter family of waveforms that can easily be computed concurrently with the instantaneous evolution.

The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$. Anticipating that resonances shall play a key role in determining the extent of waveform overlap, we choose matching times situated $5\tau_\mathrm{res}$ after each resonance of interest; we focus on the low-order 2:3 and 1:2 resonances as we expect these to be the most significant. These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral. We also include templates that match at the start of the evolution, so that we have a model that should match the portion of the evolution before the first resonance, and at the end of the evolution, as this is often where most SNR is accumulated. In addition, we include models that match exactly on each of the resonances as these may achieve partial matching both before and after the resonance and hence perform well.

In our experience, we found that these choices were not sufficient to produce a good family of adiabatic waveforms. In particular, if an adiabatic model happened to match the instantaneous model close to the edge of the envelope generated by the fast orbital evolution, then the two trajectories eventually diverged and the resulting waveform provided a poor match. We therefore adjust the values of $t_{\mathrm{match}}$ so that they correspond to times of apoapsis. This ensures that the adiabatic model intersects with the instantaneous model close to the centre of the envelope, as demonstrated by the inset plots in \figref{good-traj}. As a result, we obtain much higher overlaps in general.

We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.

This one-parameter family of adiabatic models does not explore the extrinsic parameter space nor does it consider different values of the constant intrinsic parameters (the masses of the BHs and the central BH spin). It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects. Indeed, systematic biases in parameter estimation are expected when the model waveforms do not accurately reflect the data \citep{cutler_lisa_2007}. The result of these restrictions is that the maximum overlap calculated using our family is a strict lower bound on the true maximum overlap.

\subsection{Loss of signal-to-noise ratio}
For the illustrative EMRI system detailed in \secref{effres-dephasing}, we have computed the family of adiabatic models set by $t_{\mathrm{match}}$ and the resulting trajectories are consistent for the length of the evolution. As they cannot be easily distinguished by eye, we plot the difference (in $E$, $L_z$ and $Q$) between each model and the adiabatic evolution that matches at the start, shown in \figref{res-diff-traj}. The jumps in the orbital parameters due to the 2:3 resonance can be clearly seen, as can the fast orbital oscillations present in the instantaneous evolution but absent in the adiabatics.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{res_diff_traj}
\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (upper plot), $L_z$ (middle) and $Q$ (lower) between each evolution scheme and the adiabatic model that matches at the start. The solid line shows the instantaneous evolution and the dashed lines show the different adiabatic evolutions, which match the instantaneous evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.207), at the 2:3 resonance (0.432), after the 2:3 resonance (0.258) and at the end (0.677).}
\end{figure}

None of the adiabatic models presented here give a particularly high overlap with the instantaneous evolution due to the effects of the resonance. The best-performing adiabatic model was that which matched at the end, giving $\mathcal{O} = 0.677$, while the model that matches at the start gives only $\mathcal{O} = 0.207$.

\subsection{Effect of orbital parameter jumps}
\label{sec:effres-jumps}
The magnitude of the resonance jump depends sensitively on the relative phase of the $r$ and $\theta$ motions. In order to fully understand the effect of resonances on the recovered overlap, it is instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi_0$, as carried out previously in \secref{res-orb-jumps}. Here we use our illustrative system instead, computing the trajectories and waveforms for $100$ values of $\psi_0$, uniformly sampled from $[0,2\pi)$.

For each system, we calculate the size of the resonance jump for $E$, $L_z$ and $Q$. According to \eqnref{res-jump}, these values should be oscillatory with respect to a \emph{single} phase parameter $q_0$. Therefore, plotting the jumps of one parameter against another should produce an ellipse.\footnote{We expect an ellipse rather than a simple linear dependence since the jumps in different parameters may have a phase offset with respect to each other.} Choosing an arbitrary zero-point around the ellipse, we may extract a phase parameter $q$ for each system; this is linearly offset from $q_0$ due to the choice of zero-point. We obtain the same value of $q$ irrespective of which two resonance jumps are used, lending support to the form of \eqnref{res-jump} and giving us confidence in our numerical jump calculations. The relative orbital parameter jumps are shown for each system in \figref{res-jump-vs-q}, as a function of $q$, along with sinusoidal fits to the data. The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset; this means that for every value of $q$, there is always a resonance jump in at least one parameter.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{res_jump_vs_q}
\caption{\label{fig:res-jump-vs-q}The magnitude of the relative orbital parameter jump for our illustrative system as a function of the extracted phase parameter $q$. The individual jumps as well as a sinusoidal fit are plotted for $E$ (blue circles), $L_z$ (gold squares) and $Q$ (green triangles).}
\end{figure}

After extracting the value of $q$, we compute the overlap between the instantaneous waveform and the family of adiabatics given by our set of match times $t_\mathrm{match}$ described in \secref{effres-1p-family}. The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.

The best performing adiabatic for the majority of our systems matched at the end of the inspiral. We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance. We expect the overlap to scale with the fraction of time spent post-resonance,\footnote{The overlap is linear in the inner product $\innerii{s}{h}$ and so scales linearly with time. This is in contrast to the norm of a waveform $\innerii{s}{s}^{1/2}$, which scales with the square-root of time, as shown by \eqnref{EMRI-SNR-estimate}.} the value of which is demonstrated by the red line in \figref{overlap-vs-q}. There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{overlap_vs_q}
\caption{\label{fig:overlap-vs-q}The overlap between the instantaneous and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$. Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase. The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
\end{figure}

Systems with $|q| \approx \pi$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region. The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$. A small fraction of systems have a much lower overlap than expected, at around $0.5$. This is due to a sub-optimal choice of the matching times; making small adjustments to $t_\mathrm{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.

\section{The effect of transient resonances on a population of systems}
\label{sec:effres-population}
Strong resonances can limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic regions, which may be individually undetectable. In order to assess the impact of this on future GW missions, we need to analyse the waveforms from a population of detectable EMRIs.

In \secref{EMRI-detectability}, we produced an astrophysical EMRI population, and generated an approximate waveform for each system using the analytic kludge (AK) model of \citet{barack_lisa_2004}. Calculating SNRs assuming 6 laser links and using the eLISA noise PSD, we found $513$ detectable events from a total of $6333$, assuming an SNR threshold of $15$. We now analyse these $513$ systems using our instantaneous and adiabatic self-force models. The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters, the redshift of the source, and the length of the observation $t_\mathrm{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than 15, which are also shown.

We note that eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs, due to the fact that the periapse in such systems can get much closer to the SMBH, and so the compact object experiences the strongest gravitational fields. This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in the distributions. Systems at higher redshifts start to become undetectable because the amplitude of the signal scales inversely proportional to the luminosity distance to the source, and by $z=1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished. The mass distribution for detectable EMRIs is peaked such that the typical GW frequency occurs at the base of the eLISA noise PSD.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{EMRIpar_dists}
\caption{\label{fig:EMRIpar-dists}The parameter distributions at plunge for our detectable EMRI systems (dark blue), alongside those of the undetectable systems (light pink). In each case, the y-axis values are the probability of a system being found in a particular bin, given that they are either detectable or undetectable. The exception to this is in the final plot for the SNR, where we show the number of systems in each bin. The SNRs quoted here are calculated using the AK model.}
\end{figure}

EMRI systems within our population tend to have small eccentricities at plunge due to the circularising effect of GWs: $85\%$ of detectable systems have $e < 0.1$, with a mean value of $0.05$ and a maximum of $0.4$. In \chapref{resonances}, we found a strong eccentricity dependence on the magnitude of the resonant flux enhancements. We therefore expect typical resonant jumps in these astrophysical systems to be much less than $1\%$, and so the resultant dephasing may be relatively weak.% We quantify this by calculating the loss of overlap in each system.

\subsection{Loss of signal-to-noise ratio}
After computing the SNRs of our NK waveforms for each of the $513$ systems, we can compare the values to those obtained in \secref{EMRI-detectability} using the AK formalism. The different approximation schemes are expected to produce different results, but they should be broadly similar. In \figref{pop-SNR-vs-eta}, we plot the ratio of the new SNR to the AK SNR, as a function of the mass ratio. We observe a clear trend, with systems closest to equal mass performing roughly as expected, but with the most extreme systems giving significantly lower SNRs.

This is a consequence of the PN self-force model we are using, which we have found overestimates the inspiral rate for EMRIs. Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the 2 year observation window, they evolve within the bucket of the eLISA noise curve. Increasing the rate of evolution then does not significantly change the recovered SNR since we are still sensitive to all parts of the waveform. At smaller mass ratios, the systems evolve more slowly, from lower plunge frequencies. An increased evolution rate can shift the EMRI to outside the sensitive region of the detector, thus giving a much lower SNR.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{pop_SNR_vs_eta}
\caption{\label{fig:pop-SNR-vs-eta}The ratio of SNRs as computed with waveforms generated using the NK (adopting our PN self-force model) and AK formalisms, as a function of (the base-10 logarithm of) the mass ratio. Each system has an AK SNR greater than 15, using the eLISA detector with 6 laser links.}
\end{figure}

Nevertheless, we may still study the effect of resonances, by comparing the maximum adiabatic SNR to the instantaneous SNR computed using the NK waveforms. The resulting overlap is still indicative of the effect of resonances on the population of systems.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{pop_adSNR_vs_Tad}
\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family. Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t_\mathrm{insp}$. The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t_\mathrm{ad}$, normalised by $t_\mathrm{insp}$. The solid line corresponds to the expected value if the 1:2 and 2:3 resonances cause significant waveform dephasing.}
\end{figure}

We have calculated the longest period of time, $t_\mathrm{ad}$, during each inspiral, in which neither the 1:2 nor the 2:3 resonance occurs. From the results of \secref{effres-jumps}, we expect the recovered overlap to be approximately given by the proportion of time spent in a resonance-free region, that is $t_\mathrm{ad} / t_\mathrm{insp}$. In \figref{pop-adSNR-vs-Tad}, we plot the maximum overlap between the instantaneous and adiabatic models against this ratio, highlighting the number of resonances, $N_\mathrm{res}$, each system encounters. In the majority of cases, there is no significant impact on the recovered overlap using adiabatic models.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{pop_adSNR_vs_eta}
\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t_\mathrm{ad} / t_\mathrm{insp}$, as a function of (the base-10 logarithm of) the mass ratio.}
\end{figure}

We can better understand the overlap distribution from \figref{pop-adSNR-vs-Tad} by considering the behaviour with the mass ratio. We do this in \figref{pop-adSNR-vs-eta}, plotting the difference between the computed maximum overlap and the value expected should resonances be important. A small proportion of systems have overlaps below the expected value (approximately $4\%$ have values less than $-0.05$). This might be caused by higher-order resonances disrupting the evolution, in which case $t_\mathrm{ad}$ should be smaller and the systems would lie on the expectation line. Alternatively, the low-order resonances might be long-lived, explaining why the recovered overlap is smaller than expected. However, a more likely explanation is that larger adiabatic overlaps are possible with models that we have not considered here (either with different values of $t_\mathrm{match}$ or outside our one-parameter family altogether).% We also suspect that the 2-torus averaging routine for some of these systems has not fully converged.

Roughly $30\%$ of the systems lie within $0.05$ of the expected value. However, the vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity. For the smallest mass ratios, the inspiral rate is slow, and so the systems do not encounter either the 1:2 or 2:3 resonances during their lifetime. Meanwhile for the largest mass ratios, the EMRIs encounter both resonances very close to plunge. In each case, there is a long resonance-free region, allowing a high overlap to be recovered.

The remaining $66\%$ of systems have overlaps above the level expected if resonances were to play an important role. These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t_\mathrm{ad}$ is small. For these EMRIs, resonances are not as important as expected, which is likely due to the low eccentricities of our population leading to small (and in most cases, negligible) resonant flux enhancements.

Even if we assume that all overlap reductions are due to resonances (instead of due to our adiabatic family not being extensive enough) the overall effect on the population is not significant. To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to take into account the resonance behaviour. The total number of detectable systems using a threshold SNR of $15$ decreases from $513$ to $492$. If we increase the threshold, the fractional reduction in the number of detectable systems gets even smaller. We therefore conclude that resonances do not cause sufficient waveform dephasing across a population of EMRIs for the detection rate to be appreciably diminished.

%For the smallest mass ratios (below $\sim\mathcal{O}(10^{-5})$), the inspiral rate is slow, and so the systems are less likely to encounter the 1:2 or 2:3 resonances during their lifetime. Those that do not encounter either have a maximum overlap equal to unity, demonstrating that we can neglect the effect of higher-order resonances. Most of those that only encounter one resonance (specifically this is the 1:2 resonance, which occurs at smaller values of $p$ than the 2:3 resonance) also have overlaps close to $1$, suggesting that we may also neglect the 1:2 resonance for these systems.

%At intermediate values of the mass ratio (between $10^{-5}$ and $10^{-4.4}$), the EMRIs inspiral quickly enough to encounter both resonances within the observation time, but still give overlap values approaching $1$. There is more scatter in this case, but the overlaps remain significantly above the level expected if resonances were important. At the largest values of the mass ratio, the EMRIs encounter both resonances very close to plunge, resulting in a long resonance-free region. We can therefore always recover an overlap of approximately $1$.

%\footnote{The expected value $T_\mathrm{ad} / T_\mathrm{insp}$ used in \figref{pop-adSNR-vs-eta} is calculated assuming both the 1:2 and 2:3 resonances are significant. If we assume only the 2:3 resonance is significant, the predicted values will increase. However, we still compute overlaps that are higher than expected.}. We therefore conclude that neither the 1:2 nor the 2:3 resonance cause sufficient waveform dephasing to result in a significant loss of SNR.

%A small proportion of systems are not well-described by our family of adiabatic models: there are $23$ EMRIs with maximum overlaps less than $0.9$. Some of these are likely to have larger overlaps with adiabatics that we have not considered here (either with different values of $t_\mathrm{match}$ or outside our one-parameter family altogether). We also suspect that the 2-torus averaging routine for some of these systems has not fully converged.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{pop_SNR_dist}
\caption{\label{fig:pop-SNR-dist}Histogram showing the probability distribution function for the detectable EMRI SNRs, as calculated using the AK formalism (light pink) and modified by the maximum adiabatic overlap (dark blue).}
\end{figure}

\section{Summary}
Tranisent resonances can cause severe dephasing in EMRI waveforms, with respect to adiabatic models that are indistinguishable away from the resonance. In \secref{effres-single-system}, we have shown that for one particular EMRI system with a high eccentricity, the resonant flux enhancements are large enough to shift the fundamental frequencies sufficiently that the waveforms promptly dephase. Using adiabatic models that exactly match the evolution at some time, the maximum recoverable overlap is given approximately by the fraction of time spent without encountering a resonance.

Despite this possibility of significant dephasing, we have found that the actual effect of resonances on a realistic population of EMRIs is minimal. For the most extreme mass ratios, the systems evolve too slowly to encounter resonances within the observation window. Detectable EMRIs that are the closest to equal mass evolve so quickly that the resonances occur very close to plunge. This means that the majority of the inspiral can be well-approximated by adiabatic models. Only systems with intermediate values of the mass ratio have the potential to be significantly disrupted by the presence of resonances, but we observe no such disruption. This can be attributed to the low eccentricities of our EMRI population, which give rise to small resonant flux enhancements.

Resonances may still play an important role in EMRI modelling. Our astrophysical population of EMRIs assumes one particular formation channel, which produces EMRIs that are close to circular at plunge. Different formation channels, such as the plunging orbits suggested by \citet{amaro-seoane_role_2013}, can produce populations with higher eccentricities, which may be impacted by resonance effects.

### 2015-05-29 thesis_chapters/resonances.tex
post-viva, post chop house thoughts

\chapter[Transient resonances in EMRIs]{Transient resonances in extreme-mass-ratio inspirals}\label{ch:resonances}

There are three fundamental frequencies $\Omega_k$ ($k=\{r,\,\theta,\,\phi\}$) describing bound orbits in the Kerr spacetime, corresponding to libration-type motions in the radial and polar directions, and a rotation in the azimuthal direction. At infinity, all three fundamental frequencies coincide, tending to the Keplerian frequency, but elsewhere they become distinct, with $\Omega_r < \Omega_\theta < \Omega_\phi$. Frequencies evaluated using different time coordinates can be related to each other by multiplicative factors, and so it follows that there is a well-defined notion of \textit{ratios} of frequencies. An orbit is said to exhibit a resonance when the ratio of any two frequencies is a rational number.

Resonances involving the $\phi$ frequency cannot affect the local dynamics of an EMRI system due to the axisymmetry of the Kerr spacetime. Any dynamical quantity $q$ must depend only on $r$ and $\theta$ along the worldline, and so may be Fourier expanded in coordinate time according to
\begin{equation}
\label{eq:res-dynamical-exp}
q(t) = \sum_{n,m} q_{nm} \exp\left[-\iu\left( n\Omega_r + m\Omega_\theta \right)t\right],
\end{equation}
with no dependence on $\Omega_\phi$. However, it has been shown that both $\theta\phi$-resonances \citep{hirata_resonant_2011} and $r\phi$-resonances \citep{van_de_meent_resonantly_2014} can lead to extrinsic effects. The GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host.

We shall be interested in systems that exhibit $r\theta$-resonances, such that
\begin{equation}
\label{eq:res-def}
\zeta \equiv \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r},
\end{equation}
where $n_\theta$ and $n_r$ are integers with no common multiples; we shall refer to such a system as being in a $n_\theta$:$n_r$ resonance. The effects of these resonances were first studied by \citet{flanagan_transient_2012}, who calculated the non-perturbative `jumps' in the orbital parameters as the evolution of an EMRI passes through a resonance. These occur because the two-frequency Fourier expansion in \eqnref{res-dynamical-exp} becomes invalid on resonance and should be replaced by a single-frequency series. This results in enhanced or diminished GW fluxes on resonance. \Citet{gair_resonances_2012} considered two toy models for this behaviour, differing in whether the phase or the frequency components should be continuous across the resonance. It is not fully understood which of these models provides the better description.

It is possible for resonances to capture an orbit and thus become long-lived, but such sustained resonances are likely to be very rare and so unobservable with typical GW detectors \citep{van_de_meent_conditions_2014}. In contrast, short-lived transient resonances have been found to be near-ubiquitous in EMRI systems \citep{ruangsri_census_2014} and are located in the strong-field regime close to the SMBH \citep{brink_astrophysics_2015, brink_orbital_2015}, where GW emission is significant. Typical variations of GW fluxes due to resonances are of the order of a few percent \citep{flanagan_resonantly_2014}. We extend this existing body of work by considering the effect of transient resonances on the detectability of EMRIs using GWs. However, we first discuss our method of including resonances within an evolution scheme, and provide an exploration of the relevant properties of resonances.

\section{Evolution scheme}
\label{sec:EMRI-evolution-scheme}
For generic EMRIs, there are two timescales of interest \citep{hinderer_two-timescale_2008}: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$. This makes it ideally suited to the method of osculating elements~\citep{gair_forced_2011}: on short timescales, we analyse the unperturbed system resulting in geodesic motion; while the long-term evolution is described by a sequence of instantaneous geodesics. We require, at each instant in time, that the chosen geodesic matches the true position and velocity of the particle. This amounts to a specific choice of the orbital shape parameters and the positional elements, collectively referred to as \emph{osculating elements} and denoted by $I^A$. For our evolution, we choose to consider $I^A = \{p(t),\,e(t),\,\cos\iota(t),\,\psi_0(t),\chi_0(t),\phi_0(t)\}$, making explicit the variation with time. These quantities are defined in \secref{kerr-geodesic}.

Given forced motion in a background spacetime, the acceleration
\begin{equation}
f^\alpha = \difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau},
\end{equation}
can be used to calculate the evolution of the osculating elements $\dot{I}^A(f^\alpha)$. The specific equations for Kerr are derived by \citet{gair_forced_2011}. Using these equations and the notation of \secref{kerr-geodesic}, the resulting system of differential equations to be solved takes the form
\begin{subequations}
\label{eq:kerr-osc-el-ev}
\begin{align}
\label{eq:kerr-osc-el-ev-I}
\diff{I^A}{t} &= \dot{I}^A\left(\psi(t),\,\chi(t);\,I^A\right),\quad &I^A(0) &= I^A_0,\\
\diff{\psi}{t} &= \dot{\lambda} \frac{p}{r^2 e} \sqrt{\frac{R(r)}{1-\cos^2\psi}} - \diff{\psi_0}{t},\quad &\psi(0) &= \psi_0(0),\\
\diff{\chi}{t} &= \dot{\lambda} \sqrt{\frac{(1-\cos^2\theta_\mathrm{inc}\cos^2\chi)\Theta(\theta)}{\cos^2\theta_\mathrm{inc}\sin^2\chi}} - \diff{\chi_0}{t},\quad &\chi(0) &= \chi_0(0),
\end{align}
\end{subequations}
where $\dot{\lambda} = (\dd \tau/\dd t)/\Sigma$ is the coordinate-time derivative of Mino time, which can be evaluated using \eqnref{kerr-geodesics-t}. If desired, we may also compute $\tau$ and $\phi$ by evolving them directly according to \eqnref{kerr-geodesics}. We perform the integration numerically using the standard fourth order Runge--Kutta method. We choose a step size equal to the minimum of $5\,\mathrm{s}$ and $2\pi/(8\Omega_r)$, which ensures that the fast orbital motion is modelled sufficiently accurately.\footnote{In \chapref{effect-resonances}, we compute the GWs from these systems. The fixed maximum step size of $5\,\mathrm{s}$ is chosen to ensure that the high frequency cut-off in the emitted GWs is above the most sensitive frequency of typical detectors.}

%This formalism can be used to study many kinds of small perturbations, including the effect of a distant massive object (that might be expected if an EMRI occurs during a galactic merger), the role of internal structure or spin of the compact object, and deviations of the metric from GR.

\subsection{Gravitational self-force model}
We are interested in using this evolution scheme to study the perturbation arising from the curvature generated by the particle itself. We cannot use established adiabatic techniques to describe the inspiral as the assumptions made are incompatible with transient $r\theta$-resonances. Instead, we choose to work directly with the gravitational self-force, using the same PN approximation as \citet{flanagan_transient_2012}. For comparison, \citet{flanagan_resonantly_2014} use a Teukolsky equation calculation of resonant GW fluxes.

The self-force model we adopt uses the first-order PN terms of the dissipative self-force formulated by \citet{flanagan_evolution_2007} and the conservative force formulated by \citet{iyer_post-newtonian_1993}, and \citet{kidder_coalescing_1995}. Since only the first PN terms are used, this prescription can only be of limited validity in strong fields. Both pieces of the self-force are computed assuming that the spin is small: the dissipative piece contains terms to $\order{a^2}$ and the conservative piece to $\order{a}$. This is less than ideal for high spins. While this approximate self-force is not perfect, it should serve as a guide for the behaviour of the full self-force, allowing us to assess the qualitative impact of resonances on EMRI detection.

\subsection{Adiabatic averaging}
\label{sec:res-ad-averaging}
We can study the effect of resonances by comparing an adiabatic evolution with an instantaneous self-force evolution. Using adiabatic results from the literature is useful to test the efficacy of the self-force model, but is not ideal in this case. Differences between the evolutions may arise due to the different approaches being taken, rather than as a result of transient resonances. We therefore, instead, construct an appropriate adiabatic trajectory using our self-force model, by explicitly averaging the fluxes\footnote{We use the term flux here to denote the rate of change of orbital parameters. For the case of $E$ and $L_z$, this is equivalent to (minus) the physical flux of those quantities carried by the emitted GWs, but a similar physical interpretation does not exist for the other parameters.} of our osculating elements over the $r\theta$ $2$-torus. Computing this average is trivial  if the quantity is parameterised in terms of the generalised action-angle variables $\{q_r, q_\theta\}$,
\begin{equation}
\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta} = \frac{1}{(2\pi)^2}\intd{0}{2\pi}{\intd{0}{2\pi}{\diff{X}{\lambda}}{q_r}}{q_\theta}.
\end{equation}
However, we are using $\psi$ and $\chi$, as these are simpler to evolve. Furthermore, using the osculating elements formalism, we compute instantaneous coordinate-time fluxes $\dot{X}$, not Mino-time fluxes. Changing variables gives an average of \citep{drasco_computing_2005}
\begin{align}
\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta} = {} & \frac{1}{\Lambda_r \Lambda_\theta}\intd{0}{2\pi}{\intd{0}{2\pi}{\left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-2} \diff{X}{\lambda}}{\psi}}{\chi} \nonumber\\
 = {} & \frac{1}{\Lambda_r \Lambda_\theta}\intd{0}{2\pi}{\intd{0}{2\pi}{\left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X}}{\psi}}{\chi}.
\label{eq:2torus-average}
\end{align}
This average describes the Mino-time rate of change of the quantity $X$ over an orbit. To convert to a coordinate-time flux of the averaged quantity, we simply divide by the period $\Gamma$, defining
\begin{equation}
\dot{\left\langle X\right\rangle}_{q_r,\,q_\theta} = \frac{1}{\Gamma}\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta}.
\end{equation}
It is convenient to calculate $\Gamma$ as
\begin{equation}
\Gamma = \left\langle \diff{t}{\lambda}\right\rangle_{q_r,\,q_\theta},
\end{equation}
using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.

We compute these integrals using a $N\times M$ grid of $(\psi,\chi)$ values and employing a Newton--Cotes approximation in each dimension. Specifically we take $N = M = 300$, and approximate each integral by
\begin{align}
\intd{x_1}{x_n}{f(x)}{x} \approx \delta x \bigg(&\frac{17}{48}f(x_1) + \frac{59}{48}f(x_2) + \frac{43}{48}f(x_3) + \frac{49}{48}f(x_4)\nonumber\\
& + \left\{f(x_5) + \ldots + f(x_{n-4})\right\}\nonumber\\
& + \frac{49}{48}f(x_{n-3}) + \frac{43}{48}f(x_{n-2}) + \frac{59}{48}f(x_{n-1}) +  \frac{17}{48}f(x_n)\bigg),
\end{align}
where $x_i = x_1 + (i-1)\delta x$, $x_1 = 0$, and $\delta x = 2\pi / (N-1)$. This procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform. The adiabatic derivatives vary on much longer timescales than the orbital motion, and so in practice, we can make use of interpolation. We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\dd t/\eta$ where $\dd t$ is the orbital time-step. When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives. Once the system evolves close to one of the boundaries, we update the interpolation points, taking the closest boundary to be the new central point.

%We compute these integrals using a grid of $(\psi,\chi)$ values: we split each integral into $25$ equal segments and then evaluate each of those segments using Boole's rule, which is a specific case of the Newton-Cotes method\footnote{Newton-Cotes formulae for approximating integrals are derived by fitting a polynomial to the integrand using equally spaced intermediate values. The trapezium rule and Simpson's rule are well-known examples, which use linear and quadratic fitting functions respectively. Boole's rule uses a quartic polynomial.}. The integral of $f(x)$ between $x_1$ and $x_5$ is approximated by
%\begin{equation}
%\intd{x_1}{x_5}{f(x)}{x} \approx \frac{2 h}{45}\left( 7f(x_1) + 32 f(x_2) + 12 f(x_3) + 32 f(x_4) + 7f(x_5) \right),
%\end{equation}
%where $x_i = x_1 + (i-1)h$. This procedure requires $101\times 101$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform. The adiabatic derivatives vary on much longer timescales than the orbital motion, so in practice, we can make use of interpolation. We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $dt/\sqrt{\eta}$ where $dt$ is the orbital time-step\footnote{The evolution timescale of an EMRI $\sim 1/\eta$ and so we could space the interpolation points further apart; we choose $1/\sqrt{\eta}$ instead to be on the cautious side.}. When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives. Once the system evolves close to one of the boundaries, we update the interpolation points, taking the closest boundary to be the new central point.

To demonstrate the approach, we calculate the instantaneous evolution and the 2-torus-averaged adiabatic evolution over $2$ years for a set of initial parameters chosen to avoid any significant resonances. Specifically, we choose: the mass of the compact object $\mu = 10 M_{\odot}$; the mass ratio $\eta = 3\times 10^{-6}$; the spin of the central BH $a=0.95$; the orbital shape parameters $p=7.5$, $e=0.7$ and $\cos \iota = 0.5$; and the redshift of the source $z=0.204$ (corresponding to $1\;\mathrm{Gpc}$ in our chosen $\Lambda$CDM cosmology, used in \secref{EMRI-detectability}). The remaining intrinsic (the initial phases $\psi_0$ and $\chi_0$) and extrinsic parameters are set to arbitrary values. The resulting evolutions of $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, where it can be seen that the adiabatic evolution closely matches the instantaneous evolution on longer inspiral timescales. The inset plots show the start of the evolution, on a timescale associated with the orbital motion; the 2-torus averaging explicitly smoothes out the visible structure on this scale.

%The resulting averaged fluxes successfully describe the leading-order secular evolution of the trajectory, as illustrated in \figref{good-traj} by an EMRI that does not encounter any significant resonances.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.7\textwidth]{good_traj_E}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{good_traj_Lz}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{good_traj_Q}}
\caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (upper figure), $L_z$ (middle) and $Q$ (lower) under the instantaneous (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances. The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.}
\end{figure}

%\subsection{Quasi-adiabatic averaging}
%The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$. The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurence of any transient resonances. Resonances still occur as the fundamental frequencies vary, but as the system passes through each resonance, there is no additional effect on the evolution of the system.

%Comparing instantaneous and adiabatic evolutions may be a useful exercise to calibrate and verify adiabatic models that can be generated cheaply; however it cannot be used to distinguish between dephasing effects arising from resonances and those arising from conservative corrections to the phase variables, as both are explicitly eliminated from the adiabatic prescription. We therefore take a slightly different approach, comparing the instantaneous evolution with a quasi-adiabatic evolution, generated by averaging the full available self-force over the $2$-torus, rather than just averaging the term $G_\alpha^{(1)}$.

%To be specific, returning to the action-angle formalism, we write \eqnref{EMRI-AAsf} as
%\begin{subequations}
%\label{eq:res-AAsf}
%\begin{align}
%\diff{q_\alpha}{\tau} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
%\diff{J_\alpha}{\tau} = {} & \eta \langle G_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
%\end{align}
%\end{subequations}
%where $g_\alpha$ and $G_\alpha$ encapsulate all our knowledge about the self-force and may in principle contain contributions from sub-leading terms in the mass-ratio expansion. The quasi-adiabatic evolution is then obtained by dropping the terms containing $\hat{g}_\alpha$ and $\hat{G}_\alpha$. For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an ``adiabatic evolution'' in fact refer to a ``quasi-adiabatic evolution''.

\section{Properties of transient resonances}
%The combination of an instantaneous self-force evolution and a 2-torus-averaged adiabatic evolution allows us to systematically study the effect of transient resonances on EMRIs over the course of an inspiral.% Before approaching this problem, we first investigate the properties of the resonances themselves.

\subsection{Location of resonances}
\label{sec:res-location}
In order to efficiently study transient resonances, it is essential to know where they occur in parameter space. A system exhibiting a $n_\theta$:$n_r$ resonance must obey the constraint
\begin{equation}
\label{eq:res-condition}
\Omega \equiv n_r \Omega_r(p,e,\cos\iota;a) - n_\theta \Omega_\theta(p,e,\cos\iota;a) = 0,
\end{equation}
where we have made explicit the dependence of the frequencies on $a$. The frequencies depend on the orbital shape parameters via elliptic integrals and so cannot be explicitly inverted. We instead adopt a numerical root-finding algorithm to calculate the value of $p$ such that the resonance condition holds. The mapping from the orbital shape parameters to the orbital frequencies is not one-to-one; there exist isofrequency pairings where two different sets of parameters result in the same frequencies \citep{warburton_isofrequency_2013}. One of these pairs is located very close to the SMBH in a region where we expect the orbits to be evolving very quickly. We therefore focus on the more distant resonances, as these are still within the strong-field regime of the spacetime, but are more long-lived.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=\textwidth]{resplane_a}}\\
\vspace{20pt}
\subfloat{\includegraphics[width=\textwidth]{resplane_res}}
\caption{\label{fig:res-resplane}Resonance surfaces, defined by \eqnref{res-condition}, in the orbital shape parameter space $\{p,\,e,\,\cos\iota\}$. The upper figure demonstrates the 2:3 resonance, at different values of the SMBH spin, while the lower figure fixes the spin at $a=0.95$ and depicts the surfaces for different resonances.}
\end{figure}

\Figref{res-resplane} shows the resonance surfaces in the orbital shape parameter space, for different values of the spin, and for different resonances. The surfaces are closely approximated by planes and so it is natural to consider whether some simple functional form can predict their location. For a given resonance ratio $\zeta$, the semi-latus rectum can be found using
\begin{equation}
\label{eq:res-approx-p}
p(e,\cos\iota;a,\zeta) \simeq A\frac{1 + B e + D \cos\iota}{1 - C\exp(e)},
\end{equation}
where the coefficients are given by
\begin{subequations}
\begin{align} 
A(a,\zeta) \simeq {} & a_0\frac{1 + a_1\zeta - a_2 \zeta^2 - a_3 \zeta a^2}{1 + a_4\zeta - (1 + a_4)\zeta^2}, \\
B(a,\zeta) \simeq {} & b_0(1 - b_1\zeta)\exp(-b_2\zeta)(1 - b_3 a), \\
C(a,\zeta) \simeq {} & c_0, \\
D(a,\zeta) \simeq {} & d_0\left[1 - \exp(a)\right]\left[1 - d_1\exp(\zeta)\right].
\end{align}
\end{subequations}

We generate resonance location data for all possible resonances with $2 \leq n_r \leq 7$, as well as $\zeta = 9/10$, $19/20$, $49/50$ and $99/100$, for SMBH spins between $0.01 \leq a \leq 0.999$, eccentricities in the range $0.01 \leq e \leq 0.99$ and inclinations with $-0.999999 \leq \cos\iota \leq 0.999999$. We then fit this data to the functional form in \eqnref{res-approx-p}, obtaining optimised parameters
\begin{equation}
\begin{array}{lll}
a_0 = 5.9854, & a_1 = 3.4116, & a_2 = 0.9253,\\
a_3 = 0.1959, & a_4 = 4.8846, & b_0 = 0.7692,\\
b_1 = 1.4752, & b_2 = 1.4861, & b_3 = 0.5974,\\
c_0 = 0.02332, & d_0 = 0.7968, & d_1 = 0.3115.
\end{array}
\end{equation}
This gives an approximation to the true resonant semi-latus rectum, which provides a convenient starting point for numerical root-finding algorithms. Our $12$-parameter model might be expected to overfit the data. However, we note that a simple quadratic expansion in the four variables requires $15$ free parameters and yet gives a worse fit to the data. In \figref{res-reserror}, we show the fractional error in $p$ for our fit, marginalising over $e$ and $\cos\iota$ by considering both the maximal value and the root-mean-square (RMS). Typical errors are a few percent, while the largest relative errors occur for highly spinning BHs as $\zeta\rightarrow 0$. In these cases, the value of $p$ is small such that the absolute error is less than $1$.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.92\textwidth]{reserror_max}}\\
\subfloat{\includegraphics[width=0.92\textwidth]{reserror_rms}}
\caption{\label{fig:res-reserror}The relative error in $p$ of our approximation to the location of resonances. At each value of $a$ and $\zeta$, we compute the absolute maximum (upper figure) and RMS (lower) error over $e$ and $\cos\iota$. Lighter shades correspond to larger values of the error.}
\end{figure}

%The approximant given by \eqnref{res-approx-p} should not be taken as an exact location for any given resonance, but instead as an indication of the region of parameter space in which resonances occur, providing a convenient starting point for numerical root-finding algorithms.

\subsection{Resonance timescales}
\label{sec:res-timescales}
Resonant phenomena occur on timescales intermediate to the orbital timescale $\tau_\mathrm{orb} \sim 2\pi/\Omega_k$ and the evolution timescale $\tau_\mathrm{ev} \sim \zeta/\dot{\zeta} \sim \mathcal{O}(\tau_\mathrm{orb}/\eta)$. We can describe them by considering the phase-space precession inherent to non-resonant orbits but absent for resonant trajectories. We define the precession timescale
\begin{equation}
\tau_\mathrm{pres}(t) = \frac{2\pi}{|\Omega(t)|},
\end{equation}
where $\Omega(t)$ is defined by \eqnref{res-condition} and $t$ is the coordinate time of an EMRI evolution, taken to be zero on resonance. This is a measure of how long a particular orbit takes to come close to every point within the $r\theta$ 2-torus. This timescale is infinite exactly on resonance and decreases as the orbit evolves away. A suitable resonance timescale $\tau_\mathrm{res}$ can then be determined by considering the time at which the orbit has had sufficient opportunity to significantly precess. This is given by
\begin{align}
\tau_\mathrm{res} = \tau_\mathrm{pres}(\tau_\mathrm{res}) = \frac{2\pi}{|\Omega(\tau_\mathrm{res})|}.
\end{align}
Taylor expanding $\Omega(t)$ about resonance at $t=0$ and taking an adiabatic average of $\dot{\Omega}$ as we are mainly interested in the long-term secular evolution here, we obtain
\begin{align}
\tau_\mathrm{res} \simeq \frac{2\pi}{\left|\left\langle\dot{\Omega}(0)\right\rangle_{q_r,\,q_\theta}\right| \tau_\mathrm{res}},
\end{align}
which can be rearranged to give
\begin{align}
\label{eq:res-timescale}
\tau_\mathrm{res} = \left[ \frac{2\pi}{\left|\left\langle\dot{\Omega}(0)\right\rangle_{q_r,\,q_\theta}\right|} \right]^{1/2}.
\end{align}
In terms of the mass ratio, we see that the resonance timescale $\tau_\mathrm{res} \sim \mathcal{O}(\eta^{-1/2}\tau_\mathrm{orb}) \sim \mathcal{O}(\eta^{1/2}\tau_\mathrm{ev})$ and so acts as a bridge between the other two timescales.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.33\textwidth]{geotraj_1}}
\subfloat{\includegraphics[width=0.33\textwidth]{geotraj_2}}
\subfloat{\includegraphics[width=0.33\textwidth]{geotraj_3}}
\caption{\label{fig:res-geotraj}Three geodesic trajectories for an EMRI with SMBH mass $M = 1.2\times10^6 M_\odot$, spin $a=0.95$, compact object mass $\mu = 12 M_\odot$, eccentricity $e = 0.7$, and inclination $\cos\iota = 0.2$. The left-hand plot has a semi-latus rectum $p = 9.398644$, which corresponds to a 2:3 resonance, with a resonance timescale $\tau_\mathrm{res}\approx0.02\,\mathrm{yrs}$. The central and right-hand plots have initial conditions given by an inspiralling evolution $\tau_\mathrm{res}/2$ and $\tau_\mathrm{res}$ after resonance respectively. Each geodesic is shown for a finite time $\tau_\mathrm{res}/2$ only.}
\end{figure}

We demonstrate these results by considering geodesic orbits nearby to a resonance. \Figref{res-geotraj} shows three `snapshot' orbits, each evolved for the same amount of time (in this case, we evolve for half a resonance timescale). The first trajectory is exactly on a 2:3 resonance and it can be seen that no precession takes place. The second and third trajectories have initial conditions that correspond with an inspiralling orbit at a time $T$ after passing through the 2:3 resonance (we take $T$ to be $\tau_\mathrm{res}/2$ and $\tau_\mathrm{res}$ for the two cases respectively). The resonance has an impact on the evolution for approximately a resonance timescale $\tau_\mathrm{res}$, after which the 2-torus is quickly filled by the trajectory, validating the adiabatic approach.

\subsection{Resonant flux enhancements}
Evolving through a region that does not contain a significant resonance,\footnote{A significant resonance noticeably affects the evolution of the EMRI. We intuitively expect only low-order resonances to be significant; this can be seen by considering the resonance timescale. High-order resonances have large values of $n_r + n_\theta$ and so have smaller values of $\tau_\mathrm{res}$, from \eqnref{res-timescale}. The resonance is not significant unless $\tau_\mathrm{res} > \tau_\mathrm{orb}$.} the trajectory quickly fills the $r\theta$ 2-torus and we may approximate the motion by replacing $\dd I^A/\dd t$ with $\langle\dd I^A/\dd t\rangle_{q_r,\,q_\theta}$, producing the adiabatic evolution. However, evolving through a resonance, this average is not a good approximation. With respect to the adiabatic case, there is an enhancement (or decrement) of the fluxes that depends on the relative phase of $q_r$ and $q_\theta$. After crossing the resonance region, we therefore expect to find different orbital parameters to those of an equivalent adiabatic trajectory. This difference is given by \citet{flanagan_transient_2012} as
\begin{equation}
\label{eq:res-jump}
\Delta I^A_\mathrm{jump} = \eta\sum_{s \neq 0} F^A_s(\boldsymbol{I}) \left[ \frac{2\pi}{\left|s\left\langle\dot{\Omega}\right\rangle_{q_r,\,q_\theta}\right|} \right]^{1/2} \exp\left[\iu (s q_0 + \frac{\pi}{4}\sgn s\dot{\Omega})\right],
\end{equation}
where $q_0$ is the adiabatic phase associated with the frequency $\Omega$ on resonance, and $F^A_s$ is the Fourier component of the self-force, defined such that
\begin{equation}
\diff{I^A}{t} = \eta\sum_s F^A_s(\boldsymbol{I})\exp(\iu s q).
\end{equation}
This result can be derived using matched asymptotic expansions, and can be understood as the product of the forcing function on resonance with the resonance timescale.

We use two equivalent methods to numerically demonstrate the effects of \eqnref{res-jump}. First, we evolve orbits through resonances and compute the jumps directly by considering the orbital parameters before and after the resonance. Second, we calculate the resonant fluxes along fixed geodesics.

\subsubsection{Orbital parameter jumps}
\label{sec:res-orb-jumps}
We consider a collection of EMRIs a short time before crossing the 2:3 resonance. At this point, we are not interested in the emitted GWs and so only the intrinsic parameters are relevant. For each system, we choose a mass ratio $\eta = 3\times 10^{-6}$, eccentricity $e=0.7$, and inclination $\cos\iota = 0.2$. We then vary the spin to take values $a = j/10$ for $j=[1,9]$, as well as $a = \{0.01,\,0.95,\,0.99\}$. We also vary the initial radial phase, choosing $\psi_0$ to take 50 equally spaced values between $0$ and $2\pi$. We keep $\chi_0 = 0$ fixed, and so this should span the full range of $q_0$ in \eqnref{res-jump}, enabling us to estimate the magnitude of $\Delta I^A_\mathrm{jump}$.

We evolve each of these systems through the resonance using the instantaneous self-force model and the 2-torus-averaged adiabatic model, as detailed in \secref{EMRI-evolution-scheme}. When the system encounters the 2:3 resonance, the instantaneous evolution undergoes a rapid change in the orbital parameters with respect to the adiabatic evolution, as shown for the orbital energy in \figref{res-jump-calc}. To extract the magnitude of the jump from the trajectory data, we first must account for the fast orbital oscillations as well as ensure that we include the full effect of the resonance, while minimising any contributions resulting from other sources.

To achieve this, we isolate spans of data away from the resonance. Specifically, we consider data that lies within five and ten resonance timescales away. For both the pre- and post-resonance regions, we compute $\Delta I^A(t) \equiv I^A_{\mathrm{inst}}(t) - I^A_\mathrm{ad}(t)$ for each of the orbital parameters. We then identify the most-horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations. The dashed green and yellow lines in \figref{res-jump-calc} demonstrate this bounding procedure.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{res_jump_calc}
\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and an adiabatic model, scaled by the magnitude of the 2:3 resonance jump. The thickness of the (blue) line is due to the fast orbital oscillations, which cannot be seen on this scale. The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump. The dotted red line indicates the computed size of the jump. The time axis is centred on the 2:3 resonance and is scaled by the resonance timescale.}
\end{figure}

To estimate the size of the jump $\Delta I^A_{\mathrm{jump}}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits. This allows us to take into account any linear drift in $\Delta I^A(t)$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation. %In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous\footnote{This is not a concern for the current systems being studied, but in \chapref{effect-resonances}, we study resonances that occur very near to plunge and do not follow the typical behavior described here.}. 
Our numerically computed jump appears to suitably describe the behaviour through resonance, as shown by the red dotted line in \figref{res-jump-calc}.

We compute the 2:3 resonance jump in this manner for each EMRI system in our collection. The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems. Instead, we can express it relative to the adiabatic change expected across the resonance. Using the resonance timescale from \eqnref{res-timescale} to set the length of the resonance, we compute the ratio for each orbital parameter as
\begin{equation}
\label{eq:res-jump-ratio}
\frac{\Delta I^A_\mathrm{jump}}{\left\langle\Delta I^A\right\rangle} = \frac{\Delta I^A_\mathrm{jump}}{I^A_\mathrm{ad}(t_\mathrm{res} + \tau_\mathrm{res}/2)-I^A_\mathrm{ad}(t_\mathrm{res} - \tau_\mathrm{res}/2)} \approx \frac{\Delta I^A_\mathrm{jump}}{\dot{I}^A_\mathrm{ad}(t_\mathrm{res})\tau_\mathrm{res}}.
\end{equation}
As a function of $\psi_0$, we expect these jumps to be oscillatory, as a result of \eqnref{res-jump}. In \figref{res-jump-vs-psi0}, we display the computed relative resonance jumps in energy, for three of our chosen values of the BH spin. The oscillations are sinusoidal, with a spin-dependent amplitude $|\Delta I^A_\mathrm{jump}|$ that can be calculated by taking half the difference between the maximum and minimum values. %The phase difference between the cases is due to the offset between $\psi_0$ and the phase on resonance $q_0$, which is different for each value of $a$ due to the different inspiral rates.
We plot the relative magnitudes as a function of BH spin in \figref{res-jump-vs-a}; typical jumps are of the order of a percent. The negative values indicate that the adiabatic evolution has a negative derivative.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.92\textwidth]{res_jump_vs_psi0}
\caption{\label{fig:res-jump-vs-psi0}The relative jump in energy, calculated using \eqnref{res-jump-ratio}, due to crossing the 2:3 resonance, as a function of the initial radial phase $\psi_0$, which acts as a proxy for the relative phase on resonance $q_0$. The jumps are shown for three values of the BH spin, $a = 0.1$ (blue), $0.5$ (gold) and $0.95$ (green) with thicker lines representing higher values of spin.}
\end{figure}

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.7\textwidth]{res_jump_E_vs_a}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{res_jump_Lz_vs_a}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{res_jump_Q_vs_a}}
\caption{\label{fig:res-jump-vs-a}The magnitude of the relative jump in energy (upper figure), angular momentum (middle) and Carter constant (lower), due to crossing the 2:3 resonance, as a function of BH spin $a$. Data is shown for the jump method (solid line), discussed in \secref{res-orb-jumps}, and for the flux method (dashed), discussed in \secref{res-geo-fluxes}.}
\end{figure}

This method of studying the resonance jumps is conceptually straightforward, since we are directly measuring the quantity of interest. However, for jumps that are much smaller than the width of the fast orbital oscillations, it can be difficult to extract the jump magnitude from the data. This is especially true when resonances occur close to plunge. There is an alternative way to compute the relative resonance jumps without explicitly evolving the system, which we now discuss.

\subsubsection{Fluxes on resonant geodesics}
\label{sec:res-geo-fluxes}
We have previously evolved generic Kerr inspirals by solving the differential system given by \eqnref{kerr-osc-el-ev}. We now seek to evolve a different, but closely-related, system representing a fixed geodesic trajectory, while still considering the effects of the self-force. We replace \eqnref{kerr-osc-el-ev-I} with
\begin{subequations}
\begin{align}
\diff{P^A}{t} &= \dot{I}^A\left(\psi(t),\,\chi(t);\,I^A\right),\quad &P^A(0) &= 0,\\
\diff{I^A}{t} &= 0,\quad &I^A(0) &= I^A_0,
\end{align}
\end{subequations}
where $P^A$ are auxilliary variables used to compute the integrated fluxes along an orbit but that have no impact on the evolution. We are interested in the secular accumulation of the osculating elements arising from the self-force during a resonance, and how this depends on the relative phase between the $r$ and $\theta$ motions. In order to average out the fast orbital motion, we compute the long-time average,
\begin{equation}
\label{eq:res-geo-flux-diff}
\dot{I}^A_\mathrm{inst} \equiv \left\langle \dot{P}^A \right\rangle_T = \lim_{T\rightarrow\infty} \frac{P^A(T)}{T},
\end{equation}
for each osculating element. For non-resonant geodesics, this quantity will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference $\psi_0 - \chi_0$ (as a proxy for $q_0$).

We consider a selection of different orbital shape parameters corresponding to low-order transient resonances and calculate $\dot{I}^A_\mathrm{inst}$ for $50$ different initial phases $\psi_0$. In practice, to imitate the limiting behaviour at infinite times, we evolve for an integer number of cycles of both the radial and polar motions. Specifically, when considering a $n_\theta$:$n_r$ resonance, we take $T = 2\pi N n_\theta/\Omega_r = 2\pi N n_r/\Omega_\theta$ for some integer $N$. We have found that taking $N=5$ gives sufficiently accurate results.

Averaging $\dot{I}^A_\mathrm{inst}$ over the phase difference is equivalent to evolving under the adiabatic approximation. We can therefore estimate the relative orbital parameter jumps as
\begin{equation}
\label{eq:res-flux-enhnc}
\frac{|\Delta I^A_\mathrm{jump}|}{\left\langle\Delta I^A\right\rangle} \approx \frac{\max \dot{I}^A_\mathrm{inst} - \min \dot{I}^A_\mathrm{inst} }{2 \left\langle \dot{I}^A_\mathrm{inst} \right\rangle},
\end{equation}
where the maximum, minimum and mean values on the right-hand side are taken over $\psi_0$. Computing the resonant flux enhancements in this way for the same systems as discussed in \secref{res-orb-jumps}, we find good agreement between the two methods, as demonstrated in \figref{res-jump-vs-a}.

We now compute the $p$, $e$ and $\cos\iota$ flux enhancements for the 2:3 resonance with eccentricity $e=0.3$ and varying spin and inclination, displaying the results in \figref{res-flux-2to3-e03}. Typical enhancements in $p$ and $e$ are around $0.1\%$ and $1\%$ respectively, although they can be much smaller for orbits that are nearly equatorial. Intuitively the volume of space occupied by the $r\theta$ 2-torus of allowed motions is smaller for nearly equatorial orbits. Resonant orbits therefore come close to every allowed point, thus approximating a non-resonant orbit. We note that the signs of the values for $p$ and $e$ are negative, indicating that the adiabatic fluxes are both negative in this case.

It is expected that generic inclined orbits will adiabatically evolve to larger inclinations; in the weak-field, circular limit, the 2.5PN expansion is given up through $\mathcal{O}(e^2)$ by \citet{ganz_adiabatic_2007} as
\begin{align}
\left\langle \diff{\cos\iota}{t} \right\rangle = &-\frac{244}{15} \eta p^{-11/2} a (1-e^2)^{3/2} \sin^2\iota \bigg[\left(1+\frac{189}{61}e^2 \right)\nonumber\\
 &- \left(\frac{13}{244} + \frac{277}{244}e^2\right)a p^{-1/2}\cos\iota - \left(\frac{10461}{1708}+\frac{83723}{3416}e^2\right)p^{-1}\bigg],
\end{align}
indicating that to leading order, $\dot{\iota} \propto \sin\iota$ and is positive. This is confirmed for circular orbits in the strong-field regime using Teukolsky results \citep{hughes_evolution_2000}. In contrast, we observe regions of decreasing $\iota$ (increasing $\cos\iota$) for certain retrograde orbits; \figref{res-flux-2to3-e03} demonstrates such a region for eccentric orbits, but the behaviour is found to persist as $e\rightarrow 0$. We attribute this property of our evolutions to the specific self-force model being used, and note that the changes in inclination are small across the duration of a typical observation. Neglecting the anomalous retrograde orbits on the boundary between the regions of positive and negative adiabatic flux (such that the relative flux enhancements diverge), we find typical relative flux enhancements of $\cos\iota$ at the $1\%$ level.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.7\textwidth]{res_flux_2to3_e03_p}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{res_flux_2to3_e03_e}}\\
\subfloat{\includegraphics[width=0.7\textwidth]{res_flux_2to3_e03_cosi}}
\caption{\label{fig:res-flux-2to3-e03}Resonant flux enhancements for a 2:3 resonance with $e=0.3$ and for varying values of $a$ and $\cos\iota$. We display the enhancements for $p$ (upper figure), $e$ (middle) and $\cos\iota$ (lower). Larger enhancements are shaded darker, while negative and positive regions are blue and red respectively; the only positive values are found in the lower plot for the most retrograde systems.}
\end{figure}

Around marginally spinning BHs, as $a\rightarrow 0$, the absolute enhancements become symmetric with respect to $\cos\iota$, as expected. The smallest value of spin that we consider is $a=0.01$; for $a=0$ exactly, the enhancements vanish since we may always choose coordinates such that the motion lies in the equatorial plane, with no polar oscillations. We note that features at higher values of $a$ may not be physical due to the self-force model we are using.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.5\textwidth]{res_flux_rms_p}}
\subfloat{\includegraphics[width=0.5\textwidth]{res_flux_rms_p_e095}}
\caption{\label{fig:res-flux-rms-p}Relative flux enhancements of $p$, as a function of $e$ (left figure) and $\zeta$ (right), marginalised over $a$ and $\cos\iota$ by taking the root-mean-square of a grid of values. Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red), use solid (short-dashed; long-dashed) lines and circular (square; triangular) points. The 2:3 resonance has the largest relative flux enhancement.}
\end{figure}

For the same reasoning that nearly-equatorial orbits show small flux enhancements, we expect nearly-circular orbits to exhibit similar behaviour due to the small volume of the $r\theta$ 2-torus. In figures \ref{fig:res-flux-rms-p}, \ref{fig:res-flux-rms-e} and \ref{fig:res-flux-rms-cosi}, we show the root-mean-square (over $a$ and $\cos\iota$) relative flux enhancements as a function of eccentricity, for a selection of low-order resonances. As expected, it is found that larger eccentricities give rise to larger resonant flux enhancements in general.

It is also interesting to consider the dependence on the resonance ratio $\zeta$. Fixing the eccentricity at $e=0.95$ to emphasise the variation, we compute the flux enhancements for different resonance ratios, shown in the right-hand plots of figures \ref{fig:res-flux-rms-p}, \ref{fig:res-flux-rms-e} and \ref{fig:res-flux-rms-cosi}, where we also highlight ratios arising from different values of $n_\theta$. We observe a clear relationship between the flux enhancement and $\zeta$ for all resonances with $n_\theta = 1$ or $2$, independent of which of those is considered. However, negligibly small flux enhancements are observed for systems with $n_\theta = 3$ or higher. This can be explained by the quadrupolar nature of GWs. A resonant trajectory averaged over time (and consequently $\phi$) takes $n_\theta$ (possibly repeated) values of $\theta$ at each radial turning point. Considering rings of matter at these coordinates, the dominant spherical harmonic describing the resultant mass distribution has $l=n_\theta$ (and $m=0$). Since GWs are generated by the quadrupole moment at leading order, the most significant resonant enhancements will arise from systems with $n_\theta = 2$. Due to the presence of higher harmonics, $n_\theta = 1$ systems also contribute strongly; it is perhaps more appropriate to treat $1$:$x$ resonances as \emph{de facto} $2$:$2x$ resonances, explaining the linked behaviour of the flux enhancements for $n_\theta = 1$ and $2$ systems.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.5\textwidth]{res_flux_rms_e}}
\subfloat{\includegraphics[width=0.5\textwidth]{res_flux_rms_e_e095}}
\caption{\label{fig:res-flux-rms-e}Relative flux enhancements of $e$, as a function of $e$ (left figure) and $\zeta$ (right), rms marginalised over $a$ and $\cos\iota$. Considering the low-order resonances (those that have the longest duration), the 2:3 resonance is strongest. The divergence at $\zeta \approx 0.2$ is due to the adiabatic flux $\langle\Delta e\rangle \approx 0$ close to plunge.}
\end{figure}

As the lowest-order resonance with $n_\theta = 2$, the 2:3 resonance is the most significant, providing the largest flux enhancements over a prolonged time. The divergence in the relative flux enhancement of $e$ near $\zeta = 0.2$, observed in \figref{res-flux-rms-e}, is a consequence of the adiabatic behaviour. Under GW emission, eccentricities typically decrease, apart from for a short time prior to plunge where they slightly increase. There is hence some critical $p$ such that the adiabatic rate of change of $e$ vanishes. Higher-order resonances, with smaller $\zeta$, tend to occur closer to the central BH at smaller values of $p$, and so there is a corresponding critical value of $\zeta$ where $\langle\Delta e\rangle \approx 0$ and the relative flux enhancements diverge. However, the absolute flux enhancements remain small for these resonances, and so we do not expect them to influence the evolution any more than nearby resonances with much smaller relative enhancements.% Below the divergence, the adiabatic fluxes are positive.

\begin{figure}[htbp]
\centering
\subfloat{\includegraphics[width=0.5\textwidth]{res_flux_rms_cosi1}}
\subfloat{\includegraphics[width=0.5\textwidth]{res_flux_rms_cosi1_e095}}
\caption{\label{fig:res-flux-rms-cosi}Relative flux enhancements of $\cos\iota$, as a function of $e$ (left figure) and $\zeta$ (right), rms marginalised over $a$ and $\cos\iota$, restricted to prograde orbits only. The 2:3 resonance is the most significant.}
\end{figure}

The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems by \citet{flanagan_resonantly_2014}. They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a=0.9$, either $e=0.3$ or $0.7$, and $\theta_\mathrm{inc} = 20\text{\textdegree}$ or $70$\textdegree. We compare our results to their published values (in their tables I--IV) in \figref{res-flux-FHR}. There is some qualitative agreement between the two approaches: more eccentric orbits give rise to larger enhancements; and the 3:4 resonance has a negligibly small enhancement. However, for the most significant 2:3 resonance, we typically predict larger orbital parameter jumps. Our results in \chapref{effect-resonances}, regarding the impact of resonances on a population of EMRIs, are therefore conservative predictions, as our model over-estimates the resonance effects.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{res_flux_FHR}
\caption{\label{fig:res-flux-FHR}Relative flux enhancements of $E$ (left column), $L_z$ (middle) and $Q$ (right) for resonance ratios corresponding to the 1:3, 1:2, 2:3 and 3:4 resonances. We show results as computed here with a solid red line, and results from \citet[FHR14]{flanagan_resonantly_2014} with a dashed blue line. Each row has a different eccentricity and inclination corresponding to tables I--IV in FHR14. All systems have $a=0.9$.}
\end{figure}

\section{Summary}
Transient resonances between the frequencies of motion in the $r$ and $\theta$ directions are expected to be nearly ubiquitous in observations of EMRI systems, and yet are unaccounted for by the standard adiabatic modelling techniques. These calculate the precession-averaged radiation fluxes using the Teukolsky equation and equate them to the rate of change of orbital parameters using balance arguments. However, on resonance, the orbit does not precess and fill the $r\theta$ 2-torus, and so this approach is insufficient to capture the full behaviour of the system. We instead use a gravitational self-force model to evolve EMRIs instantaneously, enabling us to study the properties of resonances.

We find the resonance surfaces in $p$--$e$--$\cos\iota$ space to be well-approximated by planes, with a fit given by \eqnref{res-approx-p}. This is sufficiently accurate to use as the starting point for numerical root-finding algorithms to determine the precise resonance location. As an EMRI evolves through a resonance, the rate of change of the orbital parameters is enhanced or diminished with respect to an adiabatic evolution, by an amount depending on the phase difference between the $r$ and $\theta$ motions. This leads to a rapid change in the parameters, often referred to as a jump. We compute the magnitude of this jump using two methods: analysing the phase space trajectory during an evolution to extract the jump directly; and calculating the flux enhancements while fixed on a resonant geodesic. Typical jumps are around $1\%$, and are greatest for the 2:3 resonance. There is a strong dependence on the eccentricity of the system, with more eccentric orbits experiencing much larger flux enhancements.

The magnitude of the orbital parameter jumps cannot tell us about the impact of resonances on GW observations of EMRIs. For that, we must calculate the emitted GWs from each system and study the loss in SNR that arises from using adiabatic templates.

### 2015-05-30 res_char.tex
Location of resonances

@@
 \email[]{cplb@star.sr.bham.ac.uk}
+\affiliation{School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
 \affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
-\affiliation{School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
 \author{Priscilla Ca\~{n}izares}
@@
 In this work, an astrophysical population of detectable extreme-mass-ratio inspirals is studied and it is found that a large proportion shall encounter a low-order resonance in the later stages of inspiral.
-The resulting effect on signal-to-noise recovery is small due to the low eccentricity of the studied population, but nevertheless at least one fifth of the systems may become undetectable if transient resonances aren't accounted for.
+The resulting effect on signal-to-noise recovery is small due to the low eccentricity of the studied population, but nevertheless at least one fifth of the systems may become undetectable if transient resonances are not accounted for.
 \end{abstract}
@@
 
-\emph{Move location stuff to appendix ... the bits here need tweaking to reflect that.}
+\emph{Move location stuff to appendix. Bits here have been tweaked.}
 
@@
 We can calculate the frequencies and so we are left with the problem of solving $\Omega = n_r \Omega_r - n_\theta \Omega_\theta = 0$ numerically.
-\Figref{res-plane-2-5-95} shows the semilatus rectum, eccentricity and (cosine of the) inclination angle of the $\nu = 2/5$ resonance surface for a BH of spin $a_\ast = 0.95$.
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_res-2-5-95-plane}
-\caption{\label{fig:res-plane-2-5-95}Location of the $2/5$ resonance surface for an $a_\ast = 0.95$ BH in terms of orbital semilatus rectum $p$, eccentricity $e$ and inclination $\iota$.}
-\end{figure}
-The resonances occur at relatively small periapses, corresponding to regions of strong-field gravity.
-Similar resonance surfaces can be found for other spin values and for other resonances.
 When considering the full parameter set of $\{p,e,\iota,a_\ast,\nu\}$, it is apparent that the search for resonances becomes expensive as a consequence of the dimensionality.
 It is therefore useful to have a guide of where to look.
-We shall attempt to build a model for the resonance plane.
+In \apref{location} we build a simple approximate model as a starting point for the numerical search.
+The resonances occur at relatively small periapses, corresponding to regions of strong-field gravity.
+Having located where in an inspiral we can expect to encounter a transient resonance, we must now consider its impact.
+In \secref{res-time} we determine the characteristic time-scales describing resonance, and in \secref{flux-enhance} we look at the impact of passing through a resonance on the evolution of the orbit.
 
@@
 
-\subsection{Resonant flux enhancement}
+\subsection{Resonant flux enhancement}\label{sec:flux-enhance}
 
@@
 
-\section{Location of resonances}
+\section{Location of resonances}\label{sec:location}
+
+We can find the location of resonances by numerically solving $\Omega = n_r \Omega_r - n_\theta \Omega_\theta = 0$.
+\Figref{res-plane-2-5-95} shows the semilatus rectum, eccentricity and (cosine of the) inclination angle of the $\nu = 2/5$ resonance surface for a BH of spin $a_\ast = 0.95$.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{Fig_res-2-5-95-plane}
+\caption{\label{fig:res-plane-2-5-95}Location of the $2/5$ resonance surface for an $a_\ast = 0.95$ BH in terms of orbital semilatus rectum $p$, eccentricity $e$ and inclination $\iota$.}
+\end{figure}
+This is almost planar, inspiring us to look for a simple description that can help guide our search for resonance locations.
 Brink, Geyer and Hinderer~\cite{Brink2013} provide series expansions for the location of resonances in the limit of equatorial orbits for small spin and eccentricity.
@@
 This should be useful for designating the region in which resonance could be expected.
-To locate them precisely, it is necessary to solve $\Omega = 0$ numerically; the approximate model should give a suitable starting point.
+To locate them precisely, it is necessary to solve $\Omega = 0$ numerically; the approximate model gives a suitable starting point.
 

### 2015-06-08 asymptotic.tex
Updating references

@@
 where we have rewritten the forcing terms as Fourier series and adapted the forcing functions to those appropriate for $q$ and $J$.
-We shall solve these before resonance and then match to solutions in the transition regime about resonance.
+We solve these before resonance and then match to solutions in the transition regime about resonance.
 
@@
 \end{equation}
-This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest order piece of) the adiabatic solution~\cite{Hinderer2008}.
+This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest-order piece of) the adiabatic solution~\cite{Hinderer2008}.
 If we similarly average \eqnref{q-1}, we find
@@
 \end{align}
-We now have expressions for the lowest order terms in the expansions.
+We now have expressions for the lowest-order terms in the expansions.
 
@@
 \end{equation}
-Using these, we can build the next order terms
+Using these, we can build the next-order terms
 \begin{align}
@@
 
-We shall be interested in the asymptotic behaviour for $|\widehat{\lambda}| \rightarrow \infty$. The complex Fresnel integral has the limit~\cite{Olver2010} % [7.5.3, 7.5.4, 7.12.2, 7.12.3]
+We are interested in the asymptotic behaviour for $|\widehat{\lambda}| \rightarrow \infty$. The complex Fresnel integral has the limit~\cite{Olver2010} % [7.5.3, 7.5.4, 7.12.2, 7.12.3]
 \begin{equation}
@@
 
-To complete our solution we must match the pre-resonance solution of \secref{before-res} with the near resonance solution of \secref{interior-res}.
-This is achieved by rewriting the pre-resonance solution in terms of the rescaled time $\widehat{\lambda}$ and comparing this with the near resonance solution expanded in the limit of $\widehat{\lambda} \rightarrow -\infty$.
+To complete our solution we must match the pre-resonance solution of \secref{before-res} with the near-resonance solution of \secref{interior-res}.
+This is achieved by rewriting the pre-resonance solution in terms of the rescaled time $\widehat{\lambda}$ and comparing this with the near-resonance solution expanded in the limit of $\widehat{\lambda} \rightarrow -\infty$.
 
@@
 \end{widetext}
-We can now compare this to the near resonance expansion with the integral replaced by the limiting form given in \eqnref{Fres-limit}.
+We can now compare this to the near-resonance expansion with the integral replaced by the limiting form given in \eqnref{Fres-limit}.
 
@@
 
-Having constructed the solution, we see that the lowest order evolution corresponds to the adiabatic solution; the deviations come in at the following order.
+Having constructed the solution, we see that the lowest-order evolution corresponds to the adiabatic solution; the deviations come in at the following order.
 When we switch from the pre-resonance solution to the post-resonance solution, there is a change in the sign of $\widehat{\lambda}$.

### 2015-06-08 res_char.tex
Updating references

@@
 \author{Robert H. Cole}
-\email[]{rhc26@ast.cam.ac.uk}
+%\email[]{rhc26@ast.cam.ac.uk}
 \affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
@@
 \author{Priscilla Ca\~{n}izares}
-\email[]{pcm@ast.cam.ac.uk}
+%\email[]{pcm@ast.cam.ac.uk}
 \author{Jonathan R. Gair}
-\email[]{jgair@ast.cam.ac.uk}
+%\email[]{jgair@ast.cam.ac.uk}
 \affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
@@
 Bodies of comparable mass can be studied using numerical relativity.
-Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}.
-These simulations shall allow us to understand BH--BH mergers.
-Stellar-mass BH mergers are targets for ground-based GW detectors, such as the currently mid-upgrade LIGO~\cite{Harry2010} and Virgo~\cite{Accadia2011}, and the in-construction KAGRA~\cite{Kuroda2010}.
+Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}; it is now possible to simulate hundreds of orbits~\cite{Szilagyi2015}.
+These simulations allow us to understand BH--BH mergers.
+Stellar-mass BH mergers are targets for ground-based GW detectors, such as the upcoming Advanced LIGO~\cite{Aasi2015} and Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Kuroda2010}.
 Massive BH mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
@@
 The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase although this can accumulate over many orbits.
-Being able to accurately model the influence of the self-force shall allow us to create reliable waveform models.
+Being able to accurately model the influence of the self-force allows us to create reliable waveform models.
 
@@
 During resonance, terms in the self-force that usually average to zero can combine coherently, significantly impacting the orbital motion~\cite{Flanagan2012a}.
+
 Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.
+However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2014} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central black hole from its host.
 
@@
 
-The leading order dissipative correction to geodesic motion is calculated following the adiabatic prescription~\cite{Hinderer2008}: by dropping the forcing term $g_\alpha^{(1)}$ (and all higher order terms) and replacing the forcing term $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$ $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
+The leading-order dissipative correction to geodesic motion is calculated following the adiabatic prescription~\cite{Hinderer2008}: by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing the forcing term $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$ $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
 For most orbits this is sufficient, $G_\alpha^{(1)}$ is given by its average value plus a rapidly oscillating component~\cite{Arnold1988}.
@@
 In this work we seek to characterise the importance of these resonances for the purposes of modelling EMRIs.
-The amplitude of expected signals shall be below the level of noise in a space-based GW detector.
+The amplitude of expected signals is below the level of noise in a space-based GW detector.
 However, systems remain in band for many hundreds of thousands of cycles and so may be detected using a matched filter, provided we have sufficiently accurate waveform templates.
@@
 One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short time-scales, for small $\eta$ there is a small disparity.
-The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e., the case for motion in Kerr) is subject to a small perturbation the form of the orbits is preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
+The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e.
+the case for motion in Kerr) is subject to a small perturbation the form of the orbits is preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
 % Chapter II, section 3.3 d)
 This should ensure that, in general, there are only small shifts in the orbital frequencies.
-However, the KAM theory only applies for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}.
+However, the KAM theory is only valid for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}.
 % Chapter V section 1.
@@
 The instantaneous orbital frequencies oscillate back and forth around their averaged values.
-
-However, there shall be a time span when the frequencies are consistently close to being commensurate.
+However, there is a time span when the frequencies are consistently close to being commensurate.
 During this time, the trajectory appears similar to a resonant trajectory, filling only a smaller region of the parameter space.
@@
 
-When analysing resonances it is useful to refer to a number of reference time-scales.
+When analysing resonances it is useful to refer to a number of characteristic time-scales.
 We always use coordinate time $t$ for these, as this corresponds to what is measured by an observer at infinity.
@@
 \end{align}
-using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \left\langle\dot{\Omega}\right\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
+using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \langle\dot{\Omega}\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
 \begin{equation}
@@
 We also wish to thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent and Leor Barack for useful conversations.
-RHC is supported by STFC; CPLB was supported by STFC and the Cambridge Philosophical Society; PC is supported by a Marie Curie Fellowship, and JRG is supported by the Royal Society.
+RHC is supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC was supported by a Marie Curie Fellowship, and JRG is supported by the Royal Society.
 \end{acknowledgments}

### 2015-06-08 res_char.tex
Overlap/order

@@
 \label{eq:overlap}
-\mathcal{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}}
+\mathbb{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}}
 \end{equation}
@@
 The adiabatic waveform shows good agreement with the full waveform in both amplitude and phase.
-Calculating the SNR for each of the models, we find an overlap $\mathcal{O} = 0.984$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
+Calculating the SNR for each of the models, we find an overlap $\mathbb{O} = 0.984$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
 
@@
 None of the adiabatic models presented here give a particularly high overlap with the full evolution, because of the effects of the resonance.
-The best-performing adiabatic model was that which matched at the end, giving $\mathcal{O} = 0.677$, while the model that matches at the start gives only $\mathcal{O} = 0.202$.
+The best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.202$.
 These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.
@@
 \includegraphics[width=\textwidth]{Fig_EMRIpar_dists}
-\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $\mathcal{O}(10^4)$ realisations.
+\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $\order{10^4}$ realisations.
 The (darker) blue bars represent the final distribution, including an SNR cut of 10.

### 2015-06-27 res_char.tex

@@
 \usepackage{hyperref}
+\usepackage{textcomp}
 
@@
 This is less than ideal for high spins.
+We also find that this particular implementation of the self-force model marginally overestimates the adiabatic inspiral rate with respect to direct PN evolutions by a factor of $\mathcal{O}(1)$, even for systems in the weak field and with low values of the spin.
 While this approximate self-force is not perfect, it should serve as a guide for the behaviour of the full self-force, allowing us to assess the qualitative impact of resonances on EMRI detection.
@@
 using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.
+We compute these integrals using a $300\times 300$ grid of $\{\psi,\chi\}$ values and employing a Newton--Cotes approximation in each dimension.
+The procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform.
+The adiabatic derivatives vary on much longer timescales than the orbital motion, and so in practice, we can make use of interpolation.
+We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\dd t/\eta$ where $\dd t$ is the orbital time-step.
+When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives.
+Once the system evolves close to one of the boundaries, we update the interpolation points, taking the closest boundary to be the new central point.
 The averaged fluxes successfully describe the leading-order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
@@
 
-\emph{Remove from here ...}
-The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$.
+%The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$.
 The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurence of any transient resonances.
@@
 
-One method of studying the effect of resonances is then to compare an adiabatic evolution with an instantaneous evolution, using the full knowledge of the self-force.
+%One method of studying the effect of resonances is then to compare an adiabatic evolution with an instantaneous evolution, using the full knowledge of the self-force.
 This may be a useful exercise to calibrate and verify adiabatic models that can be generated cheaply; however it cannot be used to distinguish between dephasing effects arising from resonances and those arising from conservative corrections to the phase variables, as both are explicitly eliminated from the adiabatic prescription.
@@
 
-To be specific, we write \eqnref{Mino-E-o-M} as
-\begin{subequations}
-\begin{align}
-\label{eq:Mino-E-o-M2}
-\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
-\diff{J_\alpha}{\lambda} = {} & \eta \langle G_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
-\end{align}
-\end{subequations}
-where $g_\alpha$ and $G_\alpha$ encapsulate all our knowledge about the self-force and may contain contributions from sub-leading terms in the mass-ratio expansion.
+%To be specific, we write \eqnref{Mino-E-o-M} as
+%\begin{subequations}
+%\begin{align}
+%\label{eq:Mino-E-o-M2}
+%\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
+%\diff{J_\alpha}{\lambda} = {} & \eta \langle G_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
+%\end{align}
+%\end{subequations}
+%where $g_\alpha$ and $G_\alpha$ encapsulate all our knowledge about the self-force and may contain contributions from sub-leading terms in the mass-ratio expansion.
 The quasi-adiabatic evolution is then obtained by dropping the terms containing $\hat{g}_\alpha$ and $\hat{G}_\alpha$.
 For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an ``adiabatic evolution'' in fact refer to a ``quasi-adiabatic evolution''.
-\emph{... to here.}
 
@@
 \label{sec:properties}
-
-\emph{Move location stuff to appendix. Bits here have been tweaked.}
 
@@
 
+We can study the predicted orbital parameter jumps numerically by considering the long-time-averaged fluxes for particular (fixed) geodesics.
+For non-resonant geodesics, this will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference.
+We consider a selection of different orbital shape parameters corresponding to low-order transient resonances and calculate the time-averaged fluxes for $50$ different initial phases.
+We evolve for an integer number of cycles of both the radial and polar motions.
+Specifically, when considering a $n_\theta$:$n_r$ resonance, we take $T = 2\pi N n_\theta/\Omega_r = 2\pi N n_r/\Omega_\theta$ for some integer $N$.
+We have found that taking $N=5$ gives sufficiently accurate results.
+We can then estimate the jumps by taking the difference between the maximum and minimum time-averaged fluxes.
+
+We expect nearly-circular orbits to show small flux enhancements due to the small volume of the $r\theta$ 2-torus; resonant orbits come close to every allowed point, thus approximating a non-resonant orbit.
+In \figref{res-flux-rms-p}, we show the root-mean-square (over $a$ and $\cos\iota$) relative flux enhancement for the semi-latus rectum as a function of eccentricity, for a selection of low-order resonances.
+As expected, it is found that larger eccentricities give rise to larger resonant flux enhancements in general.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{res_flux_rms_p}
+\caption{\label{fig:res-flux-rms-p}Relative flux enhancements of $p$, as a function of $e$, marginalised over $a$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
+Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use solid (short-dashed; long-dashed) lines.
+The 2:3 resonance has the largest relative flux enhancement.}
+\end{figure}
+
+\emph{Is this description clear, and is it actually correct? It seems reasonable ...}
+
+It is informative to consider the dependence on the resonance ratio $\nu$.
+Fixing the eccentricity at $e=0.95$ to emphasise the variation, we compute the flux enhancements for different resonance ratios, shown in \figref{fig:res-flux-rms-p-vsnu}, where we also highlight ratios arising from different values of $n_\theta$.
+We observe a clear relationship between the flux enhancement and $\nu$ for all resonances with $n_\theta = 1$ or $2$, independent of which of those is considered.
+However, negligibly small flux enhancements are observed for systems with $n_\theta = 3$ or higher.
+This can be explained by the quadrupolar nature of GWs.
+A resonant trajectory averaged over time (and consequently $\phi$) takes $n_\theta$ (possibly repeated) values of $\theta$ at each radial turning point.
+Considering rings of matter at these coordinates, the dominant spherical harmonic describing the resultant mass distribution has $l=n_\theta$ (and $m=0$).
+Since GWs are generated by the quadrupole moment at leading order, the most significant resonant enhancements will arise from systems with $n_\theta = 2$.
+Due to the presence of higher harmonics, $n_\theta = 1$ systems also contribute strongly; it is perhaps more appropriate to treat $1$:$x$ resonances as \emph{de facto} $2$:$2x$ resonances, explaining the linked behaviour of the flux enhancements for $n_\theta = 1$ and $2$ systems.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{res_flux_rms_p_e095}
+\caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements of $p$ marginalised over $a$ and $\cos\iota$ by taking the root-mean-square.
+Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use circular (square; triangular) points.}
+\end{figure}
+
+\emph{Is this bit really needed ...?}
+
+The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems~\cite{Flanagan2012a}.
+They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a=0.9$, either $e=0.3$ or $0.7$, and $\theta_\mathrm{inc} = 20\text{\textdegree}$ or $70$\textdegree.
+We compare our results to their published values (in their tables I--IV) in \figref{res-flux-FHR}.
+There is some qualitative agreement between the two approaches: more eccentric orbits give rise to larger enhancements; and the 3:4 resonance has a negligibly small enhancement.
+However, for the most significant 2:3 resonance, we typically predict larger orbital parameter jumps.
+Our results in \secref{astrophysics}, regarding the impact of resonances on a population of EMRIs, are therefore conservative predictions, as our model over-estimates the resonance effects.
+
+\begin{figure*}[htbp]
+\centering
+\includegraphics[width=\textwidth]{res_flux_FHR}
+\caption{\label{fig:res-flux-FHR}Relative flux enhancements of $E$ (left column), $L_z$ (middle) and $Q$ (right) for resonance ratios corresponding to the 1:3, 1:2, 2:3 and 3:4 resonances.
+We show results as computed here with a solid red line, and results from~\cite[FHR14]{Flanagan2012a} with a dashed blue line.
+Each row has a different eccentricity and inclination corresponding to tables I--IV in FHR14.
+All systems have $a=0.9$.}
+\end{figure*}
+
 Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on a population of inspirals.
@@
 
-To generate gravitational waveforms, we choose to employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}.
+To generate gravitational waveforms, we choose to employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}, augmented to include evolution of the positional elements.
 The principle behind the class of kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
@@
 
-Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the post-Newtonian self-force from \secref{self-force}.
+Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the dissipative part of the post-Newtonian self-force from \secref{self-force}.
 We then interpret the Boyer--Lindquist coordinates as spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique; specifically, we apply the standard quadrupole formula~\cite{Misner1973}.
@@
 
-\emph{Update this to reflect thesis.}
-
 Once the trajectories have been computed using both methods of evolution, we can calculate the corresponding waveforms.
@@
 The adiabatic waveform shows good agreement with the full waveform in both amplitude and phase.
-Calculating the SNR for each of the models, we find an overlap $\mathbb{O} = 0.984$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
+Calculating the SNR for each of the models, we find an overlap $\mathbb{O} = 0.993$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
 
@@
 
-\emph{Update this to reflect thesis.}
-
 To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral.
@@
 
-The extremely large parameter space of adiabatic waveforms renders a brute-force approach computationally impossible; it is therefore necessary to restrict the search in some way.
+The extremely large parameter space of adiabatic waveforms, coupled with the expensive nature of our 2-torus averaging routine, renders a brute-force approach computationally impossible; it is therefore necessary to restrict the search in some way.
 For this preliminary investigation into the importance of resonances, we do this by focussing on a small subset of parameters that we suspect will produce a large overlap.
@@
 
-The most restrictive aspect of this approach is that all extrinsic parameters and any intrinsic ones that remain constant throughout the evolution, such as the masses of the BHs and the central BH spin, do not get explored by our family of adiabatic models.
-It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
-Indeed, systematic biases are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.
-
 The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
-Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau_\mathrm{res}$ after each resonance of interest, namely the low-order 2:3, 1:2, 2:5 and 1:3.
+Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau_\mathrm{res}$ after each resonance of interest, namely the low-order 2:3 and 1:2.
 These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral.
@@
 
+In our experience, we found that these choices were not sufficient to produce a good family of adiabatic waveforms.
+In particular, if an adiabatic model happened to match the full model close to the edge of the envelope generated by the fast orbital evolution, then the two trajectories eventually diverged and the resulting waveform provided a poor match.
+We therefore adjust the values of $t_{\mathrm{match}}$ so that they correspond to times of apoapsis.
+This ensures that the adiabatic model intersects with the full model close to the centre of the envelope, as demonstrated by the inset plots in \figref{good-traj}.
+As a result, we obtain much higher overlaps in general.
+
 We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.
 
-In our experience, we found that these choices were not sufficient to produce a good family of adiabatic waveforms.
-In particular, if an adiabatic model happened to match the full model close to a turning point in the fast orbital evolution, then it eventually diverged from the full model and the resulting waveform provided a poor match.
-We therefore adjust the values of $t_{\mathrm{match}}$ so that they correspond to times in the inspiral when the time-derivative of a particular orbital parameter (we choose the semi-latus rectum $p$) is maximally different from the adiabatic time-derivative of that orbital parameter.
-This ensures that the adiabatic model intersects with the full model close to the centre of the envelope created by the fast oscillations, as demonstrated by the inset plots in \figref{good-traj}, and as a result gives much higher overlaps in general.
+This one-parameter family of adiabatic models does not explore the extrinsic parameter space nor does it consider different values of the constant intrinsic parameters (the masses of the BHs and the central BH spin).
+It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
+Indeed, systematic biases in parameter estimation are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.
+The result of these restrictions is that the maximum overlap calculated using our family is a strict lower bound on the true maximum overlap.
 
@@
 \caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start.
-The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.202), at the 2:3 resonance (0.458), after the 2:3 resonance (0.259) and at the end (0.677).}
+The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.207), at the 2:3 resonance (0.432), after the 2:3 resonance (0.258) and at the end (0.677).}
 \end{figure*}
@@
 None of the adiabatic models presented here give a particularly high overlap with the full evolution, because of the effects of the resonance.
-The best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.202$.
+The best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.207$.
 These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.
@@
 This allows us to take into account any linear drift in $\Delta \mathcal{I}$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation.
-In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down.
+% In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down.
 For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.
@@
 The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
+The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset; this means that for every value of $q$, there is always a resonance jump in at least one parameter.
 We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).
@@
 For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t_\mathrm{match}$ described in the previous section.
-The largest overlap in each case is plotted in \figref{overlap-vs-resjump-ill}, as a function of the percentage resonance jumps in the orbital parameters.
-We observe no strong trend with the size of the jump, although systems with small jumps in $E$ (less than $1\%$) have higher than average overlaps.
-These systems also have small jumps in the fundamental frequencies and so the adiabatic waveform dephases more slowly from the full waveform, leading to a larger overlap.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=\textwidth]{Fig_overlap_vs_resjump_ill}
-\caption{\label{fig:overlap-vs-resjump-ill}The overlap \eqnref{overlap} between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the resonance jumps in $E$ (left), $L_z$ (middle), and $Q$ (right).
+The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{overlap_vs_q}
+\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$.
 Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
 The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
-\end{figure*}
+\end{figure}
 
 The best performing adiabatic for the majority of our systems matched at the end of the inspiral.
-We can interpret the low overlaps by assuming that these adiabatic waveforms recover the entire SNR in the region post-resonance, but none pre-resonance.
-As a result, we would expect the overlap to roughly equal the fraction of time spent post-resonance, which in this case is $0.7$, demonstrated by the red line in \figref{overlap-vs-resjump-ill}.
-There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the base of the eLISA noise curve.
-
-A small fraction of systems have a much lower overlap than expected, at around $0.5$.
-This is due to a sub-optimal choice of the matching times; making small adjustments to $t_\mathrm{match}$ can lead to higher overlaps around $0.7$, but the exact adjustments required are difficult to predict in advance.
+We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance.
+We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}.
+There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
+
+Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
+The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
+A small fraction of systems have a lower overlap than expected, at around $0.5$.
+This is due to a sub-optimal choice of the matching times; making small adjustments to $t_\mathrm{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
 
@@
 
-\emph{COMPLETELY REWRITE}
-
 \emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix??}
 
-Strong resonances can severely limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic modes, which may be individually undetectable.
-To assess the impact of this on future GW missions, we need to create a sample population of detectable EMRIs.
-
-We first identify 14 parameter distributions that describe an astrophysically typical EMRI\footnote{The actual distribution in parameter space of observable EMRI systems is unknown, but we can make reasonable estimates.}.
-We choose the mass of the compact object $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs.
-We take the central object to be a typical supermassive BH at the centre of a galaxy, with mass $M$ uniformly distributed in log-space between $10^5 M_\odot$ and $10^7 M_\odot$, and dimensionless spin $a_\ast$ uniformly between its limiting values, $0$ and $1$.
+Strong resonances can limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic regions, which may be individually undetectable.
+In order to assess the impact of this on future GW missions, we need to analyse the waveforms from a population of detectable EMRIs.
+
+We generate a representative EMRI population by making reasonable assumptions about the distributions of the 14 parameters in the analytic kludge model of Barack and Cutler \cite{Barack2004}.
+The EMRI event rate is dominated by the inspiral of heavier BHs due to the larger observational volume in comparison to that for white dwarfs and neutron stars~\cite{gair_event_2004}.
+We therefore take the mass of the compact object to be $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs.
+
+We take the central object to be a typical supermassive BH at the centre of a galaxy; we construct an appropriate distribution by breaking the EMRI event rate into two pieces: the `intrinsic' event rate $\mathcal{R}$, i.e.
+the number of inspiral events per unit time for any given galaxy; and the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH.
+We are interested in galaxies that possess a SMBH with a mass in the range $(10^{4} - 10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors.
+In principle, $\mathcal{R}$ will depend on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each component, and the mass and spin of each SMBH.
+All of these properties are highly uncertain, even for our own galaxy.
+
+Despite the difficulties, simple estimates of $\mathcal{R}$ have been carried out using Monte Carlo methods to count the number of stars from isothermal distributions that spiral in to a SMBH without plunging.
+The result is a scaling law for each species of compact object of the form
+\begin{equation}
+\label{eq:EMRI-intrinsic-rate}
+\mathcal{R}(M) = \mathcal{R}_0 \left(\frac{M}{M_0}\right)^\alpha,
+\end{equation}
+where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is some fiducial mass.
+\Citet{hopman_extreme_2009} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, NSs and WDs, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}\, \mathrm{Gyr}^{-1}$ for each component.
+This neglects the effects of resonant relaxation, ignores the spin of the SMBH, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses.
+Each of these is likely to significantly impact the event rate, but \eqnref{EMRI-intrinsic-rate} can still be used as a rough guide to the expected number of events.
+
+\Citet{amaro-seoane_impact_2011} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker-Planck description for the bulk properties of the stellar distribution.
+They found a better fit for the power law spectral index for BHs of $\alpha = -0.19$.
+
+The comoving number density of galaxies is challenging to estimate due to local structure in the Universe, the evolution of that structure, and properties of the SMBHs themselves.
+We simplify the problem by assuming a homogeneous distribution that does not evolve with redshift, which is reasonable for the typical scales considered by GW detectors.
+We also neglect correlations between the SMBH mass and spin, and impose a power law scaling relation for the comoving number density
+\begin{equation}
+\diff{n}{\ln M}(M) = n_0 \left(\frac{M}{M_0}\right)^\beta.
+\end{equation}
+There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002 \, \mathrm{Mpc}^{-3}$ for SMBHs with $M < \mathcal{O}(10^7 M_\odot)$ \citep{greene_mass_2007, gair_lisa_2010}.
+
+The mass of the SMBH is then sampled from a power-law with a probability distribution function (pdf) $f(M) \propto M^{\alpha+\beta-1}$.
+We sample the dimensionless spin $a_\ast$ uniformly between its limiting values, $0$ and $1$.
 The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere.
@@
 
-We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\,\mathrm{km s^{-1}\, Mpc^{-1}}$, and sample the redshift of the source uniformly from comoving volume, up to a maximum of $z_\mathrm{max} = 0.7$.
+We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\,\mathrm{km s^{-1}\, Mpc^{-1}}$, and sample the redshift of the source uniformly from comoving volume, up to a maximum of $z_\mathrm{max} = 1.5$.
 We don't expect the exact details of the cosmology to significantly alter our results and our choice of $z_\mathrm{max}$ is justified \textit{post hoc}.
 
-The astrophysics in dense stellar clusters at the center of galaxies leading to EMRI production is highly uncertain.
-The currently favoured formation mechanism scatters compact objects onto highly eccentric orbits, which lose a burst of energy to GW emission with each periapse passage.
-This tends to circularise the orbit, as the compact object spirals inwards.
-Hopman and Alexander \cite{Hopman2005} model this scattering and give an eccentricity distribution for compact objects spiralling into a $3 \times 10^6 M_\odot$ Schwarzschild BH at the point when the orbital period enters the LISA band at $10^4\,\mathrm{s}$.
-The overall results of their work are still valid for spinning BHs and so assuming that this distribution holds for other BH masses, we can sample the orbit eccentricity and calculate the semi-latus rectum from the orbital period.
-
-These EMRIs are astrophysically motivated, but may not be detectable by eLISA.
-To assess this, we evolve the systems forward in time using the analytic kludge method of Barack and Cutler \cite{Barack2004} until they reach the last stable orbit.
-We then calculate the Peters-Matthews waveform \cite{Peters1963} for a time $t_\mathrm{insp}$ at the end of the inspiral, where $t_\mathrm{insp}$ is chosen uniformly between $0.5$ years and $2$ years.
-We focus on the late stages of inspiral as these are most likely to be detected by eLISA.
-Given the waveform, we can estimate the expected SNR by approximating the overlap with a time-domain integral
-\begin{equation}
-\overlap{s}{h} \approx 4 \intd{0}{t_\mathrm{insp}}{\frac{s(t)h(t)}{S_n(f(t))}}{t},
-\end{equation}
-where $f(t)$ is some characteristic frequency of the orbit at time $t$. We find that taking $f$ to be twice the (radial) orbital frequency gives a good approximation to the SNR.
-
-To obtain our final distribution of detectable EMRIs, shown as dark blue in \figref{EMRIpar-dists}, we quote the parameter values $t_\mathrm{insp}$ before plunge.
-Only systems with $\rho > 10$ are treated as detectable.
-Also shown in the figure, in light pink, are the systems with an estimated SNR below 10, to illustrate which systems tend to have lower SNRs.
-The acceptance rate for generating a detectable EMRI from the entire distribution is $32\%$.
-
-We note that prograde orbits (those with $\cos\iota > 0$) and systems with higher BH spins tend to be favoured; this is likely due to the fact that such systems can inspiral much closer before finally plunging, and so experience the strongest gravitational fields.
-This effect is amplified by the fact that the majority of the SNR is recovered from the final part of the inspiral, which also explains why larger values of $t_\mathrm{insp}$ are only marginally preferred over smaller values.
-
-Systems at higher redshifts start to become undetectable as the amplitude of the signal scales inversely proportional to the luminosity distance to the source.
-At $z=0.7$, the fraction of detectable systems drops to $5\%$; this is chosen as the upper limit in the redshift distribution to minimise unneccessary computations.
-We quote the masses in the system as redshifted masses, which accounts for the frequency shift due to the expansion of the Universe.
-
-The circularising effect of GW emission can be observed in the eccentricity distribution: even though the input systems from Hopman and Alexander have high eccentricites, with a peak in the distribution at $e\sim0.7$, the final distribution contains only low eccentricity systems, peaked at $e\sim0.05$ with a maximum at $e=0.31$.
+The eccentricity distribution is more complicated.
+Eccentricities for EMRIs are uncertain, and depend strongly on the formation scenario being considered.
+Here, we adopt a fit to a distribution computed using Monte Carlo simulations by \citet{Hopman2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH.
+We assume this can be extended to provide a rough estimate of the distribution around BHs of other masses and spins.
+At the point in the inspiral when the orbital period takes a fiducial value $\nu_0 = 10^4\,\mathrm{s}$, we find that the Monte Carlo eccentricity pdf is well-described by a power-law with an exponential cutoff
+\begin{widetext}\begin{equation}
+\label{eq:EMRI-e-distribution}
+f(e;\,e_\mathrm{m},\,e_\mathrm{p},\,b) \propto 
+	\begin{cases}
+		\left(e_\mathrm{m}-e\right)^{b(e_\mathrm{m}-e_\mathrm{p})} \exp\left\{b(e-e_\mathrm{m})\right\} , &\quad 0 \leq e \leq e_\mathrm{m}\\
+		0, &\quad\text{otherwise}
+	\end{cases}
+\end{equation}\end{widetext}
+where $e_\mathrm{m} = 0.81$ is the maximum observed eccentricity, $e_\mathrm{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index.
+
+We are interested in EMRI systems close to plunge, as it is here where the GW amplitude is largest.
+To create our EMRI population, we therefore sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value of $10^4\,\mathrm{s}$, and evolve the system using the AK prescription up until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $R(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.} Each of these systems is then evolved backwards for some time $t_\mathrm{insp}$, chosen uniformly from the range $[0,2]$ years, and the expected gravitational waveforms are calculated using the AK formalism.
+
+The size of the required population can be estimated by evaluating an integral of the form \citep{gair_probing_2009}
+\begin{equation}
+\label{eq:EMRI-number}
+N_\mathrm{EMRI} = \intd{z=0}{z_\mathrm{max}}{ \intd{M=M_\mathrm{min}}{M_\mathrm{max}}{   \mathcal{R}T\diff{n}{\ln M} \diff{V_\mathrm{c}}{z}   }{\ln M}  }{z},
+\end{equation}
+where $V_\mathrm{c}(z)$ is the comoving volume at redshift $z$.
+For a mission lifetime of 2 years, the integral gives a total of $6333$ EMRI events.
+This is a lower bound for $N_\mathrm{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.
+
+A given EMRI is classified as observable if its SNR exceeds some threshold value $\mathrm{SNR}_\mathrm{thres}$.
+Calculating SNRs from the generated AK waveforms, assuming 6 laser links and using the eLISA noise PSD, we find $513$ detectable events, assuming an SNR threshold of $15$.
+The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters, the redshift of the source, and the length of the observation $t_\mathrm{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than 15, which are also shown.
 
@@
 \centering
-\includegraphics[width=\textwidth]{Fig_EMRIpar_dists}
-\caption{\label{fig:EMRIpar-dists}The parameter distributions for our detectable EMRI systems, generated using $\order{10^4}$ realisations.
-The (darker) blue bars represent the final distribution, including an SNR cut of 10.
-The (lighter) pink bars stacked on top indicate the full distribution before the SNR cut.}
+\includegraphics[width=\textwidth]{EMRIpar_dists}
+\caption{\label{fig:EMRIpar-dists}The parameter distributions at plunge for our detectable EMRI systems (dark blue), alongside those of the undetectable systems (light pink).
+In each case, the y-axis values are the probability of a system being found in a particular bin, given that they are either detectable or undetectable.
+The exception to this is in the final plot for the SNR, where we show the number of systems in each bin.
+The SNRs quoted here are calculated using the AK model.}
 \end{figure*}
 
-\subsection{Population results}
+We note that eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs, due to the fact that the periapse in such systems can get much closer to the SMBH, and so the compact object experiences the strongest gravitational fields.
+This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in the distributions.
+Systems at higher redshifts start to become undetectable because the amplitude of the signal scales inversely proportional to the luminosity distance to the source, and by $z=1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
+The mass distribution for detectable EMRIs is peaked such that the typical GW frequency occurs at the base of the eLISA noise PSD.
+
+EMRI systems within our population tend to have small eccentricities at plunge due to the circularising effect of GWs: $85\%$ of detectable systems have $e < 0.1$, with a mean value of $0.05$ and a maximum of $0.4$.
+In \secref{flux-enhance}, we found a strong eccentricity dependence on the magnitude of the resonant flux enhancements.
+We therefore expect typical resonant jumps in these astrophysical systems to be much less than $1\%$, and so the resultant dephasing may be relatively weak.
+We now analyse these $513$ systems using our full and adiabatic self-force models.
+
+\subsection{Loss of signal-to-noise ratio}
 \label{sec:population}
 
-\emph{THIS IS ALL GARBAGE (oh :()}
-
-We take draws from the EMRI parameter distributions, using them as initial conditions for the full instantaneous evolution, from which we generate a family of adiabatic evolutions, as discussed in \secref{SNRloss}.
-There are broadly three categories of system regarding the effect of resonances on the evolution:
-
-\begin{enumerate}
-\item No significant resonance jumps are observed, and all the adiabatic models agree with the full evolution reasonably well: the overlaps are typically larger than $0.9$.
-The models have overlaps less than $1$ since the adiabatic approach breaks down near the end of the inspiral.
-
-\item At least one resonance jump can be observed (typically the 2:3 resonance), but the adiabatic models still produce large overlaps.
-Most of the SNR from a signal is accumulated near to the end of the inspiral, at which point there are few significant resonances.
-
-\item At least one resonance jump can be observed, and all of our chosen adiabatic models have low overlaps with the full evolution.
-This does not rule out the possibility of adiabatic models existing that would produce a high overlap, but it is difficult to predict the combination of parameters that would achieve this.
-\end{enumerate}
-
-To ensure that our chosen adiabatics are reasonable estimates to maximise the overlap, we computed many adiabatic waveforms with regular values of $t_\mathrm{match}$ for one particular system; there can be large fluctuations in the overlap, but our chosen parameters perform at least averagely well.
-
-Across the full population, of $140$ different realisations, we find that $77\%$ of systems pass through the 2:3 resonance within the later stages of inspiral.
-In addition, $99\%$ of the systems we studied encountered the 1:2 resonance.
-This suggests that transient resonances are an important consideration in the evolution of EMRIs.
-
-While resonance jumps may in principle be as large as a few percent (see the example in \figref{resjump-vs-q}), this tends to only occur in high eccentricity systems.
-The typical size of any resonance jump within our astrophysical population is less than $0.01\%$, with many systems showing no visible changes through resonance.
-\Figref{pop-SNR-vs-jump} shows the distribution of computed resonance jumps in $E$ for our population, along with the corresponding loss of overlap.
-There is no observed trend with the size of the resonance jump, although this is not surprising given the small size of the jumps.
-The large scatter on the plot is due to our choice of adiabatic family; in each case, the quoted fractional SNR is a lower bound for the actual fractional SNR across the entire adiabatic model space.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=0.92\textwidth]{Fig_pop_SNR_vs_jump}
-\caption{\label{fig:pop-SNR-vs-jump}The fraction of the full SNR recovered using our family of adiabatic models, as a function of the percentage resonance jump in $E$.
-The histograms show the marginalised distributions over each parameter.}
-\end{figure*}
-
-It is possible for EMRIs to form through alternative channels, which may result in a population of high eccentricity systems.
-As an example, compact objects can be scattered onto orbits that plunge directly into the central BH; these systems can be mistaken for slow inspirals since they still undergo many hundreds of thousands of cycles in the bandwidth of a detector like eLISA \cite{Amaro-Seoane2013}.
-This kind of event is expected to be orders of magnitude more common than classic EMRIs.
-It is expected that resonance behaviours would be much more important in this case.
-
-Even within the low eccentricity systems considered here, it is possible for some EMRIs to be undetectable using adiabatic models, given the resonances they encounter.
-These are any systems for which the full SNR is larger than some cutoff value, but the largest adiabatic SNR is smaller than the cutoff.
-The estimated fraction of systems that would be lost in this manner, as a function of cutoff SNR is shown in \figref{SNRcutoff}.
+After computing the SNRs of our NK waveforms for each of the $513$ systems, we can compare the values to those obtained using the AK formalism.
+The different approximation schemes are expected to produce different results, but they should be broadly similar.
+In \figref{pop-SNR-vs-eta}, we plot the ratio of the new SNR to the AK SNR, as a function of the mass ratio.
+We observe a clear trend, with systems closest to equal mass performing roughly as expected, but with the most extreme systems giving significantly lower SNRs.
+
+This is a consequence of the PN self-force model we are using, which we have found overestimates the inspiral rate for EMRIs.
+Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the 2 year observation window, they evolve within the bucket of the eLISA noise curve.
+Increasing the rate of evolution then does not significantly change the recovered SNR since we are still sensitive to all parts of the waveform.
+At smaller mass ratios, the systems evolve more slowly, from lower plunge frequencies.
+An increased evolution rate can shift the EMRI to outside the sensitive region of the detector, thus giving a much lower SNR.
 
@@
 \centering
-\includegraphics[width=0.46\textwidth]{Fig_SNRcutoff}
-\caption{\label{fig:SNRcutoff}The fraction of EMRI systems that would be lost using adiabatic models to analyse data that contain the effects of transient resonances.
-The error bars are estimated using the beta distribution \cite{Cameron2011}.}
+\includegraphics[width=0.46\textwidth]{pop_SNR_vs_eta}
+\caption{\label{fig:pop-SNR-vs-eta}The ratio of SNRs as computed with waveforms generated using the NK (adopting our PN self-force model) and AK formalisms, as a function of (the base-10 logarithm of) the mass ratio.
+Each system has an AK SNR greater than 15, using the eLISA detector with 6 laser links.}
+\end{figure}
+
+Nevertheless, we may still study the effect of resonances, by comparing the maximum adiabatic SNR to the full SNR computed using the NK waveforms.
+The resulting overlap is still indicative of the effect of resonances on the population of systems.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_Tad}
+\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family.
+Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t_\mathrm{insp}$.
+The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t_\mathrm{ad}$, normalised by $t_\mathrm{insp}$.
+The solid line corresponds to the expected value if the 1:2 and 2:3 resonances cause significant waveform dephasing.}
+\end{figure}
+
+We have calculated the longest period of time, $t_\mathrm{ad}$, during each inspiral, in which neither the 1:2 nor the 2:3 resonance occurs.
+From the results of \secref{effres-jumps}, we expect the recovered overlap to be approximately given by the proportion of time spent in a resonance-free region, that is $t_\mathrm{ad} / t_\mathrm{insp}$.
+In \figref{pop-adSNR-vs-Tad}, we plot the maximum overlap between the instantaneous and adiabatic models against this ratio, highlighting the number of resonances, $N_\mathrm{res}$, each system encounters.
+In the majority of cases, there is no significant impact on the recovered overlap using adiabatic models.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_eta}
+\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t_\mathrm{ad} / t_\mathrm{insp}$, as a function of (the base-10 logarithm of) the mass ratio.}
+\end{figure}
+
+We can better understand the overlap distribution from \figref{pop-adSNR-vs-Tad} by considering the behaviour with the mass ratio.
+We do this in \figref{pop-adSNR-vs-eta}, plotting the difference between the computed maximum overlap and the value expected should resonances be important.
+A small proportion of systems have overlaps below the expected value (approximately $4\%$ have values less than $-0.05$).
+This might be caused by higher-order resonances disrupting the evolution, in which case $t_\mathrm{ad}$ should be smaller and the systems would lie on the expectation line.
+Alternatively, the low-order resonances might be long-lived, explaining why the recovered overlap is smaller than expected.
+However, a more likely explanation is that larger adiabatic overlaps are possible with models that we have not considered here (either with different values of $t_\mathrm{match}$ or outside our one-parameter family altogether).
+
+Roughly $30\%$ of the systems lie within $0.05$ of the expected value.
+However, the vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity.
+For the smallest mass ratios, the inspiral rate is slow, and so the systems do not encounter either the 1:2 or 2:3 resonances during their lifetime.
+Meanwhile for the largest mass ratios, the EMRIs encounter both resonances very close to plunge.
+In each case, there is a long resonance-free region, allowing a high overlap to be recovered.
+
+The remaining $66\%$ of systems have overlaps above the level expected if resonances were to play an important role.
+These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t_\mathrm{ad}$ is small.
+For these EMRIs, resonances are not as important as expected, which is likely due to the low eccentricities of our population leading to small (and in most cases, negligible) resonant flux enhancements.
+
+Even if we assume that all overlap reductions are due to resonances (instead of due to our adiabatic family not being extensive enough) the overall effect on the population is not significant.
+To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to take into account the resonance behaviour.
+The total number of detectable systems using a threshold SNR of $15$ decreases from $513$ to $492$.
+If we increase the threshold, the fractional reduction in the number of detectable systems gets even smaller.
+We therefore conclude that resonances do not cause sufficient waveform dephasing across a population of EMRIs for the detection rate to be appreciably diminished.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.46\textwidth]{pop_SNR_dist}
+\caption{\label{fig:pop-SNR-dist}Histogram showing the probability distribution function for the detectable EMRI SNRs, as calculated using the AK formalism (light pink) and modified by the maximum adiabatic overlap (dark blue).}
 \end{figure}

### 2015-06-29 res_char.tex
Some tweaks

@@
 \begin{equation}
-\nu = \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r},
+\nu \equiv \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r},
 \end{equation}
@@
 
-In \secref{problem}, we formulate the specific problem at hand: that of geodesic motion in Kerr spacetime, perturbed by the gravitational self-force.
+In \secref{problem}, we formulate the specific problem: that of geodesic motion in Kerr spacetime, perturbed by the gravitational self-force.
 We then study generic properties of transient resonances in \secref{properties}, detailing their location in parameter space, the time-scales over which they affect the motion and the resulting GW flux enhancements.
@@
 \end{equation}
-Is in convenient to calculate $\Gamma$ as
+It is convenient to calculate $\Gamma$ as
 \begin{equation}
@@
 \end{equation}
-using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.
-We compute these integrals using a $300\times 300$ grid of $\{\psi,\chi\}$ values and employing a Newton--Cotes approximation in each dimension.
+using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.\footnote{We compute these integrals using a $300\times 300$ grid of $\{\psi,\chi\}$ values and employing a Newton--Cotes approximation in each dimension.
 The procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform.
-The adiabatic derivatives vary on much longer timescales than the orbital motion, and so in practice, we can make use of interpolation.
-We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\dd t/\eta$ where $\dd t$ is the orbital time-step.
+However, the adiabatic derivatives vary on much longer timescales than the orbital motion (see \secref{res-time}), and so in practice, we can interpolate.} %We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\delta t/\eta$ where $\delta t$ is the orbital time-step.
 When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives.
 Once the system evolves close to one of the boundaries, we update the interpolation points, taking the closest boundary to be the new central point.
-The averaged fluxes successfully describe the leading-order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
+The averaged fluxes successfully describe the leading-order secular evolution of the trajectory (as illustrated in \figref{good-traj}).
 
@@
 Having located where in an inspiral we can expect to encounter a transient resonance, we must now consider its impact.
-In \secref{res-time} we determine the characteristic time-scales describing resonance, and in \secref{flux-enhance} we look at the impact of passing through a resonance on the evolution of the orbit.
+In \secref{res-time} we determine the characteristic time-scales describing resonance, and in \secref{flux-enhance} we look at the impact of passing through a resonance on the evolution of the orbit, including some numerical results.
 
@@
 
-We can study the predicted orbital parameter jumps numerically by considering the long-time-averaged fluxes for particular (fixed) geodesics.
+We can study the predicted orbital-parameter jumps numerically by considering the long-time-average fluxes for particular (fixed) geodesics.
 For non-resonant geodesics, this will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference.
-We consider a selection of different orbital shape parameters corresponding to low-order transient resonances and calculate the time-averaged fluxes for $50$ different initial phases.
+We consider a selection of different orbital shape parameters corresponding to low-order resonances and calculate the time-averaged fluxes for $50$ different initial phases.
 We evolve for an integer number of cycles of both the radial and polar motions.
@@
 It is informative to consider the dependence on the resonance ratio $\nu$.
-Fixing the eccentricity at $e=0.95$ to emphasise the variation, we compute the flux enhancements for different resonance ratios, shown in \figref{fig:res-flux-rms-p-vsnu}, where we also highlight ratios arising from different values of $n_\theta$.
+Fixing the eccentricity at $e=0.95$ to emphasise the variation, we compute the flux enhancements for different resonance ratios, shown in \figref{res-flux-rms-p-vsnu}, where we also highlight ratios arising from different values of $n_\theta$.
 We observe a clear relationship between the flux enhancement and $\nu$ for all resonances with $n_\theta = 1$ or $2$, independent of which of those is considered.
@@
 
-The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems~\cite{Flanagan2012a}.
+The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a}.
 They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a=0.9$, either $e=0.3$ or $0.7$, and $\theta_\mathrm{inc} = 20\text{\textdegree}$ or $70$\textdegree.
@@
 \caption{\label{fig:res-flux-FHR}Relative flux enhancements of $E$ (left column), $L_z$ (middle) and $Q$ (right) for resonance ratios corresponding to the 1:3, 1:2, 2:3 and 3:4 resonances.
-We show results as computed here with a solid red line, and results from~\cite[FHR14]{Flanagan2012a} with a dashed blue line.
+We show results as computed here with a solid red line, and results from Flanagan, Hughes and Ruangeri (FHR14)~\cite{Flanagan2012a} with a dashed blue line.
 Each row has a different eccentricity and inclination corresponding to tables I--IV in FHR14.
@@
 
-\subsection{Waveform model}
+\subsection{Waveform model \& analysis}
 \emph{Can probably shorten this section a bit.}
 
-To generate gravitational waveforms, we choose to employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}, augmented to include evolution of the positional elements.
-The principle behind the class of kludge methods is to first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
-This can be much quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
+To generate gravitational waveforms, we employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}, augmented to include evolution of the positional elements.
+We first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
+This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
 
 Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the dissipative part of the post-Newtonian self-force from \secref{self-force}.
-We then interpret the Boyer--Lindquist coordinates as spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique; specifically, we apply the standard quadrupole formula~\cite{Misner1973}.
+We then map the Boyer--Lindquist coordinates to spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique, the standard quadrupole formula~\cite{Misner1973}.
 
 We expect these waveforms to be sufficiently accurate for our purposes, given the inaccuracy of the self-force model being employed.
-It would, however, be straight-forward to adopt a different waveform generation scheme, for example by including higher order multipole moments, should advances be made in determining the self-force.
+It would, however, be straight-forward to adopt a different waveform generation scheme, for example by including higher-order multipole moments, should advances be made in determining the self-force.
 
@@
 \end{equation}
-where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD).
+where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD)~\cite{Moore2014a}.
 We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept.
@@
 In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
-To take this into account, when evaluating overlaps, we first project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme, setting any values outside the domain of $\{t_j\}$ to zero.
+To take this into account, we project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme and set any values outside the domain of $\{t_j\}$ to zero.
 A Planck-taper window function~\cite{McKechan2010} is applied to the resulting time series to reduce unwanted spectral leakage in the Fourier transforms.
-
-After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
-In addition, we maximise every overlap calculation with respect to some time-offset between the models by computing the correlation between the waveforms.
-This can be performed cheaply using Fourier transforms and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.
+%After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
+We maximise every overlap calculation with respect to some time-offset between the models by computing the correlation between the waveforms, which can be performed cheaply using Fourier transforms.
+%and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.
 
@@
 We wish to test whether there exists some set of parameters $\Theta^a$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
-We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template, defined by
+We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template,
 \begin{equation}
@@
 
-To demonstrate the approach, we calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$ years for a set of initial parameters chosen to avoid any significant resonances.
-Specifically, we choose: the mass of the CO $\mu = 10 M_{\odot}$; the mass ratio $\mu / M = 3\times 10^{-6}$ where $M$ is the mass of the central BH; the spin of the central BH $a=0.95$; the initial dimensionless semi-latus rectum $p=7.5$; the initial eccentricity $e=0.7$; the initial inclination angle $\cos \iota = 0.5$; and the redshift of the source $z=0.204$.
-The remaining intrinsic (the initial phases $\psi$ and $\chi$) and extrinsic (the position of the source on the sky, the direction of the spin of the central BH and the azimuthal angle) parameters are set to arbitrary values.
-The resulting evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, where it can be seen that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
-The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the 2-torus averaging explicitly smoothes out the visible structure on this scale.
+To demonstrate the approach, we can calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$ years for a set of initial parameters chosen to avoid any significant resonances.
+Example evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, which shows that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
+The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the 2-torus averaging explicitly smooths out the visible structure on this scale.
 
@@
 \caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) under the full (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances.
-The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.}
+The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.
+This system has $\mu = 10 M_{\odot}$, $q = 3\times 10^{-6}$, $a_\ast=0.95$, initial semi-latus rectum $p_0=7.5$, initial eccentricity $e_0=0.7$, initial inclination $\cos \iota_0 = 0.5$ and redshift $z=0.204$.}
 \end{figure*}
@@
 \includegraphics[width=0.92\textwidth]{Fig_good_waveform}
-\caption{\label{fig:good-waveform}The plus-polarised waveform for an illustrative EMRI system with $p=7.5$, shown for two short segments at the start and end of a $2$ year evolution.
+\caption{\label{fig:good-waveform}The plus-polarised waveform for an illustrative EMRI system with initial semi-latus rectum $p_0=7.5$, shown for two short segments at the start and end of a $2$ year evolution.
 The full and adiabatic models are both shown, but are indistinguishable by eye.}
@@
 We generate a representative EMRI population by making reasonable assumptions about the distributions of the 14 parameters in the analytic kludge model of Barack and Cutler \cite{Barack2004}.
-The EMRI event rate is dominated by the inspiral of heavier BHs due to the larger observational volume in comparison to that for white dwarfs and neutron stars~\cite{gair_event_2004}.
+The EMRI event rate is dominated by the inspiral of heavier BHs due to the larger observational volume in comparison to that for white dwarfs and neutron stars~\cite{Gair2004}.
 We therefore take the mass of the compact object to be $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs.
@@
 where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is some fiducial mass.
-\Citet{hopman_extreme_2009} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, NSs and WDs, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}\, \mathrm{Gyr}^{-1}$ for each component.
+\Citet{Hopman2009a} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, NSs and WDs, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}\, \mathrm{Gyr}^{-1}$ for each component.
 This neglects the effects of resonant relaxation, ignores the spin of the SMBH, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses.
@@
 
-\Citet{amaro-seoane_impact_2011} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker-Planck description for the bulk properties of the stellar distribution.
+\Citet{Amaro-Seoane2011d} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker-Planck description for the bulk properties of the stellar distribution.
 They found a better fit for the power law spectral index for BHs of $\alpha = -0.19$.
@@
 \end{equation}
-There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002 \, \mathrm{Mpc}^{-3}$ for SMBHs with $M < \mathcal{O}(10^7 M_\odot)$ \citep{greene_mass_2007, gair_lisa_2010}.
+There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002 \, \mathrm{Mpc}^{-3}$ for SMBHs with $M < \mathcal{O}(10^7 M_\odot)$ \citep{Greene2007, Gair2010b}.
 
@@
 
-The size of the required population can be estimated by evaluating an integral of the form \citep{gair_probing_2009}
+The size of the required population can be estimated by evaluating an integral of the form \citep{Gair2009}
 \begin{equation}
@@
 \end{equation}
-These were fitted for all possible resonances with $n_r = 2$--$7$ as well as $\nu = 9/10$, $19/20$, $49/50$ and $99/100$; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
+These were fitted for all possible resonances with $n_r = 2$--$7$ as well as the $9:10$, $19:20$, $49:50$ and $99:100$ resonances; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
 %These were fitted for resonances with $\nu = 1/7, 1/6, 1/5, 1/4, 2/7, 1/3, 2/5, 3/7, 1/2, 4/7, 3/5, 2/3, 5/7, 3/4, 4/5, 5/6, 6/7, 9/10, 19/20, 49/50, 99/100$, with BH spins of $a_\ast = 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.93, 0.95, 0.97, 0.99, 0.999$.

### 2015-07-04 res_char.tex
Astrophysics

@@
 
-\newcommand{\sub}[1]{\ensuremath{_\text{#1}}}
-\newcommand{\super}[1]{\ensuremath{^\text{#1}}}
+\newcommand{\sub}[1]{\ensuremath{_\mathrm{#1}}}
+\newcommand{\super}[1]{\ensuremath{^\mathrm{#1}}}
 \newcommand{\dd}{\ensuremath{\mathrm{d}}}
@@
 
-\title{The importance of transient resonances in extreme-mass-ratio inspirals (new title?)}
+\title{The importance of transient resonances in extreme-mass-ratio inspirals}
 
@@
 Other effects are however neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
-During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact-object trajectory and, ultimately, the phase of the emitted radiation.
-In this work, an astrophysical population of detectable extreme-mass-ratio inspirals is studied and it is found that a large proportion shall encounter a low-order resonance in the later stages of inspiral.
-The resulting effect on signal-to-noise recovery is small due to the low eccentricity of the studied population, but nevertheless at least one fifth of the systems may become undetectable if transient resonances are not accounted for.
+During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the emitted radiation.
+In this work, an astrophysical population of detectable extreme-mass-ratio inspirals is studied.
+We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence the low eccentricity of the studied population.
+At least one fifth of the systems may become undetectable if transient resonances are not accounted for.
 \end{abstract}
@@
 Stellar-mass BH mergers are targets for ground-based GW detectors, such as the upcoming Advanced LIGO~\cite{Aasi2015} and Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Kuroda2010}.
-Massive BH mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
+Supermassive BH (SMBH) mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
 Systems of unequal masses are more challenging to evolve numerically as they complete a larger number of orbits and it is necessary to resolve two different scales.
 Calculations can instead be performed perturbatively.
-The paradigm unequal-mass system has a stellar-mass BH orbiting a massive black hole, such as those expected to be found at the centre of most galaxies.
+The paradigm unequal-mass system has a stellar-mass BH orbiting an SMBH, such as those expected to be found at the centre of most galaxies.
 These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors~\cite{Amaro-Seoane2007}.
-Despite being well-studied there still remain open questions.
+Despite being well-studied, there still remain open questions.
 
 In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}.
-In the test particle limit, the smaller body follows an exact geodesic of the BH's spacetime.
+In the test particle limit, the smaller body follows an exact geodesic of the SMBH's spacetime.
 Including the effects of the smaller body's finite mass, the background spacetime is perturbed.
@@
 The latter shifts the orbital phases inducing precession.
-The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase although this can accumulate over many orbits.
+The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase, although this can accumulate over many orbits.
 Being able to accurately model the influence of the self-force allows us to create reliable waveform models.
@@
 Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.
-However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2014} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central black hole from its host.
+However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host.
 
@@
 
-The leading-order dissipative correction to geodesic motion is calculated following the adiabatic prescription~\cite{Hinderer2008}: by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing the forcing term $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$ $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
+The leading-order dissipative correction to geodesic motion is calculated following the adiabatic prescription~\cite{Hinderer2008}: by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing the forcing term $G_\alpha^{(1)}$ with $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$, its average over the $2$-torus parametrized by $q_r$ and $q_\theta$~\cite{Drasco2005}.
 For most orbits this is sufficient, $G_\alpha^{(1)}$ is given by its average value plus a rapidly oscillating component~\cite{Arnold1988}.
@@
 We then study generic properties of transient resonances in \secref{properties}, detailing their location in parameter space, the time-scales over which they affect the motion and the resulting GW flux enhancements.
-Specific examples are considered to illustrate the effects of resonances at the start of \secref{astrophysics}, before finally turning to an astrophysical population in \secref{population}.
+Specific examples are considered to illustrate the effects of resonances in \secref{waveforms}, before finally turning to an astrophysical population in \secref{astrophysics}.
 Our conclusions can be found in \secref{conclusion}.
@@
 We use geometric units with $G = c = 1$ throughout.
-We always use $M$ for the mass of the central massive BH and $a$ as the spin parameter.
+We always use $M$ for the mass of the central SMBH and $a$ as the spin parameter.
 We also use the dimensionless spin $a_\ast \equiv a/M$; we take the convention that $0 \leq a_\ast < 1$.
-
-\section{Problem formulation}
+We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\units{km\, s^{-1}\, Mpc^{-1}}$ and do not expect the exact details of the cosmology to significantly alter our results.
+
+\section{The problem of EMRI transient resonances}
 \label{sec:problem}
-\subsection{Kerr geodesics}
 
@@
 Instantaneously, the motion of the orbiting mass can be described as geodesic, with the integrals of the motion changing on time-scales of many orbital periods.
-We analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance.
-It is therefore necessary to develop an understanding of the Kerr geodesics.
+It is therefore necessary to develop an understanding of the Kerr geodesics (\secref{geodesic}).
+Transient resonance occur when the radial and polar frequencies become commensurate (\secref{frequencies}); we analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance (\secref{forced-motion}).
+The osculating elements formalism allows for the orbital evolution to be driven by a force, here, a particular model for the self-force (\secref{self-force}) and its adiabatic average (\secref{adiabatic}).
+In following sections, we study the differences between the adiabatic and full orbital evolutions.
+
+\subsection{Kerr geodesics}
+\label{sec:geodesic}
 
@@
 \subsection{Orbital resonances}
+\label{sec:frequencies}
 
@@
 
-\subsection{Forced motion}
+\subsection{Osculating elements and forced motion}
 \label{sec:forced-motion}
 
-For generic EMRIs, there are two timescales of interest: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$.
+For generic EMRIs, there are two characteristic timescales: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$.
 These, along with the resonance timescale, are discussed more in \secref{res-time}.
@@
 Collectively, these are referred to as \emph{osculating elements} and we denote them by $I^A(t)$, making explicit the variation with time.
-
-Given geodesic motion in some background spacetime with Christoffel symbols $\Gamma^\alpha_{\beta\gamma}$, forced by some external acceleration $f^\alpha$, so that
-\begin{equation}
-\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = f^\alpha,
-\end{equation}
-we can calculate the evolution of the osculating elements $\dot{I}^A$. The specific equations for motion in Kerr are derived by Gair {\it{et al}}.~\cite{Gair2011a}.
+For a sequence of geodesics of a background spacetime, where the evolution is forced by some external acceleration, we can calculate the evolution of the osculating elements $\dot{I}^A$.
+The specific equations for motion in Kerr are derived by Gair {\it{et al}}.~\cite{Gair2011a}.
+
+%Given geodesic motion in some background spacetime with Christoffel symbols $\Gamma^\alpha_{\beta\gamma}$, forced by some external acceleration $f^\alpha$, so that
+%\begin{equation}
+%\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = f^\alpha,
+%\end{equation}
+%we can calculate the evolution of the osculating elements $\dot{I}^A$. The specific equations for motion in Kerr are derived by Gair {\it{et al}}.~\cite{Gair2011a}.
 
@@
 To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration.
-In this work we choose to work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
-For comparison, Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a} use a Teukolsky equation calculation of GW fluxes to account for the inspiral due to radiation reaction.
+We work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
+For comparison, Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a} use a Teukolsky-equation calculation of GW fluxes to account for the inspiral due to radiation reaction.
 
 The self-force model uses the first-order PN terms of the dissipative self-force formulated in Flanagan and Hinderer~\cite{Flanagan2007} and the conservative force formulated in Iyer and Will~\cite{Iyer1993}, and Kidder~\cite{Kidder1995}.
-Since only the first PN terms are used, this prescription can only be of limited validity in strong fields.
+Since only the first PN terms are used, this prescription is of limited validity in strong fields.
 Both pieces of the self-force are computed assuming that the spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$.
-This is less than ideal for high spins.
+This is suboptimal for high spins.
 We also find that this particular implementation of the self-force model marginally overestimates the adiabatic inspiral rate with respect to direct PN evolutions by a factor of $\mathcal{O}(1)$, even for systems in the weak field and with low values of the spin.
@@
 This may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
+The averaged force can be computed from the radiative field \cite{Galtsov1982,Mino2003,Sago2006,Ganz2007}.
 This piece is purely dissipative and determines how the inspiral evolves due to the radiation of GWs.
@@
 To construct an adiabatic trajectory we need the $2$-torus-averaged fluxes of our osculating elements.
+To guarantee consistency, we average our instantaneous self-force.
 Computing an average of a quantity over the $\{q_r, q_\theta\}$ is trivial if it is parametrized in terms of these variables,
@@
 \begin{equation}
-\left|\left\langle\dot{\Omega}\right\rangle_{q'}\right| \simeq \left|\frac{\Omega(t)}{t}\right|.
+\left|{\Omega(t)}\right| \simeq \left|\left\langle\dot{\Omega}(0)\right\rangle_{q'} t\right|.
+\label{eq:Taylor-Omega}
 \end{equation}
 The resonant time-scale should give an indication of the time over which we expect the effects of the resonance to be felt~\cite{Bosley1992}.
-Consider the phase of the Mino time Fourier expansion on resonance; neglecting the constant, the resonant Fourier component has form
+Consider the phase of the Mino-time Fourier expansion on resonance; neglecting the constant, the resonant Fourier component has form
 \begin{equation}
@@
 Typically, the first term is non-zero and this gives the familiar oscillation.
-On resonance, it is zero, leaving the next order term to govern the behaviour~\cite{Flanagan2012}.
+On resonance, it is zero, leaving the next-order term to govern the behaviour~\cite{Flanagan2012}.
 Only once we have moved far enough away from resonance for the first term to dominate the second do we recapture the non-resonant behaviour.
@@
 \end{align}
-using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \langle\dot{\Omega}\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
+using \eqnref{t-pres} and \eqnref{Taylor-Omega}. Rearranging and using \eqnref{t-res} gives
 \begin{equation}
@@
 
-\subsection{Resonant flux enhancement}\label{sec:flux-enhance}
+\subsection{Resonant flux enhancement}
+\label{sec:flux-enhance}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
@@
 The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a}.
-They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a=0.9$, either $e=0.3$ or $0.7$, and $\theta_\mathrm{inc} = 20\text{\textdegree}$ or $70$\textdegree.
+They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a_\ast=0.9$, either $e=0.3$ or $0.7$, and $\theta\sub{inc} = 20\text{\textdegree}$ or $70$\textdegree.
 We compare our results to their published values (in their tables I--IV) in \figref{res-flux-FHR}.
@@
 
-\begin{figure*}[htbp]
+\begin{figure*}
 \centering
@@
 Each row has a different eccentricity and inclination corresponding to tables I--IV in FHR14.
-All systems have $a=0.9$.}
+All systems have $a_\ast=0.9$.}
 \end{figure*}
@@
 
-\section{Astrophysical implications}
-\label{sec:astrophysics}
-
-To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.
-
-We discuss our chosen waveform generation scheme, giving demonstrations of its accuracy for our purposes.
-We then detail our procedure for generating a realistic distribution of detectable EMRIs, before turning to the statistical effect of resonances on the population at the end of the section.
-
-\subsection{Waveform model \& analysis}
+\section{Impact of resonances on gravitational waveforms}
+\label{sec:waveforms}
+
+Having built an understanding of the properties of transient resonances in EMRIs, we now consider their measurable consequences.
+EMRIs will be observed through GW observations, hence we are interested in how passing through a resonance impacts this signal.
+In \secref{kludge} we discuss our chosen waveform generation scheme, giving demonstrations of its accuracy; in subsequent sections we detail the impact of resonances on the measured properties of the GWs.
+In the following section, we investigate how resonances would impact real-world observations of EMRIs assuming realistic astrophysical population of EMRIs.
+
+\subsection{Waveform model and analysis}
+\label{sec:kludge}
 \emph{Can probably shorten this section a bit.}
@@
 We first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
-This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
+This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically $95\%$ or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
 
@@
 
-Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s(t)$ and $h(t)$, according to
+Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s(t)$ and $h(t)$, according to~\cite{Finn1992}
 \begin{equation}
@@
 \end{equation}
-where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD)~\cite{Moore2014a}.
-We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept.
-We use the analytic approximation of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}
-
-In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
+where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD)~\cite{Moore2014a}.\footnote{We use cubic interpolation to construct NK waveforms with identical time sampling.
+%In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
 To take this into account, we project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme and set any values outside the domain of $\{t_j\}$ to zero.
-A Planck-taper window function~\cite{McKechan2010} is applied to the resulting time series to reduce unwanted spectral leakage in the Fourier transforms.
+%This whole procedure gives more accurate results than interpolating in the Fourier domain since the waveforms are smoother in the time domain.
+A Planck-taper window function~\cite{McKechan2010} is applied to reduce unwanted spectral leakage in the Fourier transforms. 
 %After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
-We maximise every overlap calculation with respect to some time-offset between the models by computing the correlation between the waveforms, which can be performed cheaply using Fourier transforms.
+} We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept, the analytic approximation of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}.
+We maximise every inner product calculation with respect to some time-offset between the models by computing the correlation between the waveforms, which can be performed cheaply using Fourier transforms.
 %and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.
 
-%This whole procedure gives more accurate results than interpolating in the Fourier domain since the waveforms are smoother in the time domain.
-
-One of the targets of pre-eLISA research is to generate a bank of waveform templates, $h(t;\Theta^a)$ described by some set of parameters $\Theta^a$.
-Once real-world data, $s(t)$, has been collected, we can then search over the template bank to find the model that most closely matches the data.
+One of the targets of pre-eLISA research is to generate a bank of waveform templates $\{h(t;\boldsymbol{\Theta}_i)\}$ across a range of parameter space given by $\{\boldsymbol{\Theta}_i\}$.
+These can be compared to data to search for the presence of GWs.
 At the current time, with a lack of experimental data, we are interested in ensuring that the waveform templates accurately reproduce what we expect to observe in nature, without being too computationally expensive.
-
 Ideally, we would like to use waveform templates from EMRI systems that are evolved under the entire self-force model, but these are too expensive.
 In practice, we are forced to use cheaper adiabatic waveforms, but these do not include the effect of resonances.
-To assess the impact of this, we compare data $s(t;\bar{\Theta}^a)$ generated using the full self-force model to templates $h(t;\Theta^a)$ generated using the expensive 2-torus averaged self-force model (see \secref{adiabatic}).
-
-We wish to test whether there exists some set of parameters $\Theta^a$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
+To assess the impact of this, we compare data $s(t;\bar{\boldsymbol{\Theta}})$ generated using the full self-force model to templates $h(t;\boldsymbol{\Theta})$ generated using our 2-torus averaged self-force model (see \secref{adiabatic}).
+
+We wish to test whether there exists some set of parameters $\boldsymbol{\Theta}$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
 We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template,
@@
 provides an indication of how well-matched the template is to the data.
-We leave an investigation into parameter estimation biases (where the parameters of the best-matching adiabatic $\Theta^a$ do not agree with the parameters of the system $\bar{\Theta}^a$) to future work.
-
-To demonstrate the approach, we can calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$ years for a set of initial parameters chosen to avoid any significant resonances.
+We leave an investigation into parameter-estimation biases (where the parameters of the best-matching adiabatic $\boldsymbol{\Theta}$ do not agree with the parameters of the system $\bar{\boldsymbol{\Theta}}$) to future work.
+
+To demonstrate the approach, we can calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$~years for a set of initial parameters chosen to avoid any significant resonances.
 Example evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, which shows that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
@@
 The adiabatic waveform shows good agreement with the full waveform in both amplitude and phase.
-Calculating the SNR for each of the models, we find an overlap $\mathbb{O} = 0.993$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
-
-\begin{figure*}[htbp]
+We find an overlap $\mathbb{O} = 0.993$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
+
+\begin{figure*}
 \centering
@@
 \subsection{The effect of resonances: Dephasing}
+\label{sec:effres-phase}
 
@@
 Specifically, we choose the initial conditions to be the same as before but with an initial semi-latus rectum $p=7.85$.
-This system passes through the 2:3 resonance, the effect of which is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
+This system passes through the $2$:$3$ resonance, the effect of which is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
 Plotting the plus-polarised GWs at the start and end of the evolution, as before, demonstrates this dephasing, as shown in \figref{dephased-waveform}.
 
-\begin{figure*}[htbp]
+\begin{figure*}
 \centering
@@
 Before the resonance occurs, the adiabatic model provides a good match to the full evolution but the overlap is reduced near to the resonance and never fully recovers afterwards.
-This is shown in \figref{overlap-dephasing}, which is centered on the time at which the full evolution crosses the 2:3 resonance.
-
-\begin{figure}[htbp]
+This is shown in \figref{overlap-dephasing}, which is centred on the time at which the full evolution crosses the $2$:$3$ resonance.
+
+\begin{figure}
 \centering
 \includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
-\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give 25 independent (non-overlapping) calculations.
-The time axis is centred on the value at which the full evolution crosses the 2:3 resonance.}
+\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
+The time axis is centred on the value at which the full evolution crosses the $2$:$3$ resonance.}
 \end{figure}
@@
 The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
-Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau_\mathrm{res}$ after each resonance of interest, namely the low-order 2:3 and 1:2.
+Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau\sub{res}$ after each resonance of interest, namely the low-order 2:3 and 1:2.
 These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral.
@@
 As they cannot be easily distinguished by eye, we plot the difference (in $E$, $L_z$ and $Q$) between the models and the adiabatic evolution that matches at the start, shown in \figref{res-diff-traj}.
-The jumps in the orbital parameters due to the 2:3 resonance can be clearly seen, as can the fast orbital oscillations present in the full evolution but absent in the adiabatics.
-
-\begin{figure*}[htbp]
+The jumps in the orbital parameters due to the $2$:$3$ resonance can be clearly seen, as can the fast orbital oscillations present in the full evolution but absent in the adiabatics.
+
+\begin{figure*}
 \centering
 \includegraphics[width=\textwidth]{Fig_res_diff_traj}
-\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start.
-The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.207), at the 2:3 resonance (0.432), after the 2:3 resonance (0.258) and at the end (0.677).}
+\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (centre) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start.
+The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start ($0.207$), at the $2$:$3$ resonance ($0.432$), after the $2$:$3$ resonance ($0.258$) and at the end ($0.677$).}
 \end{figure*}
@@
 To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away.
-For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}_\mathrm{ad}[t_{\mathrm{match}} = t_\mathrm{res}]$ for each of the orbital parameters.
+For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}\sub{ad}[t_{\mathrm{match}} = t\sub{res}]$ for each of the orbital parameters.
 We then identify the most horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations.
@@
 \label{eq:res-jump-ratio}
-\frac{\Delta \mathcal{I}_\mathrm{jump}}{\mathcal{I}_\mathrm{ad}(t>t_\mathrm{res})-\mathcal{I}_\mathrm{ad}(t<t_\mathrm{res})} = \frac{\Delta \mathcal{I}_\mathrm{jump}}{\dot{\mathcal{I}}_\mathrm{ad}(t_\mathrm{res})\tau_\mathrm{res}}.
+\frac{\Delta \mathcal{I}\sub{jump}}{\mathcal{I}\sub{ad}(t>t\sub{res})-\mathcal{I}\sub{ad}(t<t\sub{res})} = \frac{\Delta \mathcal{I}\sub{jump}}{\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}}.
 \end{equation}
@@
 
-For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t_\mathrm{match}$ described in the previous section.
+For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section.
 The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
@@
 A small fraction of systems have a lower overlap than expected, at around $0.5$.
-This is due to a sub-optimal choice of the matching times; making small adjustments to $t_\mathrm{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
-
-\subsection{Sample EMRI population}
-
-\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix??}
+This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
+
+\section{Astrophysical implications}
+\label{sec:astrophysics}
 
@@
 In order to assess the impact of this on future GW missions, we need to analyse the waveforms from a population of detectable EMRIs.
-
-We generate a representative EMRI population by making reasonable assumptions about the distributions of the 14 parameters in the analytic kludge model of Barack and Cutler \cite{Barack2004}.
+In \secref{EMRI-population}, we detail our procedure for generating a realistic distribution of EMRIs, before turning to the effect of resonances on the detectability of EMRIs in \secref{population-SNR}.
+
+\subsection{Sample EMRI population}
+\label{sec:EMRI-population}
+
+\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix?}
+
+To generate a representative EMRI population, we need to establish plausible distributions for the parameters defining the EMRI: the properties of the central SMBH, the properties of the orbiting compact object, and the properties of the orbit itself.
 The EMRI event rate is dominated by the inspiral of heavier BHs due to the larger observational volume in comparison to that for white dwarfs and neutron stars~\cite{Gair2004}.
@@
 
-We take the central object to be a typical supermassive BH at the centre of a galaxy; we construct an appropriate distribution by breaking the EMRI event rate into two pieces: the `intrinsic' event rate $\mathcal{R}$, i.e.
-the number of inspiral events per unit time for any given galaxy; and the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH.
-We are interested in galaxies that possess a SMBH with a mass in the range $(10^{4} - 10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors.
-In principle, $\mathcal{R}$ will depend on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each component, and the mass and spin of each SMBH.
-All of these properties are highly uncertain, even for our own galaxy.
-
-Despite the difficulties, simple estimates of $\mathcal{R}$ have been carried out using Monte Carlo methods to count the number of stars from isothermal distributions that spiral in to a SMBH without plunging.
+We take the central object to be a typical SMBH at the centre of a galaxy.
+We then consider the EMRI event rate in two pieces: the intrinsic event rate per SMBH $\mathcal{R}$, i.e.
+the number of inspiral events per unit time for a given galaxy, and the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH.
+We are interested in galaxies that possess a SMBH with a mass in the range $(10^{4}$--$10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors.
+In principle, $\mathcal{R}$ will depend on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each component, and the mass and spin of each SMBH~\cite{Alexander2005}.
+All of these properties are highly uncertain, even for our own Galaxy.
+
+Despite the difficulties, simple estimates of $\mathcal{R}$ have been carried out using Monte-Carlo methods to count the number of compact objects from isothermal distributions that spiral in to a SMBH without plunging~\cite{Merritt2013}.
 The result is a scaling law for each species of compact object of the form
@@
 \end{equation}
-where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is some fiducial mass.
-\Citet{Hopman2009a} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, NSs and WDs, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}\, \mathrm{Gyr}^{-1}$ for each component.
-This neglects the effects of resonant relaxation, ignores the spin of the SMBH, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses.
+where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is a fiducial mass.
+\Citet{Hopman2009a} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, neutron stars and respectively, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}~\mathrm{Gyr}^{-1}$ for each component, again demonstrating that BHs dominate the event rate.
+This neglects the effects of resonant relaxation~\cite{Rauch1996,Merritt2011}, ignores the spin of the SMBH~\cite{Amaro-Seoane2012b}, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses.
 Each of these is likely to significantly impact the event rate, but \eqnref{EMRI-intrinsic-rate} can still be used as a rough guide to the expected number of events.
 
-\Citet{Amaro-Seoane2011d} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker-Planck description for the bulk properties of the stellar distribution.
-They found a better fit for the power law spectral index for BHs of $\alpha = -0.19$.
+\Citet{Amaro-Seoane2011d} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker--Planck description for the bulk properties of the stellar distribution.
+They found a better fit for the power-law spectral index for BHs of $\alpha = -0.19$.
 
@@
 We simplify the problem by assuming a homogeneous distribution that does not evolve with redshift, which is reasonable for the typical scales considered by GW detectors.
-We also neglect correlations between the SMBH mass and spin, and impose a power law scaling relation for the comoving number density
-\begin{equation}
-\diff{n}{\ln M}(M) = n_0 \left(\frac{M}{M_0}\right)^\beta.
-\end{equation}
-There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002 \, \mathrm{Mpc}^{-3}$ for SMBHs with $M < \mathcal{O}(10^7 M_\odot)$ \citep{Greene2007, Gair2010b}.
-
-The mass of the SMBH is then sampled from a power-law with a probability distribution function (pdf) $f(M) \propto M^{\alpha+\beta-1}$.
-We sample the dimensionless spin $a_\ast$ uniformly between its limiting values, $0$ and $1$.
-The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere.
+We also neglect correlations between the SMBH mass and spin~\cite{Volonteri2010}, and impose a power-law scaling relation for the comoving number density
+\begin{equation}
+\diff{n}{\,\ln M} = n_0 \left(\frac{M}{M_0}\right)^\beta.
+\end{equation}
+There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002~\mathrm{Mpc}^{-3}$ for SMBHs with $M < \order{10^7 M_\odot}$ \citep{Greene2007, Gair2010b}.
+
+Combining the intrinsic event rate with the comoving number density, the mass of the SMBH for the EMRI population is then sampled from a power-law with a probability distribution function $f(M) \propto M^{\alpha+\beta-1}$.
+The dimensionless spin $a_\ast$ is uniformly between its limiting values, $0$ and $1$.
+The direction of the spin of the SMBH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere.
 The angles specifying the location of pericenter and the compact object around the orbit are distributed uniformly between $0$ and $2\pi$.
-
-We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\,\mathrm{km s^{-1}\, Mpc^{-1}}$, and sample the redshift of the source uniformly from comoving volume, up to a maximum of $z_\mathrm{max} = 1.5$.
-We don't expect the exact details of the cosmology to significantly alter our results and our choice of $z_\mathrm{max}$ is justified \textit{post hoc}.
+We sample redshift of the source uniformly from comoving volume.
 
@@
 Eccentricities for EMRIs are uncertain, and depend strongly on the formation scenario being considered.
-Here, we adopt a fit to a distribution computed using Monte Carlo simulations by \citet{Hopman2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH.
-We assume this can be extended to provide a rough estimate of the distribution around BHs of other masses and spins.
-At the point in the inspiral when the orbital period takes a fiducial value $\nu_0 = 10^4\,\mathrm{s}$, we find that the Monte Carlo eccentricity pdf is well-described by a power-law with an exponential cutoff
+Here, we adopt a fit to a distribution computed using Monte-Carlo simulations by \citet{Hopman2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH.
+We assume this can be extended to provide a rough estimate of the distribution around SMBHs of other masses and spins.
+At the point in the inspiral when the orbital period takes a fiducial value $T_0 = 10^4~\mathrm{s}$, we find that the Monte-Carlo eccentricity probability distribution function is well-described by a power-law with an exponential cutoff
 \begin{widetext}\begin{equation}
 \label{eq:EMRI-e-distribution}
-f(e;\,e_\mathrm{m},\,e_\mathrm{p},\,b) \propto 
+f(e) \propto 
 	\begin{cases}
-		\left(e_\mathrm{m}-e\right)^{b(e_\mathrm{m}-e_\mathrm{p})} \exp\left\{b(e-e_\mathrm{m})\right\} , &\quad 0 \leq e \leq e_\mathrm{m}\\
-		0, &\quad\text{otherwise}
-	\end{cases}
+		\left(e\sub{m}-e\right)^{b(e\sub{m}-e\sub{p})} \exp\left\{b(e-e\sub{m})\right\} & 0 \leq e \leq e\sub{m}\\
+		0 &\mathrm{Otherwise}
+	\end{cases},
 \end{equation}\end{widetext}
-where $e_\mathrm{m} = 0.81$ is the maximum observed eccentricity, $e_\mathrm{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index.
-
-We are interested in EMRI systems close to plunge, as it is here where the GW amplitude is largest.
-To create our EMRI population, we therefore sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value of $10^4\,\mathrm{s}$, and evolve the system using the AK prescription up until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $R(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.} Each of these systems is then evolved backwards for some time $t_\mathrm{insp}$, chosen uniformly from the range $[0,2]$ years, and the expected gravitational waveforms are calculated using the AK formalism.
+where $e\sub{m} = 0.81$ is the maximum observed eccentricity, $e\sub{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index.
+
+We are interested in EMRI systems close to plunge, as it is here that the GW amplitude is largest.
+To create our EMRI population, we sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value $T_0$, and evolve the system using the analytic kludge (AK) prescription of Barack and Cutler \cite{Barack2004} until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
+This ignores the influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
 
@@
 \label{eq:EMRI-number}
-N_\mathrm{EMRI} = \intd{z=0}{z_\mathrm{max}}{ \intd{M=M_\mathrm{min}}{M_\mathrm{max}}{   \mathcal{R}T\diff{n}{\ln M} \diff{V_\mathrm{c}}{z}   }{\ln M}  }{z},
-\end{equation}
-where $V_\mathrm{c}(z)$ is the comoving volume at redshift $z$.
-For a mission lifetime of 2 years, the integral gives a total of $6333$ EMRI events.
-This is a lower bound for $N_\mathrm{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.
-
-A given EMRI is classified as observable if its SNR exceeds some threshold value $\mathrm{SNR}_\mathrm{thres}$.
-Calculating SNRs from the generated AK waveforms, assuming 6 laser links and using the eLISA noise PSD, we find $513$ detectable events, assuming an SNR threshold of $15$.
-The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters, the redshift of the source, and the length of the observation $t_\mathrm{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than 15, which are also shown.
-
-\begin{figure*}[htbp]
+N\sub{EMRI} = \intd{z\,=\,0}{z\sub{max}}{ \intd{M\,=\,M\sub{min}}{M\sub{max}}{\mathcal{R}t\sub{life}\diff{n}{\,\ln M} \diff{V\sub{c}}{z}}{\ln M}}{z},
+\end{equation}
+where $V\sub{c}(z)$ is the comoving volume at redshift $z$, and limits $z\sub{max} = 1.5$, $M\sub{min} = 10^4 M_\odot$ and $M\sub{max} = 10^7 M_\odot$ are chosen to encompass the range of detectable signals.
+For a mission lifetime of $t\sub{life} = 2\units{yr}$, the integral gives a total of $6333$ EMRI events.
+This is a lower bound for $N\sub{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.
+
+A given EMRI is classified as observable if its SNR exceeds some threshold value $\rho\sub{thres}$.
+Calculating SNRs from the generated AK waveforms, assuming $6$ laser links and using the eLISA PSD, we find $513$ detectable events for $\rho\sub{thres} = 15$.
+The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters, the redshift of the source, and the length of the observation $t\sub{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than 15, which are also shown.
+
+\begin{figure*}
 \centering
@@
 
-We note that eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs, due to the fact that the periapse in such systems can get much closer to the SMBH, and so the compact object experiences the strongest gravitational fields.
+Eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs, due to the fact that the periapse in such systems can get much closer to the SMBH, and so the compact object experiences the strongest gravitational fields.
 This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in the distributions.
-Systems at higher redshifts start to become undetectable because the amplitude of the signal scales inversely proportional to the luminosity distance to the source, and by $z=1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
+Systems at higher redshifts start to become undetectable because the GW amplitude scales inversely to the luminosity distance; by $z = 1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
 The mass distribution for detectable EMRIs is peaked such that the typical GW frequency occurs at the base of the eLISA noise PSD.
 
-EMRI systems within our population tend to have small eccentricities at plunge due to the circularising effect of GWs: $85\%$ of detectable systems have $e < 0.1$, with a mean value of $0.05$ and a maximum of $0.4$.
+EMRI systems within our population tend to have small eccentricities at plunge due to the circularising effect of GWs~\cite{Peters1964}: for detectable systems, the mean eccentricity is $0.05$ and the maximum is $0.4$, $85\%$ have $e < 0.1$.
 In \secref{flux-enhance}, we found a strong eccentricity dependence on the magnitude of the resonant flux enhancements.
@@
 \subsection{Loss of signal-to-noise ratio}
-\label{sec:population}
-
-After computing the SNRs of our NK waveforms for each of the $513$ systems, we can compare the values to those obtained using the AK formalism.
+\label{sec:population-SNR}
+
+After computing the SNRs of our NK waveforms for each of the $513$ EMRI systems, we can compare the values to those obtained using the AK formalism.
 The different approximation schemes are expected to produce different results, but they should be broadly similar.
-In \figref{pop-SNR-vs-eta}, we plot the ratio of the new SNR to the AK SNR, as a function of the mass ratio.
+In \figref{pop-SNR-vs-eta}, we plot the ratio of the new NK SNR to the AK SNR, as a function of the mass ratio.
 We observe a clear trend, with systems closest to equal mass performing roughly as expected, but with the most extreme systems giving significantly lower SNRs.
-
 This is a consequence of the PN self-force model we are using, which we have found overestimates the inspiral rate for EMRIs.
-Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the 2 year observation window, they evolve within the bucket of the eLISA noise curve.
+Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the $2\units{yr}$ observation window, they evolve within the bucket of the eLISA noise curve.
 Increasing the rate of evolution then does not significantly change the recovered SNR since we are still sensitive to all parts of the waveform.
 At smaller mass ratios, the systems evolve more slowly, from lower plunge frequencies.
-An increased evolution rate can shift the EMRI to outside the sensitive region of the detector, thus giving a much lower SNR.
-
-\begin{figure}[htbp]
+An increased evolution rate means that earlier portions of the EMRI are shifted outside the sensitive region of the detector, thus giving a much lower SNR.
+
+\begin{figure}
 \centering
@@
 \caption{\label{fig:pop-SNR-vs-eta}The ratio of SNRs as computed with waveforms generated using the NK (adopting our PN self-force model) and AK formalisms, as a function of (the base-10 logarithm of) the mass ratio.
-Each system has an AK SNR greater than 15, using the eLISA detector with 6 laser links.}
+Each system has an AK SNR greater than $15$, using the eLISA detector with 6 laser links.}
 \end{figure}
 
-Nevertheless, we may still study the effect of resonances, by comparing the maximum adiabatic SNR to the full SNR computed using the NK waveforms.
+We can study the effect of resonances, by comparing the maximum adiabatic SNR to the full SNR computed using the NK waveforms.
 The resulting overlap is still indicative of the effect of resonances on the population of systems.
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_Tad}
-\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family.
-Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t_\mathrm{insp}$.
-The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t_\mathrm{ad}$, normalised by $t_\mathrm{insp}$.
-The solid line corresponds to the expected value if the 1:2 and 2:3 resonances cause significant waveform dephasing.}
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_Tad}
+%\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family.
+Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t\sub{insp}$.
+The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t\sub{ad}$, normalised by $t\sub{insp}$.
+The solid line corresponds to the expected value if the $1$:$2$ and $2$:$3$ resonances cause significant waveform dephasing.}
+%\end{figure}
+
+For each inspiral, we denote the longest period of time $t\sub{ad}$ in which neither the $1$:$2$ nor the $2$:$3$ resonance is encountered.
+From the results of \secref{effres-phase}, we expect the recovered overlap to be approximately given by the proportion of time spent in a resonance-free region, that is $t\sub{ad} / t\sub{insp}$.
+This assumes that there is perfect overlap in the absence of a resonance and zero overlap across an overlap, with all times during the inspiral contributing equally.
+%In \figref{pop-adSNR-vs-Tad}, we plot the maximum overlap between the instantaneous and adiabatic models against this ratio, highlighting the number of resonances, $N\sub{res}$, each system encounters.
+In the majority of cases, there is no significant impact on the recovered overlap using adiabatic models.
+%We can better understand the overlap distribution from \figref{pop-adSNR-vs-Tad} by considering the behaviour with the mass ratio.
+In \figref{pop-adSNR-vs-eta} we plot the difference between the computed maximum overlap and the value expected from the time between resonances, highlighting the number of resonances $N\sub{res}$ each system encounters.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_eta}
+\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t\sub{ad} / t\sub{insp}$, as a function of the mass ratio, for our population of $513$ EMRI signals.
+Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t\sub{insp}$, with $t\sub{ad}$ the largest time spent by the inspiral without encountering any resonances.}
 \end{figure}
 
-We have calculated the longest period of time, $t_\mathrm{ad}$, during each inspiral, in which neither the 1:2 nor the 2:3 resonance occurs.
-From the results of \secref{effres-jumps}, we expect the recovered overlap to be approximately given by the proportion of time spent in a resonance-free region, that is $t_\mathrm{ad} / t_\mathrm{insp}$.
-In \figref{pop-adSNR-vs-Tad}, we plot the maximum overlap between the instantaneous and adiabatic models against this ratio, highlighting the number of resonances, $N_\mathrm{res}$, each system encounters.
-In the majority of cases, there is no significant impact on the recovered overlap using adiabatic models.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_eta}
-\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t_\mathrm{ad} / t_\mathrm{insp}$, as a function of (the base-10 logarithm of) the mass ratio.}
-\end{figure}
-
-We can better understand the overlap distribution from \figref{pop-adSNR-vs-Tad} by considering the behaviour with the mass ratio.
-We do this in \figref{pop-adSNR-vs-eta}, plotting the difference between the computed maximum overlap and the value expected should resonances be important.
 A small proportion of systems have overlaps below the expected value (approximately $4\%$ have values less than $-0.05$).
-This might be caused by higher-order resonances disrupting the evolution, in which case $t_\mathrm{ad}$ should be smaller and the systems would lie on the expectation line.
-Alternatively, the low-order resonances might be long-lived, explaining why the recovered overlap is smaller than expected.
-However, a more likely explanation is that larger adiabatic overlaps are possible with models that we have not considered here (either with different values of $t_\mathrm{match}$ or outside our one-parameter family altogether).
+This might be caused by higher-order resonances disrupting the evolution, in which case $t\sub{ad}$ should be smaller and the systems would lie on the expectation line.
+%Alternatively, the low-order resonances might be long-lived, explaining why the recovered overlap is smaller than expected.
+However, a more likely explanation is that larger adiabatic overlaps are possible with models that we have not considered here (either with different values of $t\sub{match}$ or outside our one-parameter family altogether).
 
@@
 However, the vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity.
-For the smallest mass ratios, the inspiral rate is slow, and so the systems do not encounter either the 1:2 or 2:3 resonances during their lifetime.
+For the smallest mass ratios, the inspiral rate is slow, and so the systems do not encounter either the $1$:$2$ or $2$:$3$ resonances during their lifetime.
 Meanwhile for the largest mass ratios, the EMRIs encounter both resonances very close to plunge.
@@
 The remaining $66\%$ of systems have overlaps above the level expected if resonances were to play an important role.
-These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t_\mathrm{ad}$ is small.
+These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t\sub{ad}$ is small.
 For these EMRIs, resonances are not as important as expected, which is likely due to the low eccentricities of our population leading to small (and in most cases, negligible) resonant flux enhancements.
@@
 To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to take into account the resonance behaviour.
-The total number of detectable systems using a threshold SNR of $15$ decreases from $513$ to $492$.
+The total number of detectable systems using a threshold SNR of $15$ decreases from $513$ to $492$, a loss of $4\%$.
 If we increase the threshold, the fractional reduction in the number of detectable systems gets even smaller.
@@
 If this is not possible, \eqnref{delta-I-a} can be used to join together different adiabatic models, if the phase on resonance and the magnitude of the self-force can be predicted sufficiently accurately.
-Alternatively, the jump $\Delta\mathcal{I}_\mathrm{jump}$ for each orbital parameter could be included as a free parameter in the template space, and then searched over.
+Alternatively, the jump $\Delta\mathcal{I}\sub{jump}$ for each orbital parameter could be included as a free parameter in the template space, and then searched over.
 Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.
@@
 \Figref{res-plane-2-5-95} shows the semilatus rectum, eccentricity and (cosine of the) inclination angle of the $\nu = 2/5$ resonance surface for a BH of spin $a_\ast = 0.95$.
-\begin{figure}[htbp]
+\begin{figure}
 \centering
@@
 This gives us a total of $12$ parameters for our fit.
-Whilst this may sound large, if we were fitting an expansion to quadratic order in any combination of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
+Whilst this may sound large, if we were fitting an expansion to quadratic order in combinations of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
 \begin{equation}
@@
 \end{equation}
-These were fitted for all possible resonances with $n_r = 2$--$7$ as well as the $9:10$, $19:20$, $49:50$ and $99:100$ resonances; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
+These were fitted for all possible resonances with $n_r = 2$--$7$ as well as the $9$:$10$, $19$:$20$, $49$:$50$ and $99$:$100$ resonances; with SMBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
 %These were fitted for resonances with $\nu = 1/7, 1/6, 1/5, 1/4, 2/7, 1/3, 2/5, 3/7, 1/2, 4/7, 3/5, 2/3, 5/7, 3/4, 4/5, 5/6, 6/7, 9/10, 19/20, 49/50, 99/100$, with BH spins of $a_\ast = 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.93, 0.95, 0.97, 0.99, 0.999$.
@@
 \end{figure*}
-The largest fractional error is $\sim50\%$, this is for $a\rightarrow 1$ and $\nu \rightarrow 0$, and corresponds to small $p$, such that the absolute error is still small.
+The largest fractional error is $\sim50\%$, this is for $a_\ast \rightarrow 1$ and $\nu \rightarrow 0$, and corresponds to small $p$, such that the absolute error is still small.
 Taking the root-mean-square across $e$ and $\iota$, the fractional error for a given $a_\ast$ and $\nu$ never exceeds $9\%$ and is typically less than $4\%$.

### 2016-08-03 asymptotic.tex
Up to jumps

@@
 The following derivation is reproduced from Berry~\cite{BerryThesis2013}.
-It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two time-scale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found by van de Meent~\cite{VanDeMeent2014}.
+It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two timescale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found in van de Meent~\cite{VanDeMeent2014}.
 
@@
 Similarly, we use $J'$ as the action variable conjugate to $q'$.
-The system evolves through resonance slowly, on an evolution time-scale, so we parameterize it in terms of a slow time parameter
+The system evolves through resonance slowly, on an evolution timescale, so we parameterize it in terms of a slow time parameter
 \begin{equation}
@@
 
-To find a solution away from the resonance, we decompose the problem to be a function of two time-scales~\cite{Kevorkian1971}.
+To find a solution away from the resonance, we decompose the problem to be a function of two timescales~\cite{Kevorkian1971}.
 We use the slow time $\widetilde{\lambda}$ and, as a proxy for the fast time,
@@
 The higher-order terms are identified below by matching terms in the series expansions of the equations of motion.
-Taking the two time-scales as independent, we may write the time derivative to $\order{\eta}$ as
+Taking the two timescales as independent, we may write the time derivative to $\order{\eta}$ as
 \begin{equation}
@@
 
-Using the two time-scale decomposition to replace the time derivatives in the equations of motion, and substituting in the ansatz expansions gives, to first order,
+Using the two timescale decomposition to replace the time derivatives in the equations of motion, and substituting in the ansatz expansions gives, to first order,
 \begin{widetext}
@@
 Near to resonance, we consider an interior layer expansion~\cite{Kevorkian1971}.
-As explained in \secref{res-time}, evolution near resonance proceeds on a time-scale intermediate between the slow and fast times.
+As explained in \secref{res-time}, evolution near resonance proceeds on a timescale intermediate between the slow and fast times.
 We therefore introduce a rescaled time

### 2016-08-03 cut_figures.tex
Up to jumps

--- 2015-06-29 res_char.tex
@@
-\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,showpacs,superscriptaddress,twocolumn]{revtex4}
-\usepackage{graphicx}
-\usepackage{color}
-%\usepackage{psfrag}
-\usepackage{mathrsfs}
-\usepackage{dcolumn}
-\usepackage{hyperref}
-\usepackage{textcomp}
-
-\newcommand{\eqnref}[1]{(\ref{eq:#1})}
-\newcommand{\figref}[1]{Fig.\ \ref{fig:#1}}
-\newcommand{\Figref}[1]{Fig.\ \ref{fig:#1}}
-\newcommand{\secref}[1]{Sec.\ \ref{sec:#1}}
-\newcommand{\apref}[1]{Appendix \ref{sec:#1}}
-\newcommand{\Tabref}[1]{Table \ref{tab:#1}}
-
-\newcommand{\units}[1]{\ensuremath{~\mathrm{#1}}}
-
-\newcommand{\sub}[1]{\ensuremath{_\text{#1}}}
-\newcommand{\super}[1]{\ensuremath{^\text{#1}}}
-\newcommand{\dd}{\ensuremath{\mathrm{d}}}
-\newcommand{\diff}[2]{\ensuremath{\dfrac{\dd {#1}}{\dd {#2}}}}
-\newcommand{\linediff}[2]{\ensuremath{\dd {#1}/\dd {#2}}}
-\newcommand{\diffop}[1]{\ensuremath{\dfrac{\dd}{\dd {#1}}}}
-\newcommand{\difftwo}[2]{\ensuremath{\dfrac{\dd^2 {#1}}{\dd {#2}^2}}}
-\newcommand{\linedifftwo}[2]{\ensuremath{\dd^2 {#1}/\dd {#2}^2}}
-\newcommand{\difftwoop}[1]{\ensuremath{\dfrac{\dd^2}{\dd {#1}^2}}}
-\newcommand{\diffn}[3]{\ensuremath{\dfrac{\dd^{#3} {#1}}{\dd {#2}^{#3}}}}
-\newcommand{\linediffn}[3]{\ensuremath{\dd^{#3} {#1}/\dd {#2}^{#3}}}
-\newcommand{\diffnop}[2]{\ensuremath{\dfrac{\dd^{#2}}{\dd {#1}^{#2}}}}
-\newcommand{\partialdiff}[2]{\ensuremath{\dfrac{\partial {#1}}{\partial {#2}}}}
-\newcommand{\linepartialdiff}[2]{\ensuremath{\partial {#1}/\partial {#2}}}
-\newcommand{\partialdiffop}[1]{\ensuremath{\dfrac{\partial}{\partial {#1}}}}\newcommand{\recip}[1]{\ensuremath{\frac{1}{#1}}}
-\newcommand{\grad}{\ensuremath{\boldsymbol{\nabla}}}
-\newcommand{\intd}[4]{\ensuremath{\int_{#1}^{#2}{#3}\,\dd{#4}}}
-\newcommand{\order}[1]{\ensuremath{\mathcal{O}({#1})}}
-
-\newcommand{\overlap}[2]{\ensuremath{\left(#1\middle|#2\right)}}
-
-\DeclareMathOperator{\sgn}{sgn}
-
-\begin{document}
-
-%\preprint{}
-
-\title{The importance of transient resonances in extreme-mass-ratio inspirals (new title?)}
-
-\author{Robert H. Cole}
-%\email[]{rhc26@ast.cam.ac.uk}
-\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
-\author{Christopher P.L. Berry}
-\email[]{cplb@star.sr.bham.ac.uk}
-\affiliation{School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
-\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
-\author{Priscilla Ca\~{n}izares}
-%\email[]{pcm@ast.cam.ac.uk}
-\author{Jonathan R. Gair}
-%\email[]{jgair@ast.cam.ac.uk}
-\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
-
-\date{\today}
-
-\begin{abstract}
-\emph{Update abstract}
-The inspiral of stellar-mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
-In order to detect such signals and extract the desired information, it is necessary to possess a bank of accurate waveform templates.
-For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first order dissipative gravitational self-force.
-Other effects are however neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
-During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact-object trajectory and, ultimately, the phase of the emitted radiation.
-In this work, an astrophysical population of detectable extreme-mass-ratio inspirals is studied and it is found that a large proportion shall encounter a low-order resonance in the later stages of inspiral.
-The resulting effect on signal-to-noise recovery is small due to the low eccentricity of the studied population, but nevertheless at least one fifth of the systems may become undetectable if transient resonances are not accounted for.
-\end{abstract}
-
-% 04.25.Nx 	Post-Newtonian approximation; perturbation theory; related approximations
-% 04.30.-w 	Gravitational waves
-% 04.70.-s 	Physics of black holes
-% 98.62.Js 	Galactic nuclei (including black holes), circumnuclear matter, and bulges 
-\pacs{04.25.Nx, 04.30.--w, 04.70.--s, 98.62.Js}
-
-\maketitle
-
-\section{Introduction}
-
-In the prologue to his classic monograph, Chandrasekhar~\cite{Chandrasekhar1992} celebrates the simplicity of black holes (BHs).
-The Kerr solution is defined by just two parameters: mass and spin.
-Despite the baldness of the BH metrics, great intricacies manifest in their properties.
-This is made evident when a second body is introduced.
-The two-body problem in general relativity (GR) is well studied.
-It is of paramount importance for gravitational-wave (GW) astronomy, where binary systems are the dominant source of radiation.
-Correctly modelling the dynamics of these systems is necessary to interpret and extract information from gravitational waveforms.
-
-We have made progress in understanding the general relativistic two-body problem in recent years.
-Bodies of comparable mass can be studied using numerical relativity.
-Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}; it is now possible to simulate hundreds of orbits~\cite{Szilagyi2015}.
-These simulations allow us to understand BH--BH mergers.
-Stellar-mass BH mergers are targets for ground-based GW detectors, such as the upcoming Advanced LIGO~\cite{Aasi2015} and Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Kuroda2010}.
-Massive BH mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
-Systems of unequal masses are more challenging to evolve numerically as they complete a larger number of orbits and it is necessary to resolve two different scales.
-Calculations can instead be performed perturbatively.
-The paradigm unequal-mass system has a stellar-mass BH orbiting a massive black hole, such as those expected to be found at the centre of most galaxies.
-These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors~\cite{Amaro-Seoane2007}.
-Despite being well-studied there still remain open questions.
-
-In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}.
-In the test particle limit, the smaller body follows an exact geodesic of the BH's spacetime.
-Including the effects of the smaller body's finite mass, the background spacetime is perturbed.
-The back-reaction from this deformation alters the small body's orbital trajectory, and can be modelled as a self-force that moves the body from its geodesic.
-The self-force is commonly divided into two pieces, dissipative and conservative.
-The former encapsulates the slow decay of the orbital energy and angular momentum, constants of the motion in the test particle limit, through radiation.
-The latter shifts the orbital phases inducing precession.
-The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase although this can accumulate over many orbits.
-Being able to accurately model the influence of the self-force allows us to create reliable waveform models.
-
-Flanagan and Hinderer~\cite{Flanagan2012} highlighted a previously overlooked phenomenon that occurs in the general relativistic two-body problem, that of transient resonances.
-Geodesic orbits in GR have three associated frequencies: the radial frequency $\Omega_r$, the polar frequency $\Omega_\theta$ and the azimuthal frequency $\Omega_\phi$.
-The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
-% Section 10.6
-In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field, $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
-For extreme-mass-ratio systems, the evolution time-scale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital time-scales.
-The evolution of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
-During this evolution, the frequencies may become commensurate.
-Resonances occur when the radial and polar frequencies are rational multiples of each other:
-\begin{equation}
-\nu \equiv \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r},
-\end{equation}
-where $n_r$ and $n_\theta$ are integers (with no common factors).
-During resonance, terms in the self-force that usually average to zero can combine coherently, significantly impacting the orbital motion~\cite{Flanagan2012a}.
-
-Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.
-However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2014} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central black hole from its host.
-
-Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
-We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
-These are some combination of the energy per unit mass $E$ and the axial angular momentum per unit mass $L_z$, which arise from isometries of the metric in $t$ and $\phi$, and the Carter constant per unit mass squared $Q$~\cite{Carter1968}, which is related to the separability of the equations of motion in $r$ and $\theta$.
-The system evolves following~\cite{Flanagan2012}
-\begin{subequations}
-\label{eq:Mino-E-o-M}
-\begin{align}
-\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta g_\alpha^{(1)}(q_r,q_\theta,\boldsymbol{J}) + \order{\eta^2}, \\
-\diff{J_\alpha}{\lambda} = {} & \eta G_\alpha^{(1)}(q_r,q_\theta,\boldsymbol{J}) + \order{\eta^2},
-\end{align}
-\end{subequations}
-where $\lambda$ is Mino time~\cite{Mino2003}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force.
-By working with $\lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
-At zeroth order in the mass ratio we recover the limit of purely geodesic motion: the integrals of the motion are actually constants and the angle variables evolve according to their associated frequencies $\omega_\alpha$.
-
-The leading-order dissipative correction to geodesic motion is calculated following the adiabatic prescription~\cite{Hinderer2008}: by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing the forcing term $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$ $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
-For most orbits this is sufficient, $G_\alpha^{(1)}$ is given by its average value plus a rapidly oscillating component~\cite{Arnold1988}.
-% Chapter 5, section 1
-However, this averaging fails when the ratio of frequencies is the ratio of integers.
-In this case the trajectory does not ergodically fill the $2$-torus, but instead traces out a $1$-dimensional subspace.\footnote{For illustrations, see Grossman, Levin and Perez-Giz~\cite{Grossman2012}.} There are then contributions to the self-force that no longer average out beyond $\langle G_A^{(1)}\rangle_{q_r,\,q_\theta}$.
-Intuitively, we expect that this effect is more important for ratios of small integers as when the integers are large, the orbit comes close to all points on the $2$-torus.
-
-In this work we seek to characterise the importance of these resonances for the purposes of modelling EMRIs.
-The amplitude of expected signals is below the level of noise in a space-based GW detector.
-However, systems remain in band for many hundreds of thousands of cycles and so may be detected using a matched filter, provided we have sufficiently accurate waveform templates.
-Ensuring the accuracy of EMRI templates requires calculating the impact that passing through a resonance has on the orbital evolution and discovering for which resonances this is significant.
-
-In \secref{problem}, we formulate the specific problem: that of geodesic motion in Kerr spacetime, perturbed by the gravitational self-force.
-We then study generic properties of transient resonances in \secref{properties}, detailing their location in parameter space, the time-scales over which they affect the motion and the resulting GW flux enhancements.
-Specific examples are considered to illustrate the effects of resonances at the start of \secref{astrophysics}, before finally turning to an astrophysical population in \secref{population}.
-Our conclusions can be found in \secref{conclusion}.
-
-We use geometric units with $G = c = 1$ throughout.
-We always use $M$ for the mass of the central massive BH and $a$ as the spin parameter.
-We also use the dimensionless spin $a_\ast \equiv a/M$; we take the convention that $0 \leq a_\ast < 1$.
-
-\section{Problem formulation}
-\label{sec:problem}
-\subsection{Kerr geodesics}
-
-The evolution of an extreme-mass-ratio ($\eta \ll 1$) system is slow.
-Instantaneously, the motion of the orbiting mass can be described as geodesic, with the integrals of the motion changing on time-scales of many orbital periods.
-We analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance.
-It is therefore necessary to develop an understanding of the Kerr geodesics.
-
-The geodesic equations may be written as~\cite{Carter1968, Chandrasekhar1992} % section 62
-\begin{subequations}
-\begin{align}
-\diff{t}{\lambda} = {} & a\left(L_z - aE\sin^2 \theta\right) + \frac{r^2 + a^2}{\Delta}\mathcal{T},\\
-\diff{r}{\lambda} = {} & \pm \sqrt{V_r},\\
-\diff{\theta}{\lambda} = {} & \pm \sqrt{V_\theta},\\
-\diff{\phi}{\lambda} = {} & \frac{L_z}{\sin^2 \theta} - aE + \frac{a}{\Delta}\mathcal{T},
-\end{align}
-\end{subequations}
-where $\Delta = r^2 - 2M r + a^2$; the signs of the $r$ and $\theta$ equations can be chosen independently, and we have introduced potentials
-\begin{subequations}
-\begin{align}
-\mathcal{T} = {} & E\left(r^2 +a^2\right) - aL_z,\\
-V_r = {} & \mathcal{T}^2 - \Delta\left[r^2 + \left(L_z -aE\right)^2 + Q\right],\\
-V_\theta = {} & Q - \cos^2 \theta\left[a^2\left(1 - E^2\right) + {\displaystyle \frac{L_z^2}{\sin^2\theta}}\right].
-\end{align}
-\end{subequations}
-As an affine parameter, we have used Mino time which is related to the proper time $\tau$ by~\cite{Mino2003}
-\begin{equation}
-\tau = \intd{}{}{r^2 + a^2 \cos^2\theta}{\lambda}.
-\end{equation}
-Using Mino time allows us to decouple the $r$ and $\theta$ motions.
-
-We only consider bound motion~\cite{Wilkins1972}: the radial motion covers a range $r\sub{p} \leq r \leq r\sub{a}$, where the turning points are the periapsis $r\sub{p}$ and apoapsis $r\sub{a}$.
-Drawing upon Keplerian orbits we parametrize the motion using
-\begin{equation}
-r = \frac{p M}{1+e\cos\psi},
-\end{equation}
-introducing eccentricity $e$, (dimensionless) semilatus rectum $p$ and relativistic anomaly $\psi$~\cite{Darwin1961,Drasco2004}.
-While $r$ oscillates between its maximum and minimum values, $\psi$ increases secularly, increasing by $2\pi$ across an orbit.
-The polar motion covers a range $\theta_- \leq \theta \leq \pi - \theta_-$.
-We also parametrize this motion in terms of an angular phase $\chi$, according to~\cite{Hughes2000}
-\begin{equation}
-\cos\theta = \cos\theta_-\cos\chi.
-\end{equation}
-While $\psi$ and $\chi$ are $2\pi$ periodic they are not the canonical action--angle variables~\cite{Schmidt2002}; they are, however, easy to work with.
-
-The geodesic motion can equally be described by $\{E,L_z,Q\}$ or $\{p,e,\theta_-\}$~\cite{Schmidt2002}.
-Converting between them requires finding the solutions of $V_r = 0$ and $V_\theta = 0$.
-We employ a slightly different parameter set of $\{p,e,\iota\}$ where we have introduced the inclination~\cite{Ryan1996,Glampedakis2002}
-\begin{equation}
-\tan \iota = \frac{\sqrt{Q}}{L_z}.
-\end{equation}
-This is $0 \leq \iota < \pi/2$ for prograde orbits and $\pi/2 < \iota \leq \pi$ for retrograde orbits.
-Equatorial orbits ($\theta_- = \pi/2$) have $\iota = 0$ or $\pi$ and polar orbits ($\theta_- = 0$) have $\iota = \pi/2$.
-While formulae exist for conversion between the different parameters, these are complicated and uninsightful, so we do not reproduce them here.\footnote{In practice we find turning points numerically.}
-
-\subsection{Orbital resonances}
-
-The radial and polar orbital periods in Mino time are given by
-\begin{subequations}
-\begin{align}
-\Lambda_r = {} & 2\intd{r\sub{p}}{r\sub{a}}{\recip{\sqrt{V_r}}}{r} = \intd{-\pi}{\pi}{\diff{\lambda}{\psi}}{\psi}, \\
-\Lambda_\theta = {} & 4\intd{\theta_-}{\pi/2}{\recip{\sqrt{V_\theta}}}{\theta} = \intd{-\pi}{\pi}{\diff{\lambda}{\chi}}{\chi}.
-\end{align}
-\end{subequations}
-The orbital frequencies are thus
-\begin{equation}
-\Upsilon_r = \frac{2\pi}{\Lambda_r}, \quad \Upsilon_\theta = \frac{2\pi}{\Lambda_\theta}.
-\end{equation}
-The geodesic equations for coordinate time $t$ and azimuthal angle $\phi$ are just functions of $r$ and $\theta$, hence their evolutions can be expressed as Fourier series~\cite{Drasco2004}
-\begin{subequations}
-\begin{align}
-\diff{t}{\lambda} = {} & \sum_{k_r,\,k_\theta}T_{k_r,\, k_\theta}\exp\left[-i\left(k_r\Upsilon_r + k_\theta\Upsilon_\theta\right)\lambda\right], \\
-\diff{\phi}{\lambda} = {} & \sum_{k_r,\,k_\theta}\Phi_{k_r,\, k_\theta}\exp\left[-i\left(k_r\Upsilon_r + k_\theta\Upsilon_\theta\right)\lambda\right].
-\label{eq:Mino-Fourier}
-\end{align}
-\end{subequations}
-The $(0,\,0)$ coefficients in these series give the average secular rate of increase of these quantities. We define
-\begin{equation}
-\Gamma = T_{0,\,0}, \quad \Upsilon_\phi = \Phi_{0,\,0}
-\end{equation}
-to act as Mino-time frequencies. We can now convert to coordinate-time frequencies with
-\begin{equation}
-\Omega_r = \frac{\Upsilon_r}{\Gamma}, \quad \Omega_\theta = \frac{\Upsilon_\theta}{\Gamma}, \quad \Omega_\phi = \frac{\Upsilon_\phi}{\Gamma}.
-\end{equation}
-
-Transient resonances occur when the radial and poloidal motions are commensurate, when
-\begin{equation}
-\nu = \frac{\Upsilon_r}{\Upsilon_\theta} = \frac{\Omega_r}{\Omega_\theta} = \frac{n_\theta}{n_r}
-\end{equation}
-is the ratio of small integers.
-At this point, any Fourier series like those in \eqnref{Mino-Fourier} goes from being an expansion in two frequencies to being an expansion in a single frequency~\cite{Bosley1992}.
-
-For a general non-resonant orbit there is no fixed correlation between the radial and polar coordinates.
-After a sufficiently long time, the trajectory comes arbitrarily close to every point in the range of motion (with $r\sub{p} \leq r \leq r\sub{a}$ and $\theta_- \leq \theta \leq \pi - \theta_-$); on account of the orbital precession, the whole space is densely covered.
-This does not happen on resonance, as the radial and polar motions are locked together such that we can express one as a function of the other, and so the trajectory keeps cycling over the same path.
-The points visited are controlled by the relative phases of the $r$ and $\theta$ motions.
-To represent this, we use the $r$ phase at the $\theta$ turning point $\psi_{\theta_-} = \psi(\chi = 0)$.
-Varying $\psi_{\theta_-}$ across its full range allows every point in the range of motion to be reached.
-Hence averaging over all values of $\psi_{\theta_-}$ for resonant orbits is equivalent to averaging over the $\psi$--$\chi$ $2$-torus for non-resonant orbits.
-
-One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short time-scales, for small $\eta$ there is a small disparity.
-The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e.
-the case for motion in Kerr) is subject to a small perturbation the form of the orbits is preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
-% Chapter II, section 3.3 d)
-This should ensure that, in general, there are only small shifts in the orbital frequencies.
-However, the KAM theory is only valid for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}.
-% Chapter V section 1.
-c)
-This is a further reason why resonances merit an in-depth investigation.}
-The dissipative piece causes the frequencies to evolve and, hence, the resonance cannot persist for multiple orbits (without some feedback coupling).
-In effect, we are really considering a period of time about the resonant crossing.
-The instantaneous orbital frequencies oscillate back and forth around their averaged values.
-However, there is a time span when the frequencies are consistently close to being commensurate.
-During this time, the trajectory appears similar to a resonant trajectory, filling only a smaller region of the parameter space.
-It is this time period that is of interest for transient resonances~\cite{Bosley1992}.
-
-%The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e., the case for motion in Kerr) is subject to a small perturbation the form of the orbits are preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
-% Chapter II, section 3.3 d)
-%When we average the self-force and follow the adiabatic evolution of the trajectory we are exploiting this behaviour.
-However, the KAM theory only applies for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}.
-% Chapter V section 1.
-c)
-% In these regions the adiabatic approximation breaks down and we need to follow the instantaneous evolution.
-
-\subsection{Forced motion}
-\label{sec:forced-motion}
-
-For generic EMRIs, there are two timescales of interest: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$.
-These, along with the resonance timescale, are discussed more in \secref{res-time}.
-The two-timescale nature of the problem makes it ideally suited to the method of osculating elements~\cite{Gair2011a}: on short timescales, we analyse the unperturbed system resulting in geodesic motion, and then the long-term evolution is described by a sequence of instantaneous geodesics.
-
-We require, at each instant in time, that the chosen geodesic matches the true position and velocity of the particle.
-This amounts to a specific choice of the orbital shape parameters (for example, the set $\{E,L_z,Q\}$ or the generalised action variables $J_\alpha$) and some initial phases at $t=t_0$ (for example, the set $\{\psi_0,\chi_0,\phi_0\}$).
-Collectively, these are referred to as \emph{osculating elements} and we denote them by $I^A(t)$, making explicit the variation with time.
-
-Given geodesic motion in some background spacetime with Christoffel symbols $\Gamma^\alpha_{\beta\gamma}$, forced by some external acceleration $f^\alpha$, so that
-\begin{equation}
-\difftwo{x^\alpha}{\tau} + \Gamma^\alpha_{\beta\gamma}\diff{x^\beta}{\tau}\diff{x^\gamma}{\tau} = f^\alpha,
-\end{equation}
-we can calculate the evolution of the osculating elements $\dot{I}^A$. The specific equations for motion in Kerr are derived by Gair {\it{et al}}.~\cite{Gair2011a}.
-
-\subsection{Self-force model}
-\label{sec:self-force}
-
-To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration.
-In this work we choose to work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
-For comparison, Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a} use a Teukolsky equation calculation of GW fluxes to account for the inspiral due to radiation reaction.
-
-The self-force model uses the first-order PN terms of the dissipative self-force formulated in Flanagan and Hinderer~\cite{Flanagan2007} and the conservative force formulated in Iyer and Will~\cite{Iyer1993}, and Kidder~\cite{Kidder1995}.
-Since only the first PN terms are used, this prescription can only be of limited validity in strong fields.
-Both pieces of the self-force are computed assuming that the spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$.
-This is less than ideal for high spins.
-We also find that this particular implementation of the self-force model marginally overestimates the adiabatic inspiral rate with respect to direct PN evolutions by a factor of $\mathcal{O}(1)$, even for systems in the weak field and with low values of the spin.
-While this approximate self-force is not perfect, it should serve as a guide for the behaviour of the full self-force, allowing us to assess the qualitative impact of resonances on EMRI detection.
-
-\emph{Add detail about why self-force model is rubbish. Comment on switching conservative piece off for population results.}
-
-\subsection{Adiabatic evolution}
-\label{sec:adiabatic}
-
-Beyond geodesic motion in the Kerr spacetime, a test particle follows an accelerated trajectory determined by \eqnref{Mino-E-o-M}.
-This may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
-This piece is purely dissipative and determines how the inspiral evolves due to the radiation of GWs.
-
-To construct an adiabatic trajectory we need the $2$-torus-averaged fluxes of our osculating elements.
-Computing an average of a quantity over the $\{q_r, q_\theta\}$ is trivial if it is parametrized in terms of these variables,
-\begin{equation}
-\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta} = \frac{1}{(2\pi)^2}\intd{0}{2\pi}{\intd{0}{2\pi}{\diff{X}{\lambda}}{q_r}}{q_\theta}.
-\end{equation}
-However, we are using $\psi$ and $\chi$, as these are simpler to evolve; furthermore, we compute instantaneous co-ordinate time fluxes $\dot{X}$, not Mino-time fluxes.
-Changing variables gives an average of \cite{Drasco2005}
-\begin{widetext}\begin{align}
-\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta} = {} & \frac{1}{\Lambda_r \Lambda_\theta}\intd{0}{2\pi}{\intd{0}{2\pi}{\left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-2} \diff{X}{\lambda}}{\psi}}{\chi} \\
- = {} & \frac{1}{\Lambda_r \Lambda_\theta}\intd{0}{2\pi}{\intd{0}{2\pi}{\left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X}}{\psi}}{\chi}.
-\label{eq:2torus-average}
-\end{align}\end{widetext}
-This average describes the Mino-time rate of change of the quantity $X$ over an orbit. To convert to a co-ordinate flux of the averaged quantity, we simply divide by the period $\Gamma$, defining
-\begin{equation}
-\dot{\left\langle X\right\rangle}_{q_r,\,q_\theta} = \frac{1}{\Gamma}\left\langle \diff{X}{\lambda}\right\rangle_{q_r,\,q_\theta}.
-\end{equation}
-It is convenient to calculate $\Gamma$ as
-\begin{equation}
-\Gamma = \left\langle \diff{t}{\lambda}\right\rangle_{q_r,\,q_\theta},
-\end{equation}
-using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.\footnote{We compute these integrals using a $300\times 300$ grid of $\{\psi,\chi\}$ values and employing a Newton--Cotes approximation in each dimension.
-The procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform.
-However, the adiabatic derivatives vary on much longer timescales than the orbital motion (see \secref{res-time}), and so in practice, we can interpolate.} %We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\delta t/\eta$ where $\delta t$ is the orbital time-step.
-When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives.
-Once the system evolves close to one of the boundaries, we update the interpolation points, taking the closest boundary to be the new central point.
-The averaged fluxes successfully describe the leading-order secular evolution of the trajectory (as illustrated in \figref{good-traj}).
-
-%In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform appropriate averages over this new $2$-torus.
-The osculating elements integration scheme involves computing derivatives with respect to coordinate time, and we shall often be interested in the averaged form of such derivatives.
-Given a function $X(\psi,\chi)$, we can calculate the 2-torus average \cite{Drasco2005} of the derivative with respect to Mino time in terms of precomputed quantities
-%\begin{widetext}\begin{equation}
-%\label{eq:2torus-average}
-%\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi} = \frac1{\Lambda_r\Lambda_\theta} \intd{0}{2\pi}{ \intd{0}{2\pi}{ \left(\diff{\psi}{t}\right)^{-1} \left(\diff{\chi}{t}\right)^{-1} \left(\diff{t}{\lambda}\right)^{-1} \dot{X} }{\chi} }{\psi}.
-%\end{equation}\end{widetext}
-%We then obtain a coordinate time derivative
-%\begin{equation}
-%\dot{\left\langle X\right\rangle}_{\psi,\,\chi} = \left.\left\langle \diff{X}{\lambda}\right\rangle_{\psi,\,\chi}\middle/\left\langle \diff{t}{\lambda}\right\rangle_{\psi,\,\chi}\right.,
-%\end{equation}
-%which is successful at describing the leading-order secular evolution of the trajectory, as illustrated in \figref{good-traj}.
-
-%The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$.
-The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurence of any transient resonances.
-Resonances still occur as the fundamental frequencies become commensurate, but as the system passes through each resonance, there is no additional effect on the evolution of the system.
-
-%One method of studying the effect of resonances is then to compare an adiabatic evolution with an instantaneous evolution, using the full knowledge of the self-force.
-This may be a useful exercise to calibrate and verify adiabatic models that can be generated cheaply; however it cannot be used to distinguish between dephasing effects arising from resonances and those arising from conservative corrections to the phase variables, as both are explicitly eliminated from the adiabatic prescription.
-We therefore take a different approach, comparing the instantaneous evolution with a quasi-adiabatic evolution, generated by averaging the full available self-force over the $2$-torus, rather than just averaging the term $G_\alpha^{(1)}$.
-
-%To be specific, we write \eqnref{Mino-E-o-M} as
-%\begin{subequations}
-%\begin{align}
-%\label{eq:Mino-E-o-M2}
-%\diff{q_\alpha}{\lambda} = {} & \omega_\alpha(\boldsymbol{J}) + \eta \langle g_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{g}_\alpha(q_r,q_\theta,\boldsymbol{J}), \\
-%\diff{J_\alpha}{\lambda} = {} & \eta \langle G_\alpha\rangle_{q_r,\,q_\theta}(\boldsymbol{J}) + \eta \hat{G}_\alpha(q_r,q_\theta,\boldsymbol{J}),
-%\end{align}
-%\end{subequations}
-%where $g_\alpha$ and $G_\alpha$ encapsulate all our knowledge about the self-force and may contain contributions from sub-leading terms in the mass-ratio expansion.
-The quasi-adiabatic evolution is then obtained by dropping the terms containing $\hat{g}_\alpha$ and $\hat{G}_\alpha$.
-For ease of notation, we shall drop the \emph{quasi-} identifier; all subsequent references to an ``adiabatic evolution'' in fact refer to a ``quasi-adiabatic evolution''.
-
-The combination of a full instantaneous evolution and an adiabatic evolution allows us to systematically study the effect of transient resonances on EMRIs over the course of an inspiral.
-Before approaching this problem, we first investigate the properties of the resonances themselves.
-
-\section{Properties of Transient Resonances}
-\label{sec:properties}
-
-The first step in studying the effect of transient resonances is to locate orbital parameters for which the frequencies are commensurate.
-We can calculate the frequencies and so we are left with the problem of solving $\Omega = n_r \Omega_r - n_\theta \Omega_\theta = 0$ numerically.
-When considering the full parameter set of $\{p,e,\iota,a_\ast,\nu\}$, it is apparent that the search for resonances becomes expensive as a consequence of the dimensionality.
-It is therefore useful to have a guide of where to look.
-In \apref{location} we build a simple approximate model as a starting point for the numerical search.
-The resonances occur at relatively small periapses, corresponding to regions of strong-field gravity.
-Having located where in an inspiral we can expect to encounter a transient resonance, we must now consider its impact.
-In \secref{res-time} we determine the characteristic time-scales describing resonance, and in \secref{flux-enhance} we look at the impact of passing through a resonance on the evolution of the orbit, including some numerical results.
-
-\subsection{Time-scales}\label{sec:res-time}
-
-When analysing resonances it is useful to refer to a number of characteristic time-scales.
-We always use coordinate time $t$ for these, as this corresponds to what is measured by an observer at infinity.
-Translation to Mino time can be done with an appropriate factor of $\Gamma$.
-We use the orbital period $T$, the evolution time-scale $\tau\sub{ev}$, the precession time-scale $\tau\sub{pres}$ and the resonance time-scale $\tau\sub{res}$.
-
-The simplest time-scales are the orbital periods $T_r = 2\pi/\Omega_r$, $T_\theta = 2\pi/\Omega_\theta$ and $T_\phi = 2\pi/\Omega_\phi$.
-These are the shortest in our set.
-We use $T$ to denote a time-scale of the same order as the orbital periods.
-
-We define the evolution time-scale as
-\begin{equation}
-\tau\sub{ev} = \frac{\nu}{\dot{\nu}},
-\end{equation}
-where an overdot denotes a derivative with respect to $t$.
-In general, away from resonance, we take $\nu \equiv \Omega_r/\Omega_\theta$.
-This time-scale sets the period over which there is a significant change in the frequencies.
-It acts as an inspiral time-scale.
-It is long in all cases we study, $\tau\sub{ev} \sim \order{T/\eta}$.
-It is this property which makes extreme-mass-ratio inspirals interesting, as we can follow the waveform for many cycles, accruing high signal-to-noise ratios.
-This is also what allows us to use the adiabatic prescription, as it means the trajectory moves slowly through different orbital parameters.
-
-We use the precession time-scale
-\begin{equation}
-\tau\sub{pres}(t) = \frac{2\pi}{|\Omega(t)|},
-\label{eq:t-pres}
-\end{equation}
-with $\Omega(t) = n_r \Omega_r(t) - n_\theta \Omega_\theta(t)$, where the frequencies are calculated instantaneously and the integers are for the resonance of interest.
-This time-scale becomes infinite exactly on resonance, but decreases as we get further from resonance, eventually becoming $\order{T}$.
-It measures the relative precession rate of the radial and polar motions and hence gives an indication of how long it takes to fill the entire $\psi$--$\chi$ $2$-torus.
-
-We also use the resonance time-scale
-\begin{equation}
-\tau\sub{res} = \left[\frac{2\pi}{\left|\left\langle\dot{\Omega}(0)\right\rangle_{q'}\right|}\right]^{1/2}.
-\label{eq:t-res}
-\end{equation}
-Here $\dot{\Omega}(0)$ is the rate of change of $\Omega$ at resonance, which we take to be at $t = 0$.
-The instantaneous $\dot{\Omega}$ depends upon the orbital phase and oscillates about its mean trend over an orbit.
-We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time-average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$.
-Close to resonance, $\Omega(t)$ is well approximated by a first-order Taylor expansion, decreasing linearly with time; hence we make the approximation
-\begin{equation}
-\left|\left\langle\dot{\Omega}\right\rangle_{q'}\right| \simeq \left|\frac{\Omega(t)}{t}\right|.
-\end{equation}
-The resonant time-scale should give an indication of the time over which we expect the effects of the resonance to be felt~\cite{Bosley1992}.
-Consider the phase of the Mino time Fourier expansion on resonance; neglecting the constant, the resonant Fourier component has form
-\begin{equation}
-\varphi_{n_r,\,-n_\theta} \simeq \left(n_r\Upsilon_r - n_\theta\Upsilon_\theta\right)\lambda + \left(n_r\dot{\Upsilon}_r - n_\theta\dot{\Upsilon}_\theta\right)\lambda^2 + \ldots
-\end{equation}
-Typically, the first term is non-zero and this gives the familiar oscillation.
-On resonance, it is zero, leaving the next order term to govern the behaviour~\cite{Flanagan2012}.
-Only once we have moved far enough away from resonance for the first term to dominate the second do we recapture the non-resonant behaviour.
-The first term (translating from Mino time to coordinate time) sets $\tau\sub{pres}$, the second sets $\tau\sub{res}$.
-
-Since we have argued that the effect of resonance can be thought of as a consequence of not densely covering the $\psi$--$\chi$ $2$-torus, we might expect that $\tau\sub{pres}$, as well as $\tau\sub{res}$, could be used for setting the resonance duration: the resonance ends once sufficient time has elapsed that the $2$-torus could be filled.
-This is indeed the case.
-Let $t\sub{pres}$ be the time taken to fill the torus, then
-\begin{align}
-t\sub{pres} = {} & \tau\sub{pres}(t\sub{pres}) \\
- \simeq {} & \dfrac{2\pi}{\left|\left\langle\dot{\Omega}(0)\right\rangle_{q'} t\sub{pres}\right|}, \nonumber 
-\end{align}
-using \eqnref{t-pres} and substituting $\Omega(t\sub{pres}) \simeq \langle\dot{\Omega}\rangle_{q'} t\sub{pres}$. Rearranging and using \eqnref{t-res} gives
-\begin{equation}
-t\sub{pres} \simeq \tau\sub{res}.
-\end{equation}
-The two time-scales are equivalent: we preferentially use $\tau\sub{res}$ to denote the resonance width.
-It is shorter than the inspiral time-scale, but longer than an orbital period, $\tau\sub{res} \sim \order{\eta^{1/2}\tau\sub{ev}} \sim \order{\eta^{-1/2}T}$~\cite{Flanagan2012,Gair2011a}; it therefore acts as a bridge between the two time-scales~\cite{Hinderer2008}.
-
-Since we shall be considering Fourier decompositions, in anticipation of future results, we also define a time-scale for the $s$-th resonant frequency harmonic
-\begin{align}
-%\tau_{\mathrm{pres},\,s} = {} & \frac{2\pi}{|s\Omega(t)|};\\
-\tau_{\mathrm{res},\,s} = {} & \left[\frac{2\pi}{\left|s\left\langle \dot{\Omega}(0)\right\rangle_{q'}\right|}\right]^{1/2}.
-\label{eq:T-res-s}
-\end{align}
-This assumes that $s$ is a non-zero integer.
-
-\subsection{Resonant flux enhancement}\label{sec:flux-enhance}
-
-Evolving through a resonance can lead to an enhancement (or decrement) of fluxes relative to the adiabatic prescription.
-After crossing the resonance region, the orbital parameters are different from those calculated from an adiabatic evolution.
-Flanagan and Hinderer~\cite{Flanagan2012} gave an expression for this deviation.
-If we denote the orbital parameters by $\mathcal{I}_a = \{E,L_z,Q\}$, then the change across resonance is
-\begin{align}
-\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
- {} & \times {} \exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
-\end{align}
-Here $\widehat{\kappa}_0$ is the orbital phase on resonance and $F_{a,\,s}^{(1)}$ is the $s$-th harmonic of the first-order self-force on resonance, defined such that\footnote{Since the geodesic equations decouple in Mino time rather than coordinate time, this is true only in an average sense.}
-\begin{equation}
-\diff{\mathcal{I}_a}{t} = \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
-\end{equation}
-A derivation is presented in \apref{res-asymptotic}, which contains a more comprehensive explanation of the various terms.
-This employs matched asymptotic expansions to track the evolution through resonance, following the approach of Kevorkian~\cite{Kevorkian1987}.
-
-The expression for the change can be understood using our expression for the resonance width, substituting in \eqnref{T-res-s},
-\begin{equation}
-\label{eq:delta-I-a}
-\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
-\end{equation}
-Schematically, this can be understood as the magnitude of the forcing function on resonance $\sim \eta F_{a,\,s}$ multiplied by the time on resonance $\sim \tau_{\mathrm{res},\,s}$ and a function that varies with the phase $\widehat{\kappa}_0$.
-Averaging over all values of $\widehat{\kappa}_0$ is equivalent to averaging over all values of $\psi_{\theta_-}$, and has the same effect as averaging over the $\psi$--$\chi$ $2$-torus; this gives an average discrepancy relative to the adiabatic evolution of
-\begin{equation}
-\left\langle \Delta \mathcal{I}_a \right\rangle_{\hat{\kappa}_0} = 0,
-\end{equation}
-exactly as expected.
-
-\emph{Add in here some numerical results. Definitely plot showing resonant flux enhancement vs eccentricity. Probably not much else.}
-
-We can study the predicted orbital-parameter jumps numerically by considering the long-time-average fluxes for particular (fixed) geodesics.
-For non-resonant geodesics, this will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference.
-We consider a selection of different orbital shape parameters corresponding to low-order resonances and calculate the time-averaged fluxes for $50$ different initial phases.
-We evolve for an integer number of cycles of both the radial and polar motions.
-Specifically, when considering a $n_\theta$:$n_r$ resonance, we take $T = 2\pi N n_\theta/\Omega_r = 2\pi N n_r/\Omega_\theta$ for some integer $N$.
-We have found that taking $N=5$ gives sufficiently accurate results.
-We can then estimate the jumps by taking the difference between the maximum and minimum time-averaged fluxes.
-
-We expect nearly-circular orbits to show small flux enhancements due to the small volume of the $r\theta$ 2-torus; resonant orbits come close to every allowed point, thus approximating a non-resonant orbit.
-In \figref{res-flux-rms-p}, we show the root-mean-square (over $a$ and $\cos\iota$) relative flux enhancement for the semi-latus rectum as a function of eccentricity, for a selection of low-order resonances.
-As expected, it is found that larger eccentricities give rise to larger resonant flux enhancements in general.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{res_flux_rms_p}
-\caption{\label{fig:res-flux-rms-p}Relative flux enhancements of $p$, as a function of $e$, marginalised over $a$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
-Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use solid (short-dashed; long-dashed) lines.
-The 2:3 resonance has the largest relative flux enhancement.}
-\end{figure}
-
-\emph{Is this description clear, and is it actually correct? It seems reasonable ...}
-
-It is informative to consider the dependence on the resonance ratio $\nu$.
-Fixing the eccentricity at $e=0.95$ to emphasise the variation, we compute the flux enhancements for different resonance ratios, shown in \figref{res-flux-rms-p-vsnu}, where we also highlight ratios arising from different values of $n_\theta$.
-We observe a clear relationship between the flux enhancement and $\nu$ for all resonances with $n_\theta = 1$ or $2$, independent of which of those is considered.
-However, negligibly small flux enhancements are observed for systems with $n_\theta = 3$ or higher.
-This can be explained by the quadrupolar nature of GWs.
-A resonant trajectory averaged over time (and consequently $\phi$) takes $n_\theta$ (possibly repeated) values of $\theta$ at each radial turning point.
-Considering rings of matter at these coordinates, the dominant spherical harmonic describing the resultant mass distribution has $l=n_\theta$ (and $m=0$).
-Since GWs are generated by the quadrupole moment at leading order, the most significant resonant enhancements will arise from systems with $n_\theta = 2$.
-Due to the presence of higher harmonics, $n_\theta = 1$ systems also contribute strongly; it is perhaps more appropriate to treat $1$:$x$ resonances as \emph{de facto} $2$:$2x$ resonances, explaining the linked behaviour of the flux enhancements for $n_\theta = 1$ and $2$ systems.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{res_flux_rms_p_e095}
-\caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements of $p$ marginalised over $a$ and $\cos\iota$ by taking the root-mean-square.
-Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use circular (square; triangular) points.}
-\end{figure}
-
-\emph{Is this bit really needed ...?}
-
-The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a}.
-They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a=0.9$, either $e=0.3$ or $0.7$, and $\theta_\mathrm{inc} = 20\text{\textdegree}$ or $70$\textdegree.
-We compare our results to their published values (in their tables I--IV) in \figref{res-flux-FHR}.
-There is some qualitative agreement between the two approaches: more eccentric orbits give rise to larger enhancements; and the 3:4 resonance has a negligibly small enhancement.
-However, for the most significant 2:3 resonance, we typically predict larger orbital parameter jumps.
-Our results in \secref{astrophysics}, regarding the impact of resonances on a population of EMRIs, are therefore conservative predictions, as our model over-estimates the resonance effects.
-
-\begin{figure*}[htbp]
+\begin{figure*}
 \centering
@@
 Each row has a different eccentricity and inclination corresponding to tables I--IV in FHR14.
-All systems have $a=0.9$.}
+All systems have $a_\ast=0.9$.}
 \end{figure*}
 
-Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on a population of inspirals.
-
-\section{Astrophysical implications}
-\label{sec:astrophysics}
-
-To investigate how resonances would impact real-world observations of EMRIs using GWs, we need to generate a realistic astrophysical population of EMRIs and then evolve these systems in time, including a mechanism to generate the GWs at infinity, which can then be analysed.
-
-We discuss our chosen waveform generation scheme, giving demonstrations of its accuracy for our purposes.
-We then detail our procedure for generating a realistic distribution of detectable EMRIs, before turning to the statistical effect of resonances on the population at the end of the section.
-
-\subsection{Waveform model \& analysis}
-\emph{Can probably shorten this section a bit.}
-
-To generate gravitational waveforms, we employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}, augmented to include evolution of the positional elements.
-We first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
-This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically 95\% or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
-
-Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the dissipative part of the post-Newtonian self-force from \secref{self-force}.
-We then map the Boyer--Lindquist coordinates to spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique, the standard quadrupole formula~\cite{Misner1973}.
-
-We expect these waveforms to be sufficiently accurate for our purposes, given the inaccuracy of the self-force model being employed.
-It would, however, be straight-forward to adopt a different waveform generation scheme, for example by including higher-order multipole moments, should advances be made in determining the self-force.
-
-Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s(t)$ and $h(t)$, according to
-\begin{equation}
-\label{eq:innerprod}
-\overlap{s}{h} = 2 \intd{0}{\infty}{\frac{\tilde{s}(f)\tilde{h}^*(f)+\tilde{s}^*(f)\tilde{h}(f)}{S_n(f)}}{f},
-\end{equation}
-where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD)~\cite{Moore2014a}.
-We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept.
-We use the analytic approximation of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}
-
-In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
-To take this into account, we project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme and set any values outside the domain of $\{t_j\}$ to zero.
-A Planck-taper window function~\cite{McKechan2010} is applied to the resulting time series to reduce unwanted spectral leakage in the Fourier transforms.
-%After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
-We maximise every overlap calculation with respect to some time-offset between the models by computing the correlation between the waveforms, which can be performed cheaply using Fourier transforms.
-%and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.
-
-%This whole procedure gives more accurate results than interpolating in the Fourier domain since the waveforms are smoother in the time domain.
-
-One of the targets of pre-eLISA research is to generate a bank of waveform templates, $h(t;\Theta^a)$ described by some set of parameters $\Theta^a$.
-Once real-world data, $s(t)$, has been collected, we can then search over the template bank to find the model that most closely matches the data.
-At the current time, with a lack of experimental data, we are interested in ensuring that the waveform templates accurately reproduce what we expect to observe in nature, without being too computationally expensive.
-
-Ideally, we would like to use waveform templates from EMRI systems that are evolved under the entire self-force model, but these are too expensive.
-In practice, we are forced to use cheaper adiabatic waveforms, but these do not include the effect of resonances.
-To assess the impact of this, we compare data $s(t;\bar{\Theta}^a)$ generated using the full self-force model to templates $h(t;\Theta^a)$ generated using the expensive 2-torus averaged self-force model (see \secref{adiabatic}).
-
-We wish to test whether there exists some set of parameters $\Theta^a$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
-We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template,
-\begin{equation}
-\rho\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{h}{h}}}.
-\end{equation}
-If a template were to exactly match the data, it would produce an SNR of $\sqrt{\overlap{s}{s}}$ and so the overlap
-\begin{equation}
-\label{eq:overlap}
-\mathbb{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}}
-\end{equation}
-provides an indication of how well-matched the template is to the data.
-We leave an investigation into parameter estimation biases (where the parameters of the best-matching adiabatic $\Theta^a$ do not agree with the parameters of the system $\bar{\Theta}^a$) to future work.
-
-To demonstrate the approach, we can calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$ years for a set of initial parameters chosen to avoid any significant resonances.
-Example evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, which shows that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
-The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the 2-torus averaging explicitly smooths out the visible structure on this scale.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=\textwidth]{Fig_good_traj}
-\caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) under the full (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances.
-The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.
-This system has $\mu = 10 M_{\odot}$, $q = 3\times 10^{-6}$, $a_\ast=0.95$, initial semi-latus rectum $p_0=7.5$, initial eccentricity $e_0=0.7$, initial inclination $\cos \iota_0 = 0.5$ and redshift $z=0.204$.}
-\end{figure*}
-
-Once the trajectories have been computed using both methods of evolution, we can calculate the corresponding waveforms.
-The plus-polarised GW at the start and end of the evolution is shown in \figref{good-waveform}.
-The adiabatic waveform shows good agreement with the full waveform in both amplitude and phase.
-Calculating the SNR for each of the models, we find an overlap $\mathbb{O} = 0.993$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
-
-\begin{figure*}[htbp]
+The plus-polarisation of the GW at the start and end of the evolution is shown in the top row of \figref{good-waveform}.
+\begin{figure*}
 \centering
@@
 
-\subsection{The effect of resonances: Dephasing}
-
-The illustrative EMRI system of the previous section does not encounter a significant resonance and the adiabatic evolution provides a good match to the full evolution.
-We now study a system that does pass through a resonance during its $2$ year evolution.
-Specifically, we choose the initial conditions to be the same as before but with an initial semi-latus rectum $p=7.85$.
-This system passes through the 2:3 resonance, the effect of which is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
 Plotting the plus-polarised GWs at the start and end of the evolution, as before, demonstrates this dephasing, as shown in \figref{dephased-waveform}.
-
-\begin{figure*}[htbp]
+\begin{figure*}
 \centering
@@
 \end{figure*}
-
-To highlight the dephasing more quantitatively, we calculate the shortened overlap between the two models as a function of time $t$, defined as the overlap obtained by including only the part of the waveform within some $\Delta t$ of $t$.
-For our illustrative system, we choose $\Delta t$ such that we can calculate 25 non-overlapping values of this shortened overlap.
-Before the resonance occurs, the adiabatic model provides a good match to the full evolution but the overlap is reduced near to the resonance and never fully recovers afterwards.
-This is shown in \figref{overlap-dephasing}, which is centered on the time at which the full evolution crosses the 2:3 resonance.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
-\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give 25 independent (non-overlapping) calculations.
-The time axis is centred on the value at which the full evolution crosses the 2:3 resonance.}
-\end{figure}
-
-Also shown is the shortened overlap computed between the full evolution and a different adiabatic evolution that is chosen to match the full evolution at the end of the integration.
-In this case, we see similar behaviour as before: the adiabatic waveform has a high overlap in the region where it is constructed to match the full evolution and the boundary of that region is set by the presence of an important resonance.
-
-\subsection{The effect of resonances: Loss of SNR}
-\label{sec:SNRloss}
-
-To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral.
-The dephasing means that an adiabatic evolution with the same initial conditions as a full evolution is likely to have a low overlap if a significant resonance is encountered.
-However, this does not necessarily mean that all adiabatic evolutions have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.
-
-The extremely large parameter space of adiabatic waveforms, coupled with the expensive nature of our 2-torus averaging routine, renders a brute-force approach computationally impossible; it is therefore necessary to restrict the search in some way.
-For this preliminary investigation into the importance of resonances, we do this by focussing on a small subset of parameters that we suspect will produce a large overlap.
-We make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t_{\mathrm{match}}$.
-This reduces the search to a one parameter family of waveforms that can easily be computed concurrently with the full evolution.
-
-The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
-Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau_\mathrm{res}$ after each resonance of interest, namely the low-order 2:3 and 1:2.
-These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral.
-We also include templates that match at the start of the evolution, so that we have a model that should match the portion of the evolution before the first resonance, and at the end of the evolution, as this is often where most SNR is accumulated.
-In addition, we include models that match exactly on each of the resonances as these may achieve partial matching both before and after the resonance and hence perform well.
-
-In our experience, we found that these choices were not sufficient to produce a good family of adiabatic waveforms.
-In particular, if an adiabatic model happened to match the full model close to the edge of the envelope generated by the fast orbital evolution, then the two trajectories eventually diverged and the resulting waveform provided a poor match.
-We therefore adjust the values of $t_{\mathrm{match}}$ so that they correspond to times of apoapsis.
-This ensures that the adiabatic model intersects with the full model close to the centre of the envelope, as demonstrated by the inset plots in \figref{good-traj}.
-As a result, we obtain much higher overlaps in general.
-
-We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.
-
-This one-parameter family of adiabatic models does not explore the extrinsic parameter space nor does it consider different values of the constant intrinsic parameters (the masses of the BHs and the central BH spin).
-It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
-Indeed, systematic biases in parameter estimation are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.
-The result of these restrictions is that the maximum overlap calculated using our family is a strict lower bound on the true maximum overlap.
-
-For the illustrative EMRI system detailed in the previous section, we have computed the family of adiabatic models set by $t_{\mathrm{match}}$ and the resulting trajectories are consistent for the length of the evolution.
-As they cannot be easily distinguished by eye, we plot the difference (in $E$, $L_z$ and $Q$) between the models and the adiabatic evolution that matches at the start, shown in \figref{res-diff-traj}.
-The jumps in the orbital parameters due to the 2:3 resonance can be clearly seen, as can the fast orbital oscillations present in the full evolution but absent in the adiabatics.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=\textwidth]{Fig_res_diff_traj}
-\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start.
-The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start (0.207), at the 2:3 resonance (0.432), after the 2:3 resonance (0.258) and at the end (0.677).}
-\end{figure*}
-
-None of the adiabatic models presented here give a particularly high overlap with the full evolution, because of the effects of the resonance.
-The best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.207$.
-These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.
-
-\subsection{The effect of resonances: Jump sizes}
-
-\emph{Update this to reflect thesis.}
-
-As demonstrated in \figref{res-diff-traj} and computed analytically in \apref{res-asymptotic}, the full evolution undergoes a rapid change in the orbital parameters with respect to the adiabatic evolutions.
-To extract the magnitude of this jump from the trajectory data, we first must account for the fast orbital oscillations as well as ensuring that we include the full effect of the resonance, while minimising any changes resulting from other sources.
-
-To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away.
-For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}_\mathrm{ad}[t_{\mathrm{match}} = t_\mathrm{res}]$ for each of the orbital parameters.
-We then identify the most horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations.
-The resulting data and bounding lines are shown for our illustrative system in \figref{res-jump-calc}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_res_jump_calc}
-\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the 2:3 resonance, scaled by the magnitude of the 2:3 resonance jump.
-The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump.
-The dotted red line indicates the computed size of the jump.
-The time axis is centred on the 2:3 resonance and is scaled by the appropriate resonance timescale.}
-\end{figure}
-
-To estimate the size of the jump $\Delta \mathcal{I}_{\mathrm{jump}}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits.
-This allows us to take into account any linear drift in $\Delta \mathcal{I}$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation.
-% In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down.
-For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.
-
-The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems.
-Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance.
-For the orbital parameter $\mathcal{I}$,
-\begin{equation}
-\label{eq:res-jump-ratio}
-\frac{\Delta \mathcal{I}_\mathrm{jump}}{\mathcal{I}_\mathrm{ad}(t>t_\mathrm{res})-\mathcal{I}_\mathrm{ad}(t<t_\mathrm{res})} = \frac{\Delta \mathcal{I}_\mathrm{jump}}{\dot{\mathcal{I}}_\mathrm{ad}(t_\mathrm{res})\tau_\mathrm{res}}.
-\end{equation}
-
-The resonance jump depends sensitively on the relative phase of radial and poloidal motions.
-It is therefore instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi$, which we do for our illustrative system parameters.
-Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqnref{delta-I-a}.
-
-Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump.
-We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$.
-The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
-The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset; this means that for every value of $q$, there is always a resonance jump in at least one parameter.
-We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
-\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$.
-The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqnref{res-jump-ratio}.
-The individual jumps as well as a sinusoidal fit are plotted for the three orbital shape parameters (solid lines) and the three orbital frequencies (dotted).}
-\end{figure}
-
-For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t_\mathrm{match}$ described in the previous section.
-The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{overlap_vs_q}
-\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$.
-Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
-The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
-\end{figure}
-
-The best performing adiabatic for the majority of our systems matched at the end of the inspiral.
-We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance.
-We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}.
-There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
-
-Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
-The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
-A small fraction of systems have a lower overlap than expected, at around $0.5$.
-This is due to a sub-optimal choice of the matching times; making small adjustments to $t_\mathrm{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
-
-\subsection{Sample EMRI population}
-
-\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix??}
-
-Strong resonances can limit our ability to recover SNR from waveforms using adiabatic templates; they partition the inspiral, splitting up the total SNR into different adiabatic regions, which may be individually undetectable.
-In order to assess the impact of this on future GW missions, we need to analyse the waveforms from a population of detectable EMRIs.
-
-We generate a representative EMRI population by making reasonable assumptions about the distributions of the 14 parameters in the analytic kludge model of Barack and Cutler \cite{Barack2004}.
-The EMRI event rate is dominated by the inspiral of heavier BHs due to the larger observational volume in comparison to that for white dwarfs and neutron stars~\cite{Gair2004}.
-We therefore take the mass of the compact object to be $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs.
-
-We take the central object to be a typical supermassive BH at the centre of a galaxy; we construct an appropriate distribution by breaking the EMRI event rate into two pieces: the `intrinsic' event rate $\mathcal{R}$, i.e.
-the number of inspiral events per unit time for any given galaxy; and the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH.
-We are interested in galaxies that possess a SMBH with a mass in the range $(10^{4} - 10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors.
-In principle, $\mathcal{R}$ will depend on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each component, and the mass and spin of each SMBH.
-All of these properties are highly uncertain, even for our own galaxy.
-
-Despite the difficulties, simple estimates of $\mathcal{R}$ have been carried out using Monte Carlo methods to count the number of stars from isothermal distributions that spiral in to a SMBH without plunging.
-The result is a scaling law for each species of compact object of the form
-\begin{equation}
-\label{eq:EMRI-intrinsic-rate}
-\mathcal{R}(M) = \mathcal{R}_0 \left(\frac{M}{M_0}\right)^\alpha,
-\end{equation}
-where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is some fiducial mass.
-\Citet{Hopman2009a} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, NSs and WDs, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}\, \mathrm{Gyr}^{-1}$ for each component.
-This neglects the effects of resonant relaxation, ignores the spin of the SMBH, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses.
-Each of these is likely to significantly impact the event rate, but \eqnref{EMRI-intrinsic-rate} can still be used as a rough guide to the expected number of events.
-
-\Citet{Amaro-Seoane2011d} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker-Planck description for the bulk properties of the stellar distribution.
-They found a better fit for the power law spectral index for BHs of $\alpha = -0.19$.
-
-The comoving number density of galaxies is challenging to estimate due to local structure in the Universe, the evolution of that structure, and properties of the SMBHs themselves.
-We simplify the problem by assuming a homogeneous distribution that does not evolve with redshift, which is reasonable for the typical scales considered by GW detectors.
-We also neglect correlations between the SMBH mass and spin, and impose a power law scaling relation for the comoving number density
-\begin{equation}
-\diff{n}{\ln M}(M) = n_0 \left(\frac{M}{M_0}\right)^\beta.
-\end{equation}
-There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002 \, \mathrm{Mpc}^{-3}$ for SMBHs with $M < \mathcal{O}(10^7 M_\odot)$ \citep{Greene2007, Gair2010b}.
-
-The mass of the SMBH is then sampled from a power-law with a probability distribution function (pdf) $f(M) \propto M^{\alpha+\beta-1}$.
-We sample the dimensionless spin $a_\ast$ uniformly between its limiting values, $0$ and $1$.
-The direction of the spin of the central BH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere.
-The angles specifying the location of pericenter and the compact object around the orbit are distributed uniformly between $0$ and $2\pi$.
-
-We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\,\mathrm{km s^{-1}\, Mpc^{-1}}$, and sample the redshift of the source uniformly from comoving volume, up to a maximum of $z_\mathrm{max} = 1.5$.
-We don't expect the exact details of the cosmology to significantly alter our results and our choice of $z_\mathrm{max}$ is justified \textit{post hoc}.
-
-The eccentricity distribution is more complicated.
-Eccentricities for EMRIs are uncertain, and depend strongly on the formation scenario being considered.
-Here, we adopt a fit to a distribution computed using Monte Carlo simulations by \citet{Hopman2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH.
-We assume this can be extended to provide a rough estimate of the distribution around BHs of other masses and spins.
-At the point in the inspiral when the orbital period takes a fiducial value $\nu_0 = 10^4\,\mathrm{s}$, we find that the Monte Carlo eccentricity pdf is well-described by a power-law with an exponential cutoff
-\begin{widetext}\begin{equation}
-\label{eq:EMRI-e-distribution}
-f(e;\,e_\mathrm{m},\,e_\mathrm{p},\,b) \propto 
-	\begin{cases}
-		\left(e_\mathrm{m}-e\right)^{b(e_\mathrm{m}-e_\mathrm{p})} \exp\left\{b(e-e_\mathrm{m})\right\} , &\quad 0 \leq e \leq e_\mathrm{m}\\
-		0, &\quad\text{otherwise}
-	\end{cases}
-\end{equation}\end{widetext}
-where $e_\mathrm{m} = 0.81$ is the maximum observed eccentricity, $e_\mathrm{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index.
-
-We are interested in EMRI systems close to plunge, as it is here where the GW amplitude is largest.
-To create our EMRI population, we therefore sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value of $10^4\,\mathrm{s}$, and evolve the system using the AK prescription up until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $R(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.} Each of these systems is then evolved backwards for some time $t_\mathrm{insp}$, chosen uniformly from the range $[0,2]$ years, and the expected gravitational waveforms are calculated using the AK formalism.
-
-The size of the required population can be estimated by evaluating an integral of the form \citep{Gair2009}
-\begin{equation}
-\label{eq:EMRI-number}
-N_\mathrm{EMRI} = \intd{z=0}{z_\mathrm{max}}{ \intd{M=M_\mathrm{min}}{M_\mathrm{max}}{   \mathcal{R}T\diff{n}{\ln M} \diff{V_\mathrm{c}}{z}   }{\ln M}  }{z},
-\end{equation}
-where $V_\mathrm{c}(z)$ is the comoving volume at redshift $z$.
-For a mission lifetime of 2 years, the integral gives a total of $6333$ EMRI events.
-This is a lower bound for $N_\mathrm{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.
-
-A given EMRI is classified as observable if its SNR exceeds some threshold value $\mathrm{SNR}_\mathrm{thres}$.
-Calculating SNRs from the generated AK waveforms, assuming 6 laser links and using the eLISA noise PSD, we find $513$ detectable events, assuming an SNR threshold of $15$.
-The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters, the redshift of the source, and the length of the observation $t_\mathrm{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than 15, which are also shown.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=\textwidth]{EMRIpar_dists}
-\caption{\label{fig:EMRIpar-dists}The parameter distributions at plunge for our detectable EMRI systems (dark blue), alongside those of the undetectable systems (light pink).
-In each case, the y-axis values are the probability of a system being found in a particular bin, given that they are either detectable or undetectable.
-The exception to this is in the final plot for the SNR, where we show the number of systems in each bin.
-The SNRs quoted here are calculated using the AK model.}
-\end{figure*}
-
-We note that eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs, due to the fact that the periapse in such systems can get much closer to the SMBH, and so the compact object experiences the strongest gravitational fields.
-This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in the distributions.
-Systems at higher redshifts start to become undetectable because the amplitude of the signal scales inversely proportional to the luminosity distance to the source, and by $z=1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
-The mass distribution for detectable EMRIs is peaked such that the typical GW frequency occurs at the base of the eLISA noise PSD.
-
-EMRI systems within our population tend to have small eccentricities at plunge due to the circularising effect of GWs: $85\%$ of detectable systems have $e < 0.1$, with a mean value of $0.05$ and a maximum of $0.4$.
-In \secref{flux-enhance}, we found a strong eccentricity dependence on the magnitude of the resonant flux enhancements.
-We therefore expect typical resonant jumps in these astrophysical systems to be much less than $1\%$, and so the resultant dephasing may be relatively weak.
-We now analyse these $513$ systems using our full and adiabatic self-force models.
-
-\subsection{Loss of signal-to-noise ratio}
-\label{sec:population}
-
-After computing the SNRs of our NK waveforms for each of the $513$ systems, we can compare the values to those obtained using the AK formalism.
-The different approximation schemes are expected to produce different results, but they should be broadly similar.
-In \figref{pop-SNR-vs-eta}, we plot the ratio of the new SNR to the AK SNR, as a function of the mass ratio.
-We observe a clear trend, with systems closest to equal mass performing roughly as expected, but with the most extreme systems giving significantly lower SNRs.
-
-This is a consequence of the PN self-force model we are using, which we have found overestimates the inspiral rate for EMRIs.
-Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the 2 year observation window, they evolve within the bucket of the eLISA noise curve.
-Increasing the rate of evolution then does not significantly change the recovered SNR since we are still sensitive to all parts of the waveform.
-At smaller mass ratios, the systems evolve more slowly, from lower plunge frequencies.
-An increased evolution rate can shift the EMRI to outside the sensitive region of the detector, thus giving a much lower SNR.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{pop_SNR_vs_eta}
-\caption{\label{fig:pop-SNR-vs-eta}The ratio of SNRs as computed with waveforms generated using the NK (adopting our PN self-force model) and AK formalisms, as a function of (the base-10 logarithm of) the mass ratio.
-Each system has an AK SNR greater than 15, using the eLISA detector with 6 laser links.}
-\end{figure}
-
-Nevertheless, we may still study the effect of resonances, by comparing the maximum adiabatic SNR to the full SNR computed using the NK waveforms.
-The resulting overlap is still indicative of the effect of resonances on the population of systems.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_Tad}
-\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family.
-Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t_\mathrm{insp}$.
-The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t_\mathrm{ad}$, normalised by $t_\mathrm{insp}$.
-The solid line corresponds to the expected value if the 1:2 and 2:3 resonances cause significant waveform dephasing.}
-\end{figure}
-
-We have calculated the longest period of time, $t_\mathrm{ad}$, during each inspiral, in which neither the 1:2 nor the 2:3 resonance occurs.
-From the results of \secref{effres-jumps}, we expect the recovered overlap to be approximately given by the proportion of time spent in a resonance-free region, that is $t_\mathrm{ad} / t_\mathrm{insp}$.
-In \figref{pop-adSNR-vs-Tad}, we plot the maximum overlap between the instantaneous and adiabatic models against this ratio, highlighting the number of resonances, $N_\mathrm{res}$, each system encounters.
-In the majority of cases, there is no significant impact on the recovered overlap using adiabatic models.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_eta}
-\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t_\mathrm{ad} / t_\mathrm{insp}$, as a function of (the base-10 logarithm of) the mass ratio.}
-\end{figure}
-
-We can better understand the overlap distribution from \figref{pop-adSNR-vs-Tad} by considering the behaviour with the mass ratio.
-We do this in \figref{pop-adSNR-vs-eta}, plotting the difference between the computed maximum overlap and the value expected should resonances be important.
-A small proportion of systems have overlaps below the expected value (approximately $4\%$ have values less than $-0.05$).
-This might be caused by higher-order resonances disrupting the evolution, in which case $t_\mathrm{ad}$ should be smaller and the systems would lie on the expectation line.
-Alternatively, the low-order resonances might be long-lived, explaining why the recovered overlap is smaller than expected.
-However, a more likely explanation is that larger adiabatic overlaps are possible with models that we have not considered here (either with different values of $t_\mathrm{match}$ or outside our one-parameter family altogether).
-
-Roughly $30\%$ of the systems lie within $0.05$ of the expected value.
-However, the vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity.
-For the smallest mass ratios, the inspiral rate is slow, and so the systems do not encounter either the 1:2 or 2:3 resonances during their lifetime.
-Meanwhile for the largest mass ratios, the EMRIs encounter both resonances very close to plunge.
-In each case, there is a long resonance-free region, allowing a high overlap to be recovered.
-
-The remaining $66\%$ of systems have overlaps above the level expected if resonances were to play an important role.
-These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t_\mathrm{ad}$ is small.
-For these EMRIs, resonances are not as important as expected, which is likely due to the low eccentricities of our population leading to small (and in most cases, negligible) resonant flux enhancements.
-
-Even if we assume that all overlap reductions are due to resonances (instead of due to our adiabatic family not being extensive enough) the overall effect on the population is not significant.
-To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to take into account the resonance behaviour.
-The total number of detectable systems using a threshold SNR of $15$ decreases from $513$ to $492$.
-If we increase the threshold, the fractional reduction in the number of detectable systems gets even smaller.
-We therefore conclude that resonances do not cause sufficient waveform dephasing across a population of EMRIs for the detection rate to be appreciably diminished.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{pop_SNR_dist}
-\caption{\label{fig:pop-SNR-dist}Histogram showing the probability distribution function for the detectable EMRI SNRs, as calculated using the AK formalism (light pink) and modified by the maximum adiabatic overlap (dark blue).}
-\end{figure}
-
-\section{Concluding remarks}
-\label{sec:conclusion}
-
-\emph{Need to reconclude}
-
-Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral.
-We have presented an evolution scheme appropriate for studying the effects of resonances on inspirals, which has been applied to an astrophysically realistic population.
-The location and duration of resonances, as well as their effect on orbital parameters, can all be calculated.
-
-Amongst a population of sources, unmodelled resonances could diminish detection prospects.
-However, due to the low eccentricity of detectable sources, the overall effect on SNR recovery is small; it is therefore likely to be sufficient to use adiabatic models to detect EMRIs.
-An unresolved question here is how using these adiabatic models would affect parameter estimation: systematic biases may be introduced if we neglect the presence of important resonances.
-One method of providing an answer to this problem is to perform an MCMC parameter estimation study using our adiabatic waveforms as templates; we leave this to future work.
-
-High eccentricity EMRIs may be possible, in which case transient resonances will play an important role in their detectability; systems that undergo a jump of a few percent dephase very rapidly from the adiabatic approximation.
-In such scenarios, adiabatic models can still be used, but there must be sufficient SNR in each non-resonant region that they can be detected as individual sources and then identified as arising from the same inspiral.
-If this is not possible, \eqnref{delta-I-a} can be used to join together different adiabatic models, if the phase on resonance and the magnitude of the self-force can be predicted sufficiently accurately.
-Alternatively, the jump $\Delta\mathcal{I}_\mathrm{jump}$ for each orbital parameter could be included as a free parameter in the template space, and then searched over.
-Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.
-
-This work is highly dependent on the chosen self-force model, and so the results should be taken as qualitative estimates, given the unsuitability of a PN self-force to the problem of fast motion near to highly-spinning BHs.
-In the future, with a more accurate self-force model, more exact quantitative results can be derived.
-
-\begin{acknowledgments}
-The authors extend our sincere gratitude to \'{E}anna Flanagan and Tanja Hinderer for providing their PN self-force code, without which this work would not be possible.
-We also wish to thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent and Leor Barack for useful conversations.
-RHC is supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC was supported by a Marie Curie Fellowship, and JRG is supported by the Royal Society.
-\end{acknowledgments}
-
-\appendix
-
-\section{Location of resonances}\label{sec:location}
-
-We can find the location of resonances by numerically solving $\Omega = n_r \Omega_r - n_\theta \Omega_\theta = 0$.
-\Figref{res-plane-2-5-95} shows the semilatus rectum, eccentricity and (cosine of the) inclination angle of the $\nu = 2/5$ resonance surface for a BH of spin $a_\ast = 0.95$.
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_res-2-5-95-plane}
-\caption{\label{fig:res-plane-2-5-95}Location of the $2/5$ resonance surface for an $a_\ast = 0.95$ BH in terms of orbital semilatus rectum $p$, eccentricity $e$ and inclination $\iota$.}
-\end{figure}
-This is almost planar, inspiring us to look for a simple description that can help guide our search for resonance locations.
-Brink, Geyer and Hinderer~\cite{Brink2013} provide series expansions for the location of resonances in the limit of equatorial orbits for small spin and eccentricity.
-We do not follow this approach of trying to find analytic expressions for the resonance surface.
-The expressions become complicated when venturing away from limiting cases.
-Instead, we build an approximate phenomenological model and fit this to the resonance plane.
-%\footnote{A stopped clock is precisely correct twice a day, while a clock that is five minutes slow is never right but will often be close enough.
-Within this metaphor, our approximation would be a clock that varies between telling the correct time and being fifteen minutes off throughout the day: it is generally going to be good enough to give you a sense of the time but if you have an important event you should check with something better.}
-This should be useful for designating the region in which resonance could be expected.
-To locate them precisely, it is necessary to solve $\Omega = 0$ numerically; the approximate model gives a suitable starting point.
-
-The resonant semilatus rectum for any particular spin and resonance ratio can be well approximated as
-\begin{equation}
-p(e,\iota;a_\ast,\nu) \simeq A\frac{1 + B e + D \cos\iota}{1 - C\exp(e)}.
-\end{equation}
-The coefficients $\{A,B,C,D\}$ depend upon the spin and the particular resonance; they can be approximated as
-\begin{align} 
-A(a_\ast,\nu) \simeq {} & a_0\frac{1 + a_1\nu - a_2 \nu^2 - a_3 \nu a_\ast^2}{1 + a_4\nu - (1 + a_4)\nu^2}, \\
-B(a_\ast,\nu) \simeq {} & b_0(1 - b_1\nu)\exp(-b_2\nu)(1 - b_3 a_\ast), \\
-C(a_\ast,\nu) \simeq {} & c_0, \\
-D(a_\ast,\nu) \simeq {} & d_0\left[1 - \exp(a_\ast)\right]\left[1 - d_1\exp(\nu)\right].
-\end{align}
-This gives us a total of $12$ parameters for our fit.
-Whilst this may sound large, if we were fitting an expansion to quadratic order in any combination of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
-\begin{equation}
-\begin{array}{lll}
-a_0 = 5.9854, & a_1 = 3.4116, & a_2 = 0.9253,\\
-a_3 = 0.1959, & a_4 = 4.8846, & b_0 = 0.7692,\\
-b_1 = 1.4752, & b_2 = 1.4861, & b_3 = 0.5974,\\
-c_0 = 0.02332, & d_0 = 0.7968, & d_1 = 0.3115.
-\end{array}
-\end{equation}
-These were fitted for all possible resonances with $n_r = 2$--$7$ as well as the $9:10$, $19:20$, $49:50$ and $99:100$ resonances; with MBH spins of $a_\ast = 0.01$--$0.999$; for orbits with eccentricities $e = 0.01$--$0.99$, and inclinations $\cos\iota = -0.999999$--$0.999999$.
-%These were fitted for resonances with $\nu = 1/7, 1/6, 1/5, 1/4, 2/7, 1/3, 2/5, 3/7, 1/2, 4/7, 3/5, 2/3, 5/7, 3/4, 4/5, 5/6, 6/7, 9/10, 19/20, 49/50, 99/100$, with BH spins of $a_\ast = 0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.93, 0.95, 0.97, 0.99, 0.999$.
-
-Using this approximation, the maximum error in $p$ for a given $a_\ast$ and $\nu$ is typically $\sim10\%$ and less that $1$ in absolute terms.
-The relative error in the semilatus rectum is illustrated in \figref{p-error}.
-\begin{figure*}[htp]
-\centering
-\centerline{\includegraphics[width=0.47\textwidth]{Fig_fit-error-max-plane}\quad\includegraphics[width=0.47\textwidth]{Fig_fit-error-RMS-plane}}
-\caption{\label{fig:p-error}Relative error in the approximate semilatus rectum compared to the accurate numerical result as a function of BH spin $a_\ast$ and resonance ratio $\nu$.
-The left panel shows the maximum relative error and the right shows the root-mean-square error; in both cases we are marginalising over eccentricity and inclination.}
-\end{figure*}
-The largest fractional error is $\sim50\%$, this is for $a\rightarrow 1$ and $\nu \rightarrow 0$, and corresponds to small $p$, such that the absolute error is still small.
-Taking the root-mean-square across $e$ and $\iota$, the fractional error for a given $a_\ast$ and $\nu$ never exceeds $9\%$ and is typically less than $4\%$.
-
-\section{Asymptotic solution for passage through resonance}\label{sec:res-asymptotic}
-
-\input{./asymptotic.tex}
-
-\bibliography{Resonances}
-
-\end{document}

### 2016-08-03 res_char.tex
Up to jumps

@@
 \newcommand{\eqnref}[1]{(\ref{eq:#1})}
-\newcommand{\figref}[1]{Fig.\ \ref{fig:#1}}
-\newcommand{\Figref}[1]{Fig.\ \ref{fig:#1}}
-\newcommand{\secref}[1]{Sec.\ \ref{sec:#1}}
-\newcommand{\apref}[1]{Appendix \ref{sec:#1}}
-\newcommand{\Tabref}[1]{Table \ref{tab:#1}}
+\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
+\newcommand{\Figref}[1]{Fig.~\ref{fig:#1}}
+\newcommand{\secref}[1]{Sec.~\ref{sec:#1}}
+\newcommand{\apref}[1]{Appendix~\ref{sec:#1}}
+\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}
 
@@
 \newcommand{\linepartialdiff}[2]{\ensuremath{\partial {#1}/\partial {#2}}}
-\newcommand{\partialdiffop}[1]{\ensuremath{\dfrac{\partial}{\partial {#1}}}}\newcommand{\recip}[1]{\ensuremath{\frac{1}{#1}}}
+\newcommand{\partialdiffop}[1]{\ensuremath{\dfrac{\partial}{\partial {#1}}}}
+\newcommand{\recip}[1]{\ensuremath{\frac{1}{#1}}}
 \newcommand{\grad}{\ensuremath{\boldsymbol{\nabla}}}
@@
 \author{Robert H. Cole}
-%\email[]{rhc26@ast.cam.ac.uk}
 \affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
@@
 \author{Priscilla Ca\~{n}izares}
-%\email[]{pcm@ast.cam.ac.uk}
+\affiliation{Institute of Mathematics, Astrophysics and Particle Physics, Radboud University, Heyendaalseweg 135, 6525 AJ Nijmegen, The Netherlands}
 \author{Jonathan R. Gair}
-%\email[]{jgair@ast.cam.ac.uk}
-\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
+\affiliation{School of Mathematics, University of Edinburgh, Peter Guthrie Tait Road, Edinburgh EH9 3FD, UK}
 
@@
 In order to detect such signals and extract the desired information, it is necessary to possess a bank of accurate waveform templates.
-For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first order dissipative gravitational self-force.
-Other effects are however neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order dissipative gravitational self-force.
+Other effects are, however, neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
 During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the emitted radiation.
@@
 Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}; it is now possible to simulate hundreds of orbits~\cite{Szilagyi2015}.
-These simulations allow us to understand BH--BH mergers.
-Stellar-mass BH mergers are targets for ground-based GW detectors, such as the upcoming Advanced LIGO~\cite{Aasi2015} and Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Kuroda2010}.
-Supermassive BH (SMBH) mergers, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
+These simulations allow us to understand binary BH coalescences.
+Stellar-mass BH mergers are targets for ground-based GW detectors, such as Advanced LIGO~\cite{Aasi2015} and Advanced Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Aso2013}.
+The first direct observations of GWs came from the coalescences of two stellar-mass BHs~\cite{Abbott2016,Abbott2016e,Abbott2016d}, and analysis of their properties relied upon our understanding of binary BH waveforms~\cite{Abbott2016f,Abbott2016h,Abbott2016d}.
+%Supermassive BH (SMBH) coalescences, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
+
 Systems of unequal masses are more challenging to evolve numerically as they complete a larger number of orbits and it is necessary to resolve two different scales.
 Calculations can instead be performed perturbatively.
-The paradigm unequal-mass system has a stellar-mass BH orbiting an SMBH, such as those expected to be found at the centre of most galaxies.
-These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors~\cite{Amaro-Seoane2007}.
-Despite being well-studied, there still remain open questions.
+The paradigm unequal-mass system has a stellar-mass BH orbiting a supermassive BH (SMBH), such as those expected to be found at the centres of galaxies.
+These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors like the evolving Laser Interferometer Space Antenna (eLISA)~\cite{Amaro-Seoane2007,Amaro-Seoane2012a}.
+Despite being well-studied, there still remain open questions in modelling generic binaries.
 
@@
 % Section 10.6
-In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field, $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
-For extreme-mass-ratio systems, the evolution time-scale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital time-scales.
+In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
+For extreme-mass-ratio systems, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
 The evolution of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
@@
 In \secref{problem}, we formulate the specific problem: that of geodesic motion in Kerr spacetime, perturbed by the gravitational self-force.
-We then study generic properties of transient resonances in \secref{properties}, detailing their location in parameter space, the time-scales over which they affect the motion and the resulting GW flux enhancements.
+We then study generic properties of transient resonances in \secref{properties}, detailing their location in parameter space, the timescales over which they affect the motion and the resulting GW flux enhancements.
 Specific examples are considered to illustrate the effects of resonances in \secref{waveforms}, before finally turning to an astrophysical population in \secref{astrophysics}.
@@
 We also use the dimensionless spin $a_\ast \equiv a/M$; we take the convention that $0 \leq a_\ast < 1$.
-We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega_m = 0.3$ and $H_0 = 70\units{km\, s^{-1}\, Mpc^{-1}}$ and do not expect the exact details of the cosmology to significantly alter our results.
+We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega\sub{m} = 0.3$ and $H_0 = 70\units{km\, s^{-1}\, Mpc^{-1}}$ and do not expect the exact details of the cosmology to significantly alter our results.
 
@@
 The evolution of an extreme-mass-ratio ($\eta \ll 1$) system is slow.
-Instantaneously, the motion of the orbiting mass can be described as geodesic, with the integrals of the motion changing on time-scales of many orbital periods.
+Instantaneously, the motion of the orbiting mass can be described as geodesic, with the integrals of the motion changing on timescales of many orbital periods.
 It is therefore necessary to develop an understanding of the Kerr geodesics (\secref{geodesic}).
@@
 
+Central to understanding transient resonances in a knowledge of orbits in Kerr spacetime, and hence we begin with details of evolving Kerr geodesics.
+Those familiar with calculating orbits in Kerr may skip this section.
 The geodesic equations may be written as~\cite{Carter1968, Chandrasekhar1992} % section 62
@@
 
-One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short time-scales, for small $\eta$ there is a small disparity.
+One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short timescales, for small $\eta$ there is a small disparity.
 The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e.
@@
 
-%The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e., the case for motion in Kerr) is subject to a small perturbation the form of the orbits are preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
-% Chapter II, section 3.3 d)
-%When we average the self-force and follow the adiabatic evolution of the trajectory we are exploiting this behaviour.
-However, the KAM theory only applies for sufficiently incommensurate orbits: close to resonance it does not apply~\cite{Moser1973}.
-% Chapter V section 1.
-c)
-% In these regions the adiabatic approximation breaks down and we need to follow the instantaneous evolution.
-
 \subsection{Osculating elements and forced motion}
@@
 
-For generic EMRIs, there are two characteristic timescales: the \emph{fast} orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the \emph{slow} inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$.
+For generic EMRIs, there are two characteristic timescales: the fast orbital motion, related to the fundamental frequencies $\sim1/\Omega$, and the slow inspiral, related to the change in fundamental frequencies $\sim\Omega/\dot{\Omega}$, where an overdot denotes a derivative with respect to coordinate time $t$.
 These, along with the resonance timescale, are discussed more in \secref{res-time}.
-The two-timescale nature of the problem makes it ideally suited to the method of osculating elements~\cite{Gair2011a}: on short timescales, we analyse the unperturbed system resulting in geodesic motion, and then the long-term evolution is described by a sequence of instantaneous geodesics.
+The two-timescale nature of the problem makes it ideally suited to the method of osculating elements~\cite{Pound2008,Gair2011a}: on short timescales, we analyse the unperturbed system resulting in geodesic motion, and then the long-term evolution is described by a sequence of instantaneous geodesics.
 
@@
 Collectively, these are referred to as \emph{osculating elements} and we denote them by $I^A(t)$, making explicit the variation with time.
-For a sequence of geodesics of a background spacetime, where the evolution is forced by some external acceleration, we can calculate the evolution of the osculating elements $\dot{I}^A$.
+For a sequence of geodesics of a background spacetime, where the evolution is forced by some external acceleration (in our case from the self-force), we can calculate the evolution of the osculating elements $\dot{I}^A$.
 The specific equations for motion in Kerr are derived by Gair {\it{et al}}.~\cite{Gair2011a}.
@@
 
-\subsection{Self-force model}
+\subsection{Gravitational self-force model}
 \label{sec:self-force}
@@
 The procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform.
-However, the adiabatic derivatives vary on much longer timescales than the orbital motion (see \secref{res-time}), and so in practice, we can interpolate.} %We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\delta t/\eta$ where $\delta t$ is the orbital time-step.
+However, the adiabatic derivatives vary on much longer timescales than the orbital motion (see \secref{res-time}), and so in practice, we can interpolate.}
+%We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\delta t/\eta$ where $\delta t$ is the orbital time-step.
 When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives.
@@
 
-%In our parameterisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform appropriate averages over this new $2$-torus.
+%In our parametrisation of the EMRI system, using $\{\psi,\chi\}$ rather than $\{q_r,q_\theta\}$, we must perform appropriate averages over this new $2$-torus.
 The osculating elements integration scheme involves computing derivatives with respect to coordinate time, and we shall often be interested in the averaged form of such derivatives.
@@
 %The effect of resonances must be completely contained within the pieces of the self-force left after subtraction of the $2$-torus-averaged pieces, $\hat{X}\equiv X-\langle X\rangle_{q_r,\,q_\theta}$.
-The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurence of any transient resonances.
+The adiabatic evolution, which contains only a single $2$-torus-averaged forcing term, will therefore not be affected by the occurrence of any transient resonances.
 Resonances still occur as the fundamental frequencies become commensurate, but as the system passes through each resonance, there is no additional effect on the evolution of the system.
@@
 
-\section{Properties of Transient Resonances}
+\section{Properties of transient resonances}
 \label{sec:properties}
@@
 Having located where in an inspiral we can expect to encounter a transient resonance, we must now consider its impact.
-In \secref{res-time} we determine the characteristic time-scales describing resonance, and in \secref{flux-enhance} we look at the impact of passing through a resonance on the evolution of the orbit, including some numerical results.
-
-\subsection{Time-scales}\label{sec:res-time}
-
-When analysing resonances it is useful to refer to a number of characteristic time-scales.
+In \secref{res-time} we determine the characteristic timescales describing resonance, and in \secref{flux-enhance} we calculate the impact of passing through a resonance on the evolution of the orbit.
+
+\subsection{timescales}\label{sec:res-time}
+
+When analysing resonances it is useful to refer to a number of characteristic timescales.
 We always use coordinate time $t$ for these, as this corresponds to what is measured by an observer at infinity.
 Translation to Mino time can be done with an appropriate factor of $\Gamma$.
-We use the orbital period $T$, the evolution time-scale $\tau\sub{ev}$, the precession time-scale $\tau\sub{pres}$ and the resonance time-scale $\tau\sub{res}$.
-
-The simplest time-scales are the orbital periods $T_r = 2\pi/\Omega_r$, $T_\theta = 2\pi/\Omega_\theta$ and $T_\phi = 2\pi/\Omega_\phi$.
+We use the orbital period $T$, the evolution timescale $\tau\sub{ev}$, the precession timescale $\tau\sub{pres}$ and the resonance timescale $\tau\sub{res}$.
+
+The simplest timescales are the orbital periods $T_r = 2\pi/\Omega_r$, $T_\theta = 2\pi/\Omega_\theta$ and $T_\phi = 2\pi/\Omega_\phi$.
 These are the shortest in our set.
-We use $T$ to denote a time-scale of the same order as the orbital periods.
-
-We define the evolution time-scale as
+We use $T$ to denote a timescale of the same order as the orbital periods.
+
+We define the evolution timescale as
 \begin{equation}
@@
 In general, away from resonance, we take $\nu \equiv \Omega_r/\Omega_\theta$.
-This time-scale sets the period over which there is a significant change in the frequencies.
-It acts as an inspiral time-scale.
+This timescale sets the period over which there is a significant change in the frequencies.
+It acts as an inspiral timescale.
 It is long in all cases we study, $\tau\sub{ev} \sim \order{T/\eta}$.
-It is this property which makes extreme-mass-ratio inspirals interesting, as we can follow the waveform for many cycles, accruing high signal-to-noise ratios.
+It is this property which makes EMRIs interesting, as we can follow the waveform for many cycles, accruing high signal-to-noise ratios.
 This is also what allows us to use the adiabatic prescription, as it means the trajectory moves slowly through different orbital parameters.
 
-We use the precession time-scale
+We use the precession timescale
 \begin{equation}
@@
 with $\Omega(t) = n_r \Omega_r(t) - n_\theta \Omega_\theta(t)$, where the frequencies are calculated instantaneously and the integers are for the resonance of interest.
-This time-scale becomes infinite exactly on resonance, but decreases as we get further from resonance, eventually becoming $\order{T}$.
+This timescale becomes infinite exactly on resonance, but decreases as we get further from resonance, eventually becoming $\order{T}$.
 It measures the relative precession rate of the radial and polar motions and hence gives an indication of how long it takes to fill the entire $\psi$--$\chi$ $2$-torus.
 
-We also use the resonance time-scale
+We also use the resonance timescale
 \begin{equation}
@@
 \end{equation}
-The resonant time-scale should give an indication of the time over which we expect the effects of the resonance to be felt~\cite{Bosley1992}.
+The resonant timescale should give an indication of the time over which we expect the effects of the resonance to be felt~\cite{Bosley1992}.
 Consider the phase of the Mino-time Fourier expansion on resonance; neglecting the constant, the resonant Fourier component has form
@@
 \end{equation}
-The two time-scales are equivalent: we preferentially use $\tau\sub{res}$ to denote the resonance width.
-It is shorter than the inspiral time-scale, but longer than an orbital period, $\tau\sub{res} \sim \order{\eta^{1/2}\tau\sub{ev}} \sim \order{\eta^{-1/2}T}$~\cite{Flanagan2012,Gair2011a}; it therefore acts as a bridge between the two time-scales~\cite{Hinderer2008}.
-
-Since we shall be considering Fourier decompositions, in anticipation of future results, we also define a time-scale for the $s$-th resonant frequency harmonic
+The two timescales are equivalent: we preferentially use $\tau\sub{res}$ to denote the resonance width.
+It is shorter than the inspiral timescale, but longer than an orbital period, $\tau\sub{res} \sim \order{\eta^{1/2}\tau\sub{ev}} \sim \order{\eta^{-1/2}T}$~\cite{Flanagan2012,Gair2011a}; it therefore acts as a bridge between the two timescales~\cite{Hinderer2008}.
+
+Since we shall be considering Fourier decompositions, in anticipation of future results, we also define a timescale for the $s$-th resonant frequency harmonic
 \begin{align}
@@
 
-The expression for the change can be understood using our expression for the resonance width, substituting in \eqnref{T-res-s},
+To explain the form of this expression we substitute in our expression for the resonance width from \eqnref{T-res-s},
 \begin{equation}
@@
 \end{equation}
-Schematically, this can be understood as the magnitude of the forcing function on resonance $\sim \eta F_{a,\,s}$ multiplied by the time on resonance $\sim \tau_{\mathrm{res},\,s}$ and a function that varies with the phase $\widehat{\kappa}_0$.
+Schematically, this can then be understood as the magnitude of the forcing function on resonance $\sim \eta F_{a,\,s}$ multiplied by the time on resonance $\sim \tau_{\mathrm{res},\,s}$ and a function that varies with the phase $\widehat{\kappa}_0$.
 Averaging over all values of $\widehat{\kappa}_0$ is equivalent to averaging over all values of $\psi_{\theta_-}$, and has the same effect as averaging over the $\psi$--$\chi$ $2$-torus; this gives an average discrepancy relative to the adiabatic evolution of
@@
 
+Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on EMRI waveforms (\secref{waveforms}) and the population of observable inspirals (\secref{astrophysics}).
+
+\section{The impact of resonances on EMRIs}
+\label{sec:waveforms}
+
+Having built an understanding of the properties of transient resonances, we now consider their impact on GW signals.
+In \secref{kludge} we discuss our chosen waveform generation scheme, giving a demonstration of its accuracy for evolutions which avoid low-order resonances in \secref{nonres}.
+In \secref{effres-phase} we detail the impact of resonances on the match between waveforms computed from adiabatic and fully instantaneous evolutions, and in \secref{effres-jump} we look at the changes in orbital parameters across resonance.
+
+\subsection{Waveform model and analysis}
+\label{sec:kludge}
+
+One of the targets of pre-eLISA research is to generate a bank of waveform templates $\{h(t;\boldsymbol{\Theta}_i)\}$ across a range of parameter space $\{\boldsymbol{\Theta}_i\}$.
+These can be compared to data to search for the presence of GWs.
+Templates must accurately reproduce what we expect to observe in nature, without being too computationally expensive.
+Ideally, we would like to use waveform templates from EMRI systems that are evolved under the instantaneous self-force model, but these are computationally challenging.
+The alternative is to use cheaper adiabatic waveforms, but these do not include the effect of resonances.
+To assess the impact of this choice, we compare data $s(t;\bar{\boldsymbol{\Theta}})$ generated using the full self-force model (\secref{forced-motion}) to templates $h(t;\boldsymbol{\Theta})$ generated using our $2$-torus averaged self-force model (\secref{adiabatic}).
+
+To generate gravitational waveforms, we employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}, augmented to include evolution of the positional elements.
+We first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
+This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; agreement between Teukolsky-based and the best NK waveforms are typically $95\%$ or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
+
+Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH with evolution driven by the dissipative part of self-force either calculated instantaneously or following the adiabatic prescription.
+We then map the Boyer--Lindquist coordinates to spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique, the standard quadrupole formula~\cite{Misner1973}.
+
+We expect these NK waveforms to be sufficiently accurate for our purposes, given the approximate (post-Newtonian, low-spin) self-force model (\secref{self-force}).
+Results can be straight-forwardly refined as developments are made in computing more comprehensive self-force models.
+
+The similarity of two waveforms, $s(t)$ and $h(t)$, can be evaluated using the noise-weighted inner product~\cite{Finn1992}
+\begin{equation}
+\label{eq:innerprod}
+\overlap{s}{h} = 2 \intd{0}{\infty}{\frac{\tilde{s}(f)\tilde{h}^*(f)+\tilde{s}^*(f)\tilde{h}(f)}{S_n(f)}}{f},
+\end{equation}
+where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD)~\cite{Moore2014a}.\footnote{We use cubic interpolation to construct NK waveforms with identical time sampling.
+%In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
+To take this into account, we project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme and set any values outside the domain of $\{t_j\}$ to zero.
+%This whole procedure gives more accurate results than interpolating in the Fourier domain since the waveforms are smoother in the time domain.
+A Planck-taper window function~\cite{McKechan2010} is applied to reduce unwanted spectral leakage in the Fourier transforms. 
+%After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
+} We use the analytic approximation for eLISA of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}.
+Following the success for the LISA Pathfinder mission~\cite{Armano2016}, this sensitivity should be achievable in a future mission.
+
+We wish to test whether there exists some set of parameters $\boldsymbol{\Theta}$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
+We do this by evaluating the signal-to-noise ratio (SNR) for each (normalised) waveform template,
+\begin{equation}
+\rho\left[h\right] = \max_t \frac{\overlap{s}{h}}{\sqrt{\overlap{h}{h}}}.
+\end{equation}
+We maximise over the time offset for the template to find the best fit to the data. If a template exactly matches the data, it would produce an SNR of $\sqrt{\overlap{s}{s}}$, hence the overlap
+\begin{equation}
+\label{eq:overlap}
+\mathbb{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}},
+\end{equation}
+which ranges from $0$ to $1$ provides a convenient indication of how well-matched the template is to the data.
+
+\subsection{The non-resonant case}
+\label{sec:nonres}
+
+Before investigating the impact of resonances on an EMRI signal, we first compare results from the full instantaneous evolution and the $2$-torus averaged adiabatic evolution over $2~\mathrm{yr}$ for an inspiral which avoids any significant resonances.
+Example evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, which shows that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
+The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the $2$-torus averaging explicitly smooths out the visible structure on this scale.
+The two appraches are in good agreement.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{Fig_good_traj}
+\caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) under the full (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances.
+The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.
+This system has $\mu = 10 M_{\odot}$, $q = 3\times 10^{-6}$, $a_\ast=0.95$, initial semilatus rectum $p_0 = 7.5$, initial eccentricity $e_0 = 0.7$, initial inclination $\cos \iota_0 = 0.5$ and redshift $z=0.204$.}
+\end{figure*}
+
+Using the two trajectories, we can calculate the corresponding NK waveforms.
+The two waveforms exhibit good agreement in both amplitude and phase across the entire duration.
+We find an overlap $\mathbb{O} = 0.993$, illustrating that adiabatic models can safely be used when resonances are not encountered.
+
+\subsection{The effect of resonances: Dephasing and overlap}
+\label{sec:effres-phase}
+
+We now study a system that does pass through a resonance during its $2~\mathrm{yr}$ evolution.
+Specifically, we choose the initial conditions to be the same as in \secref{nonres}, but with an initial semilatus rectum $p_0 = 7.85$.
+This system passes through the $2$:$3$ resonance.
+The effect of passing through resonance is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
+
+To illustrate the dephasing, we calculate as a function of time $t$ the shortened overlap between the two models, defined as the overlap obtained by including only the part of the waveform within $\Delta t$ of $t$.
+We choose $\Delta t$ such that we can calculate $25$ non-overlapping shortened overlaps.
+Before the resonance occurs, the adiabatic model provides a good match to the full evolution, but the overlap is reduced near to the resonance and never fully recovers afterwards.
+This is shown in \figref{overlap-dephasing}, which is centred on the time at which the full evolution crosses the $2$:$3$ resonance.
+Also shown is the shortened overlap computed between the full evolution and a different adiabatic evolution that is chosen to match the full evolution at the end of the integration.
+In this case, we see similar behaviour: the adiabatic waveform has a high overlap where it is constructed to match the full evolution, but this is disrupted by the resonance.
+Passing through resonance can adversely affect the overlap of adiabatic templates.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
+\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
+The time axis is centred on the value at which the full evolution crosses the $2$:$3$ resonance.}
+\end{figure}
+
+To be able to detect signals, we must have templates which match the signals.
+We have seen that overlaps between adiabatic and full instantaneous evolutions dephase following a resonance.
+However, this does not necessarily mean that no adiabatic evolution will have a low overlap with the observed signal.
+It is possible that a difference between the instantaneous and adiabatic waveforms could be ameliorated by changing the parameters of the template $\boldsymbol{\Theta}$.
+In this case, the waveform mismatch would not limit detectability of the signals, but would lead to errors in parameter estimation, a stealth bias caused by incorrect waveforms~\cite{Cutler2007}.
+We leave an investigation into the impact of transient resonances on parameter estimation to future work.\footnote{If resonances were successfully included in the waveforms used for parameter estimation, the sharp nature of jumps may help construct precise inferences of the source parameters~\cite{Mandel2014}.} However, we consider the possibility of obtaining better waveform matches by varying the parameters of the EMRI.
+
+The large parameter space of adiabatic waveforms, coupled with the expensive nature of our $2$-torus averaging routine, renders a brute-force approach prohibitively expensive.
+For this preliminary investigation, we focus on a small subset of parameters that we suspect will produce a large overlap, and make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t_{\mathrm{match}}$.
+This reduces the search to a $1$-parameter family of waveforms that can easily be computed concurrently with the full evolution.
+
+%This $1$-parameter family of adiabatic models does not explore the extrinsic parameter space of different orientations, nor does it consider different values of the constant intrinsic parameters (the masses of the BHs and the SMBH spin).
+It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
+Exploring this parameter space would also reveal degeneracies between this parameters.
+The result of these restrictions is that the maximum overlap calculated using our family is a strict lower bound on the true maximum overlap.
+
+The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
+We choose $5\tau\sub{res}$ after each resonance of interest, namely the low-order $1$:$2$ and $2$:$3$ resonances.\footnote{We adjust the values of $t_{\mathrm{match}}$ so that they correspond to times of apoapsis.
+This ensures that the adiabatic model intersects with the full instantaneous model close to the centre of it oscillatory envelope, as demonstrated by the inset plots in \figref{good-traj}.
+The generally obtains better matches than if we match close to the extrema of the envelope, which are less representative of the average behaviour.} These matching times lie in a portion of the evolution that is not affected by a resonance, and so should allow for a large overlap with the adiabatic model for that region of the inspiral.
+For comparison, we also consider templates that match at the start and end of the evolution, and that match exactly on each of the resonances.
+
+%We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.
+
+We have computed this family of adiabatic evolutions for our illustrative resonance.
+In \figref{res-diff-traj}, we plot the difference in the orbital parameters ($E$, $L_z$ and $Q$) between the various trajectories and the adiabatic evolution that matches at the start.
+The jumps in the orbital parameters due to the $2$:$3$ resonance can be clearly seen, as can the fast orbital oscillations present in the full instanteous evolution but absent in the adiabatic evolutions.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{Fig_res_diff_traj}
+\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (centre) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start.
+The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start ($0.207$), at the $2$:$3$ resonance ($0.432$), after the $2$:$3$ resonance ($0.258$) and at the end ($0.677$).}
+\end{figure*}
+
+None of the adiabatic models presented here give a particularly high overlap with the entire signal, because of the effects of the resonance.
+In this case, the best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.207$.
+These are similar to the adiabatic matched before and after resonance respectively.
+These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.
+If we could construct an adiabatic model that includes the jump across resonances, it may give a good overall fit to the signal.
+
+\subsection{The effect of resonances: Jump sizes}
+\label{sec:effres-jump}
+
+\emph{Update this to reflect thesis.}
+
+As demonstrated in \figref{res-diff-traj} and computed analytically in \apref{res-asymptotic}, the full evolution undergoes a rapid change in the orbital parameters with respect to the adiabatic evolutions.
+To extract the magnitude of this jump from the trajectory data, we first must account for the fast orbital oscillations as well as ensuring that we include the full effect of the resonance, while minimising any changes resulting from other sources.
+
+To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away.
+For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}\sub{ad}[t_{\mathrm{match}} = t\sub{res}]$ for each of the orbital parameters~\cite{ColeThesis2015}.
+We then identify the most horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations.
+The resulting data and bounding lines are shown for our illustrative system in \figref{res-jump-calc}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{Fig_res_jump_calc}
+\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the 2:3 resonance, scaled by the magnitude of the $2$:$3$ resonance jump.
+The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump.
+The dotted red line indicates the computed size of the jump.
+The time axis is centred on the 2:3 resonance and is scaled by the appropriate resonance timescale.}
+\end{figure}
+
+To estimate the size of the jump $\Delta \mathcal{I}_{\mathrm{jump}}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits.
+This allows us to take into account any linear drift in $\Delta \mathcal{I}$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation.
+% In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down.
+For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.
+
+The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems.
+Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance.
+For the orbital parameter $\mathcal{I}$,
+\begin{equation}
+\label{eq:res-jump-ratio}
+\frac{\Delta \mathcal{I}\sub{jump}}{\mathcal{I}\sub{ad}(t>t\sub{res})-\mathcal{I}\sub{ad}(t<t\sub{res})} = \frac{\Delta \mathcal{I}\sub{jump}}{\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}}.
+\end{equation}
+
+The resonance jump depends sensitively on the relative phase of radial and poloidal motions.
+It is therefore instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi$, which we do for our illustrative system parameters.
+Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqnref{delta-I-a}.
+
+Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump.
+We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$.
+The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
+The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset; this means that for every value of $q$, there is always a resonance jump in at least one parameter.
+We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
+\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$.
+The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqnref{res-jump-ratio}.
+The individual jumps as well as a sinusoidal fit are plotted for the three orbital shape parameters (solid lines) and the three orbital frequencies (dotted).}
+\end{figure}
+
+For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section.
+The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{overlap_vs_q}
+\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$.
+Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
+The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
+\end{figure}
+
+The best performing adiabatic for the majority of our systems matched at the end of the inspiral.
+We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance.
+We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}.
+There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
+
+Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
+The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
+A small fraction of systems have a lower overlap than expected, at around $0.5$.
+This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
+
+%%%%%%%
 \emph{Add in here some numerical results. Definitely plot showing resonant flux enhancement vs eccentricity. Probably not much else.}
@@
 We expect nearly-circular orbits to show small flux enhancements due to the small volume of the $r\theta$ 2-torus; resonant orbits come close to every allowed point, thus approximating a non-resonant orbit.
-In \figref{res-flux-rms-p}, we show the root-mean-square (over $a$ and $\cos\iota$) relative flux enhancement for the semi-latus rectum as a function of eccentricity, for a selection of low-order resonances.
+In \figref{res-flux-rms-p}, we show the root-mean-square (over $a$ and $\cos\iota$) relative flux enhancement for the semilatus rectum as a function of eccentricity, for a selection of low-order resonances.
 As expected, it is found that larger eccentricities give rise to larger resonant flux enhancements in general.
@@
 The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a}.
-They consider the 1:2, 1:3, 2:3 and 3:4 resonances for EMRIs with $a_\ast=0.9$, either $e=0.3$ or $0.7$, and $\theta\sub{inc} = 20\text{\textdegree}$ or $70$\textdegree.
-We compare our results to their published values (in their tables I--IV) in \figref{res-flux-FHR}.
+They consider the $1$:$2$, $1$:$3$, $2$:$3$ and $3$:$4$ resonances for EMRIs with $a_\ast=0.9$, either $e=0.3$ or $0.7$, and $\theta\sub{inc} = 20^\circ$ or $70^\circ$.
 There is some qualitative agreement between the two approaches: more eccentric orbits give rise to larger enhancements; and the 3:4 resonance has a negligibly small enhancement.
-However, for the most significant 2:3 resonance, we typically predict larger orbital parameter jumps.
+However, for the most significant $2$:$3$ resonance, we typically predict larger orbital parameter jumps.
 Our results in \secref{astrophysics}, regarding the impact of resonances on a population of EMRIs, are therefore conservative predictions, as our model over-estimates the resonance effects.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{res_flux_FHR}
-\caption{\label{fig:res-flux-FHR}Relative flux enhancements of $E$ (left column), $L_z$ (middle) and $Q$ (right) for resonance ratios corresponding to the 1:3, 1:2, 2:3 and 3:4 resonances.
-We show results as computed here with a solid red line, and results from Flanagan, Hughes and Ruangeri (FHR14)~\cite{Flanagan2012a} with a dashed blue line.
-Each row has a different eccentricity and inclination corresponding to tables I--IV in FHR14.
-All systems have $a_\ast=0.9$.}
-\end{figure*}
-
-Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on a population of inspirals.
-
-\section{Impact of resonances on gravitational waveforms}
-\label{sec:waveforms}
-
-Having built an understanding of the properties of transient resonances in EMRIs, we now consider their measurable consequences.
-EMRIs will be observed through GW observations, hence we are interested in how passing through a resonance impacts this signal.
-In \secref{kludge} we discuss our chosen waveform generation scheme, giving demonstrations of its accuracy; in subsequent sections we detail the impact of resonances on the measured properties of the GWs.
-In the following section, we investigate how resonances would impact real-world observations of EMRIs assuming realistic astrophysical population of EMRIs.
-
-\subsection{Waveform model and analysis}
-\label{sec:kludge}
-\emph{Can probably shorten this section a bit.}
-
-To generate gravitational waveforms, we employ the numerical kludge (NK) method of Babak {\it{et al}}.~\cite{Babak2007}, augmented to include evolution of the positional elements.
-We first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
-This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; overlaps between Teukolsky-based and the best NK waveforms are typically $95\%$ or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
-
-Following the NK method, we first compute the inspiral trajectory of the compact object around a central Kerr BH using the method of osculating elements detailed in \secref{forced-motion}, with the dissipative part of the post-Newtonian self-force from \secref{self-force}.
-We then map the Boyer--Lindquist coordinates to spherical polar coordinates in Minkowski, facilitating the use of a flat-spacetime waveform generation technique, the standard quadrupole formula~\cite{Misner1973}.
-
-We expect these waveforms to be sufficiently accurate for our purposes, given the inaccuracy of the self-force model being employed.
-It would, however, be straight-forward to adopt a different waveform generation scheme, for example by including higher-order multipole moments, should advances be made in determining the self-force.
-
-Once we have established a waveform generation procedure, we can define a natural inner product between two waveforms, $s(t)$ and $h(t)$, according to~\cite{Finn1992}
-\begin{equation}
-\label{eq:innerprod}
-\overlap{s}{h} = 2 \intd{0}{\infty}{\frac{\tilde{s}(f)\tilde{h}^*(f)+\tilde{s}^*(f)\tilde{h}(f)}{S_n(f)}}{f},
-\end{equation}
-where $\tilde{s}(f)$ represents the Fourier transform of $s(t)$ and similarly for $\tilde{h}(f)$, and $S_n(f)$ is the one-sided noise power spectral density (PSD)~\cite{Moore2014a}.\footnote{We use cubic interpolation to construct NK waveforms with identical time sampling.
-%In practice, the continuous functions $s(t)$ and $h(t)$ are not available, instead replaced by the time series $s(t_i)$ and $h(t_j)$ where $\{t_i\}$ and $\{t_j\}$ are equally-spaced sets of time values, but which do not necessarily have the same length or spacing.
-To take this into account, we project the $h(t_j)$ onto $\{t_i\}$ using a cubic interpolation scheme and set any values outside the domain of $\{t_j\}$ to zero.
-%This whole procedure gives more accurate results than interpolating in the Fourier domain since the waveforms are smoother in the time domain.
-A Planck-taper window function~\cite{McKechan2010} is applied to reduce unwanted spectral leakage in the Fourier transforms. 
-%After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
-} We are interested in the detectability of EMRIs using future space-based GW detectors and so we use a PSD appropriate for describing the eLISA mission concept, the analytic approximation of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}.
-We maximise every inner product calculation with respect to some time-offset between the models by computing the correlation between the waveforms, which can be performed cheaply using Fourier transforms.
-%and can provide a significant increase to the overlap, even though typical offset times are shorter than the orbital timescale.
-
-One of the targets of pre-eLISA research is to generate a bank of waveform templates $\{h(t;\boldsymbol{\Theta}_i)\}$ across a range of parameter space given by $\{\boldsymbol{\Theta}_i\}$.
-These can be compared to data to search for the presence of GWs.
-At the current time, with a lack of experimental data, we are interested in ensuring that the waveform templates accurately reproduce what we expect to observe in nature, without being too computationally expensive.
-Ideally, we would like to use waveform templates from EMRI systems that are evolved under the entire self-force model, but these are too expensive.
-In practice, we are forced to use cheaper adiabatic waveforms, but these do not include the effect of resonances.
-To assess the impact of this, we compare data $s(t;\bar{\boldsymbol{\Theta}})$ generated using the full self-force model to templates $h(t;\boldsymbol{\Theta})$ generated using our 2-torus averaged self-force model (see \secref{adiabatic}).
-
-We wish to test whether there exists some set of parameters $\boldsymbol{\Theta}$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
-We do this by evaluating the signal-to-noise ratio (SNR) for each waveform template,
-\begin{equation}
-\rho\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{h}{h}}}.
-\end{equation}
-If a template were to exactly match the data, it would produce an SNR of $\sqrt{\overlap{s}{s}}$ and so the overlap
-\begin{equation}
-\label{eq:overlap}
-\mathbb{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}}
-\end{equation}
-provides an indication of how well-matched the template is to the data.
-We leave an investigation into parameter-estimation biases (where the parameters of the best-matching adiabatic $\boldsymbol{\Theta}$ do not agree with the parameters of the system $\bar{\boldsymbol{\Theta}}$) to future work.
-
-To demonstrate the approach, we can calculate the full evolution and the 2-torus averaged adiabatic evolution over $2$~years for a set of initial parameters chosen to avoid any significant resonances.
-Example evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, which shows that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
-The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the 2-torus averaging explicitly smooths out the visible structure on this scale.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{Fig_good_traj}
-\caption{\label{fig:good-traj}The evolution of the orbital parameters $E$ (left), $L_z$ (center) and $Q$ (right) under the full (solid line) and adiabatic (dashed) models for an illustrative EMRI system that does not encounter any significant resonances.
-The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.
-This system has $\mu = 10 M_{\odot}$, $q = 3\times 10^{-6}$, $a_\ast=0.95$, initial semi-latus rectum $p_0=7.5$, initial eccentricity $e_0=0.7$, initial inclination $\cos \iota_0 = 0.5$ and redshift $z=0.204$.}
-\end{figure*}
-
-Once the trajectories have been computed using both methods of evolution, we can calculate the corresponding waveforms.
-The plus-polarised GW at the start and end of the evolution is shown in \figref{good-waveform}.
-The adiabatic waveform shows good agreement with the full waveform in both amplitude and phase.
-We find an overlap $\mathbb{O} = 0.993$, illustrating that adiabatic models can be used in place of the full evolution in cases where resonances are not encountered.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.92\textwidth]{Fig_good_waveform}
-\caption{\label{fig:good-waveform}The plus-polarised waveform for an illustrative EMRI system with initial semi-latus rectum $p_0=7.5$, shown for two short segments at the start and end of a $2$ year evolution.
-The full and adiabatic models are both shown, but are indistinguishable by eye.}
-\end{figure*}
-
-\subsection{The effect of resonances: Dephasing}
-\label{sec:effres-phase}
-
-The illustrative EMRI system of the previous section does not encounter a significant resonance and the adiabatic evolution provides a good match to the full evolution.
-We now study a system that does pass through a resonance during its $2$ year evolution.
-Specifically, we choose the initial conditions to be the same as before but with an initial semi-latus rectum $p=7.85$.
-This system passes through the $2$:$3$ resonance, the effect of which is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
-Plotting the plus-polarised GWs at the start and end of the evolution, as before, demonstrates this dephasing, as shown in \figref{dephased-waveform}.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.92\textwidth]{Fig_dephased_waveform}
-\caption{\label{fig:dephased-waveform}The plus-polarised waveform for an illustrative EMRI system with $p=7.85$, shown for two short segments at the start and end of a $2$ year evolution.
-The full and adiabatic models are both shown; there is a significant dephasing between them by the end of the evolution.}
-\end{figure*}
-
-To highlight the dephasing more quantitatively, we calculate the shortened overlap between the two models as a function of time $t$, defined as the overlap obtained by including only the part of the waveform within some $\Delta t$ of $t$.
-For our illustrative system, we choose $\Delta t$ such that we can calculate 25 non-overlapping values of this shortened overlap.
-Before the resonance occurs, the adiabatic model provides a good match to the full evolution but the overlap is reduced near to the resonance and never fully recovers afterwards.
-This is shown in \figref{overlap-dephasing}, which is centred on the time at which the full evolution crosses the $2$:$3$ resonance.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
-\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
-The time axis is centred on the value at which the full evolution crosses the $2$:$3$ resonance.}
-\end{figure}
-
-Also shown is the shortened overlap computed between the full evolution and a different adiabatic evolution that is chosen to match the full evolution at the end of the integration.
-In this case, we see similar behaviour as before: the adiabatic waveform has a high overlap in the region where it is constructed to match the full evolution and the boundary of that region is set by the presence of an important resonance.
-
-\subsection{The effect of resonances: Loss of SNR}
-\label{sec:SNRloss}
-
-To be able to quantify the effect that the resonance has on an evolution, we turn to the SNR and the overlap between full and adiabatic models computed for the entire length of the observed inspiral.
-The dephasing means that an adiabatic evolution with the same initial conditions as a full evolution is likely to have a low overlap if a significant resonance is encountered.
-However, this does not necessarily mean that all adiabatic evolutions have low overlaps; we need to search for optimal adiabatic parameters, such that the overlap is maximal.
-
-The extremely large parameter space of adiabatic waveforms, coupled with the expensive nature of our 2-torus averaging routine, renders a brute-force approach computationally impossible; it is therefore necessary to restrict the search in some way.
-For this preliminary investigation into the importance of resonances, we do this by focussing on a small subset of parameters that we suspect will produce a large overlap.
-We make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t_{\mathrm{match}}$.
-This reduces the search to a one parameter family of waveforms that can easily be computed concurrently with the full evolution.
-
-The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
-Anticipating that resonances shall play a key role in determining the degree of waveform overlap, we choose matching times situated at $5\tau\sub{res}$ after each resonance of interest, namely the low-order 2:3 and 1:2.
-These matching times lie in a portion of the evolution that is not affected by a resonance and so should result in a large overlap for that region of the inspiral.
-We also include templates that match at the start of the evolution, so that we have a model that should match the portion of the evolution before the first resonance, and at the end of the evolution, as this is often where most SNR is accumulated.
-In addition, we include models that match exactly on each of the resonances as these may achieve partial matching both before and after the resonance and hence perform well.
-
-In our experience, we found that these choices were not sufficient to produce a good family of adiabatic waveforms.
-In particular, if an adiabatic model happened to match the full model close to the edge of the envelope generated by the fast orbital evolution, then the two trajectories eventually diverged and the resulting waveform provided a poor match.
-We therefore adjust the values of $t_{\mathrm{match}}$ so that they correspond to times of apoapsis.
-This ensures that the adiabatic model intersects with the full model close to the centre of the envelope, as demonstrated by the inset plots in \figref{good-traj}.
-As a result, we obtain much higher overlaps in general.
-
-We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.
-
-This one-parameter family of adiabatic models does not explore the extrinsic parameter space nor does it consider different values of the constant intrinsic parameters (the masses of the BHs and the central BH spin).
-It is possible that small deviations from these true parameters give rise to better matching adiabatic models, to partially correct for the (lack of) resonance effects.
-Indeed, systematic biases in parameter estimation are expected when the model waveforms do not accurately reflect the data~\cite{Cutler2007}.
-The result of these restrictions is that the maximum overlap calculated using our family is a strict lower bound on the true maximum overlap.
-
-For the illustrative EMRI system detailed in the previous section, we have computed the family of adiabatic models set by $t_{\mathrm{match}}$ and the resulting trajectories are consistent for the length of the evolution.
-As they cannot be easily distinguished by eye, we plot the difference (in $E$, $L_z$ and $Q$) between the models and the adiabatic evolution that matches at the start, shown in \figref{res-diff-traj}.
-The jumps in the orbital parameters due to the $2$:$3$ resonance can be clearly seen, as can the fast orbital oscillations present in the full evolution but absent in the adiabatics.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{Fig_res_diff_traj}
-\caption{\label{fig:res-diff-traj}The differences in orbital parameters $E$ (left), $L_z$ (centre) and $Q$ (right) between each evolution scheme and the adiabatic model that matches at the start.
-The solid line shows the full evolution and the dashed lines show the different adiabatic evolutions, which match the full evolution at different times throughout the inspiral (numbers in parentheses give the overlap with the full evolution): at the start ($0.207$), at the $2$:$3$ resonance ($0.432$), after the $2$:$3$ resonance ($0.258$) and at the end ($0.677$).}
-\end{figure*}
-
-None of the adiabatic models presented here give a particularly high overlap with the full evolution, because of the effects of the resonance.
-The best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.207$.
-These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.
-
-\subsection{The effect of resonances: Jump sizes}
-
-\emph{Update this to reflect thesis.}
-
-As demonstrated in \figref{res-diff-traj} and computed analytically in \apref{res-asymptotic}, the full evolution undergoes a rapid change in the orbital parameters with respect to the adiabatic evolutions.
-To extract the magnitude of this jump from the trajectory data, we first must account for the fast orbital oscillations as well as ensuring that we include the full effect of the resonance, while minimising any changes resulting from other sources.
-
-To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away.
-For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}\sub{ad}[t_{\mathrm{match}} = t\sub{res}]$ for each of the orbital parameters.
-We then identify the most horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations.
-The resulting data and bounding lines are shown for our illustrative system in \figref{res-jump-calc}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_res_jump_calc}
-\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the 2:3 resonance, scaled by the magnitude of the 2:3 resonance jump.
-The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump.
-The dotted red line indicates the computed size of the jump.
-The time axis is centred on the 2:3 resonance and is scaled by the appropriate resonance timescale.}
-\end{figure}
-
-To estimate the size of the jump $\Delta \mathcal{I}_{\mathrm{jump}}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits.
-This allows us to take into account any linear drift in $\Delta \mathcal{I}$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation.
-% In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down.
-For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.
-
-The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems.
-Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance.
-For the orbital parameter $\mathcal{I}$,
-\begin{equation}
-\label{eq:res-jump-ratio}
-\frac{\Delta \mathcal{I}\sub{jump}}{\mathcal{I}\sub{ad}(t>t\sub{res})-\mathcal{I}\sub{ad}(t<t\sub{res})} = \frac{\Delta \mathcal{I}\sub{jump}}{\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}}.
-\end{equation}
-
-The resonance jump depends sensitively on the relative phase of radial and poloidal motions.
-It is therefore instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi$, which we do for our illustrative system parameters.
-Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqnref{delta-I-a}.
-
-Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump.
-We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$.
-The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
-The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset; this means that for every value of $q$, there is always a resonance jump in at least one parameter.
-We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
-\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$.
-The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqnref{res-jump-ratio}.
-The individual jumps as well as a sinusoidal fit are plotted for the three orbital shape parameters (solid lines) and the three orbital frequencies (dotted).}
-\end{figure}
-
-For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section.
-The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{overlap_vs_q}
-\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$.
-Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
-The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
-\end{figure}
-
-The best performing adiabatic for the majority of our systems matched at the end of the inspiral.
-We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance.
-We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}.
-There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
-
-Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
-The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
-A small fraction of systems have a lower overlap than expected, at around $0.5$.
-This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
+%%%%
 
@@
 To create our EMRI population, we sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value $T_0$, and evolve the system using the analytic kludge (AK) prescription of Barack and Cutler \cite{Barack2004} until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
-This ignores the influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
+This ignores the (small) influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
 
@@
 \label{eq:EMRI-number}
-N\sub{EMRI} = \intd{z\,=\,0}{z\sub{max}}{ \intd{M\,=\,M\sub{min}}{M\sub{max}}{\mathcal{R}t\sub{life}\diff{n}{\,\ln M} \diff{V\sub{c}}{z}}{\ln M}}{z},
+N\sub{EMRI} = \intd{z\,=\,0}{z\sub{max}}{ \intd{M\,=\,M\sub{min}}{M\sub{max}}{\mathcal{R}t\sub{life}\diff{n}{\,\ln M} \diff{V\sub{c}}{z}}{\,\ln M}}{z},
 \end{equation}
@@
 This gives us a total of $12$ parameters for our fit.
-Whilst this may sound large, if we were fitting an expansion to quadratic order in combinations of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{This does not give as good a fit as our function.} Our optimised parameters are
+Whilst this may sound large, if we were fitting an expansion to quadratic order in combinations of $\{e,\iota,a_\ast,\nu\}$ we would have $15$ parameters.\footnote{We find that this does not give as good a fit as our function.} Our optimised parameters are
 \begin{equation}

### 2016-08-06 asymptotic.tex
Jumps done

@@
 We are not particularly interested in the deviation in $J$, of greater concern is the change in the orbital parameters $\{E,L_z,Q\}$.
-Assuming that there is a smooth transformation that maps between $J$ and these, then, to lowest order, we can calculate the deviation relative to the adiabatic prescription by substituting the forcing functions $G^{(1)} \rightarrow G_a^{(1)}$, where $G_a^{(1)}$ describes the evolution of $\mathcal{I}_a$ through the effects of the self-force.
+Assuming that there is a smooth transformation that maps between $J$ and these, then, to lowest order, we can calculate the deviation relative to the adiabatic prescription by substituting the forcing functions $G^{(1)} \rightarrow G_a^{(1)}$, where $G_a^{(1)}$ describes the evolution of $\mathcal{I}^a$ through the effects of the self-force.
 This result is quoted by Flanagan and Hinderer~\cite{Flanagan2012}.
@@
 \begin{equation}
-\left\langle \diff{\mathcal{I}_a}{t}\right\rangle_{q'} =  \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
+\left\langle \diff{\mathcal{I}^a}{t}\right\rangle_{q'} =  \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
 \end{equation}
@@
 \begin{align}
-\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
+\Delta \mathcal{I}^a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
  {} & \times \left.\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\dot{\Omega}\right)\right]\right. \\

### 2016-08-06 cut_figures.tex
Jumps done

@@
 \end{figure*}
+
+For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section.
+The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{overlap_vs_q}
+\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$.
+Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
+The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
+\end{figure}
+The best performing adiabatic for the majority of our systems matched at the end of the inspiral.
+We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance.
+We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}.
+There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
+Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
+The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
+A small fraction of systems have a lower overlap than expected, at around $0.5$.
+This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.

### 2016-08-06 res_char.tex
Jumps done

@@
 \newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
-\newcommand{\Figref}[1]{Fig.~\ref{fig:#1}}
+\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
 \newcommand{\secref}[1]{Sec.~\ref{sec:#1}}
@@
 Flanagan and Hinderer~\cite{Flanagan2012} highlighted a previously overlooked phenomenon that occurs in the general relativistic two-body problem, that of transient resonances.
-Geodesic orbits in GR have three associated frequencies: the radial frequency $\Omega_r$, the polar frequency $\Omega_\theta$ and the azimuthal frequency $\Omega_\phi$.
-The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
+Geodesic orbits in GR have three associated frequencies: the radial frequency $\Omega_r$, the polar frequency $\Omega_\theta$ and the azimuthal frequency $\Omega_\phi$.\footnote{In the strong-field regime, it is possible to have isofrequency pairings, where two different orbits share the same orbital frequencies~\cite{Warburton2013}.
+The evolution of the frequencies will still differ, such that orbital trajectories can be reconstructed.} The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
 % Section 10.6
-In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
+In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field regime $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
 For extreme-mass-ratio systems, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
@@
 \end{subequations}
-The orbital frequencies are thus
+The orbital frequencies are thus~\cite{Fujita2009}
 \begin{equation}
@@
 \end{equation}
-to act as Mino-time frequencies. We can now convert to coordinate-time frequencies with
+to act as Mino-time frequencies. We can now convert to coordinate-time frequencies with~\cite{Drasco2004}
 \begin{equation}
@@
 Beyond geodesic motion in the Kerr spacetime, a test particle follows an accelerated trajectory determined by \eqnref{Mino-E-o-M}.
-This may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005}.
-The averaged force can be computed from the radiative field \cite{Galtsov1982,Mino2003,Sago2006,Ganz2007}.
+This may be approximated by the adiabatic prescription~\cite{Hinderer2008} by dropping the forcing term $g_\alpha^{(1)}$ (and all higher-order terms) and replacing $G_\alpha^{(1)}$ with its average over the $2$-torus parametrized by $q_r$ and $q_\theta$, $\langle G_\alpha^{(1)}\rangle_{q_r,\,q_\theta}$~\cite{Drasco2005,Grossman2011}.
+The averaged force can be computed from the radiative field~\cite{Galtsov1982,Mino2003,Sago2006,Ganz2007}.
 This piece is purely dissipative and determines how the inspiral evolves due to the radiation of GWs.
@@
 \end{align}\end{widetext}
-This average describes the Mino-time rate of change of the quantity $X$ over an orbit. To convert to a co-ordinate flux of the averaged quantity, we simply divide by the period $\Gamma$, defining
+This average describes the Mino-time rate of change of the quantity $X$ over an orbit.
+To convert to a co-ordinate flux of the averaged quantity, we simply divide by the period $\Gamma$~\cite{Flanagan2012a}, defining
 \begin{equation}
@@
 Flanagan and Hinderer~\cite{Flanagan2012} gave an expression for this deviation.
-If we denote the orbital parameters by $\mathcal{I}_a = \{E,L_z,Q\}$, then the change across resonance is
+If we denote the orbital parameters by $\mathcal{I}^a = \{E,L_z,Q\}$, then the change across resonance is
 \begin{align}
-\Delta \mathcal{I}_a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
+\Delta \mathcal{I}^a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
  {} & \times {} \exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
@@
 \begin{equation}
-\diff{\mathcal{I}_a}{t} = \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
+\diff{\mathcal{I}^a}{t} = \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
 \end{equation}
@@
 \label{eq:delta-I-a}
-\Delta \mathcal{I}_a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
+\Delta \mathcal{I}^a = \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega}\right)\right]. 
 \end{equation}
 Schematically, this can then be understood as the magnitude of the forcing function on resonance $\sim \eta F_{a,\,s}$ multiplied by the time on resonance $\sim \tau_{\mathrm{res},\,s}$ and a function that varies with the phase $\widehat{\kappa}_0$.
-Averaging over all values of $\widehat{\kappa}_0$ is equivalent to averaging over all values of $\psi_{\theta_-}$, and has the same effect as averaging over the $\psi$--$\chi$ $2$-torus; this gives an average discrepancy relative to the adiabatic evolution of
-\begin{equation}
-\left\langle \Delta \mathcal{I}_a \right\rangle_{\hat{\kappa}_0} = 0,
+Averaging over all values of $\widehat{\kappa}_0$ is equivalent to averaging over all values of $\psi_{\theta_-}$, and has the same effect as averaging over the $\psi$--$\chi$ $2$-torus~\cite{Grossman2011}; this gives an average discrepancy relative to the adiabatic evolution of
+\begin{equation}
+\left\langle \Delta \mathcal{I}^a \right\rangle_{\hat{\kappa}_0} = 0,
 \end{equation}
@@
 The inset plots show the behaviour on short timescales, where the fast orbital oscillations can be seen.
-This system has $\mu = 10 M_{\odot}$, $q = 3\times 10^{-6}$, $a_\ast=0.95$, initial semilatus rectum $p_0 = 7.5$, initial eccentricity $e_0 = 0.7$, initial inclination $\cos \iota_0 = 0.5$ and redshift $z=0.204$.}
+This system has $\mu = 10 M_{\odot}$, $\eta = 3\times 10^{-6}$, $a_\ast=0.95$, initial semilatus rectum $p_0 = 7.5$, initial eccentricity $e_0 = 0.7$, initial inclination $\cos \iota_0 = 0.5$ and redshift $z=0.204$.}
 \end{figure*}
@@
 The large parameter space of adiabatic waveforms, coupled with the expensive nature of our $2$-torus averaging routine, renders a brute-force approach prohibitively expensive.
-For this preliminary investigation, we focus on a small subset of parameters that we suspect will produce a large overlap, and make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t_{\mathrm{match}}$.
+For this preliminary investigation, we focus on a small subset of parameters that we suspect will produce a large overlap, and make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t\sub{match}$.
 This reduces the search to a $1$-parameter family of waveforms that can easily be computed concurrently with the full evolution.
@@
 
-The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t_{\mathrm{match}}$.
-We choose $5\tau\sub{res}$ after each resonance of interest, namely the low-order $1$:$2$ and $2$:$3$ resonances.\footnote{We adjust the values of $t_{\mathrm{match}}$ so that they correspond to times of apoapsis.
+The problem of searching over adiabatic templates now reduces to the task of choosing appropriate values of $t\sub{match}$.
+To demonstrate how changing the matching time affects the overlap, we use $5\tau\sub{res}$ after each resonance of interest, namely the low-order $1$:$2$ and $2$:$3$ resonances.\footnote{We adjust the values of $t\sub{match}$ so that they correspond to times of apoapsis.
 This ensures that the adiabatic model intersects with the full instantaneous model close to the centre of it oscillatory envelope, as demonstrated by the inset plots in \figref{good-traj}.
@@
 
-%We have tried a more rigid system of choosing the values of $t_{\mathrm{match}}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.
+%We have tried a more rigid system of choosing the values of $t\sub{match}$, producing an adiabatic evolution regularly at fixed intervals, but these generally perform no better than our adaptive match points and require many more adiabatic evolutions to be calculated.
 
@@
 These are similar to the adiabatic matched before and after resonance respectively.
-These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution.
+These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution ($69.8\%$ of the inspiral is post-resonance): there is not an exact equivalence because of the frequency dependence of the PSD.
+
 If we could construct an adiabatic model that includes the jump across resonances, it may give a good overall fit to the signal.
@@
 
-\emph{Update this to reflect thesis.}
-
-As demonstrated in \figref{res-diff-traj} and computed analytically in \apref{res-asymptotic}, the full evolution undergoes a rapid change in the orbital parameters with respect to the adiabatic evolutions.
-To extract the magnitude of this jump from the trajectory data, we first must account for the fast orbital oscillations as well as ensuring that we include the full effect of the resonance, while minimising any changes resulting from other sources.
-
-To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away.
-For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I} \equiv \mathcal{I}_{\mathrm{full}} - \mathcal{I}\sub{ad}[t_{\mathrm{match}} = t\sub{res}]$ for each of the orbital parameters~\cite{ColeThesis2015}.
+As explained in \secref{flux-enhance} and illustrated in \figref{res-diff-traj}, the full instantaneous evolution undergoes a rapid change in the orbital parameters (with respect to the adiabatic evolution) when passing through resonance.
+The size of the jump influences the subsequent orbital evolution.
+
+To extract the magnitude of this jump from the trajectory data, we must account for the fast orbital oscillations as well as the general average (adiabatic) evolution~\cite{ColeThesis2015}.
+We first computed the difference $\Delta \mathcal{I}^a \equiv \mathcal{I}^a\sub{full} - \mathcal{I}^a\sub{ad}[t\sub{match} = t\sub{res}]$ between the orbital parameters calculated using from the full instantaneous evolution and from an adiabatic evolution matched to the parameters at resonance.
+We then fit linear bounds to the oscillating envelope, both before and after the resonance, using data $5\tau\sub{res}$ to $10\tau\sub{res}$ away.
+These are averaged to give general pre- and post-resonance trends, which are extrapolated to the time of resonance.
+The difference at the time of resonance gives an estimate for the jump $\Delta \mathcal{I}^a\sub{jump}$ in orbital parameter $\mathcal{I}^a$.\footnote{As a cross-check, comparable values for the jump are obtained by averaging (over an integer number of radial and poloidal orbits) the flux on a resonant geodesic~\cite{ColeThesis2015}.} \Figref{res-jump-calc} illustrates how the size of the jump is calculated.
+
+%To achieve this, we isolate spans of data away from the resonance; specifically, looking at data that lies within five and ten resonance timescales away.
+For both the pre- and post-resonance datasets, we compute $\Delta \mathcal{I}^a \equiv \mathcal{I}^a\sub{full} - \mathcal{I}^a\sub{ad}[t\sub{match} = t\sub{res}]$ for each of the orbital parameters~\cite{ColeThesis2015}.
 We then identify the most horizontal straight lines that bound the data, which give an approximation to the behaviour of the fast orbital oscillations.
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_res_jump_calc}
-\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the 2:3 resonance, scaled by the magnitude of the $2$:$3$ resonance jump.
-The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump.
-The dotted red line indicates the computed size of the jump.
-The time axis is centred on the 2:3 resonance and is scaled by the appropriate resonance timescale.}
-\end{figure}
-
-To estimate the size of the jump $\Delta \mathcal{I}_{\mathrm{jump}}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits.
+%To estimate the size of the jump $\Delta \mathcal{I}\sub{jump}$, we extrapolate the bounding lines to the resonance time and calculate the difference between the pre- and post-resonance values, quoting the final result as the average of that obtained using the upper and lower limits.
 This allows us to take into account any linear drift in $\Delta \mathcal{I}$ that can occur over the resonance timescale due to the inaccuracy of the adiabatic approximation.
 % In addition, by using information from both the upper and lower limits, we can veto jump calculations that appear erroneous: if one jump is positive and the other negative, then we ignore that resonance; this often occurs very near to plunge, where the adiabatic approximation breaks down.
-For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.
-
-The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems.
-Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance.
-For the orbital parameter $\mathcal{I}$,
-\begin{equation}
-\label{eq:res-jump-ratio}
-\frac{\Delta \mathcal{I}\sub{jump}}{\mathcal{I}\sub{ad}(t>t\sub{res})-\mathcal{I}\sub{ad}(t<t\sub{res})} = \frac{\Delta \mathcal{I}\sub{jump}}{\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}}.
-\end{equation}
-
-The resonance jump depends sensitively on the relative phase of radial and poloidal motions.
-It is therefore instructive to study the resonance problem using an ensemble of systems differing only in the initial value of the radial phase $\psi$, which we do for our illustrative system parameters.
-Choosing random values for $\psi$ corresponds to random values of the orbital phase on resonance $\widehat{\kappa}_0$, which should map out the full range of resonance jumps for these particular orbital shape parameters, according to \eqnref{delta-I-a}.
-
-Given our computed resonance jumps for all of the orbital parameters, we extract a single phase parameter $q$ for each system that specifies the magnitude of each jump.
-We expect such a phase parameter to be linearly offset from $\widehat{\kappa}_0$.
-The resonance jumps are sinusoidal with respect to $q$, as shown by the fits in \figref{resjump-vs-q}, lending support to \eqnref{delta-I-a} and giving us confidence in our numerical jump calculations.
-The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset; this means that for every value of $q$, there is always a resonance jump in at least one parameter.
-We can also compute the jumps in the fundamental frequencies for each system (shown as dashed lines in \figref{resjump-vs-q}).
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
-\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$.
-The jump is expressed as a percentage of the adiabatic change in each parameter, given by \eqnref{res-jump-ratio}.
-The individual jumps as well as a sinusoidal fit are plotted for the three orbital shape parameters (solid lines) and the three orbital frequencies (dotted).}
-\end{figure}
-
-For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section.
-The largest overlap in each case, plotted in \figref{overlap-vs-q}, appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{overlap_vs_q}
-\caption{\label{fig:overlap-vs-q}The overlap between the full and adiabatic evolutions, maximised over the family of generated adiabatics, as a function of the extracted phase parameter $q$.
-Each dot represents a system with our illustrative parameters and a random choice of the initial radial phase.
-The red line indicates the expected overlap assuming that the waveforms match exactly post-resonance, but have zero overlap pre-resonance.}
-\end{figure}
-
-The best performing adiabatic for the majority of our systems matched at the end of the inspiral.
-We can therefore estimate the value of the overlap by assuming that the adiabatic waveforms recover the entire SNR in the region post-resonance, but none at all in the region pre-resonance.
-We expect the overlap to scale with the fraction of time spent post-resonance, the value of which is demonstrated by the red line in \figref{overlap-vs-q}.
-There is not an exact equivalence because of the frequency-dependent noise of eLISA; cycles near the end of the inspiral accumulate proportionally more SNR at a frequency closer to the minimum of the eLISA noise curve.
-
-Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
-The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
-A small fraction of systems have a lower overlap than expected, at around $0.5$.
-This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
-
-%%%%%%%
-\emph{Add in here some numerical results. Definitely plot showing resonant flux enhancement vs eccentricity. Probably not much else.}
-
-We can study the predicted orbital-parameter jumps numerically by considering the long-time-average fluxes for particular (fixed) geodesics.
+%For typical systems, this numerically computed jump appears to well describe the resonance, as shown by the red dotted line in \figref{res-jump-calc}.
+
+%We can study the predicted orbital-parameter jumps numerically by considering the long-time-average fluxes for particular (fixed) geodesics.
 For non-resonant geodesics, this will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference.
@@
 
-We expect nearly-circular orbits to show small flux enhancements due to the small volume of the $r\theta$ 2-torus; resonant orbits come close to every allowed point, thus approximating a non-resonant orbit.
-In \figref{res-flux-rms-p}, we show the root-mean-square (over $a$ and $\cos\iota$) relative flux enhancement for the semilatus rectum as a function of eccentricity, for a selection of low-order resonances.
-As expected, it is found that larger eccentricities give rise to larger resonant flux enhancements in general.
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{Fig_res_jump_calc}
+\caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the $2$:$3$ resonance, scaled by the magnitude of the $2$:$3$ resonance jump.
+The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump.
+The dotted red line indicates the computed size of the jump.
+The time axis is centred on the $2$:$3$ resonance and is scaled by the resonance timescale $\tau\sub{res}$.}
+\end{figure}
+
+%The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems.
+Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance.
+For the orbital parameter $\mathcal{I}$,
+%\begin{equation}
+%\label{eq:res-jump-ratio}
+%\frac{\Delta \mathcal{I}\sub{jump}}{\mathcal{I}\sub{ad}(t>t\sub{res})-\mathcal{I}\sub{ad}(t<t\sub{res})} = \frac{\Delta \mathcal{I}\sub{jump}}{\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}}.
+%\end{equation}
+
+The jump in orbital parameters depends sensitively on the relative phase of radial and poloidal motions.
+This can be illustrated using an ensemble of orbits with different orbital phases $\psi_{\theta_-}$.
+According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.
+Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
+We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta L_{z\,\mathrm{jump}}$ ellipse.
+The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset~\cite{Flanagan2012a}; this means that for every value of $q$, there is always a resonance jump in at least one parameter.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$ as expected from matching the post-resonance region of the the waveform.}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
+\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$ (cf.\ figure 3 of \cite{Flanagan2012a}).
+The jump is expressed as a percentage of the adiabatic change in each parameter across resonance $\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}$.
+The individual jumps as well as a sinusoidal fit are plotted for the orbital parameters (solid) and the three orbital frequencies (dotted).}
+\end{figure}
+
+%For each system, we compute the overlap between the full evolution and the family of adiabatics given by our set of match times $t\sub{match}$ described in the previous section.
+The largest overlap in each case appears to be roughly independent of the value of $q$ and hence with the relative sizes of the orbital parameter jumps.
+Systems with $|q| \approx \pi/2$ (corresponding to small jumps in $E$ and $L_z$) have higher than average overlaps due to additional accumulation in the pre-resonance region.
+The adiabatic models of these systems dephase more slowly, but still do not produce very high overlaps due to the jump in $Q$.
+We expect the overlap to scale with the fraction of time spent post-resonance.
+
+Let us now consider jumps for systems other than our illustrative $2$:$3$ resonance.
+We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
+In \figref{res-flux-rms-p}, we show the root-mean-square (over $a_\ast$ and $\cos\iota$) relative resonant jump for the semilatus rectum as a function of eccentricity for a selection of low-order resonances.
+The other orbital shape parameters show a similar trend with increasing eccentircity~\cite{ColeThesis2015}.
+Larger eccentricities do give rise to larger jumps, matching our expectations and Teukolsky-based calculations by Flanagan, Hughes and Ruangsri~\cite{Flanagan2012a}.
 
@@
 Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use solid (short-dashed; long-dashed) lines.
-The 2:3 resonance has the largest relative flux enhancement.}
+The $2$:$3$ resonance has the largest relative flux enhancement.}
 \end{figure}
 
-\emph{Is this description clear, and is it actually correct? It seems reasonable ...}
-
-It is informative to consider the dependence on the resonance ratio $\nu$.
-Fixing the eccentricity at $e=0.95$ to emphasise the variation, we compute the flux enhancements for different resonance ratios, shown in \figref{res-flux-rms-p-vsnu}, where we also highlight ratios arising from different values of $n_\theta$.
-We observe a clear relationship between the flux enhancement and $\nu$ for all resonances with $n_\theta = 1$ or $2$, independent of which of those is considered.
-However, negligibly small flux enhancements are observed for systems with $n_\theta = 3$ or higher.
-This can be explained by the quadrupolar nature of GWs.
-A resonant trajectory averaged over time (and consequently $\phi$) takes $n_\theta$ (possibly repeated) values of $\theta$ at each radial turning point.
-Considering rings of matter at these coordinates, the dominant spherical harmonic describing the resultant mass distribution has $l=n_\theta$ (and $m=0$).
-Since GWs are generated by the quadrupole moment at leading order, the most significant resonant enhancements will arise from systems with $n_\theta = 2$.
-Due to the presence of higher harmonics, $n_\theta = 1$ systems also contribute strongly; it is perhaps more appropriate to treat $1$:$x$ resonances as \emph{de facto} $2$:$2x$ resonances, explaining the linked behaviour of the flux enhancements for $n_\theta = 1$ and $2$ systems.
+We also expect that the effects of passing through resonance will dependent upon the particular resonance.
+Intuitively, we would expect that when $n_\theta$ and $n_r$ are large, the effects of resonance will be small, since the orbit comes close to all points on the $2$-torus.
+We have already seen that this is the case, as the adiabatic evolution is a good match to the instantaneous evolution up until it hits a resonance which is the ratio of small integers (like $2$:$3$).
+In \figref{res-flux-rms-p-vsnu} we plot the root-mean-square (over $a_\ast$ and $\cos\iota$) relative resonant jump for the semilatus rectum as a function of the resonance ratio $\nu$ for various $n_\theta$, the eccentricity is set to $e=0.95$ to emphasise the variation.
+
+While we might naively expect jumps with $n_\theta = 1$ to be greatest, we see that this is not the case.
+Instead the $n_\theta = 1$ and $n_\theta = 2$ jumps form a single continuum: we can treat $1$:$x$ resonances as \emph{de facto} $2$:$2x$ resonances.
+This suggests that more insight into resonance jumps could be gained from considering the motion across two radial periods, even for resonances with $n_\theta = 1$.
+Moving beyond $n_\theta = 2$, we see that the magnitude of jumps decreases rapidly for $n_\theta = 3$~\cite{Flanagan2012a}.
+For larger $n_\theta$, jumps are so small that cannot accurately calculate them.
 
@@
 
-\emph{Is this bit really needed ...?}
-
-The resonant enhancements to the fluxes of $E$, $L_z$ and $Q$ have been calculated previously using Teukolsky-based methods for a handful of systems Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a}.
-They consider the $1$:$2$, $1$:$3$, $2$:$3$ and $3$:$4$ resonances for EMRIs with $a_\ast=0.9$, either $e=0.3$ or $0.7$, and $\theta\sub{inc} = 20^\circ$ or $70^\circ$.
-There is some qualitative agreement between the two approaches: more eccentric orbits give rise to larger enhancements; and the 3:4 resonance has a negligibly small enhancement.
-However, for the most significant $2$:$3$ resonance, we typically predict larger orbital parameter jumps.
-Our results in \secref{astrophysics}, regarding the impact of resonances on a population of EMRIs, are therefore conservative predictions, as our model over-estimates the resonance effects.
-%%%%
-
 \section{Astrophysical implications}
@@
 
-\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix?}
+%\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix?}
 
@@
 We are interested in EMRI systems close to plunge, as it is here that the GW amplitude is largest.
-To create our EMRI population, we sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value $T_0$, and evolve the system using the analytic kludge (AK) prescription of Barack and Cutler \cite{Barack2004} until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r_p \leq r_a$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
+To create our EMRI population, we sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value $T_0$, and evolve the system using the analytic kludge (AK) prescription of Barack and Cutler \cite{Barack2004} until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
 This ignores the (small) influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.

### 2016-08-10 asymptotic.tex
First draft of text

@@
 The following derivation is reproduced from Berry~\cite{BerryThesis2013}.
-It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two timescale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found in van de Meent~\cite{VanDeMeent2014}.
+It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two timescale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found in van de Meent~\cite{VanDeMeent2013}.
 
@@
 We are only concerned with the $r$ and $\theta$ motions, so we have a $2$-dimensional system.
-We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992,VanDeMeent2014}.
+We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992,VanDeMeent2013}.
 This becomes one of the new angle variables, the other variable $q'$ can be either $q_r$ or $q_\theta$ (as, on resonance, varying one necessarily changes the other).

### 2016-08-10 cut_figures.tex
First draft of text

@@
 This is due to a sub-optimal choice of the matching times; making small adjustments to $t\sub{match}$ can lead to higher overlaps, but the exact adjustments required are difficult to predict in advance.
+
+After computing the SNRs of our NK waveforms for each of the $513$ EMRI systems, we can compare the values to those obtained using the AK formalism.
+The different approximation schemes are expected to produce different results, but they should be broadly similar.
+In \figref{pop-SNR-vs-eta}, we plot the ratio of the new NK SNR to the AK SNR, as a function of the mass ratio.
+We observe a clear trend, with systems closest to equal mass performing roughly as expected, but with the most extreme systems giving significantly lower SNRs.
+This is a consequence of the PN self-force model we are using, which we have found overestimates the inspiral rate for EMRIs.
+Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the $2\units{yr}$ observation window, they evolve within the bucket of the eLISA noise curve.
+Increasing the rate of evolution then does not significantly change the recovered SNR since we are still sensitive to all parts of the waveform.
+At smaller mass ratios, the systems evolve more slowly, from lower plunge frequencies.
+An increased evolution rate means that earlier portions of the EMRI are shifted outside the sensitive region of the detector, thus giving a much lower SNR.
+\begin{figure}
+\centering
+\includegraphics[width=0.46\textwidth]{pop_SNR_vs_eta}
+\caption{\label{fig:pop-SNR-vs-eta}The ratio of SNRs as computed with waveforms generated using the NK (adopting our PN self-force model) and AK formalisms, as a function of (the base-10 logarithm of) the mass ratio.
+Each system has an AK SNR greater than $15$, using the eLISA detector with 6 laser links.}
+\end{figure}
+We can study the effect of resonances, by comparing the maximum adiabatic SNR to the full SNR computed using the NK waveforms.
+The resulting overlap is still indicative of the effect of resonances on the population of systems.
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_Tad}
+%\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family.
+Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t\sub{insp}$.
+The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t\sub{ad}$, normalised by $t\sub{insp}$.
+The solid line corresponds to the expected value if the $1$:$2$ and $2$:$3$ resonances cause significant waveform dephasing.}
+%\end{figure}

### 2016-08-10 res_char.tex
First draft of text

@@
 
-\newcommand{\eqnref}[1]{(\ref{eq:#1})}
+\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
 \newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
@@
 \begin{abstract}
-\emph{Update abstract}
 The inspiral of stellar-mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
-In order to detect such signals and extract the desired information, it is necessary to possess a bank of accurate waveform templates.
+In order to detect and analyse these signals accurate waveform templates are required.
 For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order dissipative gravitational self-force.
@@
 During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the emitted radiation.
-In this work, an astrophysical population of detectable extreme-mass-ratio inspirals is studied.
-We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence the low eccentricity of the studied population.
-At least one fifth of the systems may become undetectable if transient resonances are not accounted for.
+We investigate the properties of transient resonances.
+We have present an evolution scheme for studying the effects of resonances and applied to an astrophysically realistic population.
+We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence the low eccentricity of the inspirals.
+Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals.
 \end{abstract}
@@
 We also use the dimensionless spin $a_\ast \equiv a/M$; we take the convention that $0 \leq a_\ast < 1$.
-We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega\sub{m} = 0.3$ and $H_0 = 70\units{km\, s^{-1}\, Mpc^{-1}}$ and do not expect the exact details of the cosmology to significantly alter our results.
+We assume a standard cosmology with $\Omega_\Lambda = 0.7$, $\Omega\sub{m} = 0.3$ and $H_0 = 70\units{km\, s^{-1}\, Mpc^{-1}}$ and do not expect the exact details of the cosmology to significantly alter our results~\cite{Mapelli2012}.
 
@@
 In order to assess the impact of this on future GW missions, we need to analyse the waveforms from a population of detectable EMRIs.
-In \secref{EMRI-population}, we detail our procedure for generating a realistic distribution of EMRIs, before turning to the effect of resonances on the detectability of EMRIs in \secref{population-SNR}.
+In \secref{EMRI-population}, we detail our procedure for generating an astrophysically motivated distribution of EMRIs, before turning to the effect of resonances on the detectability of EMRIs in \secref{population-SNR}.
 
@@
 
-%\emph{Basically summarise the GOAT population. Maybe include event rate estimates in appendix?}
-
-To generate a representative EMRI population, we need to establish plausible distributions for the parameters defining the EMRI: the properties of the central SMBH, the properties of the orbiting compact object, and the properties of the orbit itself.
-The EMRI event rate is dominated by the inspiral of heavier BHs due to the larger observational volume in comparison to that for white dwarfs and neutron stars~\cite{Gair2004}.
-We therefore take the mass of the compact object to be $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs.
-
-We take the central object to be a typical SMBH at the centre of a galaxy.
-We then consider the EMRI event rate in two pieces: the intrinsic event rate per SMBH $\mathcal{R}$, i.e.
-the number of inspiral events per unit time for a given galaxy, and the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH.
-We are interested in galaxies that possess a SMBH with a mass in the range $(10^{4}$--$10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors.
-In principle, $\mathcal{R}$ will depend on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each component, and the mass and spin of each SMBH~\cite{Alexander2005}.
-All of these properties are highly uncertain, even for our own Galaxy.
-
-Despite the difficulties, simple estimates of $\mathcal{R}$ have been carried out using Monte-Carlo methods to count the number of compact objects from isothermal distributions that spiral in to a SMBH without plunging~\cite{Merritt2013}.
+The EMRI event rate depends on the exact composition of the population of compact objects around each SMBH, the stellar density profile for each species of compact object, and the mass and spin of each SMBH~\cite{Alexander2005}, all of these properties are highly uncertain, even for our own Galaxy.
+Here, we sketch out a model which includes most of the important effects, which should suffice for illustrating the potential impact of transient resonances on the rate of detection.
+
+\subsubsection{Population model}
+
+To generate a representative EMRI population, we need to establish plausible distributions for the parameters defining the EMRI: the properties of the orbiting compact objects, the properties of the central SMBHs, and the properties of the orbits.
+
+EMRIs require a compact object such as a white dwarf, a neutron star or a stellar-mass BH to orbit around a SMBH.
+Main-sequence stars are tidally disrupted before they can complete the inspiral~\cite{Rees1988,Sigurdsson1997}.
+We expect the EMRI event rate to be dominated by BHs as the most massive species of compact object~\cite{Gair2004}.
+First, as BHs are more massive than white dwarfs or neutron stars, their GW signal is louder~\cite{Peters1963,Barack2004} and hence detectable EMRIs can come from a larger volume.
+Second, dynamical friction in the dense nuclear star clusters~\cite{Chandrasekhar1960,Antonini2011} also leads to mass segregation, causing the most massive objects concentrate closer to the SMBH~\cite{Bahcall1977,Freitag2006,Alexander2009}.
+We therefore expect BHs to dominate the inner regions of nuclear star clusters, making them the most probable candidate to inspiral.
+We adopt a fiducial mass for the compact object of $\mu = 10 M_\odot$, corresponding to a typical mass for stellar BHs~\cite{Casares2014,Corral-Santana2016,Tetarenko2016,Abbott2016d}.\footnote{If the $\sim30 M_\odot$ BHs of GW150914~\cite{Abbott2016f,Abbott2016d} are common, then this would enhance the EMRI rate.}
+
+We take the central object to be a typical SMBH at the centre of a galaxy~\cite{Kormendy1995,Ferrarese2005}; we are interested in galaxies that possess a SMBH with a mass in the range $(10^{4}$--$10^7) M_\odot$, as these give rise to EMRIs in the frequency band of space-based GW detectors~\cite{Amaro-Seoane2007,Amaro-Seoane2012a}.
+We consider the EMRI event rate as the combination of two pieces: the comoving number density of SMBHs in the Universe, which is the same as that of galaxies if we assume all galaxies host a single SMBH~\cite{Lynden-Bell1971,Soltan1982}, and the intrinsic event rate per SMBH $\mathcal{R}$, the number of inspiral events per unit time for a given galaxy.
+
+The comoving number density of galaxies is challenging to estimate because of the effects of local structure in the Universe, the evolution of that structure, and properties of the SMBHs themselves.
+We simplify the problem by assuming a homogeneous distribution that does not evolve with redshift, which is reasonable for the typical scales considered by GW detectors.
+We also neglect correlations between the SMBH mass and spin~\cite{Volonteri2010,Dotti2013,Sesana2014}, and impose a power-law scaling relation for the comoving number density
+\begin{equation}
+\diff{n}{\,\ln M} = n_0 \left(\frac{M}{M_0}\right)^\beta.
+\end{equation}
+There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; we use $\beta = 0$ and $n_0 = 0.002~\mathrm{Mpc}^{-3}$ for SMBHs with $M < \order{10^7 M_\odot}$~\cite{Greene2007, Gair2010b}.
+
+Simple estimates of the intrinsic rate $\mathcal{R}$ have been carried out using Monte-Carlo methods to count the number of compact objects from isothermal distributions that spiral in to a SMBH without plunging~\cite{Merritt2013}.
 The result is a scaling law for each species of compact object of the form
@@
 where $M$ is the mass of the SMBH and $M_0 = 3\times 10^6 M_\odot$ is a fiducial mass.
-\Citet{Hopman2009a} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, neutron stars and respectively, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}~\mathrm{Gyr}^{-1}$ for each component, again demonstrating that BHs dominate the event rate.
-This neglects the effects of resonant relaxation~\cite{Rauch1996,Merritt2011}, ignores the spin of the SMBH~\cite{Amaro-Seoane2012b}, and assumes that the $M$--$\sigma$ relation holds for all SMBH masses.
-Each of these is likely to significantly impact the event rate, but \eqnref{EMRI-intrinsic-rate} can still be used as a rough guide to the expected number of events.
-
-\Citet{Amaro-Seoane2011d} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker--Planck description for the bulk properties of the stellar distribution.
-They found a better fit for the power-law spectral index for BHs of $\alpha = -0.19$.
-
-The comoving number density of galaxies is challenging to estimate due to local structure in the Universe, the evolution of that structure, and properties of the SMBHs themselves.
-We simplify the problem by assuming a homogeneous distribution that does not evolve with redshift, which is reasonable for the typical scales considered by GW detectors.
-We also neglect correlations between the SMBH mass and spin~\cite{Volonteri2010}, and impose a power-law scaling relation for the comoving number density
-\begin{equation}
-\diff{n}{\,\ln M} = n_0 \left(\frac{M}{M_0}\right)^\beta.
-\end{equation}
-There is significant uncertainty in the SMBH mass function, but this simple functional form is found to be in good agreement with observations from the Sloan Digital Sky Survey for the mass range of interest; it is sufficient to set $\beta = 0$ and $n_0 = 0.002~\mathrm{Mpc}^{-3}$ for SMBHs with $M < \order{10^7 M_\odot}$ \citep{Greene2007, Gair2010b}.
+Hopman~\cite{Hopman2009a} finds that $\alpha = \{-0.15,\,-0.25,\,-0.25\}$ for BHs, neutron stars and respectively, with respective event rates $\mathcal{R}_0 = \{400,\, 7,\, 20\}~\mathrm{Gyr}^{-1}$ for each component, showing how BHs dominate the event rate.
+Amaro-Seoane and Preto~\cite{Amaro-Seoane2011d} studied the effects of mass segregation on the intrinsic EMRI event rate, using direct-summation $N$-body simulations to calibrate a Fokker--Planck description for the bulk properties of the stellar distribution.
+They found a better fit for the power-law spectral index for BHs of $\alpha = -0.19$, which we use here.
+The simple power-law description does not incorporate the effects of either resonant relaxation~\cite{Rauch1996,Rauch1998,Merritt2011} or anomalous relaxation~\cite{Hamers2014,Merritt2015c}, ignores the spin of the SMBH~\cite{Amaro-Seoane2012b}, and assumes that the $M$--$\sigma$~\cite{Ferrarese2000,Graham2016} relation holds for all SMBH masses (cf.\ \cite{Hlavacek-Larrondo2012,Ferre-Mateu2015,VanLoon2015,Trakhtenbrot2015,King2016}).
+Each of these is likely to  impact the event rate, but \eqnref{EMRI-intrinsic-rate} can still be used as a rough guide to the expected number of events.
 
 Combining the intrinsic event rate with the comoving number density, the mass of the SMBH for the EMRI population is then sampled from a power-law with a probability distribution function $f(M) \propto M^{\alpha+\beta-1}$.
-The dimensionless spin $a_\ast$ is uniformly between its limiting values, $0$ and $1$.
-The direction of the spin of the SMBH, the direction of the angular momentum of the orbit and the direction of the source on the sky are all chosen uniformly on the sphere.
-The angles specifying the location of pericenter and the compact object around the orbit are distributed uniformly between $0$ and $2\pi$.
-We sample redshift of the source uniformly from comoving volume.
-
+
+We distribute the dimensionless spin $a_\ast$ of the SMBH is uniformly between its limiting values of $0$ and $1$.
+X-ray measurements show that SMBH spins can take a range of values~\cite{Reynolds2013a,Patrick2012,Walton2013,BerryThesis2013}; there is an observed preference for larger ($> 0.9$) spin values, but this may be a selection effect~\cite{Brenneman2011}.
+Therefore, the uniform prior is a safe choice given our ignorance of the true distribution.
+The direction of the spin is uniformly distributed across the surface of a sphere.
+
+We distribute SMBHs uniformly throughout the Universe.
+We sample redshift of the source uniformly from comoving volume out to maximum redshift of $z\sub{max} = 1.5$, beyond which we cannot detect EMRIs.
+Sources are uniformly distributed across the sky.
+
+To describe the orbit we need to specify the inclination distribution, the eccentricity distribution and the orbital phase at periapsis.
+The inclination is uniformly distributed across all orientations (uniform in $\cos \iota$); the poloidal and azimuthal phases at periapsis are uniform between $0$ and $2\pi$.
 The eccentricity distribution is more complicated.
+
 Eccentricities for EMRIs are uncertain, and depend strongly on the formation scenario being considered.
-Here, we adopt a fit to a distribution computed using Monte-Carlo simulations by \citet{Hopman2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH.
+Here, we adopt a fit to a distribution computed using Monte-Carlo simulations by Hopman and Alexander~\cite{Hopman2005}, who model the scattering process of compact objects onto inspiral orbits around a $3 \times 10^6 M_\odot$ Schwarzschild BH.
 We assume this can be extended to provide a rough estimate of the distribution around SMBHs of other masses and spins.
 At the point in the inspiral when the orbital period takes a fiducial value $T_0 = 10^4~\mathrm{s}$, we find that the Monte-Carlo eccentricity probability distribution function is well-described by a power-law with an exponential cutoff
-\begin{widetext}\begin{equation}
+%\begin{widetext}
+\begin{equation}
 \label{eq:EMRI-e-distribution}
@@
 	\begin{cases}
-		\left(e\sub{m}-e\right)^{b(e\sub{m}-e\sub{p})} \exp\left\{b(e-e\sub{m})\right\} & 0 \leq e \leq e\sub{m}\\
-		0 &\mathrm{Otherwise}
+		\left(e\sub{m}-e\right)^{b(e\sub{m}-e\sub{p})} \exp\left[b(e-e\sub{m})\right] & 0 \leq e \leq e\sub{m}\\
+		0 & \mathrm{Otherwise}
 	\end{cases},
-\end{equation}\end{widetext}
-where $e\sub{m} = 0.81$ is the maximum observed eccentricity, $e\sub{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index.
-
-We are interested in EMRI systems close to plunge, as it is here that the GW amplitude is largest.
-To create our EMRI population, we sample the eccentricity from the distribution given by \eqnref{EMRI-e-distribution}, set the orbital period equal to the fiducial value $T_0$, and evolve the system using the analytic kludge (AK) prescription of Barack and Cutler \cite{Barack2004} until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r)=0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
-This ignores the (small) influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
-
-The size of the required population can be estimated by evaluating an integral of the form \citep{Gair2009}
+\end{equation}
+%\end{widetext}
+where $e\sub{m} = 0.81$ is the maximum observed eccentricity, $e\sub{p} = 0.69$ is the peak of the distribution, and $b=11$ is the exponential index~\cite{ColeThesis2015}.
+The mean eccentricity at this period is $0.60$, slightly below that expected for a thermal distribution.
+Significant eccentricity is retained as EMRIs enter the eLISA frequency band~\cite{Merritt2015c}.
+
+To evolve the orbits, we start with orbital periods of $T_0$ and then use the analytic kludge (AK) prescription of Barack and Cutler~\cite{Barack2004}.
+This is similar to the NK approach, but uses a series of Keplerian ellipses rather than Kerr geodesics.
+Evolution of the orbit includes the effects of periapse precession, Lense--Thirring
+precession, and radiation reaction calculated using PN expressions.
+AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
+We follow inspirals until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r) = 0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
+This ignores the (small) influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for a mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
+
+\subsubsection{Population results}
+
+The estimated size of the can be by evaluating from an integral of the form~\cite{Gair2009}
 \begin{equation}
@@
 where $V\sub{c}(z)$ is the comoving volume at redshift $z$, and limits $z\sub{max} = 1.5$, $M\sub{min} = 10^4 M_\odot$ and $M\sub{max} = 10^7 M_\odot$ are chosen to encompass the range of detectable signals.
-For a mission lifetime of $t\sub{life} = 2\units{yr}$, the integral gives a total of $6333$ EMRI events.
+For a mission lifetime of $t\sub{life} = 2\units{yr}$, the integral gives a total of $6333$ EMRI systems.
 This is a lower bound for $N\sub{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.
@@
 A given EMRI is classified as observable if its SNR exceeds some threshold value $\rho\sub{thres}$.
-Calculating SNRs from the generated AK waveforms, assuming $6$ laser links and using the eLISA PSD, we find $513$ detectable events for $\rho\sub{thres} = 15$.
-The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters, the redshift of the source, and the length of the observation $t\sub{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than 15, which are also shown.
+Calculating SNRs from the generated AK waveforms, assuming $6$ laser links and using the eLISA PSD\cite{Amaro-Seoane2012a}, we find $513$ detectable events across the mission for $\rho\sub{thres} = 15$ (cf.\ \cite{Gair2004,Amaro-Seoane2012a,Mapelli2012}).
+The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters at plunge, the redshift of the source, and the length of the observation $t\sub{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than $15$, which are also shown.
 
@@
 
-Eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs, due to the fact that the periapse in such systems can get much closer to the SMBH, and so the compact object experiences the strongest gravitational fields.
-This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in the distributions.
-Systems at higher redshifts start to become undetectable because the GW amplitude scales inversely to the luminosity distance; by $z = 1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
 The mass distribution for detectable EMRIs is peaked such that the typical GW frequency occurs at the base of the eLISA noise PSD.
-
-EMRI systems within our population tend to have small eccentricities at plunge due to the circularising effect of GWs~\cite{Peters1964}: for detectable systems, the mean eccentricity is $0.05$ and the maximum is $0.4$, $85\%$ have $e < 0.1$.
+Systems at higher redshifts start to tail off because the GW amplitude scales inversely to the luminosity distance; by $z = 1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
+Eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs because the periapsis in such systems can get much closer to the SMBH, and so the GWs are intrinsically loader.
+This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in its distribution.
+
+EMRI systems within our populations have small eccentricities at plunge than initially due to the circularising effect of GWs~\cite{Peters1964}.
+For detectable systems, the mean eccentricity is $0.05$ and the maximum is $0.4$, $85\%$ have $e < 0.1$.
 In \secref{flux-enhance}, we found a strong eccentricity dependence on the magnitude of the resonant flux enhancements.
-We therefore expect typical resonant jumps in these astrophysical systems to be much less than $1\%$, and so the resultant dephasing may be relatively weak.
-We now analyse these $513$ systems using our full and adiabatic self-force models.
+We therefore expect typical resonant jumps in these astrophysical systems to be much less than $1\%$, and so the resultant dephasing to be relatively weak.
+We now analyse the $513$ systems using our NK models to check for the impact of resonances.
 
@@
 
-After computing the SNRs of our NK waveforms for each of the $513$ EMRI systems, we can compare the values to those obtained using the AK formalism.
-The different approximation schemes are expected to produce different results, but they should be broadly similar.
-In \figref{pop-SNR-vs-eta}, we plot the ratio of the new NK SNR to the AK SNR, as a function of the mass ratio.
-We observe a clear trend, with systems closest to equal mass performing roughly as expected, but with the most extreme systems giving significantly lower SNRs.
-This is a consequence of the PN self-force model we are using, which we have found overestimates the inspiral rate for EMRIs.
-Detectable systems with larger mass ratios (closer to equal mass) plunge at higher frequencies, and so during the $2\units{yr}$ observation window, they evolve within the bucket of the eLISA noise curve.
-Increasing the rate of evolution then does not significantly change the recovered SNR since we are still sensitive to all parts of the waveform.
-At smaller mass ratios, the systems evolve more slowly, from lower plunge frequencies.
-An increased evolution rate means that earlier portions of the EMRI are shifted outside the sensitive region of the detector, thus giving a much lower SNR.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46\textwidth]{pop_SNR_vs_eta}
-\caption{\label{fig:pop-SNR-vs-eta}The ratio of SNRs as computed with waveforms generated using the NK (adopting our PN self-force model) and AK formalisms, as a function of (the base-10 logarithm of) the mass ratio.
-Each system has an AK SNR greater than $15$, using the eLISA detector with 6 laser links.}
-\end{figure}
-
-We can study the effect of resonances, by comparing the maximum adiabatic SNR to the full SNR computed using the NK waveforms.
-The resulting overlap is still indicative of the effect of resonances on the population of systems.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_Tad}
-%\caption{\label{fig:pop-adSNR-vs-Tad}The maximum overlap for $513$ detectable EMRI systems obtained by comparing the instantaneous evolution with each of the adiabatic models in our one-parameter family.
-Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t\sub{insp}$.
-The overlap is plotted against the largest time spent by the inspiral without encountering any resonances $t\sub{ad}$, normalised by $t\sub{insp}$.
-The solid line corresponds to the expected value if the $1$:$2$ and $2$:$3$ resonances cause significant waveform dephasing.}
-%\end{figure}
-
+We can study the effect of resonances, by comparing the adiabatic evolution to the full instantaneous evolution; a loss in SNR will reduce the number of detected events.
 For each inspiral, we denote the longest period of time $t\sub{ad}$ in which neither the $1$:$2$ nor the $2$:$3$ resonance is encountered.
@@
 This might be caused by higher-order resonances disrupting the evolution, in which case $t\sub{ad}$ should be smaller and the systems would lie on the expectation line.
-%Alternatively, the low-order resonances might be long-lived, explaining why the recovered overlap is smaller than expected.
-However, a more likely explanation is that larger adiabatic overlaps are possible with models that we have not considered here (either with different values of $t\sub{match}$ or outside our one-parameter family altogether).
+However, a more likely explanation is that larger a sub-optimal matching time $t\sub{match}$ was used, and a larger overlap is achievable with a different choice of adiabatic evolution.
 
@@
 However, the vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity.
-For the smallest mass ratios, the inspiral rate is slow, and so the systems do not encounter either the $1$:$2$ or $2$:$3$ resonances during their lifetime.
-Meanwhile for the largest mass ratios, the EMRIs encounter both resonances very close to plunge.
+For the smallest (most extreme) mass ratios, the inspiral rate is slow, and so the systems do not encounter either the $1$:$2$ or $2$:$3$ resonances during their lifetime.
+Meanwhile, for the largest (least extreme) mass ratios, the EMRIs encounter both resonances close to plunge.
 In each case, there is a long resonance-free region, allowing a high overlap to be recovered.
 
-The remaining $66\%$ of systems have overlaps above the level expected if resonances were to play an important role.
+The remaining $66\%$ of systems have overlaps above the level expected if resonances lead to significant dephasing.
 These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t\sub{ad}$ is small.
@@
 
-Even if we assume that all overlap reductions are due to resonances (instead of due to our adiabatic family not being extensive enough) the overall effect on the population is not significant.
-To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to take into account the resonance behaviour.
-The total number of detectable systems using a threshold SNR of $15$ decreases from $513$ to $492$, a loss of $4\%$.
-If we increase the threshold, the fractional reduction in the number of detectable systems gets even smaller.
+Even assuming that all overlap reductions are due to transient resonances (neglecting contributions from imperfectly selected adiabatic waveforms), the overall effect on the population is not significant.
+To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to account for the loss in SNR caused by passing throungh resonance.
+The total number of detectable systems decreases from $513$ to $492$, a loss of $4\%$.
+If we increase the threshold $\rho\sub{thres}$, the fractional reduction in the number of detectable systems gets even smaller.
 We therefore conclude that resonances do not cause sufficient waveform dephasing across a population of EMRIs for the detection rate to be appreciably diminished.
@@
 
-\section{Concluding remarks}
+\section{Conclusion}
 \label{sec:conclusion}
 
-\emph{Need to reconclude}
-
 Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral.
-We have presented an evolution scheme appropriate for studying the effects of resonances on inspirals, which has been applied to an astrophysically realistic population.
-The location and duration of resonances, as well as their effect on orbital parameters, can all be calculated.
+Passing through resonance can lead to an enhancement or decrement in the radiated fluxes associated with orbital evolution, which in turn leads to a jump in the orbital parameters across the resonance.
+Including the signature of resonances is necessary for optimal analysis on EMRI signals.
+
+The duration of resonances and their effect on orbital parameters, can be calculated.
+However, this requires a self-force model.
+We have made use of a low spin, first-order PN self force.
+This is of limited validity and so the results should be taken as qualitative estimates; however, the self-force model should suffice for illustrating the potential effects of resonances.
+
+The resonant jump in the orbital parameters depends upon the orbital phase on resonance.
+This makes it difficult to predict, without detailed calculation, the evolution of an inspiral.
+The magnitude of the jump depends upon the order of resonance and the orbital eccentricity.
+The $2$:$3$ and $1$:$2$ resonances have the largest effects, and higher order resonance are less important.
+Crucially, jumps are smaller for lower eccentricity orbits.
+High eccentricity EMRIs may encounter jumps in their orbital parameters of a few percent, leading to rapid dephasing of their waveforms compared to those from an adiabatic evolution.
 
 Amongst a population of sources, unmodelled resonances could diminish detection prospects.
-However, due to the low eccentricity of detectable sources, the overall effect on SNR recovery is small; it is therefore likely to be sufficient to use adiabatic models to detect EMRIs.
-An unresolved question here is how using these adiabatic models would affect parameter estimation: systematic biases may be introduced if we neglect the presence of important resonances.
-One method of providing an answer to this problem is to perform an MCMC parameter estimation study using our adiabatic waveforms as templates; we leave this to future work.
-
-High eccentricity EMRIs may be possible, in which case transient resonances will play an important role in their detectability; systems that undergo a jump of a few percent dephase very rapidly from the adiabatic approximation.
-In such scenarios, adiabatic models can still be used, but there must be sufficient SNR in each non-resonant region that they can be detected as individual sources and then identified as arising from the same inspiral.
-If this is not possible, \eqnref{delta-I-a} can be used to join together different adiabatic models, if the phase on resonance and the magnitude of the self-force can be predicted sufficiently accurately.
-Alternatively, the jump $\Delta\mathcal{I}\sub{jump}$ for each orbital parameter could be included as a free parameter in the template space, and then searched over.
+However, because of the circularising effects of GW emission, by the time that the most important resonances are encountered, the orbital eccentricity is low.
+Therefore, the overall effect on SNR recovery is small, and there is not a significant reduction in the number of detectable EMRIs.
+
+While it may not be essential to model resonances to detect (at least a subpopulation of) EMRIs, an unresolved question here is how resonances would affect parameter estimation.
+Systematic biases may be introduced if inaccurate templates are used; equally, the distinctive features of resonances may allow more precise measurements to be made.
+Developing a more accurate self-force model is required for a complete quantitative understanding of the effects of transient resonances.
+
+Adiabatic waveform models can still be used for EMRIs away from resonance.
+Therefore it may be possible to stitch together waveforms from a sequence of adiabatic evolutions, if the phase on resonance and the magnitude of the self-force can be predicted with sufficient accurately.
 Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.
-
-This work is highly dependent on the chosen self-force model, and so the results should be taken as qualitative estimates, given the unsuitability of a PN self-force to the problem of fast motion near to highly-spinning BHs.
-In the future, with a more accurate self-force model, more exact quantitative results can be derived.
 
@@
 The authors extend our sincere gratitude to \'{E}anna Flanagan and Tanja Hinderer for providing their PN self-force code, without which this work would not be possible.
-We also wish to thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent and Leor Barack for useful conversations.
-RHC is supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC was supported by a Marie Curie Fellowship, and JRG is supported by the Royal Society.
+We thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent, Leor Barack, Scott Hughes and Nico Yunes for useful conversations; Stanislav Babak for examining both \cite{BerryThesis2013} and \cite{ColeThesis2015}, and Christopher Moore with help preparing this manuscript.
+RHC was supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC was supported by a Marie Curie Fellowship, and JRG was supported by the Royal Society.
 \end{acknowledgments}

### 2016-08-15 res_char.tex
New figures

@@
 
-\author{Robert H. Cole}
-\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
 \author{Christopher P.L. Berry}
@@
 \affiliation{School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
+\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
+\author{Robert H. Cole}
 \affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
@@
 The inspiral of stellar-mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
-In order to detect and analyse these signals accurate waveform templates are required.
+In order to detect and analyse these signals, accurate waveform templates are required.
 For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order dissipative gravitational self-force.
-Other effects are, however, neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+Other effects are, neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
 During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the emitted radiation.
 We investigate the properties of transient resonances.
-We have present an evolution scheme for studying the effects of resonances and applied to an astrophysically realistic population.
+We present an evolution scheme for studying the effects of resonances and apply this to an astrophysically motivatedc population.
 We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence the low eccentricity of the inspirals.
@@
 
-Systems of unequal masses are more challenging to evolve numerically as they complete a larger number of orbits and it is necessary to resolve two different scales.
+Systems of unequal masses are more challenging to evolve numerically as they complete a larger number of orbits, and it is necessary to resolve two different scales.
 Calculations can instead be performed perturbatively.
-The paradigm unequal-mass system has a stellar-mass BH orbiting a supermassive BH (SMBH), such as those expected to be found at the centres of galaxies.
+The paradigm unequal-mass system has a stellar-mass BH orbiting a supermassive BH (SMBH), such as those expected to be found at the centres of galaxies~\cite{Kormendy1995,Ferrarese2005,Boehle2016}.
 These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors like the evolving Laser Interferometer Space Antenna (eLISA)~\cite{Amaro-Seoane2007,Amaro-Seoane2012a}.
@@
 In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}.
-In the test particle limit, the smaller body follows an exact geodesic of the SMBH's spacetime.
+In the test-particle limit, the smaller body follows an exact geodesic of the SMBH's spacetime.
 Including the effects of the smaller body's finite mass, the background spacetime is perturbed.
@@
 Geodesic orbits in GR have three associated frequencies: the radial frequency $\Omega_r$, the polar frequency $\Omega_\theta$ and the azimuthal frequency $\Omega_\phi$.\footnote{In the strong-field regime, it is possible to have isofrequency pairings, where two different orbits share the same orbital frequencies~\cite{Warburton2013}.
-The evolution of the frequencies will still differ, such that orbital trajectories can be reconstructed.} The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
+The evolution of the frequencies still differ, such that orbital trajectories can be reconstructed.} The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
 % Section 10.6
 In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field regime $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
-For extreme-mass-ratio systems, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
+For EMRIs, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
 The evolution of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
@@
 We use geometric units with $G = c = 1$ throughout.
-We always use $M$ for the mass of the central SMBH and $a$ as the spin parameter.
+We always use $M$ for the mass of the central SMBH and $a$ as its Kerr spin parameter.
 We also use the dimensionless spin $a_\ast \equiv a/M$; we take the convention that $0 \leq a_\ast < 1$.
@@
 Instantaneously, the motion of the orbiting mass can be described as geodesic, with the integrals of the motion changing on timescales of many orbital periods.
-It is therefore necessary to develop an understanding of the Kerr geodesics (\secref{geodesic}).
+It is therefore necessary to develop an understanding of the Kerr geodesics (\secref{geodesic}; those familiar with calculating orbits in Kerr may skip this section.).
 Transient resonance occur when the radial and polar frequencies become commensurate (\secref{frequencies}); we analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance (\secref{forced-motion}).
@@
 Central to understanding transient resonances in a knowledge of orbits in Kerr spacetime, and hence we begin with details of evolving Kerr geodesics.
-Those familiar with calculating orbits in Kerr may skip this section.
 The geodesic equations may be written as~\cite{Carter1968, Chandrasekhar1992} % section 62
@@
 %After tapering the projected waveforms, we then calculate overlaps using \eqnref{innerprod} but with discrete Fourier transforms and the integral over frequency replaced with an appropriate sum over Fourier modes.
-} We use the analytic approximation for eLISA of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}.
+}
+
+We use the analytic approximation for the eLISA PSD of Amaro-Seoane {\it{et al}}.~\cite{Amaro-Seoane2012a}.
 Following the success for the LISA Pathfinder mission~\cite{Armano2016}, this sensitivity should be achievable in a future mission.
@@
 Example evolutions of the orbital parameters $E$, $L_z$ and $Q$ are shown in \figref{good-traj}, which shows that the adiabatic evolution closely matches the full evolution on longer inspiral timescales.
-The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the CO; the $2$-torus averaging explicitly smooths out the visible structure on this scale.
+The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the compact object; the $2$-torus averaging explicitly smooths out the visible structure on this scale.
 The two appraches are in good agreement.
@@
 Specifically, we choose the initial conditions to be the same as in \secref{nonres}, but with an initial semilatus rectum $p_0 = 7.85$.
-This system passes through the $2$:$3$ resonance.
-The effect of passing through resonance is to cause a shift in the orbital parameters (and hence in the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
+This system passes through the $2$:$3$ resonance; the effect is to cause a shift in the orbital parameters (and hence the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
 
@@
 \centering
-\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
+%\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
+\includegraphics[width=0.46\textwidth]{Fig_overlap}
 \caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
-The time axis is centred on the value at which the full evolution crosses the $2$:$3$ resonance.}
+The time $t\sub{res} = 0$ is when the full evolution crosses the $2$:$3$ resonance.}
 \end{figure}
@@
 We have seen that overlaps between adiabatic and full instantaneous evolutions dephase following a resonance.
-However, this does not necessarily mean that no adiabatic evolution will have a low overlap with the observed signal.
+However, this does not necessarily mean that no adiabatic evolution has a low overlap with the observed signal.
 It is possible that a difference between the instantaneous and adiabatic waveforms could be ameliorated by changing the parameters of the template $\boldsymbol{\Theta}$.
@@
 The large parameter space of adiabatic waveforms, coupled with the expensive nature of our $2$-torus averaging routine, renders a brute-force approach prohibitively expensive.
-For this preliminary investigation, we focus on a small subset of parameters that we suspect will produce a large overlap, and make the assumption that a good adiabatic model will \emph{exactly} match the full model at some time $t\sub{match}$.
+For this preliminary investigation, we focus on a small subset of parameters that we suspect will produce a large overlap, and make the assumption that a good adiabatic model \emph{exactly} matches the full model at some time $t\sub{match}$.
 This reduces the search to a $1$-parameter family of waveforms that can easily be computed concurrently with the full evolution.
@@
 In this case, the best-performing adiabatic model was that which matched at the end, giving $\mathbb{O} = 0.677$, while the model that matches at the start gives only $\mathbb{O} = 0.207$.
-These are similar to the adiabatic matched before and after resonance respectively.
+These are similar overlaps from to the adiabatic evolutions matched before and after resonance respectively.
 These values can be explained qualitatively by the relative lengths of the adiabatic-like regions on either side of the resonance in the full evolution ($69.8\%$ of the inspiral is post-resonance): there is not an exact equivalence because of the frequency dependence of the PSD.
@@
 This can be illustrated using an ensemble of orbits with different orbital phases $\psi_{\theta_-}$.
-According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.
-Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
+According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$, as expected from matching the post-resonance region of the the waveform.} Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
 We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta L_{z\,\mathrm{jump}}$ ellipse.
-The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset~\cite{Flanagan2012a}; this means that for every value of $q$, there is always a resonance jump in at least one parameter.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$ as expected from matching the post-resonance region of the the waveform.}
+The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset~\cite{Flanagan2012a}; this means that for every value of $q$, there is always a resonance jump in at least one parameter.
 
@@
 \centering
-\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
+%\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
+\includegraphics[width=0.46\textwidth]{Fig_q_jump}
 \caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$ (cf.\ figure 3 of \cite{Flanagan2012a}).
-The jump is expressed as a percentage of the adiabatic change in each parameter across resonance $\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}$.
-The individual jumps as well as a sinusoidal fit are plotted for the orbital parameters (solid) and the three orbital frequencies (dotted).}
+The jump is relative to the adiabatic change in each parameter across resonance.
+The individual jumps as well as a sinusoidal fit are plotted for the energy $E$, angular momentum $L_z$ and Carter constant $Q$.}
 \end{figure}
@@
 We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
-In \figref{res-flux-rms-p}, we show the root-mean-square (over $a_\ast$ and $\cos\iota$) relative resonant jump for the semilatus rectum as a function of eccentricity for a selection of low-order resonances.
+In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum as a function of eccentricity for a selection of low-order resonances.
 The other orbital shape parameters show a similar trend with increasing eccentircity~\cite{ColeThesis2015}.
@@
 \centering
-\includegraphics[width=0.46\textwidth]{res_flux_rms_p}
-\caption{\label{fig:res-flux-rms-p}Relative flux enhancements of $p$, as a function of $e$, marginalised over $a$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
-Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use solid (short-dashed; long-dashed) lines.
+%\includegraphics[width=0.46\textwidth]{res_flux_rms_p}
+\includegraphics[width=0.46\textwidth]{Fig_e_rms_p}
+\caption{\label{fig:res-flux-rms-p}Relative flux enhancements of $p$, as a function of $e$, marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
+Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red).
 The $2$:$3$ resonance has the largest relative flux enhancement.}
@@
 \centering
-\includegraphics[width=0.46\textwidth]{res_flux_rms_p_e095}
-\caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements of $p$ marginalised over $a$ and $\cos\iota$ by taking the root-mean-square.
+%\includegraphics[width=0.46\textwidth]{res_flux_rms_p_e095}
+\includegraphics[width=0.46\textwidth]{Fig_nu_rms_p}
+\caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements of $p$ marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square.
 Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use circular (square; triangular) points.}
@@
 Therefore, the uniform prior is a safe choice given our ignorance of the true distribution.
-The direction of the spin is uniformly distributed across the surface of a sphere.
+The direction of the spin is uniformly distributed across the surface of the unit sphere.
 
@@
 
-To describe the orbit we need to specify the inclination distribution, the eccentricity distribution and the orbital phase at periapsis.
+To describe the orbit we need to specify the inclination distribution, the eccentricity distribution and the   phase at periapsis.
 The inclination is uniformly distributed across all orientations (uniform in $\cos \iota$); the poloidal and azimuthal phases at periapsis are uniform between $0$ and $2\pi$.

### 2016-08-15 res_char.tex
Relaitve flux

@@
 
-%The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems.
-Instead, we can express the resonance jump as a fraction of the adiabatic change expected across the resonance.
-For the orbital parameter $\mathcal{I}$,
-%\begin{equation}
-%\label{eq:res-jump-ratio}
-%\frac{\Delta \mathcal{I}\sub{jump}}{\mathcal{I}\sub{ad}(t>t\sub{res})-\mathcal{I}\sub{ad}(t<t\sub{res})} = \frac{\Delta \mathcal{I}\sub{jump}}{\dot{\mathcal{I}}\sub{ad}(t\sub{res})\tau\sub{res}}.
-%\end{equation}
+The absolute size of the resonance jump is often not particularly useful, especially when comparing different systems. Instead, we use the fractional enhancement relative to the adiabatic evolution,
+\begin{equation}
+\label{eq:res-jump-ratio}
+\delta \mathcal{I}^a = \frac{\Delta \mathcal{I}^a\sub{jump}}{\dot{\mathcal{I}}^a\sub{ad}(t\sub{res})\tau\sub{res}}.
+\end{equation}
+Relative enhancements may be of the order of a few percent.
 
@@
 \includegraphics[width=0.46\textwidth]{Fig_e_rms_p}
-\caption{\label{fig:res-flux-rms-p}Relative flux enhancements of $p$, as a function of $e$, marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
+\caption{\label{fig:res-flux-rms-p}Relative flux enhancements for the semilatus rectum $p$, as a function of $e$, marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
 Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red).
@@
 \includegraphics[width=0.46\textwidth]{Fig_nu_rms_p}
-\caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements of $p$ marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square.
+\caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements for the semilatus rectum $p$ marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square.
 Resonances with $n_\theta = 1$ ($2$; $3$) are coloured blue (gold; red) and use circular (square; triangular) points.}

### 2016-08-15 res_char.tex
RevTeX 4.1

@@
-\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,showpacs,superscriptaddress,twocolumn]{revtex4}
+\documentclass[aps,prd,amsfonts,amssymb,amsmath,nofootinbib,showpacs,superscriptaddress,twocolumn,floatfix]{revtex4-1}
 \usepackage{graphicx}
@@
 \usepackage{dcolumn}
-\usepackage{hyperref}
+\usepackage[breaklinks=true]{hyperref}
 \usepackage{textcomp}
@@
 A given EMRI is classified as observable if its SNR exceeds some threshold value $\rho\sub{thres}$.
-Calculating SNRs from the generated AK waveforms, assuming $6$ laser links and using the eLISA PSD\cite{Amaro-Seoane2012a}, we find $513$ detectable events across the mission for $\rho\sub{thres} = 15$ (cf.\ \cite{Gair2004,Amaro-Seoane2012a,Mapelli2012}).
+Calculating SNRs from the generated AK waveforms, assuming $6$ laser links and using the eLISA PSD~\cite{Amaro-Seoane2012a}, we find $513$ detectable events across the mission for $\rho\sub{thres} = 15$ (cf.\ \cite{Gair2004,Amaro-Seoane2012a,Mapelli2012}).
 The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters at plunge, the redshift of the source, and the length of the observation $t\sub{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than $15$, which are also shown.

### 2016-08-16 asymptotic.tex
RevTeX 4.1
Typos

@@
 The following derivation is reproduced from Berry~\cite{BerryThesis2013}.
-It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two timescale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found in van de Meent~\cite{VanDeMeent2013}.
+It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two-timescale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found in van de Meent~\cite{VanDeMeent2013}.
 
@@
 To rewrite the pre-resonance solution, we begin with the fast phase parameter
+\pagebreak
 \begin{equation}

### 2016-08-16 res_char.tex
Typos

@@
 For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order dissipative gravitational self-force.
-Other effects are, neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+Other effects are neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
 During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the emitted radiation.
 We investigate the properties of transient resonances.
-We present an evolution scheme for studying the effects of resonances and apply this to an astrophysically motivatedc population.
-We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence the low eccentricity of the inspirals.
+We present an evolution scheme for studying the effects of resonances and apply this to an astrophysically motivated population.
+We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence of the low eccentricity of the inspirals.
 Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals.
@@
 In this case the trajectory does not ergodically fill the $2$-torus, but instead traces out a $1$-dimensional subspace.\footnote{For illustrations, see Grossman, Levin and Perez-Giz~\cite{Grossman2012}.} There are then contributions to the self-force that no longer average out beyond $\langle G_A^{(1)}\rangle_{q_r,\,q_\theta}$.
-Intuitively, we expect that this effect is more important for ratios of small integers as when the integers are large, the orbit comes close to all points on the $2$-torus.
+Intuitively, we expect that this effect is more important for ratios of small integers since when the integers are large the orbit comes close to all points on the $2$-torus.
 
@@
 
-Central to understanding transient resonances in a knowledge of orbits in Kerr spacetime, and hence we begin with details of evolving Kerr geodesics.
+Central to understanding transient resonances is a knowledge of orbits in Kerr spacetime, and hence we begin with details of evolving Kerr geodesics.
 The geodesic equations may be written as~\cite{Carter1968, Chandrasekhar1992} % section 62
@@
 
-\emph{Add detail about why self-force model is rubbish. Comment on switching conservative piece off for population results.}
+%\emph{Add detail about why self-force model is rubbish. Comment on switching conservative piece off for population results.}
 
@@
 
-Knowing where resonances are found in parameter space, how long they last, and how big an effect they are likely to have, enables us to study and interpret the observable effects of resonances on EMRI waveforms (\secref{waveforms}) and the population of observable inspirals (\secref{astrophysics}).
+Knowing where resonances are found in parameter space, how long they last, and how great an effect they are likely to have, enables us to study and interpret the observable effects of resonances on EMRI waveforms (\secref{waveforms}) and the population of observable inspirals (\secref{astrophysics}).
 
@@
 To demonstrate how changing the matching time affects the overlap, we use $5\tau\sub{res}$ after each resonance of interest, namely the low-order $1$:$2$ and $2$:$3$ resonances.\footnote{We adjust the values of $t\sub{match}$ so that they correspond to times of apoapsis.
-This ensures that the adiabatic model intersects with the full instantaneous model close to the centre of it oscillatory envelope, as demonstrated by the inset plots in \figref{good-traj}.
-The generally obtains better matches than if we match close to the extrema of the envelope, which are less representative of the average behaviour.} These matching times lie in a portion of the evolution that is not affected by a resonance, and so should allow for a large overlap with the adiabatic model for that region of the inspiral.
+This ensures that the adiabatic model intersects with the full instantaneous model close to the centre of its oscillatory envelope, as demonstrated by the inset plots in \figref{good-traj}.
+This generally obtains better matches than if we match close to the extrema of the envelope, which are less representative of the average behaviour.} These matching times lie in a portion of the evolution that is not affected by a resonance, and so should allow for a large overlap with the adiabatic model for that region of the inspiral.
 For comparison, we also consider templates that match at the start and end of the evolution, and that match exactly on each of the resonances.
@@
 This can be illustrated using an ensemble of orbits with different orbital phases $\psi_{\theta_-}$.
-According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$, as expected from matching the post-resonance region of the the waveform.} Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
+According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$, as expected from matching the post-resonance region of the waveform.} Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
 We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta L_{z\,\mathrm{jump}}$ ellipse.
@@
 
-We also expect that the effects of passing through resonance will dependent upon the particular resonance.
+We also expect that the effects of passing through resonance depend upon the particular resonance.
 Intuitively, we would expect that when $n_\theta$ and $n_r$ are large, the effects of resonance will be small, since the orbit comes close to all points on the $2$-torus.
@@
 
-We distribute the dimensionless spin $a_\ast$ of the SMBH is uniformly between its limiting values of $0$ and $1$.
+We distribute the SMBH's dimensionless spin $a_\ast$ uniformly between its limiting values of $0$ and $1$.
 X-ray measurements show that SMBH spins can take a range of values~\cite{Reynolds2013a,Patrick2012,Walton2013,BerryThesis2013}; there is an observed preference for larger ($> 0.9$) spin values, but this may be a selection effect~\cite{Brenneman2011}.
@@
 
-The estimated size of the can be by evaluating from an integral of the form~\cite{Gair2009}
+The estimated size of the EMRI population can be by evaluating from~\cite{Gair2009}
 \begin{equation}
@@
 Systems at higher redshifts start to tail off because the GW amplitude scales inversely to the luminosity distance; by $z = 1.5$, the distribution of detectable EMRIs with eLISA has essentially vanished.
-Eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs because the periapsis in such systems can get much closer to the SMBH, and so the GWs are intrinsically loader.
+Eccentric prograde orbits around SMBHs with larger spins tend to produce larger SNRs because the periapsis in such systems can get much closer to the SMBH, and so the GWs are intrinsically louder.
 This effect also causes the detectable EMRIs to have smaller values of $p$ at plunge, as observed in its distribution.
@@
 Even assuming that all overlap reductions are due to transient resonances (neglecting contributions from imperfectly selected adiabatic waveforms), the overall effect on the population is not significant.
-To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to account for the loss in SNR caused by passing throungh resonance.
+To illustrate this, we plot the AK SNRs in \figref{pop-SNR-dist}, multiplied by the maximum recovered overlap to account for the loss in SNR caused by passing through resonance.
 The total number of detectable systems decreases from $513$ to $492$, a loss of $4\%$.
@@
 
-The duration of resonances and their effect on orbital parameters, can be calculated.
+The duration of resonances and their effect on orbital parameters can be calculated.
 However, this requires a self-force model.
@@
 Adiabatic waveform models can still be used for EMRIs away from resonance.
-Therefore it may be possible to stitch together waveforms from a sequence of adiabatic evolutions, if the phase on resonance and the magnitude of the self-force can be predicted with sufficient accurately.
+Therefore it may be possible to stitch together waveforms from a sequence of adiabatic evolutions, if the phase on resonance and the magnitude of the self-force can be predicted with sufficient accuracy.
 Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.

### 2016-08-22 res_char.tex
Replotting figures

@@
 \usepackage{mathrsfs}
+\usepackage[caption=false]{subfig}
 \usepackage{dcolumn}
@@
 %\includegraphics[width=0.46\textwidth]{Fig_overlap_vs_time}
-\includegraphics[width=0.46\textwidth]{Fig_overlap}
+\includegraphics[width=0.4\textwidth]{Fig_overlap}
 \caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
@@
 %\includegraphics[width=0.46\textwidth]{Fig_jump_vs_q}
-\includegraphics[width=0.46\textwidth]{Fig_q_jump}
+\includegraphics[width=0.4\textwidth]{Fig_q_jump}
 \caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$ (cf.\ figure 3 of \cite{Flanagan2012a}).
@@
 %\includegraphics[width=0.46\textwidth]{res_flux_rms_p}
-\includegraphics[width=0.46\textwidth]{Fig_e_rms_p}
+\includegraphics[width=0.4\textwidth]{Fig_e_rms_p}
 \caption{\label{fig:res-flux-rms-p}Relative flux enhancements for the semilatus rectum $p$, as a function of $e$, marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square of a grid of values.
@@
 %\includegraphics[width=0.46\textwidth]{res_flux_rms_p_e095}
-\includegraphics[width=0.46\textwidth]{Fig_nu_rms_p}
+\includegraphics[width=0.4\textwidth]{Fig_nu_rms_p}
 \caption{\label{fig:res-flux-rms-p-vsnu}Relative flux enhancements for the semilatus rectum $p$ marginalised over $a_\ast$ and $\cos\iota$ by taking the root-mean-square.
@@
 where $V\sub{c}(z)$ is the comoving volume at redshift $z$, and limits $z\sub{max} = 1.5$, $M\sub{min} = 10^4 M_\odot$ and $M\sub{max} = 10^7 M_\odot$ are chosen to encompass the range of detectable signals.
-For a mission lifetime of $t\sub{life} = 2\units{yr}$, the integral gives a total of $6333$ EMRI systems.
+For a mission lifetime of $t\sub{life} = 2\units{yr}$, the integral gives a total of $6330$ EMRI systems.
 This is a lower bound for $N\sub{EMRI}$ as we are neglecting EMRIs that merge outside the observation window but nevertheless accumulate sufficient SNR during this time to be observable.
@@
 Calculating SNRs from the generated AK waveforms, assuming $6$ laser links and using the eLISA PSD~\cite{Amaro-Seoane2012a}, we find $513$ detectable events across the mission for $\rho\sub{thres} = 15$ (cf.\ \cite{Gair2004,Amaro-Seoane2012a,Mapelli2012}).
-The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters at plunge, the redshift of the source, and the length of the observation $t\sub{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than $15$, which are also shown.
+The parameter distributions for the mass and spin of the SMBH, the orbital shape parameters at plunge, the redshift of the source, and the length of the observation $t\sub{insp}$ are shown in \figref{EMRIpar-dists}, to be contrasted with the distributions of the $5820$ systems with an SNR less than $15$, which are also shown.%Total = 5819 + 513 = 6332
 
@@
 \centering
-\includegraphics[width=\textwidth]{EMRIpar_dists}
-\caption{\label{fig:EMRIpar-dists}The parameter distributions at plunge for our detectable EMRI systems (dark blue), alongside those of the undetectable systems (light pink).
-In each case, the y-axis values are the probability of a system being found in a particular bin, given that they are either detectable or undetectable.
-The exception to this is in the final plot for the SNR, where we show the number of systems in each bin.
-The SNRs quoted here are calculated using the AK model.}
+%\includegraphics[width=\textwidth]{EMRIpar_dists}
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_M} \quad
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_a} \\ \vspace{0.1cm}
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_z} \quad
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_p} \\ \vspace{0.1cm}
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_e} \quad
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_iota} \\ \vspace{0.1cm}
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_t} \quad
+\includegraphics[width=0.37\textwidth]{Fig_res_pop_SNR}
+\caption{\label{fig:EMRIpar-dists}Parameter distributions at plunge for our detectable EMRI systems (dashed outline), alongside those of the undetectable systems (solid outline).
+For the system parameters, the ordinate-axis values are the probability of a system being found in a particular bin, given that they are either detectable or undetectable.
+In the final plot for the SNR, we show the number of systems in each bin.
+The SNRs quoted here are calculated using the AK model and we assume a detection threshold of $\rho\sub{thres} = 15$.}
 \end{figure*}
@@
 \centering
-\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_eta}
-\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t\sub{ad} / t\sub{insp}$, as a function of the mass ratio, for our population of $513$ EMRI signals.
-Each system encounters either $0$ (blue squares), $1$ (gold triangles) or $2$ (green circles) resonances during the observation window $t\sub{insp}$, with $t\sub{ad}$ the largest time spent by the inspiral without encountering any resonances.}
+%\includegraphics[width=0.46\textwidth]{pop_adSNR_vs_eta}
+\includegraphics[width=0.4\textwidth]{Fig_res_overlap_t}
+\caption{\label{fig:pop-adSNR-vs-eta}The difference between the maximum overlap and the expected value $t\sub{ad} / t\sub{insp}$, as a function of the mass ratio for our population of $513$ EMRI signals.
+Each system encounters either $0$ (circles), $1$ (squares) or $2$ (triangles) resonances during the observation window $t\sub{insp}$, with $t\sub{ad}$ the largest time spent by the inspiral without encountering any resonances.}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.46\textwidth]{pop_SNR_dist}
-\caption{\label{fig:pop-SNR-dist}Histogram showing the probability distribution function for the detectable EMRI SNRs, as calculated using the AK formalism (light pink) and modified by the maximum adiabatic overlap (dark blue).}
+%\includegraphics[width=0.46\textwidth]{pop_SNR_dist}
+\includegraphics[width=0.4\textwidth]{Fig_res_pop_SNR_overlap}
+\caption{\label{fig:pop-SNR-dist}Histogram showing the probability distribution function for the detectable EMRI SNRs, as calculated using the AK formalism (solid outline) and modified by the maximum adiabatic overlap (dashed outline).
+The tail below $\rho\sub{thres} = 15$ indicates the detections lost because of resonances.}
 \end{figure}
@@
 RHC was supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC was supported by a Marie Curie Fellowship, and JRG was supported by the Royal Society.
+This is LIGO Document P1600251.
 \end{acknowledgments}

### 2016-08-25 res_char.tex
afiliaions and grants

@@
 \affiliation{Institute of Mathematics, Astrophysics and Particle Physics, Radboud University, Heyendaalseweg 135, 6525 AJ Nijmegen, The Netherlands}
+\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
 \author{Jonathan R. Gair}
 \affiliation{School of Mathematics, University of Edinburgh, Peter Guthrie Tait Road, Edinburgh EH9 3FD, UK}
+\affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
 
@@
 We thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent, Leor Barack, Scott Hughes and Nico Yunes for useful conversations; Stanislav Babak for examining both \cite{BerryThesis2013} and \cite{ColeThesis2015}, and Christopher Moore with help preparing this manuscript.
-RHC was supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC was supported by a Marie Curie Fellowship, and JRG was supported by the Royal Society.
+RHC was supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC's work was supported by a Marie Curie Intra-European Fellowship within the 7th European Community Framework Programme (PIEF-GA-2011-299190), and JRG was supported by the Royal Society.
 This is LIGO Document P1600251.

### 2016-08-25 res_char.tex
included some suggestions

@@
 
+\newcommand{\pcm}[1]{{\color{blue}~\textsf{[P: #1]}}}
+\newcommand{\cm}[2]{{\color{red}~\textsf{[CITE ME: #2]}}}
+
 \begin{document}
@@
 We present an evolution scheme for studying the effects of resonances and apply this to an astrophysically motivated population.
-We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence of the low eccentricity of the inspirals.
+We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence of the low eccentricity of the inspirals .
 Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals.
+
+\pcm{The inspiral of stellar-mass compact objects, like neutron stars or stellar mass black holes, into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
+In order to detect and analyse these signals, waveform templates describing the adiabatic inspiral of the compact star via self-force effects are required.
+However, other effects are neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the waveform.
+We present an evolution scheme for studying the effects of transient resonances and apply this to an astrophysically motivated population.
+We find that a large proportion of systems encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence of the low eccentricity of the inspirals.
+Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals.
+}
 \end{abstract}
@@
 \section{Introduction}
+\pcm{The paper is very nice and the result is very important bcs it closes and open topic about the possible bias introduced by resonances in EMRI PE.
+General comments: (1)check acronyms along the text, in particular gravitational wave and GW, EMRI, etc....!!, (2) we don't mention anything about the astrophysical implications in the intro.
+It would be nice to include a couple of sentences.
+(3) I'll emphasise more the final result in the conclusions.
+}
 
@@
 
+\pcm{to put things in a historical context, maybe we could include something like this somewhere, : The first detection of GWs represents a paradigm shift in GW astronomy, revolutionising astrophysics and theoretical physics  and changing our understanding of the Universe, perhaps in an even more dramatic way than happened when the X-ray, radio or infrared electromagnetic bands were opened.}
+
 We have made progress in understanding the general relativistic two-body problem in recent years.
@@
 Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}; it is now possible to simulate hundreds of orbits~\cite{Szilagyi2015}.
-These simulations allow us to understand binary BH coalescences.
-Stellar-mass BH mergers are targets for ground-based GW detectors, such as Advanced LIGO~\cite{Aasi2015} and Advanced Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Aso2013}.
+\pcm{what about EOB+NR???
+it's playing an important role...} These simulations allow us to understand binary BH coalescences.
+Stellar-mass BH mergers are targets for ground-based GW detectors, such as Advanced LIGO~\cite{Aasi2015} and Advanced Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Aso2013} \pcm{ET...}.
 The first direct observations of GWs came from the coalescences of two stellar-mass BHs~\cite{Abbott2016,Abbott2016e,Abbott2016d}, and analysis of their properties relied upon our understanding of binary BH waveforms~\cite{Abbott2016f,Abbott2016h,Abbott2016d}.
 %Supermassive BH (SMBH) coalescences, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
+
+\pcm{We have made progress in understanding the general relativistic two-body problem in recent years.
+Bodies of comparable mass can be studied using numerical relativity.
+Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}.
+The length of these waveforms is limited by their high computational cost, and to generate sufficiently many NR waveforms, long enough to be used directly in GW analysis, is still work in progress ~\cite{Szilagyi2015}.
+Reliable low-frequency (of the order v/c, where v is the velocity of the system) inspiral waveforms can be obtained using the post-Newtonian (PN) formalism.
+But PN waveforms are valid for slow motion and hence they are not valid in the strong field regime close to merger.
+For this reason, the effective-one-body (EOB)formalism, a hybrid of PN and NR, is gaining popularity.
+On the one hand, it provides a complete description of the coalescence of compact object binaries and, on the other hand, it is computationally cheaper compared to full NR simulations.
+EOB produces waveforms for the inspiral, merger and ringdown stages using PN and black-hole perturbation theory results.
+Hence, the EOB model reduces to the PN approximation at low frequency.
+In the strong- field regime, however, EOB models are calibrated to existing NR simulations.
+These waveforms allow us to understand binary BH coalescences....}
 
@@
 
-In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}.
+In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}\pcm{need up-to-date references, like Isoyama et al 2014,  Pound 2012, Hughes et al 2015}.
 In the test-particle limit, the smaller body follows an exact geodesic of the SMBH's spacetime.
@@
 The back-reaction from this deformation alters the small body's orbital trajectory, and can be modelled as a self-force that moves the body from its geodesic.
-The self-force is commonly divided into two pieces, dissipative and conservative.
+The self-force is commonly divided into two pieces, dissipative and conservative \pcm{cites???}.
 The former encapsulates the slow decay of the orbital energy and angular momentum, constants of the motion in the test particle limit, through radiation.
@@
 In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field regime $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
-For EMRIs, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
-The evolution of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
-During this evolution, the frequencies may become commensurate.
+For EMRIs, the evolution \pcm{inspiral} timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
+The evolution  \pcm{inspiral}  of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
+During this evolution, the frequencies may become commensurate\pcm{, and resonances occur...}.
 Resonances occur when the radial and polar frequencies are rational multiples of each other:
@@
 Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.
-However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host.
+However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host \pcm{cite??}.
 
 Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
+We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ \pcm{cite} using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
 We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
-These are some combination of the energy per unit mass $E$ and the axial angular momentum per unit mass $L_z$, which arise from isometries of the metric in $t$ and $\phi$, and the Carter constant per unit mass squared $Q$~\cite{Carter1968}, which is related to the separability of the equations of motion in $r$ and $\theta$.
+These are some combination of the energy per unit mass $E$ and the axial angular momentum per unit mass $L_z$ \pcm{of the orbit}, which arise from isometries of the metric in $t$ and $\phi$, and the Carter constant per unit mass squared $Q$~\cite{Carter1968}, which is related to the separability of the equations of motion in $r$ and $\theta$.
 The system evolves following~\cite{Flanagan2012}
@@
 \end{subequations}
-where $\lambda$ is Mino time~\cite{Mino2003}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force.
+where $\lambda$ is Mino time~\cite{Mino2003} \pcm{not everyone is familiar with this term.
+Besides, Mino does't like it.
+We could consider calling it "effective time" or similar}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force \pcm{this implies that there are  second, etc orders, we need to mention it and cite appropriately}.
 By working with $\lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
@@
 
+\pcm{we are including a lot of material that can be found elsewhere and that we are not actually using directly.
+This makes the paper unecessary long.
+We could consider reducing the prevous part or including it in the appendix.}
 One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short timescales, for small $\eta$ there is a small disparity.
@@
 
-To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration.
+To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration \pcm{that deviates the orbit from a geodesic motion}.
 We work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
@@
 Since only the first PN terms are used, this prescription is of limited validity in strong fields.
-Both pieces of the self-force are computed assuming that the spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$.
+Both pieces of the self-force are computed assuming that the \pcm{BH} spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$.
 This is suboptimal for high spins.
@@
 This system passes through the $2$:$3$ resonance; the effect is to cause a shift in the orbital parameters (and hence the fundamental frequencies) that is not replicated by the adiabatic evolution, thus resulting in a rapid dephasing of the waveforms.
+\pcm{perhaps we could mention why we choose this one in particular, i.e.
+it's the strongest one that we found from the start of the evolutions at p =xxx}
 
@@
 The dotted red line indicates the computed size of the jump.
-The time axis is centred on the $2$:$3$ resonance and is scaled by the resonance timescale $\tau\sub{res}$.}
+The time axis is centred on the $2$:$3$ resonance and is scaled by the resonance timescale $\tau\sub{res}$.
+\pcm{plots too small}}
 \end{figure}
@@
 We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
-In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum as a function of eccentricity for a selection of low-order resonances.
+In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum \pcm{p} as a function of eccentricity \pcm{e} for a selection of low-order resonances.
 The other orbital shape parameters show a similar trend with increasing eccentircity~\cite{ColeThesis2015}.

### 2016-08-25 res_char.tex
Adding ref

--- 2016-08-25 res_char.tex
@@
 Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$ \citep{Hinderer2008}.
+We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$~\cite{Hinderer2008}.
 We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
@@
 It is long in all cases we study, $\tau\sub{ev} \sim \order{T/\eta}$.
-It is this property which makes EMRIs interesting, as we can follow the waveform for many cycles, accruing high signal-to-noise ratios.
+It is this property which makes EMRIs interesting, as we can follow the waveform for many cycles, accruing high signal-to-noise ratios (SNRs).
 This is also what allows us to use the adiabatic prescription, as it means the trajectory moves slowly through different orbital parameters.
@@
 
-We also use the resonance timescale
+We also use the resonance timescale (cf.\ \cite{Ruangsri2014})
 \begin{equation}
@@
 Typically, the first term is non-zero and this gives the familiar oscillation.
-On resonance, it is zero, leaving the next-order term to govern the behaviour~\cite{Flanagan2012}.
+On resonance, it is zero, leaving the next-order term to govern the behaviour~\cite{Flanagan2012,Ruangsri2014}.
 Only once we have moved far enough away from resonance for the first term to dominate the second do we recapture the non-resonant behaviour.
@@
 We wish to test whether there exists some set of parameters $\boldsymbol{\Theta}$ such that the resulting adiabatic waveform is sufficiently similar to the full waveform.
-We do this by evaluating the signal-to-noise ratio (SNR) for each (normalised) waveform template,
+We do this by evaluating the SNR for each (normalised) waveform template,
 \begin{equation}
@@
 Let us now consider jumps for systems other than our illustrative $2$:$3$ resonance.
-We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
+We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits~\cite{Flanagan2012} because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
 In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum as a function of eccentricity for a selection of low-order resonances.
@@
 
-Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral.
+Transient resonances in EMRIs are an important consideration in waveform modelling due to the high proportion of expected systems encountering a low-order resonance in the later stages of inspiral~\cite{Ruangsri2014}.
 Passing through resonance can lead to an enhancement or decrement in the radiated fluxes associated with orbital evolution, which in turn leads to a jump in the orbital parameters across the resonance.

### 2016-08-25 res_char.tex

@@
 
+\newcommand{\pcm}[1]{{\color{blue}~\textsf{[P: #1]}}}
+\newcommand{\cplb}[1]{{\color{cyan}~\textsf{[CPLB: #1]}}}
+\newcommand{\cm}[2]{{\color{red}~\textsf{[CITE ME: #2]}}}
+
 \begin{document}
@@
 \begin{abstract}
-The inspiral of stellar-mass compact objects into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
-In order to detect and analyse these signals, accurate waveform templates are required.
-For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order dissipative gravitational self-force.
-Other effects are neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
-During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the emitted radiation.
-We investigate the properties of transient resonances.
-We present an evolution scheme for studying the effects of resonances and apply this to an astrophysically motivated population.
-We find that a large proportion encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence of the low eccentricity of the inspirals.
+The inspiral of stellar-mass compact objects, like neutron stars or stellar-mass black holes, into supermassive black holes provides a wealth of information about the strong gravitational-field regime via the emission of gravitational waves.
+In order to detect and analyse these signals, accurate waveform templates which include the effects of the compact object's gravitational self-force are required.
+For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order gravitational self-force.
+However, other effects are neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the waveform.
+We present an evolution scheme for studying the effects of transient resonances and apply this to an astrophysically motivated population.
+We find that a large proportion of systems encounter a low-order resonance in the later stages of inspiral; however, the resulting effect on signal-to-noise recovery is small as a consequence of the low eccentricity of the inspirals.
 Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals.
+
 \end{abstract}
@@
 Rapid advances in this field have been made following breakthroughs in 2005~\cite{Pretorius2005,Campanelli2006,Baker2006}; it is now possible to simulate hundreds of orbits~\cite{Szilagyi2015}.
-These simulations allow us to understand binary BH coalescences.
-Stellar-mass BH mergers are targets for ground-based GW detectors, such as Advanced LIGO~\cite{Aasi2015} and Advanced Virgo~\cite{Acernese2015}, and the in-construction KAGRA~\cite{Aso2013}.
-The first direct observations of GWs came from the coalescences of two stellar-mass BHs~\cite{Abbott2016,Abbott2016e,Abbott2016d}, and analysis of their properties relied upon our understanding of binary BH waveforms~\cite{Abbott2016f,Abbott2016h,Abbott2016d}.
+However, the computational cost of numerical-relativity simulations means that other approaches must be used to generate the large number of waveforms required for GW detection and analysis.
+Analytic relativity approaches such as post-Newtonian theory~\cite{Blanchet2014,Buonanno2009}, which can be used to model the early inspiral where the gravitational field is still relatively weak, and the effective-one-body formalism~\cite{Buonanno1999,Buonanno2000,Damour2009,Barausse2010}, which can incorporate merger and ringdown, allow us to generate less expensive waveform approximants.
+These approximants can be calibrated to match numerical relativity results for improved accuracy~\cite{Taracchini2014,Pan2014,Husa2015,Khan2015,Schmidt2015}, and the resulting waveforms allow us to understand comparable-mass binary BHs.
+
+Stellar-mass BH mergers are targets for ground-based GW detectors, such as Advanced LIGO~\cite{Aasi2015} and Advanced Virgo~\cite{Acernese2015}, the in-construction KAGRA~\cite{Aso2013}, and the proposed Einstein Telescope~\cite{Punturo2010a}.
+The first direct observations of GWs came from the coalescences of two stellar-mass BHs~\cite{Abbott2016,Abbott2016e,Abbott2016d}, and analysis of their properties~\cite{Abbott2016f,Abbott2016h,Abbott2016d} (plus subsequent inferences about their astrophysical origin~\cite{Abbott2016i,Abbott2016d} and tests of GR~\cite{Abbott2016j,Yunes2016,Abbott2016d}) relied upon our knowledge of binary BH waveforms.
 %Supermassive BH (SMBH) coalescences, expected to be the result of galaxy mergers, are potential sources for a space-borne detector such as eLISA~\cite{Amaro-Seoane2012a}.
@@
 These extreme-mass-ratio inspirals (EMRIs) produce GWs that are a promising signal for space-borne detectors like the evolving Laser Interferometer Space Antenna (eLISA)~\cite{Amaro-Seoane2007,Amaro-Seoane2012a}.
-Despite being well-studied, there still remain open questions in modelling generic binaries.
-
-In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004}.
+EMRIs provide a chance to measure the properties of SMBHs~\cite{Barack2004,Arun2009}, their evolution~\cite{Gair2010b,Gair2010a} and environment~\cite{Yunes2011a,Barausse2014b}, and also test for deviations from the predictions of GR~\cite{Barack2007,Gair2012a}.
+To detect and analyse EMRI signals we must have waveforms for generic orbits which are accurate for the $\sim10^4$--$10^5$ cycles of the inspiral.
+%Despite being well-studied, there still remain open questions in modelling generic binaries.
+
+%In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004,Pound2015}
+To improve our understanding of extreme-mass-ratio systems, efforts are concentrated on modelling the gravitational self-force~\cite{Barack2009,Poisson2004,Pound2015}\pcm{need up-to-date references, like Isoyama et al 2014,  Pound 2012, Hughes et al 2015}\cplb{References are currently just reviews, let me know specific places to add latest results}.
 In the test-particle limit, the smaller body follows an exact geodesic of the SMBH's spacetime.
@@
 The back-reaction from this deformation alters the small body's orbital trajectory, and can be modelled as a self-force that moves the body from its geodesic.
-The self-force is commonly divided into two pieces, dissipative and conservative.
+The self-force can be divided into two pieces, dissipative and conservative~\cite{Sago2008,Barack2009}.
 The former encapsulates the slow decay of the orbital energy and angular momentum, constants of the motion in the test particle limit, through radiation.
@@
 For EMRIs, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
-The evolution of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
-During this evolution, the frequencies may become commensurate.
-Resonances occur when the radial and polar frequencies are rational multiples of each other:
+The evolution inspiral of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
+During this evolution, the frequencies may become commensurate: resonances occur when the radial and polar frequencies are rational multiples of each other:
 \begin{equation}
@@
 Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.
-However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host.
+However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host~\cite{VanDeMeent2014}.
 
 Geodesic motion in Kerr spacetime can be described by use of the action--angle formalism~\cite{Goldstein2002}. % Chapter 10
-We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$ using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$~\cite{Hinderer2008}.
+We consider a body of mass $\mu$ orbiting a BH of mass $M$, with $\eta = \mu/M \ll 1$,\footnote{To first order, the mass ratio $\eta$ is the same as the symmetric mass ratio $\mu M/(\mu+M)^2$.} and describe the motion in the directions of the standard Boyer--Lindquist coordinates $\{t,r,\theta,\phi\}$~\cite{Boyer1967} using generalised angle variables $q_\alpha = \{q_t,q_r,q_\theta,q_\phi\}$~\cite{Hinderer2008}.
 We denote the first integrals of the geodesic motion, the generalised action variables, by $J_\alpha$.
-These are some combination of the energy per unit mass $E$ and the axial angular momentum per unit mass $L_z$, which arise from isometries of the metric in $t$ and $\phi$, and the Carter constant per unit mass squared $Q$~\cite{Carter1968}, which is related to the separability of the equations of motion in $r$ and $\theta$.
+These are some combination of the energy per unit mass $E$ and the axial angular momentum per unit mass $L_z$ of the orbit, which arise from isometries of the metric in $t$ and $\phi$, and the Carter constant per unit mass squared $Q$~\cite{Carter1968}, which is related to the separability of the equations of motion in $r$ and $\theta$.
 The system evolves following~\cite{Flanagan2012}
@@
 \end{subequations}
-where $\lambda$ is Mino time~\cite{Mino2003}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force.
+where $\lambda$ is Mino time~\cite{Mino2003} \pcm{not everyone is familiar with this term.
+Besides, Mino does't like it.
+We could consider calling it "effective time" or similar} \cplb{with apologies to Mino, I think we're stuck with it}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force \pcm{this implies that there are  second, etc orders, we need to mention it and cite appropriately}\cplb{Suggestions for citations?}.
 By working with $\lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
@@
 Ensuring the accuracy of EMRI templates requires calculating the impact that passing through a resonance has on the orbital evolution and discovering for which resonances this is significant.
+
+We show how the properties of resonances can be understood from the properties of the orbit.
+The effects of passing through resonance depend sensitively on the phase at resonance, making them difficult to predict without detailed calculation.
+The low-order resonances, such as the $1$:$2$ and $2$:$3$ resonances, can leave a noticeable imprint on the waveform.
+However, since most EMRIs have a low eccentricity when passing through these resonances, we find that for an astrophysical population of EMRIs there should not be a significant reduction in detectable signals when using adiabatic waveforms.
+The effect of resonances on parameter estimation is yet to be investigated.
 
@@
 
+\pcm{we are including a lot of material that can be found elsewhere and that we are not actually using directly.
+This makes the paper unecessary long.
+We could consider reducing the previous part or including it in the appendix.}
+\cplb{Most is necessary to introduce notation}
+
 One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short timescales, for small $\eta$ there is a small disparity.
@@
 
-To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration.
+To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration which causes the orbit to deviate from a single geodesic.
 We work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
@@
 Since only the first PN terms are used, this prescription is of limited validity in strong fields.
-Both pieces of the self-force are computed assuming that the spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$.
+Both pieces of the self-force are computed assuming that the SMBH's spin is small: the dissipative piece contains terms to $\order{a_\ast^2}$ and the conservative piece to $\order{a_\ast}$.
 This is suboptimal for high spins.
@@
 The averaged force can be computed from the radiative field~\cite{Galtsov1982,Mino2003,Sago2006,Ganz2007}.
-This piece is purely dissipative and determines how the inspiral evolves due to the radiation of GWs.
+This piece is purely dissipative~\cite{Pound2005} and determines how the inspiral evolves due to the radiation of GWs.
 
@@
 
-\subsection{timescales}\label{sec:res-time}
+\subsection{Timescales}\label{sec:res-time}
 
@@
 We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits~\cite{Flanagan2012} because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
-In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum as a function of eccentricity for a selection of low-order resonances.
+In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum $p$ as a function of eccentricity $e$ for a selection of low-order resonances.
 The other orbital shape parameters show a similar trend with increasing eccentircity~\cite{ColeThesis2015}.
@@
 We follow inspirals until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r) = 0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
-This ignores the (small) influence of the self-force.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for a mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
+This ignores the (small) influence of the self-force~\cite{Isoyama2014}.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for a mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
 

### 2016-08-26 res_char.tex
Final figure

@@
 According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$, as expected from matching the post-resonance region of the waveform.} Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
-We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta L_{z\,\mathrm{jump}}$ ellipse.
+We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta Q{\mathrm{jump}}$ ellipse.
 The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset~\cite{Flanagan2012a}; this means that for every value of $q$, there is always a resonance jump in at least one parameter.

### 2016-08-26 plotting/plot_e_nu_trends.m
Adding plotting scripts

@@ -1,0 +1,67 @@
+function plot_e_nu_trends
+
+    Colour_cube = [0 0 0
+            213/255 94/255 0 % Vermillion
+            230/255 159/255 0 % Orange
+            240/255 228/255 66/255 % Yellow
+            0 158/255 115/255 % Bluish green
+            86/255 180/255 233/255 % Sky blue
+            0 114/255 178/255 % Blue
+            204/255 121/255 167/255 % Reddish purple
+            0.5 0.5 0.5]; % Gray
+
+    % Data from resonance_fluxes.nb
+    e_data12 = [0.05, 5.858e-6; 0.3, 0.000348937; 0.5, 0.00212336; 0.7, 0.00606875; 0.95, 0.0134315; 0.99, 0.0151526];
+    e_data13 = [0.05, 0.0000117947; 0.3, 0.000100109; 0.5, 0.00122221; 0.7, 0.00493451; 0.95, 0.0131254; 0.99, 0.0150047];
+    e_data23 = [0.05, 8.21449e-6; 0.3, 0.00141763; 0.5, 0.00529747; 0.7, 0.0111033; 0.95, 0.018834; 0.99, 0.0205877];
+    e_data25 = [0.05, 8.42378e-6; 0.3, 0.000153722; 0.5, 0.00141111; 0.7, 0.00503526; 0.95, 0.0127566; 0.99, 0.0145667];
+    e_data34 = [0.05, 1.07717e-6; 0.3, 2.94956e-6; 0.5, 5.80487e-6; 0.7, 0.0000126471; 0.95, 0.0000278118; 0.99, 0.0000428484];
+    
+    figure;
+    hold on
+    plot(e_data12(:,1),e_data12(:,2),'-','Color',Colour_cube(7,:),'LineWidth',2.2)
+    plot(e_data13(:,1),e_data13(:,2),'--','Color',Colour_cube(7,:),'LineWidth',2.2)
+    plot(e_data23(:,1),e_data23(:,2),'-','Color',Colour_cube(3,:),'LineWidth',1.6)
+    plot(e_data25(:,1),e_data25(:,2),'--','Color',Colour_cube(3,:),'LineWidth',1.6)
+    plot(e_data34(:,1),e_data34(:,2),'-','Color',Colour_cube(2,:),'LineWidth',1.0)
+    hold off
[…]

### 2016-08-26 plotting/plot_emri_pop.m

@@ -1,0 +1,212 @@
+function plot_emri_pop
+
+    Colour_cube = [0 0 0
+        213/255 94/255 0 % Vermillion
+        230/255 159/255 0 % Orange
+        240/255 228/255 66/255 % Yellow
+        0 158/255 115/255 % Bluish green
+        86/255 180/255 233/255 % Sky blue
+        0 114/255 178/255 % Blue
+        204/255 121/255 167/255 % Reddish purple
+        0.5 0.5 0.5]; % Gray
+
+    % From generate_population.nb
+    lowSNR = load('low.dat');
+    highSNR = load('high.dat');
+    
+    lowCount = length(lowSNR)
+    highCount = length(highSNR)
+
+    figure;
+    hold on
+    t_bins = logspace(4,7,31);
+    countL = histc(lowSNR(1,:),t_bins)/(lowCount);
+    countH = histc(highSNR(1,:),t_bins)/(highCount);
+    bar(t_bins,countL,'histc');
+    bar(t_bins,countH,'histc');
+    delete(findobj('marker','*'))
+    h = findobj(gca,'Type','patch');
+    set(h(1),'FaceColor',Colour_cube(2,:),'EdgeColor',Colour_cube(2,:));
+    set(h(2),'FaceColor',Colour_cube(6,:),'EdgeColor',Colour_cube(6,:));
+    stairs([0, t_bins],[0, countL],'k','LineWidth',0.8);
+    stairs([0, t_bins],[0, countH],'k--','LineWidth',1.6);
+    hold off
+    box on
+    set(gca, 'XScale','log')
+    set(gca, 'Layer','top')
+    xlim([1e4, 1e7])
+    ylabel('\Large{}Fraction per bin')
+    xlabel('\Large{}Mass $M/M_\odot$')
+    set(gca,'YTickLabel',{'$0.00$','','$0.02$','','$0.04$','','$0.06$','','$0.08$','','$0.10$'})
+    %set(gca,'XTickLabel',{'$10^4$','$10^5$','$10^6$','$10^7$'})
+    laprint(gcf, 'res_pop_M');
+    
+    figure;
+    hold on
+    t_bins = linspace(0,1,31);
+    countL = histc(lowSNR(2,:),t_bins)/(lowCount);
+    countH = histc(highSNR(2,:),t_bins)/(highCount);
[…]

### 2016-08-26 plotting/plot_overlap.m

@@ -1,0 +1,87 @@
+function plot_overlap
+
+    Colour_cube = [0 0 0
+            213/255 94/255 0 % Vermillion
+            230/255 159/255 0 % Orange
+            240/255 228/255 66/255 % Yellow
+            0 158/255 115/255 % Bluish green
+            86/255 180/255 233/255 % Sky blue
+            0 114/255 178/255 % Blue
+            204/255 121/255 167/255 % Reddish purple
+            0.5 0.5 0.5]; % Gray
+
+    % Data taken from resonance_plots_waveforms.nb
+    tres = 1.9074056026223514e7;
+    %Tyear = 31536000;
+    Tyear = 31556925.9747;
+    
+    shortoverlaps2 = [1.262638369553906e6, 0.9959581179457297;
+        3.7879151086617177e6, 0.995979808290894;
+        6.31319184776953e6, 0.9962610114991012;
+        8.838468586877342e6, 0.9956082392986677;
+        1.1363745325985152e7, 0.9962209434766633;
+        1.3889022065092966e7, 0.9962077957413803;
+        1.641429880420078e7, 0.9887587007193637;
+        1.893957554330859e7, 0.8585962772514422;
+        2.14648522824164e7, 0.5916570646464226;
+        2.399012902152421e7, 0.586969393597272;
+        2.6515405760632027e7, 0.6116767153032336;
+        2.9040682499739837e7, 0.6264554791168533;
+        3.1565959238847647e7, 0.6407217430372012;
+        3.409123597795545e7, 0.6521028020071359;
+        3.661651271706327e7, 0.6646461745668794;
+        3.914178945617108e7, 0.6807448717069252;
+        4.166706619527889e7, 0.6934237630774438;
+        4.41923429343867e7, 0.7064992117680112;
+        4.671761967349451e7, 0.7193224996684443;
+        4.924289641260233e7, 0.7298055247471343;
+        5.176817315171014e7, 0.7405060303276473;
+        5.429344989081795e7, 0.7500907663599096;
+        5.6818726629925765e7, 0.7616223073087071;
+        5.9344003369033575e7, 0.7732180769432753;
+        6.1869280108141385e7, 0.7813632435078454];
+    
+    shortoverlaps2END = [1.262638369553906e6, 0.5355648732928917;
[…]

### 2016-08-26 plotting/plot_pop_overlap.m

@@ -1,0 +1,71 @@
+function plot_pop_overlap
+
+    Colour_cube = [0 0 0
+        213/255 94/255 0 % Vermillion
+        230/255 159/255 0 % Orange
+        240/255 228/255 66/255 % Yellow
+        0 158/255 115/255 % Bluish green
+        86/255 180/255 233/255 % Sky blue
+        0 114/255 178/255 % Blue
+        204/255 121/255 167/255 % Reddish purple
+        0.5 0.5 0.5]; % Gray
+
+    % From analyse_population.nb
+    SNR = [70.1875, 51.4284, 56.6448, 60.9731, 21.8433, 52.7982, 23.7042,  36.2833, 17.4335, 76.4502, 36.1407, 25.1984, 17.0621, 36.5038,  31.6015, 26.3492, 17.6434, 18.1633, 20.0794, 55.1473, 44.3521,  18.3533, 15.8701, 15.0657, 17.5525, 28.2467, 26.2184, 20.8776,  17.0778, 25.3732, 24.7565, 26.5562, 26.1024, 19.2594, 27.2579,  17.8678, 20.7458, 24.965, 15.0387, 17.3408, 19.9272, 29.3477,  23.0111, 29.2082, 23.5387, 28.8501, 32.9321, 16.2862, 21.0746,  17.7186, 16.8952, 15.5229, 26.5321, 27.8968, 26.0581, 31.0297,  19.6141, 16.9867, 18.077, 21.3162, 15.8422, 15.3012, 20.5882,  17.7982, 17.9511, 23.5375, 20.9114, 27.5673, 15.8872, 24.2043,  20.8464, 16.0371, 17.7657, 25.4534, 24.9518, 18.6827, 17.0871,  18.6833, 20.442, 19.8649, 17.2197, 18.8884, 28.986, 30.9798,  19.5682, 19.2796, 17.9669, 16.6977, 16.6941, 17.5346, 16.2181,  16.3699, 15.3976, 20.207, 15.679, 18.8524, 15.0583, 18.1429,  19.5645, 17.6559, 18.908, 16.8141, 17.0581, 20.2352, 16.6646,  17.7126, 19.3223, 21.1969, 17.105, 20.7919, 16.3018, 15.3686,  20.6951, 20.2427, 18.9497, 28.4021, 18.2772, 16.0311, 17.9048,  15.8581, 18
[…]

### 2016-08-26 plotting/plot_sinusoid.m

@@ -1,0 +1,60 @@
+function plot_sinusoid
+
+    Colour_cube = [0 0 0
+            213/255 94/255 0 % Vermillion
+            230/255 159/255 0 % Orange
+            240/255 228/255 66/255 % Yellow
+            0 158/255 115/255 % Bluish green
+            86/255 180/255 233/255 % Sky blue
+            0 114/255 178/255 % Blue
+            204/255 121/255 167/255 % Reddish purple
+            0.5 0.5 0.5]; % Gray
+
+    % Taken from globalpar_analysis.nb
+    qlin = linspace(-pi,pi,201);
+    qs = [-1.2977e+00, -5.7704e-01, 1.4589e-01, -6.2399e-01, -1.5435e+00, -1.1571e+00, -1.2257e+00, -2.2139e+00, 6.4122e-01, -4.2559e+00, -4.6586e+00, 2.6969e+00, -2.4462e+00, 4.1755e+00, 3.3515e+00, -4.6394e+00, -2.3992e+00, 4.0486e-01, -2.3098e+00, 9.4632e-01, 1.1747e+00, -1.6414e+00, 2.2971e-01, 4.3113e-01, 2.5457e+00, -1.5805e+00, -1.3057e-01, 3.3858e+00, -2.8316e+00, -5.1256e-01, -4.5870e+00, -7.5156e-01, 5.9797e-01, -7.7956e-01, 1.4353e+00, -6.4838e-02, 9.2073e-01, 1.5137e-01, 4.1759e+00, -1.5526e+00, -1.3687e+00, 3.6658e-02, 1.0859e+00, -4.7903e-01, -1.4892e+00, 3.6222e+00, 1.3355e+00, -8.1718e-01, 3.0435e+00, 3.5121e+00, -5.0988e-01, -2.3025e+00, -3.9020e-02, 3.1936e+00, -3.1434e+00, -2.4255e+00, 2.0789e+00, -2.3374e+00, -1.4393e+00, 1.5602e+00, -2.8709e+00, 1.3821e+00, -1.3241e+00, 1.5328e+00, 3.0532e+00, -5.0289e-01, 4.1445e+00, 2.6411e+00, 6.7835e-01, 1.5215e+00, 2.1367e+00, -1.8208e+00, -4.0846e+00, -4.1386e-01, -4.3664e+00, 1.5826e+00, -1.1286e+00, -1.0235e+00, -1.7166e+00, -2.4903e+00, 5.2756e-01, 4.1695e+00, 3.0405e+00, -2.5
[…]

### 2016-08-26 plotting/plot_pop_overlap.m
Fixing figure 10

@@ -60,7 +60,7 @@ function plot_pop_overlap
     ylabel('\Large{}Count per bin')
     xlabel('\Large{}Signal-to-noise ratio')
     set(gca,'YTickLabel',{'$0$','$10$','$20$','$30$','$40$','$50$','$60$','$70$','$80$'})
-    set(gca,'XTickLabel',{'$0$','$10$','$20$','$30$','$40$','$50$'})
+    set(gca,'XTickLabel',{'$5$','$10$','$15$','$20$','$25$','$30$','$35$','$40$','$45$','$50$'})
     l = legend('\large{}Adaibatic SNR $\rho_\mathrm{AK}$', '\large{}Overlap-weighted SNR $\mathbb{O} \rho_\mathrm{AK}$');
     set(l,'Box', 'off','Color','none');
     set(l, 'Position', [0.65 0.75 0.25 0.1]);

### 2016-08-27 res_char.tex
Adding self-force refs

@@
 \section{Introduction}
-\pcm{The paper is very nice and the result is very important bcs it closes and open topic about the possible bias introduced by resonances in EMRI PE.
-General comments: (1)check acronyms along the text, in particular gravitational wave and GW, EMRI, etc....!!, (2) we don't mention anything about the astrophysical implications in the intro.
-It would be nice to include a couple of sentences.
-(3) I'll emphasise more the final result in the conclusions.
-}
 
@@
 \end{subequations}
-where $\lambda$ is Mino time~\cite{Mino2003} \pcm{not everyone is familiar with this term.
-Besides, Mino does't like it.
-We could consider calling it "effective time" or similar} \cplb{with apologies to Mino, I think we're stuck with it}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force \pcm{this implies that there are  second, etc orders, we need to mention it and cite appropriately}\cplb{Suggestions for citations?}.
-By working with $\lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
+where $\lambda$ is Mino time~\cite{Mino2003}, and the forcing functions $g_\alpha^{(1)}$ and $G_A^{(1)}$ originate from the first-order self-force.\footnote{For a discussion of the second-order self-force, see \cite{Rosenthal2006,Pound2012,Gralla2012}.} By working with $\lambda$ instead of proper time $\tau$, the radial and polar motions decouple.
 At zeroth order in the mass ratio we recover the limit of purely geodesic motion: the integrals of the motion are actually constants and the angle variables evolve according to their associated frequencies $\omega_\alpha$.
@@
 According to \eqnref{delta-I-a}, the jumps should oscillate as a function of the phase $q$.\footnote{The overlaps for all of these different phases are similar, clustering around $0.7$, as expected from matching the post-resonance region of the waveform.} Therefore, assuming that the lowest harmonic dominates, plotting a jump in one parameter against another should trace out an ellipse.
-We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta Q{\mathrm{jump}}$ ellipse.
+We find that this is the case; in \figref{resjump-vs-q} we plot resonant jumps as a function of the phase $q$ constructed from the $\Delta E\sub{jump}$--$\Delta Q\sub{jump}$ ellipse.
 The jumps in $E$ and $L_z$ are approximately in phase, but those in $Q$ are found to be offset~\cite{Flanagan2012a}; this means that for every value of $q$, there is always a resonance jump in at least one parameter.

### 2016-08-29 res_char.tex
Proof-reading introduction

@@
 In order to detect and analyse these signals, accurate waveform templates which include the effects of the compact object's gravitational self-force are required.
-For computational efficiency, adiabatic templates are often used, which accurately reproduce orbit-averaged trajectories arising from the first-order gravitational self-force.
-However, other effects are neglected, in particular that of transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
-During such resonances, the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and, ultimately, the phase of the waveform.
+For computational efficiency, adiabatic templates are often used.
+These accurately reproduce orbit-averaged trajectories arising from the first-order gravitational self-force, but neglect other effects, such as transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+During such resonances the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and the phase of the waveform.
 We present an evolution scheme for studying the effects of transient resonances and apply this to an astrophysically motivated population.
@@
 %In the case of extreme-mass-ratio systems, efforts are concentrated on understanding the gravitational self-force~\cite{Barack2009,Poisson2004,Pound2015}
-To improve our understanding of extreme-mass-ratio systems, efforts are concentrated on modelling the gravitational self-force~\cite{Barack2009,Poisson2004,Pound2015}\pcm{need up-to-date references, like Isoyama et al 2014,  Pound 2012, Hughes et al 2015}\cplb{References are currently just reviews, let me know specific places to add latest results}.
+To improve our understanding of extreme-mass-ratio systems, efforts are concentrated on modelling the gravitational self-force~\cite{Barack2009,Poisson2004,Pound2015}.
 In the test-particle limit, the smaller body follows an exact geodesic of the SMBH's spacetime.
@@
 The self-force can be divided into two pieces, dissipative and conservative~\cite{Sago2008,Barack2009}.
-The former encapsulates the slow decay of the orbital energy and angular momentum, constants of the motion in the test particle limit, through radiation.
+The former encapsulates the slow decay of the orbital energy and angular momentum (constants of the motion in the test particle limit) through radiation of GWs.
 The latter shifts the orbital phases inducing precession.
@@
 
-\pcm{we are including a lot of material that can be found elsewhere and that we are not actually using directly.
-This makes the paper unecessary long.
-We could consider reducing the previous part or including it in the appendix.}
-\cplb{Most is necessary to introduce notation}
-
 One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short timescales, for small $\eta$ there is a small disparity.

### 2016-08-31 res_char.tex
First comments from Stas

@@
 Resonances involving the azimuthal motion do not produce a comparable effect because of the axisymmetry of the background spacetime.
-However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases, sufficient to eject the central BH from its host~\cite{VanDeMeent2014}.
+However, both $\theta$--$\phi$ resonances~\cite{Hirata2011} and $r$--$\phi$ resonances~\cite{VanDeMeent2013} can lead to extrinsic effects; the GWs from such systems are not emitted isotropically and the imbalance produces a kick velocity that is, in some cases with moderate mass ratios, sufficient to eject the central BH from its host~\cite{VanDeMeent2014}.
 
@@
 We work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
-For comparison, Flanagan, Hughes and Ruangeri~\cite{Flanagan2012a} use a Teukolsky-equation calculation of GW fluxes to account for the inspiral due to radiation reaction.
+For comparison, Flanagan, Hughes and Ruangsri~\cite{Flanagan2012a} use a Teukolsky-equation calculation of GW fluxes to account for the inspiral due to radiation reaction.
 
@@
 The instantaneous $\dot{\Omega}$ depends upon the orbital phase and oscillates about its mean trend over an orbit.
-We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time-average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$.
-Close to resonance, $\Omega(t)$ is well approximated by a first-order Taylor expansion, decreasing linearly with time; hence we make the approximation
+We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time-average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$.\footnote{On resonance, we are interested in the relative $r$--$\theta$ orbital phase ($\psi_{\theta_-}$ or equivalent), which sets the resonant trajectory in the $r$--$\theta$ plane, but not the exact phase of the orbit around this loop.} Close to resonance, $\Omega(t)$ is well approximated by a first-order Taylor expansion, decreasing linearly with time; hence we make the approximation
 \begin{equation}
@@
 \end{align}
-Here $\widehat{\kappa}_0$ is the orbital phase on resonance and $F_{a,\,s}^{(1)}$ is the $s$-th harmonic of the first-order self-force on resonance, defined such that\footnote{Since the geodesic equations decouple in Mino time rather than coordinate time, this is true only in an average sense.}
+Here $\widehat{\kappa}_0$ is the phase on resonance, which sets the resonant trajectory in the $r$--$\theta$ plane similarly to $\psi_{\theta_-}$, and $F_{a,\,s}^{(1)}$ is the $s$-th harmonic of the first-order self-force on resonance, defined such that\footnote{Since the geodesic equations decouple in Mino time rather than coordinate time, this is true only in an average sense.}
 \begin{equation}
@@
 \label{eq:overlap}
-\mathbb{O}\left[h\right] = \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}},
+\mathbb{O}\left[h\right] = \max_t \frac{\overlap{s}{h}}{\sqrt{\overlap{s}{s}}\sqrt{\overlap{h}{h}}},
 \end{equation}
@@
 The inset plots show the start of the evolution, on a timescale associated with the orbital motion of the compact object; the $2$-torus averaging explicitly smooths out the visible structure on this scale.
-The two appraches are in good agreement.
+The two approaches are in good agreement.
 
@@
 \caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
-The time $t\sub{res} = 0$ is when the full evolution crosses the $2$:$3$ resonance.}
+The time $t = t\sub{res}$ is when the full evolution crosses the $2$:$3$ resonance.}
 \end{figure}
@@
 In \figref{res-diff-traj}, we plot the difference in the orbital parameters ($E$, $L_z$ and $Q$) between the various trajectories and the adiabatic evolution that matches at the start.
-The jumps in the orbital parameters due to the $2$:$3$ resonance can be clearly seen, as can the fast orbital oscillations present in the full instanteous evolution but absent in the adiabatic evolutions.
+The jumps in the orbital parameters due to the $2$:$3$ resonance can be clearly seen, as can the fast orbital oscillations present in the full instantaneous evolution but absent in the adiabatic evolutions.
 
@@
 In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum $p$ as a function of eccentricity $e$ for a selection of low-order resonances.
-The other orbital shape parameters show a similar trend with increasing eccentircity~\cite{ColeThesis2015}.
+The other orbital shape parameters show a similar trend with increasing eccentricity~\cite{ColeThesis2015}.
 Larger eccentricities do give rise to larger jumps, matching our expectations and Teukolsky-based calculations by Flanagan, Hughes and Ruangsri~\cite{Flanagan2012a}.
@@
 Roughly $30\%$ of the systems lie within $0.05$ of the expected value.
-However, the vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity.
+The vast majority of these are not significantly disrupted by resonances, and produce overlaps approaching unity.
 For the smallest (most extreme) mass ratios, the inspiral rate is slow, and so the systems do not encounter either the $1$:$2$ or $2$:$3$ resonances during their lifetime.
@@
 These occur at low and intermediate values of the mass ratio, where the inspiral rate is slow enough that the low-order resonances are encountered in the middle of the observation window, and the resulting value of $t\sub{ad}$ is small.
-For these EMRIs, resonances are not as important as expected, which is likely due to the low eccentricities of our population leading to small (and in most cases, negligible) resonant flux enhancements.
+For these EMRIs, resonances are not as important as expected.
+Some of these could be because of fortuitous phases on resonance corresponding to small resonance jumps.
+The most likely explanation for the lack of impact is because of the low eccentricities of our population, this means that the magnitude of the resonant flux enhancements are small (and in most cases negligible).
 
@@
 The authors extend our sincere gratitude to \'{E}anna Flanagan and Tanja Hinderer for providing their PN self-force code, without which this work would not be possible.
-We thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent, Leor Barack, Scott Hughes and Nico Yunes for useful conversations; Stanislav Babak for examining both \cite{BerryThesis2013} and \cite{ColeThesis2015}, and Christopher Moore with help preparing this manuscript.
+We are grateful to Stanislav Babak for examining both \cite{BerryThesis2013} and \cite{ColeThesis2015}, and for providing comments on this draft.
+We thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent, Leor Barack, Scott Hughes and Nico Yunes for useful conversations, and Christopher Moore with help preparing this manuscript.
 RHC was supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC's work was supported by a Marie Curie Intra-European Fellowship within the 7th European Community Framework Programme (PIEF-GA-2011-299190), and JRG was supported by the Royal Society.

### 2016-08-31 res_char.tex
Combinig asymptotic.tex with main text

@@
 
-\input{./asymptotic.tex}
+The impact of passing through resonance on the evolution can be modelled analytically using asymptotic expansions~\cite{Gair2012}.
+Solutions for the motion are constructed far away from resonance and these are matched to a transition region in the vicinity of resonance~\cite{Kevorkian1971,Bosley1992}.
+By comparing the matched solution, which incorporates the effects of resonance, with the results of an adiabatic evolution, it is possible to estimate the discrepancy in the orbital parameters.
+This determines the difference in the orbital phase between the two approaches.
+If this error is sufficiently small, then it is safe to ignore the effects of the resonance; however, only a small difference is needed to impact the subsequent waveform, since the error accumulates over the subsequent observation of $\sim \order{\eta^{-1}}$ cycles~\cite{Flanagan2012}.
+We derive formulae which can be used to calculate the discrepancy in the orbital parameters.
+
+The following derivation is reproduced from Berry~\cite{BerryThesis2013}.
+It is is based upon the analysis of Kevorkian~\cite{Kevorkian1987}; small adjustments have been made to adapt to the specific problem of GW inspiral, but the general argument is unchanged.\footnote{The same two-timescale theory underpins the analysis of Hinderer and Flanagan~\cite{Hinderer2008}, but this explicitly ignores resonances.} A similar derivation can be found in van de Meent~\cite{VanDeMeent2013}.
+
+We model the system using action--angle variables.
+We are only concerned with the $r$ and $\theta$ motions, so we have a $2$-dimensional system.
+We perform a canonical transformation  to isolate the resonant combination $q = n_r q_r - n_\theta q_\theta$~\cite{Bosley1992,VanDeMeent2013}.
+This becomes one of the new angle variables, the other variable $q'$ can be either $q_r$ or $q_\theta$ (as, on resonance, varying one necessarily changes the other).
+We use $J$ as the conjugate action variable to $q$ and $\omega = n_r \omega_r - n_\theta \omega_\theta$ as its frequency.
+Similarly, we use $J'$ as the action variable conjugate to $q'$.
+The system evolves through resonance slowly, on an evolution timescale, so we parametrize it in terms of a slow time parameter
+\begin{equation}
+\widetilde{\lambda} = \eta\lambda.
+\end{equation}
+The orbits of $q'$ proceed with the fast time $\lambda$; since this is much more rapid than the evolution we are interested in, it is safe to average over it.
+We are not interested in the fine-grained fast oscillations caused by changes in $q'$.
+For this analysis we consider the reduced problem of evolving $q$ and $J$.
+
+At resonance $\widetilde{\lambda} = \widetilde{\lambda}_\star$ and $\omega\left(\widetilde{\lambda}_\star\right) = 0$. We assume that the frequency has a simple zero and can be expanded as
+\begin{equation}
+\omega\left(\widetilde{\lambda}\right) = \varpi_1\left(\widetilde{\lambda} - \widetilde{\lambda}_\star\right) + \varpi_2\left(\widetilde{\lambda} - \widetilde{\lambda}_\star\right)^2 + \ldots
+\label{eq:omega-series}
+\end{equation}
+The frequency is actually a function of the angle variables, but since these evolve with $\widetilde{\lambda}$ it is safe to write it as a function of the slow time.\footnote{In effect we are defining $\omega\left(\tilde{\lambda}\right) \equiv \omega\left[J\left(\tilde{\lambda}\right)\right]$.}
+
+Using the slow time, the equations of motion (\ref{eq:Mino-E-o-M}) become
+\begin{subequations}
+\begin{align}
+\diff{q}{\widetilde{\lambda}} = {} & \dfrac{\omega(J)}{\eta} + \sum_s g_s^{(1)}(J)\exp(is q)  + \order{\eta}, \\
+\diff{J}{\widetilde{\lambda}} = {} & \sum_s G_s^{(1)}(J)\exp(is q) + \order{\eta},
+\end{align}
+\end{subequations}
+where we have rewritten the forcing terms as Fourier series and adapted the forcing functions to those appropriate for $q$ and $J$.
+We solve these before resonance and then match to solutions in the transition regime about resonance.
+
+\subsection{Solution before resonance}\label{sec:before-res}
+
+To find a solution away from the resonance, we decompose the problem to be a function of two timescales~\cite{Kevorkian1971}.
+We use the slow time $\widetilde{\lambda}$ and, as a proxy for the fast time,
+\begin{equation}
+\Psi = \intd{0}{\lambda}{\omega(\eta\tau)}{\tau} = \recip{\eta}\intd{0}{\tilde{\lambda}}{\omega(\widetilde{\tau})}{\widetilde{\tau}}.
+\end{equation}
+From this
+\begin{equation}
+\omega = \diff{\Psi}{\lambda}.
+\end{equation}
+In terms of these two variables, we can build ansatz solutions
+\begin{subequations}
+\begin{align}
+\label{eq:q-series}
+q(\lambda;\,\eta) = {} & \Psi + q_0\left(\widetilde{\lambda}\right) + \eta q_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}, \\
+J(\lambda;\,\eta) = {} & J_0\left(\widetilde{\lambda}\right) + \eta J_1\left(\Psi,\widetilde{\lambda}\right) + \order{\eta^2}.
+\label{eq:J-series}
+\end{align}
+\end{subequations}
+We can also write a series expansion for the frequency, since it is affected by the self-force too,
+\begin{equation}
+\omega(\lambda;\,\eta) = \omega_0\left(\widetilde{\lambda}\right) + \eta \omega_1\left(\widetilde{\lambda}\right) + \order{\eta^2}.
+\end{equation}
+In the limit of $\eta \rightarrow 0$ we are left with a constant frequency $\omega_0(0)$.
+The higher-order terms are identified below by matching terms in the series expansions of the equations of motion.
+Taking the two timescales as independent, we may write the time derivative to $\order{\eta}$ as
+\begin{equation}
+\diffop{\lambda} = \omega_0\partialdiffop{\Psi} + \eta\omega_1\partialdiffop{\Psi} + \eta\partialdiffop{\widetilde{\lambda}}.
+\end{equation}
+
+Using the two timescale decomposition to replace the time derivatives in the equations of motion, and substituting in the ansatz expansions gives, to first order,
+\begin{widetext}
+\begin{subequations}
+\begin{align}
+\label{eq:q-1}
+\omega_0 + \eta\omega_1 + \eta\partialdiff{q_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{q_1}{\Psi} = {} & \omega(J_0) + \eta\diff{\omega}{J}J_1 + \eta \sum_s g_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right], \\
+\eta\partialdiff{J_0}{\widetilde{\lambda}} + \eta\omega_0\partialdiff{J_1}{\Psi} = {} & \eta \sum_s G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right].
+\label{eq:J-1}
+\end{align}
+\end{subequations}
+\end{widetext}
+Averaging \eqnref{J-1} over $\Psi$ gives\footnote{The ansatz is constructed such that $J_0 \equiv \langle J_0\rangle_\Psi$ and $q_0 \equiv \langle q_0\rangle_\Psi$.}
+\begin{equation}
+\partialdiff{J_0}{\widetilde{\lambda}} = G_0^{(1)}(J_0).
+\label{eq:J-ad}
+\end{equation}
+This describes the adiabatic evolution, hence we can identify $J_0\left(\widetilde{\lambda}\right)$ with (the lowest-order piece of) the adiabatic solution~\cite{Hinderer2008}.
+If we similarly average \eqnref{q-1}, we find
+\begin{equation}
+\omega_0 + \eta\omega_1 +\eta\partialdiff{q_0}{\widetilde{\lambda}} = \omega(J_0) + \eta\partialdiff{\omega}{J}\langle J_1\rangle_\Psi + \eta g_0^{(1)}(J_0).
+\end{equation}
+From this we can identify the terms that originate from the frequency and, matching by order in $\eta$, obtain
+\begin{subequations}
+\begin{align}
+\omega_0 = {} & \omega(J_0), \\
+\omega_1 = {} & \partialdiff{\omega}{J}\langle J_1\rangle_\Psi.
+\end{align}
+\end{subequations}
+This leaves
+\begin{align}
+\partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0), \\
+q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau}.
+\label{eq:q-0-sol}
+\end{align}
+We now have expressions for the lowest-order terms in the expansions.
+
+Subtracting the $s = 0$ components from \eqnref{J-1} leaves
+\begin{equation}
+\omega_0\partialdiff{J_1}{\Psi} = \sum_{s\,\neq\,0} G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right].
+\end{equation}
+This can be solved to give
+\begin{equation}
+J_1 = \langle J_1\rangle_\Psi + \recip{\omega_0}\sum_{s\,\neq\,0} \dfrac{G_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right]}{is}.
+\label{eq:J-1-sol}
+\end{equation}
+We can do the same for \eqnref{q-1} to obtain
+\begin{equation}
+q_1 = \langle q_1\rangle_\Psi + \recip{\omega_0}\sum_{s\,\neq\,0} \dfrac{g_s^{(1)}(J_0)\exp\left[is(\Psi + q_0)\right]}{is}.
+\label{eq:q-1-sol}
+\end{equation}
+To find the constants of integration, $\langle q_1\rangle_\Psi$ and $\langle J_1\rangle_\Psi$, it is necessary to extend the analysis to second order in $\eta$.
+This shows that $\langle J_1\rangle_\Psi$ is the first-order component of the adiabatic solution.
+We do not need explicit forms for later calculations, so we will not proceed further.
+We have successfully constructed the pre-resonance solution.
+
+\subsection{Solution near resonance}\label{sec:interior-res}
+
+Near to resonance, we consider an interior layer expansion~\cite{Kevorkian1971}.
+As explained in \secref{res-time}, evolution near resonance proceeds on a timescale intermediate between the slow and fast times.
+We therefore introduce a rescaled time
+\begin{equation}
+\widehat{\lambda} = \dfrac{\widetilde{\lambda} - \widetilde{\lambda}_\star}{\eta^{1/2}} = \eta^{1/2}(\lambda - \lambda_\star).
+\end{equation}
+As for the before resonance solution, we can create a series expansion; however, now we expand in terms of $\eta^{1/2}$~\cite{Flanagan2012}
+\begin{subequations}
+\begin{align}
+q\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{q}_0\left(\widehat{\lambda}\right) + \eta^{1/2} \widehat{q}_{1/2}\left(\widehat{\lambda}\right) + \order{\eta}, \\
+J\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{J}_0 + \eta^{1/2} \widehat{J}_1\left(\widehat{\lambda}\right) + \order{\eta}.
+\end{align}
+\end{subequations}
+The series expansion for the frequency, \eqnref{omega-series}, can be rewritten as
+\begin{equation}
+\omega\left(\widehat{\lambda}\right) = \eta^{1/2}\varpi_1\widehat{\lambda} + \eta\varpi_2\widehat{\lambda}^2 + \order{\eta^{3/2}}.
+\label{eq:omega-hat}
+\end{equation}
+Proceeding to write the equations of motion in terms of the rescaled time gives
+\begin{subequations}
+\begin{align}
+\diff{q}{\widehat{\lambda}} = {} & \varpi_1\widehat{\lambda} + \eta^{1/2}\varpi_2\widehat{\lambda}^2 \nonumber \\ 
+ {} & + \left. \eta^{1/2}\sum_s g_s^{(1)}\left(\widehat{J}_0,\widetilde{\lambda}_\star\right)\exp(is \widehat{q}_0)  + \order{\eta}, \right. \\
+\diff{J}{\widehat{\lambda}} = {} & \eta^{1/2}\sum_s G_s^{(1)}\left(\widehat{J}_0,\widetilde{\lambda}_\star\right)\exp(is \widehat{q}_0) + \order{\eta}.
+\end{align}
+\end{subequations}
+
+From the equations of motion we can match terms by their order in $\eta^{1/2}$. At zeroth order we find
+\begin{equation}
+\widehat{J}_0 = \widehat{\varrho}_0
+\end{equation}
+is constant, and
+\begin{equation}
+\widehat{q}_0 = \widehat{\kappa}_0 + \dfrac{\varpi_1\widehat{\lambda}^2}{2}.
+\end{equation}
+Using these, we can build the next-order terms
+\begin{align}
+\widehat{q}_{1/2} = {} & \widehat{\kappa}_{1/2} + \dfrac{\varpi_2\widehat{\lambda}^3}{3} + g_0^{(1)}(\widehat{\varrho}_0)\widehat{\lambda} \nonumber \\ 
+ {} & + \sum_{s\,\neq\,0}g_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}, \\
+\widehat{J}_{1/2} = {} & \widehat{\varrho}_{1/2} + G_0^{(1)}(\widehat{\varrho}_0)\widehat{\lambda} \nonumber \\
+ {} & + \sum_{s\,\neq\,0}G_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}.
+\end{align}
+Both of these involve the complex Fresnel integral~\cite{Olver2010}, the details of which are given in the following section. We have now constructed the interior solution. % chapter 7
+
+\subsection{The complex Fresnel integral}
+
+The solution for the motion in the interior region near to resonance involves the integral
+\begin{equation}
+W\left(\widehat{\lambda}\right) = \intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}.
+\end{equation}
+The complex Fresnel integral is
+\begin{equation}
+\mathcal{Y}(z) = \intd{0}{z}{\exp\left(\dfrac{i\pi x^2}{2}\right)}{x} = \mathcal{C}(z) + i\mathcal{S}(z),
+\end{equation}
+where $\mathcal{C}(z)$ and $\mathcal{S}(z)$ are the cosine and sine Fresnel integrals~\cite{Olver2010}, and hence % [7.2.7, 7.2.8]
+\begin{equation}
+W\left(\widehat{\lambda}\right) = \sqrt{\dfrac{\pi}{s\varpi_1}}\mathcal{Y}\left(\sqrt{\dfrac{s\varpi_1}{\pi}}\widehat{\lambda}\right).
+\end{equation}
+
+We are interested in the asymptotic behaviour for $|\widehat{\lambda}| \rightarrow \infty$. The complex Fresnel integral has the limit~\cite{Olver2010} % [7.5.3, 7.5.4, 7.12.2, 7.12.3]
+\begin{equation}
+\lim_{|z|\,\rightarrow\,\infty} \mathcal{Y}(z) \sim \dfrac{\sgn z}{\sqrt{2}} \exp\left(\dfrac{i\pi}{4}\right) - \dfrac{i}{\pi z}\exp\left(-\dfrac{i\pi z^2}{2}\right),
+\end{equation}
+where 
+\begin{equation}
+\sgn z = \begin{cases}
+1 & z > 0 \\
+-1 & z < 0
+\end{cases}\,.
+\end{equation}
+Hence,
+\begin{align}
+\lim_{|\widehat{\lambda}|\,\rightarrow\,\infty}W\left(\widehat{\lambda}\right) \sim {} & \dfrac{\sgn \widehat{\lambda}}{\sqrt{2}}\sqrt{\dfrac{\pi}{|s\varpi_1|}}\exp\left[\sgn(s\varpi_1)\dfrac{i\pi}{4}\right] \nonumber \\
+  {} & + \recip{is\varpi_1}\exp\left(\dfrac{is \varpi_1 \widehat{\lambda}^2}{2}\right).
+\label{eq:Fres-limit}
+\end{align}
+
+\subsection{Matching solutions}
+
+To complete our solution we must match the pre-resonance solution of \secref{before-res} with the near-resonance solution of \secref{interior-res}.
+This is achieved by rewriting the pre-resonance solution in terms of the rescaled time $\widehat{\lambda}$ and comparing this with the near-resonance solution expanded in the limit of $\widehat{\lambda} \rightarrow -\infty$.
+
+To rewrite the pre-resonance solution, we begin with the fast phase parameter
+%\pagebreak
+\begin{equation}
+\Psi\left(\widehat{\lambda}\right) = \dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \eta^{1/2}\dfrac{\varpi_2\widehat{\lambda}^3}{3} + \order{\eta}.
+\end{equation}
+Using this together with equations (\ref{eq:q-0-sol}) and (\ref{eq:q-1-sol}) in \eqnref{q-series}, the angle variable is
+\begin{widetext}
+\begin{align}
+q\left(\widehat{\lambda};\,\eta\right) = {} & \dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \kappa_\star + \eta^{1/2}\dfrac{\varpi_2\widehat{\lambda}^3}{3} + \eta^{1/2}g_0^{(1)}(J_\star)\widehat{\lambda} + \dfrac{\eta^{1/2}}{\varpi_1\widehat{\lambda}}\sum_{s\,\neq\,0}\recip{is}g_s^{(1)}(J_\star)\exp\left[is\left(\dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \kappa_\star\right)\right] + \order{\eta},
+\end{align}
+where we have defined $J_\star \equiv J_0\left(\widetilde{\lambda}_\star\right)$ and $\kappa_\star = \kappa_0 + \intd{0}{\tilde{\lambda}_\star}{g_0^{(1)}[J_0(\tau)]}{\tau}$, and used \eqnref{omega-hat} to substitute for $\omega$.
+The action variable is similarly determined by using equations (\ref{eq:J-ad}) and (\ref{eq:J-1-sol}) with \eqnref{J-series} to give
+\begin{align}
+J\left(\widehat{\lambda};\,\eta\right) = {} & J_\star + \eta^{1/2}G_0^{(1)}(J_\star)\widehat{\lambda} + \dfrac{\eta^{1/2}}{\varpi_1\widehat{\lambda}}\sum_{s\,\neq\,0}\recip{is}G_s^{(1)}(J_\star)\exp\left[is\left(\dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2} + \kappa_\star\right)\right] + \order{\eta}.
+\end{align}
+\end{widetext}
+We can now compare this to the near-resonance expansion with the integral replaced by the limiting form given in \eqnref{Fres-limit}.
+
+At zeroth order, we immediately obtain
+\begin{align}
+\widehat{\kappa}_0 = {} & \dfrac{\Psi_\star}{\eta} + \kappa_\star, \\
+\widehat{\varrho}_0 = {} & J_\star.
+\end{align}
+These fix the integration constants. The more interesting result is now found by comparing the $\order{\eta^{1/2}}$ terms. Equating the angle variable expressions and cancelling terms gives
+\begin{align}
+%\widehat{\kappa}_{1/2} + \sum_{s\,\neq\,0}g_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\left\{\dfrac{\sgn \widehat{\lambda}}{\sqrt{2}}\sqrt{\dfrac{\pi}{|s\varpi_1}}\exp\left[\sgn(s\varpi_1)\dfrac{\pi}{4}\right] + \recip{is\varpi_1}\exp\left(\dfrac{is \varpi_1 \widehat{\lambda}^2}{2}\right)\right\} = {} & \recip{is\varpi_1\widehat{\lambda}}g_s^{(1)}(\widehat{\varrho}_0)\exp\left[is\left(\dfrac{\Psi_\star}{\eta} + \dfrac{\varpi_1\widehat{\lambda}^2}{2}\right)\right] \nonumber \\
+\widehat{\kappa}_{1/2} = {} & \sum_{s\,\neq\,0}g_s^{(1)}(\widehat{\varrho}_0)\sqrt{\dfrac{\pi}{2|s\varpi_1|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\varpi_1\right)\right].
+\end{align}
+Similarly, for the action variables
+\begin{equation}
+\widehat{\varrho}_{1/2} = \sum_{s\,\neq\,0}G_s^{(1)}(\widehat{\varrho}_0)\sqrt{\dfrac{\pi}{2|s\varpi_1|}}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\varpi_1\right)\right].
+\label{eq:J-1/2}
+\end{equation}
+We now have a matched solution through resonance.
+
+Having constructed the solution, we see that the lowest-order evolution corresponds to the adiabatic solution; the deviations come in at the following order.
+When we switch from the pre-resonance solution to the post-resonance solution, there is a change in the sign of $\widehat{\lambda}$.
+Therefore, when matching the post-resonance solution $\widehat{\varrho}_{1/2}$ and $\widehat{\kappa}_{1/2}$ also change sign: there is a change of
+\begin{align}
+\Delta q = {} & 2 \eta^{1/2}\widehat{\kappa}_{1/2}, \\
+\Delta J = {} & 2 \eta^{1/2}\widehat{\varrho}_{1/2}
+\label{eq:jumps}
+\end{align}
+across the resonance~\cite{Kevorkian1987}.
+We are not particularly interested in the deviation in $J$, of greater concern is the change in the orbital parameters $\{E,L_z,Q\}$.
+Assuming that there is a smooth transformation that maps between $J$ and these, then, to lowest order, we can calculate the deviation relative to the adiabatic prescription by substituting the forcing functions $G^{(1)} \rightarrow G_a^{(1)}$, where $G_a^{(1)}$ describes the evolution of $\mathcal{I}^a$ through the effects of the self-force.
+This result is quoted by Flanagan and Hinderer~\cite{Flanagan2012}.
+The change in the orbital parameters is determined by the forcing functions, hence it is essential to have an accurate self-force model.
+
+As a final step in understanding our result, we switch from Mino time to coordinate time. An appropriate redefinition of the forcing functions can be done by scaling by $\Gamma$, we define
+\begin{equation}
+F_a^{(1)} = \dfrac{G_a^{(1)}}{\Gamma},
+\end{equation}
+such that the equation of motion becomes
+\begin{equation}
+\left\langle \diff{\mathcal{I}^a}{t}\right\rangle_{q'} =  \eta\sum_s F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}})\exp(is q) + \order{\eta^2}.
+\end{equation}
+Here we have made the averaging over $q'$ explicit to show that the equation is only defined as an orbital average: not only does our asymptotic expansion average out oscillations over an orbit in $q'$, but in converting from $\lambda$ to $t$ we have used $\Gamma$ which is an orbital average.
+From \eqnref{omega-series}, we recognise that
+\begin{equation}
+\varpi_1 = \partialdiff{\omega}{\widetilde{\lambda}} = \dfrac{\Gamma^2}{\eta}\left\langle\dot{\Omega}\right\rangle_{q'}.
+\end{equation}
+We have used the averaged form of $\dot{\Omega}(t)$ as this is appropriate. Using these to adapt equations (\ref{eq:J-1/2}) and (\ref{eq:jumps}), we obtain
+\begin{align}
+\Delta \mathcal{I}^a = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\left[\dfrac{2\pi}{\left|s \left\langle\dot{\Omega}\right\rangle_{q'}\right|}\right]^{1/2} \nonumber \\*
+ {} & \times \left.\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4}\sgn s\dot{\Omega}\right)\right]\right. \\
+ = {} & \eta\sum_{s\,\neq\,0}F_{a,\,s}^{(1)}(\boldsymbol{\mathcal{I}}_\star)\tau_{\mathrm{res},\,s}\exp\left[i\left(s \widehat{\kappa}_0 + \dfrac{\pi}{4} \sgn s\dot{\Omega} \right)\right],
+\end{align}
+using \eqnref{T-res-s} and representing the values on resonance of $E$, $L_z$ and $Q$ with $\boldsymbol{\mathcal{I}}_\star$.
 

### 2016-08-31 assemble-things.sh
Adding assembly script

@@ -1,0 +1,34 @@
+#!/bin/bash/
+
+mkdir -p submit
+
+rm -r res-char.tar.gz
+rm -f ./submit/*
+
+cp res_char.tex ./submit/
+cp res_char.bbl ./submit/
+
+cp Resonances.bib ./submit/
+
+cp Fig_good_traj.* ./submit/
+cp Fig_overlap.* ./submit/
+cp Fig_res_diff_traj.* ./submit/
+cp Fig_res_jump_calc.* ./submit/
+cp Fig_q_jump.* ./submit/
+cp Fig_e_rms_p.* ./submit/
+cp Fig_nu_rms_p.* ./submit/
+cp Fig_res_pop_M.* ./submit/
+cp Fig_res_pop_a.* ./submit/
+cp Fig_res_pop_z.* ./submit/
+cp Fig_res_pop_p.* ./submit/
+cp Fig_res_pop_e.* ./submit/
+cp Fig_res_pop_iota.* ./submit/
+cp Fig_res_pop_t.* ./submit/
+cp Fig_res_pop_SNR.* ./submit/
+cp Fig_res_overlap_t.* ./submit/
+cp Fig_res_pop_SNR_overlap.* ./submit/
+cp Fig_res-2-5-95-plane.* ./submit/
+cp Fig_fit-error-max-plane.* ./submit/
+cp Fig_fit-error-RMS-plane.* ./submit/
+
+tar -zcvf res-char.tar.gz submit

### 2016-08-31 res_char.tex
Adding clarification on Fig. 4
Minor changes for presentation purposes

@@
 For computational efficiency, adiabatic templates are often used.
-These accurately reproduce orbit-averaged trajectories arising from the first-order gravitational self-force, but neglect other effects, such as transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
+These accurately reproduce orbit-averaged trajectories arising from the first-order self-force, but neglect other effects, such as transient resonances, where the radial and poloidal fundamental frequencies become commensurate.
 During such resonances the flux of gravitational waves can be diminished or enhanced, leading to a shift in the compact object's trajectory and the phase of the waveform.
@@
 Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals.
-
 \end{abstract}
@@
 \caption{\label{fig:res-jump-calc}The difference in energy between the instantaneous model and adiabatic model that matches at the $2$:$3$ resonance, scaled by the magnitude of the $2$:$3$ resonance jump.
+The apparent thickness of the line is because of oscillations on the short orbital timescale, which are too fast to be resolved on on the plotted scale.
 The dashed green (yellow) lines show the bounding fits to the data before (after) the resonance, used to numerically estimate the size of the jump.
@@
 \Psi = \intd{0}{\lambda}{\omega(\eta\tau)}{\tau} = \recip{\eta}\intd{0}{\tilde{\lambda}}{\omega(\widetilde{\tau})}{\widetilde{\tau}}.
+\vspace{0.000cm}
 \end{equation}

### 2016-11-08 referee_response.txt
Starting on referee comments

We thank the referee for their careful reading of our document, especially given its length! We have implemented almost all of their suggestions, as indicated below, and we feel that the manuscript has been improved by these changes.

I addition to the changes commented below, we have updated some of the references and added the referee to the acknowledgements.

#1
In Fig.2, is it possible for authors to add the values of semilatus
rectum p and eccentricity e at t=t_{res}? I would like to compare the
values with those in Fig.8 as a reference.

>> Figure 8 shows the orbital parameters at plunge, rather than at resonance...

#2
In Sec.V.A the distribution of inclination angle is assume to be
uniform. Do the authors know any reference to justify this assumption?

For example, in Ref.[78], Flanagan and Hinderer (2007), Ref.[83], Sago
et al, and Ref.[84], Ganz et al, of this manuscript, it is found that
gravitational radiation reaction drives the orbital plane to
anti-align with the symmetry axis. The effect would be very small, but
it might be commented on the effect (ex. in a footnote).

>> The true inclination distribution is currently uncertain, as we are not sure of the dynamics at the very centre of galaxies. We used a uniform distribution as the least constraining proposal in absence of better information. It is true that radiation reaction will lead to anti-alignment, but we are considering the initial distribution of orbits from which inspirals begin, this is the same reason why not all the orbits are circular, despite the circularising effect of radiation reaction. Adding a comment on potential alignment is a sensible recommendation, so we have added a footnote to explain this...

#3
In the last paragraph in Sec.VI, there is the sentence ends with "if
the phase on resonance and the magnitude of the self-force can be
predicted with sufficient accuracy.".

If the authors would think that interpolation formulas of the
gravitational self-force in Kerr spacetime, which might be an
extension of Ref.[75], Warburton et al (2012), could be used for this
purpose, it might be commented in the same paragraph in Sec.VI.

#4
In Appendix B it might be mentioned that \kappa_0 in Eq.(B14),
\hat\varrho_0 in Eq.(B22), \hat\kappa_0 in Eq.(B23), \hat\kappa_{1/2}
in Eq.(B24) and \hat\varrho_{1/2} in Eq.(B25), are the constants of
integration when they appear for the first time although it might be
understood if readers would follow equations.

>> This is a good suggestion, we have added clarifications after each of the equations (or pairs of equations).

#5
I might find some typos. Please check if the authors could agree with
them:

a. In Eq.(5), parenthesis might be added to the integrand.

>> Yes, but we think it is clear as it is.

b. In the first paragraph in Sec.V.B, is "across an overlap" replaced
by "across a resonance"?

>> Yes, this has been corrected.

c. In Eq.(B19b), \hat{J_{1}} may be replaced by \hat{J_{1/2}}.

>> Yes, this has been fixed.

d. In Eq.(B29), exp(-i\pi z^2/2) may be replaced by exp(i\pi z^2/2).

>> Yes, this is extremely well spotted!

e. In Eq.(B31), 1/(is\varpi_1) may be replaced by
1/(is\varpi_1\hat{\lambda}).

>> Yes, this has been corrected.

### 2016-11-08 res_char.tex
Final Stas comment
Starting on referee comments

@@
 The other orbital shape parameters show a similar trend with increasing eccentricity~\cite{ColeThesis2015}.
+Across the grid of $a_\ast$ and $\cos\iota$ values there may be an order of magnitude variation, but the same general trends are observed.
 Larger eccentricities do give rise to larger jumps, matching our expectations and Teukolsky-based calculations by Flanagan, Hughes and Ruangsri~\cite{Flanagan2012a}.
@@
 From the results of \secref{effres-phase}, we expect the recovered overlap to be approximately given by the proportion of time spent in a resonance-free region, that is $t\sub{ad} / t\sub{insp}$.
-This assumes that there is perfect overlap in the absence of a resonance and zero overlap across an overlap, with all times during the inspiral contributing equally.
+This assumes that there is perfect overlap in the absence of a resonance and zero overlap across a resonance, with all times during the inspiral contributing equally.
 %In \figref{pop-adSNR-vs-Tad}, we plot the maximum overlap between the instantaneous and adiabatic models against this ratio, highlighting the number of resonances, $N\sub{res}$, each system encounters.
@@
 We thank Tanja Hinderer, Jeandrew Brink, Maarten van de Meent, Leor Barack, Scott Hughes and Nico Yunes for useful conversations, and Christopher Moore with help preparing this manuscript.
+We are grateful to the anonymous referee for their exceptional careful reading of the paper.
 RHC was supported by STFC; CPLB thanks STFC and the Cambridge Philosophical Society; PC's work was supported by a Marie Curie Intra-European Fellowship within the 7th European Community Framework Programme (PIEF-GA-2011-299190), and JRG was supported by the Royal Society.
@@
 \partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0), \\
-q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau}.
+q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau},
 \label{eq:q-0-sol}
 \end{align}
-We now have expressions for the lowest-order terms in the expansions.
+where $\kappa_0$ is the constant of integration. We now have expressions for the lowest-order terms in the expansions.
 
@@
 q\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{q}_0\left(\widehat{\lambda}\right) + \eta^{1/2} \widehat{q}_{1/2}\left(\widehat{\lambda}\right) + \order{\eta}, \\
-J\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{J}_0 + \eta^{1/2} \widehat{J}_1\left(\widehat{\lambda}\right) + \order{\eta}.
+J\left(\widehat{\lambda};\,\eta\right) = {} & \widehat{J}_0 + \eta^{1/2} \widehat{J}_{1/2}\left(\widehat{\lambda}\right) + \order{\eta}.
 \end{align}
@@
 \begin{equation}
-\widehat{q}_0 = \widehat{\kappa}_0 + \dfrac{\varpi_1\widehat{\lambda}^2}{2}.
-\end{equation}
-Using these, we can build the next-order terms
+\widehat{q}_0 = \widehat{\kappa}_0 + \dfrac{\varpi_1\widehat{\lambda}^2}{2},
+\end{equation}
+where $\widehat{\varrho}_0$ and $\widehat{\kappa}_0$ are the constants of integration. Using these, we can build the next-order terms
 \begin{align}
@@
 \widehat{J}_{1/2} = {} & \widehat{\varrho}_{1/2} + G_0^{(1)}(\widehat{\varrho}_0)\widehat{\lambda} \nonumber \\
- {} & + \sum_{s\,\neq\,0}G_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau}.
-\end{align}
-Both of these involve the complex Fresnel integral~\cite{Olver2010}, the details of which are given in the following section. We have now constructed the interior solution. % chapter 7
+ {} & + \sum_{s\,\neq\,0}G_s^{(1)}(\widehat{\varrho}_0)\exp(is \widehat{\kappa}_0)\intd{0}{\hat{\lambda}}{\exp\left(\dfrac{is \varpi_1\tau^2}{2}\right)}{\tau},
+\end{align}
+introducing integration constants $\widehat{q}_{1/2}$ and $\widehat{\varrho}_{1/2}$.
+Both the above expressions involve the complex Fresnel integral~\cite{Olver2010}, the details of which are given in the following section.
+We have now constructed the interior solution.
+% chapter 7
 
@@
 \begin{equation}
-\lim_{|z|\,\rightarrow\,\infty} \mathcal{Y}(z) \sim \dfrac{\sgn z}{\sqrt{2}} \exp\left(\dfrac{i\pi}{4}\right) - \dfrac{i}{\pi z}\exp\left(-\dfrac{i\pi z^2}{2}\right),
+\lim_{|z|\,\rightarrow\,\infty} \mathcal{Y}(z) \sim \dfrac{\sgn z}{\sqrt{2}} \exp\left(\dfrac{i\pi}{4}\right) - \dfrac{i}{\pi z}\exp\left(\dfrac{i\pi z^2}{2}\right),
 \end{equation}
@@
 \lim_{|\widehat{\lambda}|\,\rightarrow\,\infty}W\left(\widehat{\lambda}\right) \sim {} & \dfrac{\sgn \widehat{\lambda}}{\sqrt{2}}\sqrt{\dfrac{\pi}{|s\varpi_1|}}\exp\left[\sgn(s\varpi_1)\dfrac{i\pi}{4}\right] \nonumber \\
-  {} & + \recip{is\varpi_1}\exp\left(\dfrac{is \varpi_1 \widehat{\lambda}^2}{2}\right).
+  {} & + \recip{is\varpi_1 \widehat{\lambda}}\exp\left(\dfrac{is \varpi_1 \widehat{\lambda}^2}{2}\right).
 \label{eq:Fres-limit}

### 2016-11-08 plotting/plot_emri_pop.m

@@ -154,7 +154,7 @@ function plot_emri_pop
     ylabel('\Large{}Fraction per bin')
     xlabel('\Large{}Inclination $\cos \iota$')
     set(gca,'YTickLabel',{'$0.00$','$0.01$','$0.02$','$0.03$','$0.04$','$0.05$','$0.06$','$0.07$'})
-    set(gca,'XTickLabel',{'$-1.0$','$-0.5$','$0.0$','$0.5$''$1.0$'})
+    set(gca,'XTickLabel',{'$-1.0$','$-0.5$','$0.0$','$0.5$','$1.0$'})
     laprint(gcf, 'res_pop_iota')
     
     figure;

### 2016-11-10 referee_response.txt
Referee comments

@@
 
-I addition to the changes commented below, we have updated some of the references and added the referee to the acknowledgements.
+In addition to the changes commented below, we have updated some of the references and added the referee to the acknowledgements.
+We have also removed the mention in the Introduction that $\Omega_r < \Omega_\theta < \Omega_\phi$ as Maarten van de Meent has pointed out to us that this hierarchy is not always true; this does not effect anything else in the paper.
 
@@
 
->> Figure 8 shows the orbital parameters at plunge, rather than at resonance...
+>> The values on resonance are p = 7.67, e = 0.674 and cos(iota) = 0.497.
+Figure 8 shows the orbital parameters at plunge, rather than at resonance, so these can't be compared directly.
+This example doesn't reach plunge during it's two-year window, so we think it would be confusing to add these numbers to the text.
+This example is to illustrate what resonances can be like, and the calculation of what they are typically like is left for later on in the paper.
 
@@
 We used a uniform distribution as the least constraining proposal in absence of better information.
-It is true that radiation reaction will lead to anti-alignment, but we are considering the initial distribution of orbits from which inspirals begin, this is the same reason why not all the orbits are circular, despite the circularising effect of radiation reaction.
-Adding a comment on potential alignment is a sensible recommendation, so we have added a footnote to explain this...
+It is true that radiation reaction will lead to anti-alignment, but we are considering the initial distribution of orbits from which inspirals begin; this is the same reason why not all the orbits are circular, despite the circularising effect of radiation reaction.
+Adding a comment on potential alignment is a sensible recommendation, so we have added a footnote to explain this while discussing the evolution of the analytic kludge model.
 
@@
 purpose, it might be commented in the same paragraph in Sec.VI.
+
+>> We think that interpolation formulae could certainly be used if there is sufficient numerical data to calibrate these. We have added a footnote to highlight this as suggested.
 

### 2016-11-10 res_char.tex
Referee comments

@@
 % Section 10.6
-In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field regime $\Omega_r < \Omega_\theta < \Omega_\phi$, and they may differ significantly.
+In the weak-field limit, these all tend towards the Keplerian frequency; in the strong-field regime they may differ significantly.
 For EMRIs, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
@@
 where an overdot denotes a derivative with respect to $t$.
-In general, away from resonance, we take $\nu \equiv \Omega_r/\Omega_\theta$.
+In general, away from resonance, we take $\nu \equiv \Omega_r/\Omega_\theta < 1$.
 This timescale sets the period over which there is a significant change in the frequencies.
@@
 Evolution of the orbit includes the effects of periapse precession, Lense--Thirring
-precession, and radiation reaction calculated using PN expressions.
-AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
+precession, and radiation reaction calculated using PN expressions.\footnote{While the AK approach does include these relativistic effects, it does not capture the full nature of the evolution: for example, it assumes that the angle between the orbital angular momentum and the SMBH spin is constant~\cite{Barack2004}, whereas radiation reaction should push the orbital plane towards being anitaligned~\cite{Flanagan2007}.
+The approximate nature of the AK evolution should not effect our results more than the uncertainty in the initial conditions for EMRI orbits.} AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
 We follow inspirals until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r) = 0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
@@
 Adiabatic waveform models can still be used for EMRIs away from resonance.
-Therefore it may be possible to stitch together waveforms from a sequence of adiabatic evolutions, if the phase on resonance and the magnitude of the self-force can be predicted with sufficient accuracy.
-Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.
+Therefore, it may be possible to stitch together waveforms from a sequence of adiabatic evolutions, if the phase on resonance and the magnitude of the self-force can be predicted with sufficient accuracy.\footnote{This may be possible using interpolation schemes~\cite{Warburton2012} if there is sufficient numerical data available for calibration.} Such hybrid models merit further study as relatively simple ways of incorporating resonance effects into adiabatic models.
 

### 2016-11-21 res_char.tex
Jon's suggestions

@@
 \caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
-The time $t = t\sub{res}$ is when the full evolution crosses the $2$:$3$ resonance.}
+The time $t = t\sub{res}$ is when the full evolution crosses the $2$:$3$ resonance ($p \simeq 7.67$, $e \simeq 0.674$, $\cos(\iota) \simeq 0.497$).}
 \end{figure}
@@
 precession, and radiation reaction calculated using PN expressions.\footnote{While the AK approach does include these relativistic effects, it does not capture the full nature of the evolution: for example, it assumes that the angle between the orbital angular momentum and the SMBH spin is constant~\cite{Barack2004}, whereas radiation reaction should push the orbital plane towards being anitaligned~\cite{Flanagan2007}.
-The approximate nature of the AK evolution should not effect our results more than the uncertainty in the initial conditions for EMRI orbits.} AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
+The approximate nature of the AK evolution should not effect our results more than the uncertainty in the initial conditions for EMRI orbits (for example, how loss-cone dynamics are modified by the SMBH's spin~\cite{Amaro-Seoane2012b}).} AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
 We follow inspirals until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r) = 0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.

### 2016-11-22 referee_response.txt
Jon's suggestions
Typos

@@
 In addition to the changes commented below, we have updated some of the references and added the referee to the acknowledgements.
-We have also removed the mention in the Introduction that $\Omega_r < \Omega_\theta < \Omega_\phi$ as Maarten van de Meent has pointed out to us that this hierarchy is not always true; this does not effect anything else in the paper.
+We have also removed the mention in the Introduction that $\Omega_r < \Omega_\theta < \Omega_\phi$ as Maarten van de Meent has pointed out to us that this hierarchy is not always true; this does not affect anything else in the paper.
 
@@
 
->> The values on resonance are p = 7.67, e = 0.674 and cos(iota) = 0.497.
+>> The values on resonance are p = 7.67, e = 0.674 and cos(iota) = 0.497, and we have added these to the figure caption.
 Figure 8 shows the orbital parameters at plunge, rather than at resonance, so these can't be compared directly.
-This example doesn't reach plunge during it's two-year window, so we think it would be confusing to add these numbers to the text.
+This example doesn't reach plunge during its two-year window, so we think it would be confusing to add these numbers to the text.
 This example is to illustrate what resonances can be like, and the calculation of what they are typically like is left for later on in the paper.

### 2016-11-22 res_char.tex
Typos

@@
 However, the computational cost of numerical-relativity simulations means that other approaches must be used to generate the large number of waveforms required for GW detection and analysis.
-Analytic relativity approaches such as post-Newtonian theory~\cite{Blanchet2014,Buonanno2009}, which can be used to model the early inspiral where the gravitational field is still relatively weak, and the effective-one-body formalism~\cite{Buonanno1999,Buonanno2000,Damour2009,Barausse2010}, which can incorporate merger and ringdown, allow us to generate less expensive waveform approximants.
+Analytic relativity approaches such as post-Newtonian (PN) theory~\cite{Blanchet2014,Buonanno2009}, which can be used to model the early inspiral where the gravitational field is still relatively weak, and the effective-one-body formalism~\cite{Buonanno1999,Buonanno2000,Damour2009,Barausse2010}, which can incorporate merger and ringdown, allow us to generate less expensive waveform approximants.
 These approximants can be calibrated to match numerical relativity results for improved accuracy~\cite{Taracchini2014,Pan2014,Husa2015,Khan2015,Schmidt2015}, and the resulting waveforms allow us to understand comparable-mass binary BHs.
@@
 Geodesic orbits in GR have three associated frequencies: the radial frequency $\Omega_r$, the polar frequency $\Omega_\theta$ and the azimuthal frequency $\Omega_\phi$.\footnote{In the strong-field regime, it is possible to have isofrequency pairings, where two different orbits share the same orbital frequencies~\cite{Warburton2013}.
-The evolution of the frequencies still differ, such that orbital trajectories can be reconstructed.} The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
+The evolution of the frequencies still differ, such that orbital trajectories can be reconstructed from the frequencies.} The first two describe libration and the third rotation (except in the case of polar orbits where $\Omega_\theta$ also describes rotation)~\cite{Goldstein2002}.
 % Section 10.6
@@
 For EMRIs, the evolution timescale is much longer than the orbital period such that the motion of the smaller body is approximately geodesic over orbital timescales.
-The evolution inspiral of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
+The inspiral of the orbit can be approximated as a series of geodesics using the osculating element formalism~\cite{Pound2008,Gair2011a}.
 During this evolution, the frequencies may become commensurate: resonances occur when the radial and polar frequencies are rational multiples of each other:
@@
 One might be concerned about the nature of resonances following the inclusion of the self-force: true geodesic motion only exists at zeroth order in $\eta$ and, while it is a good approximation over short timescales, for small $\eta$ there is a small disparity.
-The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e.
-the case for motion in Kerr) is subject to a small perturbation the form of the orbits is preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
+The conservative piece of the self-force induces extra precession which leads to a slight shift in the orbital frequencies~\cite{Warburton2012}.\footnote{The Kolmogorov--Arnold--Moser (KAM) theorem states that when an integrable Hamiltonian (i.e.\ the case for motion in Kerr) is subject to a small perturbation the form of the orbits is preserved albeit slightly deformed~\cite{Arnold1963,Moser1973}.
 % Chapter II, section 3.3 d)
@@
 To follow the evolution of the inspiral we must have a means of prescribing the forcing acceleration which causes the orbit to deviate from a single geodesic.
-We work directly with the gravitational self-force, using the same post-Newtonian (PN) approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
+We work directly with the gravitational self-force, using the same PN approximation as Flanagan and Hinderer~\cite{Flanagan2012}.
 For comparison, Flanagan, Hughes and Ruangsri~\cite{Flanagan2012a} use a Teukolsky-equation calculation of GW fluxes to account for the inspiral due to radiation reaction.
@@
 We first compute inspiral trajectories and then separately (and not necessarily consistently) calculate the GW emission sourced by a compact object moving along that trajectory.
-This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; agreement between Teukolsky-based and the best NK waveforms are typically $95\%$ or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
+This is quicker and easier to calculate than waveforms using Teukolsky-based methods (currently the most accurate prescription available) and yet gives similar results; agreement between Teukolsky-based and the best NK waveforms is typically $95\%$ or higher for a variety of orbits~\cite{Babak2007, Berry2013}.
 
@@
 
-We expect these NK waveforms to be sufficiently accurate for our purposes, given the approximate (post-Newtonian, low-spin) self-force model (\secref{self-force}).
+We expect these NK waveforms to be sufficiently accurate for our purposes, given the approximate (PN, low-spin) self-force model (\secref{self-force}).
 Results can be straight-forwardly refined as developments are made in computing more comprehensive self-force models.
@@
 We have seen that overlaps between adiabatic and full instantaneous evolutions dephase following a resonance.
-However, this does not necessarily mean that no adiabatic evolution has a low overlap with the observed signal.
+However, this does not necessarily mean that no adiabatic evolution has a high overlap with the observed signal.
 It is possible that a difference between the instantaneous and adiabatic waveforms could be ameliorated by changing the parameters of the template $\boldsymbol{\Theta}$.
@@
 The approximate nature of the AK evolution should not effect our results more than the uncertainty in the initial conditions for EMRI orbits (for example, how loss-cone dynamics are modified by the SMBH's spin~\cite{Amaro-Seoane2012b}).} AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
-We follow inspirals until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r) = 0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r_p$, which designates the orbit as marginally stable.
+We follow inspirals until the last stable orbit (LSO).\footnote{The LSO is determined numerically by calculating the roots of $V_r(r) = 0$, which we denote in ascending order by $r_4 \leq r_3 \leq r\sub{p} \leq r\sub{a}$, and stopping the evolution when $r_3 = r\sub{p}$, which designates the orbit as marginally stable.
 This ignores the (small) influence of the self-force~\cite{Isoyama2014}.} Each of these systems is then evolved backwards for some time $t\sub{insp}$, chosen uniformly from the range $[0,t\sub{life}]$ for a mission lifetime $t\sub{life}$, and the expected GWs are calculated using the AK formalism.
@@
 This might be caused by higher-order resonances disrupting the evolution, in which case $t\sub{ad}$ should be smaller and the systems would lie on the expectation line.
-However, a more likely explanation is that larger a sub-optimal matching time $t\sub{match}$ was used, and a larger overlap is achievable with a different choice of adiabatic evolution.
+However, a more likely explanation is that a sub-optimal matching time $t\sub{match}$ was used, and a larger overlap is achievable with a different choice of adiabatic evolution.
 
@@
 Passing through resonance can lead to an enhancement or decrement in the radiated fluxes associated with orbital evolution, which in turn leads to a jump in the orbital parameters across the resonance.
-Including the signature of resonances is necessary for optimal analysis on EMRI signals.
+Including the signature of resonances is necessary for optimal analysis of EMRI signals.
 
@@
 Brink, Geyer and Hinderer~\cite{Brink2013} provide series expansions for the location of resonances in the limit of equatorial orbits for small spin and eccentricity.
-We do not follow this approach of trying to find analytic expressions for the resonance surface.
-The expressions become complicated when venturing away from limiting cases.
+We do not follow this approach of trying to find analytic expressions for the resonance surface; the expressions become complicated when venturing away from limiting cases.
 Instead, we build an approximate phenomenological model and fit this to the resonance plane.

### 2016-12-15 res_char.tex
Copy edit

@@
 
-\title{The importance of transient resonances in extreme-mass-ratio inspirals}
+\title{Importance of transient resonances in extreme-mass-ratio inspirals}
 
@@
 \author{Jonathan R. Gair}
-\affiliation{School of Mathematics, University of Edinburgh, Peter Guthrie Tait Road, Edinburgh EH9 3FD, UK}
+\affiliation{School of Mathematics, University of Edinburgh, Peter Guthrie Tait Road, Edinburgh EH9 3FD, United Kingdom}
 \affiliation{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
@@
 Including the effects of the smaller body's finite mass, the background spacetime is perturbed.
-The back-reaction from this deformation alters the small body's orbital trajectory, and can be modelled as a self-force that moves the body from its geodesic.
+The backreaction from this deformation alters the small body's orbital trajectory, and can be modelled as a self-force that moves the body from its geodesic.
 The self-force can be divided into two pieces, dissipative and conservative~\cite{Sago2008,Barack2009}.
@@
 The latter shifts the orbital phases inducing precession.
-The dissipative piece is time-asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time-symmetric and has a smaller influence on the phase, although this can accumulate over many orbits.
+The dissipative piece is time asymmetric and has the larger effect on the evolution of the orbital phase; the conservative piece is time symmetric and has a smaller influence on the phase, although this can accumulate over many orbits.
 Being able to accurately model the influence of the self-force allows us to create reliable waveform models.
@@
 It is therefore necessary to develop an understanding of the Kerr geodesics (\secref{geodesic}; those familiar with calculating orbits in Kerr may skip this section.).
-Transient resonance occur when the radial and polar frequencies become commensurate (\secref{frequencies}); we analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance (\secref{forced-motion}).
+Transient resonances occur when the radial and polar frequencies become commensurate (\secref{frequencies}); we analyse the behaviour of resonances within the osculating element framework, where the trajectory is described by a sequence of geodesics that each match onto the motion at a particular instance (\secref{forced-motion}).
 The osculating elements formalism allows for the orbital evolution to be driven by a force, here, a particular model for the self-force (\secref{self-force}) and its adiabatic average (\secref{adiabatic}).
@@
 The geodesic equations for coordinate time $t$ and azimuthal angle $\phi$ are just functions of $r$ and $\theta$, hence their evolutions can be expressed as Fourier series~\cite{Drasco2004}
-\begin{subequations}
+\begin{subequations}\label{eq:Mino-Fourier}
 \begin{align}
@@
 \diff{\phi}{\lambda} = {} & \sum_{k_r,\,k_\theta}\Phi_{k_r,\, k_\theta}\exp\left[-i\left(k_r\Upsilon_r + k_\theta\Upsilon_\theta\right)\lambda\right].
-\label{eq:Mino-Fourier}
 \end{align}
@@
 
-For a general non-resonant orbit there is no fixed correlation between the radial and polar coordinates.
+For a general nonresonant orbit there is no fixed correlation between the radial and polar coordinates.
 After a sufficiently long time, the trajectory comes arbitrarily close to every point in the range of motion (with $r\sub{p} \leq r \leq r\sub{a}$ and $\theta_- \leq \theta \leq \pi - \theta_-$); on account of the orbital precession, the whole space is densely covered.
@@
 Varying $\psi_{\theta_-}$ across its full range allows every point in the range of motion to be reached.
-Hence averaging over all values of $\psi_{\theta_-}$ for resonant orbits is equivalent to averaging over the $\psi$--$\chi$ $2$-torus for non-resonant orbits.
+Hence averaging over all values of $\psi_{\theta_-}$ for resonant orbits is equivalent to averaging over the $\psi$--$\chi$ $2$-torus for nonresonant orbits.
 
@@
 \end{equation}
-However, we are using $\psi$ and $\chi$, as these are simpler to evolve; furthermore, we compute instantaneous co-ordinate time fluxes $\dot{X}$, not Mino-time fluxes.
+However, we are using $\psi$ and $\chi$, as these are simpler to evolve; furthermore, we compute instantaneous coordinate-time fluxes $\dot{X}$, not Mino-time fluxes.
 Changing variables gives an average of \cite{Drasco2005}
@@
 This average describes the Mino-time rate of change of the quantity $X$ over an orbit.
-To convert to a co-ordinate flux of the averaged quantity, we simply divide by the period $\Gamma$~\cite{Flanagan2012a}, defining
+To convert to a coordinate flux of the averaged quantity, we simply divide by the period $\Gamma$~\cite{Flanagan2012a}, defining
 \begin{equation}
@@
 using \eqnref{2torus-average}, as this allows us to eliminate $\Lambda_r$ and $\Lambda_\theta$ from the calculation.\footnote{We compute these integrals using a $300\times 300$ grid of $\{\psi,\chi\}$ values and employing a Newton--Cotes approximation in each dimension.
-The procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time-step of an evolution, and so is computationally expensive to perform.
+The procedure requires $\mathcal{O}(10^5)$ separate evaluations of the derivatives at each time step of an evolution, and so is computationally expensive to perform.
 However, the adiabatic derivatives vary on much longer timescales than the orbital motion (see \secref{res-time}), and so in practice, we can interpolate.}
-%We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\delta t/\eta$ where $\delta t$ is the orbital time-step.
+%We calculate the adiabatic derivatives, as above, at three points in the inspiral spaced by $\delta t/\eta$ where $\delta t$ is the orbital time step.
 When the system is close to the central point, we use a quadratic fit to estimate the adiabatic derivatives.
@@
 The instantaneous $\dot{\Omega}$ depends upon the orbital phase and oscillates about its mean trend over an orbit.
-We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time-average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$.\footnote{On resonance, we are interested in the relative $r$--$\theta$ orbital phase ($\psi_{\theta_-}$ or equivalent), which sets the resonant trajectory in the $r$--$\theta$ plane, but not the exact phase of the orbit around this loop.} Close to resonance, $\Omega(t)$ is well approximated by a first-order Taylor expansion, decreasing linearly with time; hence we make the approximation
+We are interested in the averaged behaviour, not the periodic modulations about this, which is why we use the time average $\langle\dot{\Omega}\rangle_{q'}$; here we use $q'$ to represent a phase that varies over an orbit with period of order $T$.\footnote{On resonance, we are interested in the relative $r$--$\theta$ orbital phase ($\psi_{\theta_-}$ or equivalent), which sets the resonant trajectory in the $r$--$\theta$ plane, but not the exact phase of the orbit around this loop.} Close to resonance, $\Omega(t)$ is well approximated by a first-order Taylor expansion, decreasing linearly with time; hence we make the approximation
 \begin{equation}
@@
 \end{equation}
-Typically, the first term is non-zero and this gives the familiar oscillation.
+Typically, the first term is nonzero and this gives the familiar oscillation.
 On resonance, it is zero, leaving the next-order term to govern the behaviour~\cite{Flanagan2012,Ruangsri2014}.
-Only once we have moved far enough away from resonance for the first term to dominate the second do we recapture the non-resonant behaviour.
+Only once we have moved far enough away from resonance for the first term to dominate the second do we recapture the nonresonant behaviour.
 The first term (translating from Mino time to coordinate time) sets $\tau\sub{pres}$, the second sets $\tau\sub{res}$.
@@
 \end{align}
-This assumes that $s$ is a non-zero integer.
+This assumes that $s$ is a nonzero integer.
 
@@
 We expect these NK waveforms to be sufficiently accurate for our purposes, given the approximate (PN, low-spin) self-force model (\secref{self-force}).
-Results can be straight-forwardly refined as developments are made in computing more comprehensive self-force models.
+Results can be straightforwardly refined as developments are made in computing more comprehensive self-force models.
 
@@
 \end{equation}
-which ranges from $0$ to $1$ provides a convenient indication of how well-matched the template is to the data.
-
-\subsection{The non-resonant case}
+which ranges from $0$ to $1$ provides a convenient indication of how well matched the template is to the data.
+
+\subsection{The nonresonant case}
 \label{sec:nonres}
@@
 To illustrate the dephasing, we calculate as a function of time $t$ the shortened overlap between the two models, defined as the overlap obtained by including only the part of the waveform within $\Delta t$ of $t$.
-We choose $\Delta t$ such that we can calculate $25$ non-overlapping shortened overlaps.
+We choose $\Delta t$ such that we can calculate $25$ nonoverlapping shortened overlaps.
 Before the resonance occurs, the adiabatic model provides a good match to the full evolution, but the overlap is reduced near to the resonance and never fully recovers afterwards.
@@
 \includegraphics[width=0.4\textwidth]{Fig_overlap}
-\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (non-overlapping) calculations.
+\caption{\label{fig:overlap-dephasing}The overlap computed between the full evolution and an adiabatic evolution for an illustrative EMRI system with $p=7.85$, as a function of time, including only the parts of the waveforms within some $\Delta t$ of $t$, arbitrarily chosen to give $25$ independent (nonoverlapping) calculations.
 The time $t = t\sub{res}$ is when the full evolution crosses the $2$:$3$ resonance ($p \simeq 7.67$, $e \simeq 0.674$, $\cos(\iota) \simeq 0.497$).}
@@
 %We can study the predicted orbital-parameter jumps numerically by considering the long-time-average fluxes for particular (fixed) geodesics.
-For non-resonant geodesics, this will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference.
+For nonresonant geodesics, this will depend only on the orbital shape parameters, but for resonant orbits there will be an additional dependence on the initial phase difference.
 We consider a selection of different orbital shape parameters corresponding to low-order resonances and calculate the time-averaged fluxes for $50$ different initial phases.
@@
 \includegraphics[width=0.4\textwidth]{Fig_q_jump}
-\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$ (cf.\ figure 3 of \cite{Flanagan2012a}).
+\caption{\label{fig:resjump-vs-q}The magnitude of the resonance jump for our illustrative system as a function of the extracted phase parameter $q$ (cf.\ Fig.~3 of \cite{Flanagan2012a}).
 The jump is relative to the adiabatic change in each parameter across resonance.
@@
 Let us now consider jumps for systems other than our illustrative $2$:$3$ resonance.
-We expect nearly-circular orbits to encounter smaller jumps than more eccentric orbits~\cite{Flanagan2012} because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a non-resonant orbit.
+We expect nearly circular orbits to encounter smaller jumps than more eccentric orbits~\cite{Flanagan2012} because they have a smaller $r$--$\theta$ $2$-torus, meaning that resonant orbits come closer to every allowed point, nearer approximating a nonresonant orbit.
 In \figref{res-flux-rms-p}, we show the root-mean-square (over a grid of $a_\ast$ and $\cos\iota$ values) relative resonant jump for the semilatus rectum $p$ as a function of eccentricity $e$ for a selection of low-order resonances.
@@
 
-Combining the intrinsic event rate with the comoving number density, the mass of the SMBH for the EMRI population is then sampled from a power-law with a probability distribution function $f(M) \propto M^{\alpha+\beta-1}$.
+Combining the intrinsic event rate with the comoving number density, the mass of the SMBH for the EMRI population is then sampled from a power law with a probability distribution function $f(M) \propto M^{\alpha+\beta-1}$.
 
@@
 We assume this can be extended to provide a rough estimate of the distribution around SMBHs of other masses and spins.
-At the point in the inspiral when the orbital period takes a fiducial value $T_0 = 10^4~\mathrm{s}$, we find that the Monte-Carlo eccentricity probability distribution function is well-described by a power-law with an exponential cutoff
+At the point in the inspiral when the orbital period takes a fiducial value $T_0 = 10^4~\mathrm{s}$, we find that the Monte-Carlo eccentricity probability distribution function is well described by a power law with an exponential cutoff
 %\begin{widetext}
@@
 Evolution of the orbit includes the effects of periapse precession, Lense--Thirring
-precession, and radiation reaction calculated using PN expressions.\footnote{While the AK approach does include these relativistic effects, it does not capture the full nature of the evolution: for example, it assumes that the angle between the orbital angular momentum and the SMBH spin is constant~\cite{Barack2004}, whereas radiation reaction should push the orbital plane towards being anitaligned~\cite{Flanagan2007}.
+precession, and radiation reaction calculated using PN expressions.\footnote{While the AK approach does include these relativistic effects, it does not capture the full nature of the evolution: for example, it assumes that the angle between the orbital angular momentum and the SMBH spin is constant~\cite{Barack2004}, whereas radiation reaction should push the orbital plane towards being antialigned~\cite{Flanagan2007}.
 The approximate nature of the AK evolution should not effect our results more than the uncertainty in the initial conditions for EMRI orbits (for example, how loss-cone dynamics are modified by the SMBH's spin~\cite{Amaro-Seoane2012b}).} AK waveforms are less computationally expensive than NK waveforms, allowing us to simulate a large population of EMRIs.
@@
 This might be caused by higher-order resonances disrupting the evolution, in which case $t\sub{ad}$ should be smaller and the systems would lie on the expectation line.
-However, a more likely explanation is that a sub-optimal matching time $t\sub{match}$ was used, and a larger overlap is achievable with a different choice of adiabatic evolution.
+However, a more likely explanation is that a suboptimal matching time $t\sub{match}$ was used, and a larger overlap is achievable with a different choice of adiabatic evolution.
 
@@
 The magnitude of the jump depends upon the order of resonance and the orbital eccentricity.
-The $2$:$3$ and $1$:$2$ resonances have the largest effects, and higher order resonance are less important.
+The $2$:$3$ and $1$:$2$ resonances have the largest effects, and higher-order resonances are less important.
 Crucially, jumps are smaller for lower eccentricity orbits.
@@
 
-Using this approximation, the maximum error in $p$ for a given $a_\ast$ and $\nu$ is typically $\sim10\%$ and less that $1$ in absolute terms.
+Using this approximation, the maximum error in $p$ for a given $a_\ast$ and $\nu$ is typically $\sim10\%$ and less than $1$ in absolute terms.
 The relative error in the semilatus rectum is illustrated in \figref{p-error}.
@@
 \begin{align}
-\partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0), \\
+\partialdiff{q_0}{\widetilde{\lambda}} = {} & g_0^{(1)}(J_0) \\
 q_0 = {} & \kappa_0 + \intd{0}{\tilde{\lambda}}{g_0^{(1)}[J_0(\tau)]}{\tau},
@@
 \end{equation}
-Using this together with equations (\ref{eq:q-0-sol}) and (\ref{eq:q-1-sol}) in \eqnref{q-series}, the angle variable is
+Using this together with Eqs~(\ref{eq:q-0-sol}) and (\ref{eq:q-1-sol}) in \eqnref{q-series}, the angle variable is
 \begin{widetext}
@@
 where we have defined $J_\star \equiv J_0\left(\widetilde{\lambda}_\star\right)$ and $\kappa_\star = \kappa_0 + \intd{0}{\tilde{\lambda}_\star}{g_0^{(1)}[J_0(\tau)]}{\tau}$, and used \eqnref{omega-hat} to substitute for $\omega$.
-The action variable is similarly determined by using equations (\ref{eq:J-ad}) and (\ref{eq:J-1-sol}) with \eqnref{J-series} to give
+The action variable is similarly determined by using Eqs~(\ref{eq:J-ad}) and (\ref{eq:J-1-sol}) with \eqnref{J-series} to give
 \begin{align}
@@
 \end{align}
-Similarly, for the action variables
+Similarly, for the action variable
 \begin{equation}
@@
 \end{equation}
-We have used the averaged form of $\dot{\Omega}(t)$ as this is appropriate. Using these to adapt equations (\ref{eq:J-1/2}) and (\ref{eq:jumps}), we obtain
+We have used the averaged form of $\dot{\Omega}(t)$ as this is appropriate. Using these to adapt Eqs~(\ref{eq:J-1/2}) and (\ref{eq:jumps}), we obtain
 \begin{align}

### 2016-12-15 Proceedings/Berry_LISA11.tex
First LISA Symposium proceedings draft

\documentclass[a4paper]{jpconf}
\usepackage{amssymb,amsfonts,amsmath}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{hyperref}
\bibliographystyle{iopart-num}
\usepackage[numbers,square,sort&compress]{natbib}

\DeclareMathOperator{\sgn}{sgn}

\begin{document}
\title{Understanding the importance of transient resonances in extreme mass ratio inspirals}

\author{C~P~L~Berry$^{1,2}$, R~H~Cole$^{2}$, P~Ca\~{n}izares$^{3,2}$ and J~R~Gair$^{4,2}$}

\address{$^1$ School of Physics \& Astronomy, University of Birmingham, Birmingham, B15 2TT, UK}
\address{$^2$ Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, UK}
\address{$^3$ Institute of Mathematics, Astrophysics and Particle Physics, Radboud University, Heyendaalseweg 135, 6525 AJ Nijmegen, Netherlands}
\address{$^4$ School of Mathematics, University of Edinburgh, Peter Guthrie Tait Road, Edinburgh EH9 3FD, UK}

\ead{cplb@star.sr.bham.ac.uk}

\begin{abstract}
Extreme mass ratio inspirals (EMRIs) occur when a compact object orbits a much larger one, like a solar-mass black hole around a massive black hole. The orbit has $3$ frequencies which evolve through the inspiral. If the orbital radial frequency and polar frequency become commensurate, the system passes through a \emph{transient resonance}. Evolving through resonance causes a jump in the evolution of the orbital parameters. We study these jumps and their impact on EMRI gravitational-wave detection. Jumps are smaller for lower eccentricity orbits; since most EMRIs have small eccentricities when passing through resonances, we expect that the impact of detection will be small. Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals for an astrophysically motivated population of EMRIs.
\end{abstract}

%\section{Extreme mass ratio inspirals and transient resonances}

The first observations of gravitational waves came from comparable mass stellar-mass binary black holes~\cite{Abbott2016d}. The evolving Laser Interferometer Space Antenna (eLISA) provides the chance to observe gravitational waves from stellar-mass black holes orbiting supermassive black holes~\cite{Amaro-Seoane2007}. These extreme mass ratio inspirals (EMRIs) emit $\sim10^4$--$10^5$ gravitational-wave cycles in eLISA's frequency band, allowing exquisite measurements to be made~\cite{Barack2004}. To detect and analyse EMRIs accurate waveform templates are required.

Flanagan and Hinderer~\cite{Flanagan2012} highlighted a previously overlooked phenomenon that complicates EMRI waveform modelling, that of transient resonances. During an inspiral, a resonance occurs when the radial frequency $\Omega_r$ and polar frequency $\Omega_\theta$ are a rational ratio of each other. Then, the orbit does not cover the whole allowed $r$--$\theta$ plane, but cycles over a single loop~\cite{Grossman2012}. The orbital phase can be expressed as
\begin{equation}
\varphi {} \sim {} \left(n_r \Omega_r - n_\theta \Omega_\theta\right) t + \left(n_r \dot{\Omega}_r - n_\theta \dot{\Omega}_\theta\right)t^2 + \ldots = \Omega t + 2\pi \frac{t^2}{\tau_\mathrm{res}^2} + \ldots
\end{equation}
On resonance $\Omega = 0$ so the first term, which normally dominates, goes to zero; the second term then governs dynamics for a duration set by the resonance time $\tau_\mathrm{res}$~\cite{Ruangsri2014,Berry2016a}. The evolution of the inspiral is determined by the gravitational self-force~\cite{Barack2009}, across a resonance terms that usually average to zero can combine coherently, significantly impacting the orbital motion~\cite{Flanagan2012}.

For previous EMRI studies, adiabatic waveforms have been used. These average the self-force over the $r$--$\theta$ plane (or equivalently the $r$--$\theta$ phase) and do not include resonances. Adiabatic waveforms quickly dephase from true EMRI waveforms if passing through resonance causes a jump in the orbital parameters. The mismatch between adiabatic and the true waveforms could prevent signals from being detected or lead to biased parameter estimates.

In \cite{Berry2016a}, we investigated jumps in the orbital parameters from passing through a resonance. We used the same approximate self-force model as Flanagan and Hinderer~\cite{Flanagan2012}. The jumps depend upon the magnitude of the self-force, the resonance time and the relative $r$--$\theta$ phase~\cite{Kevorkian1987,VanDeMeent2013,Berry2016a}. Jumps are most significant for low-order resonances, like the $2$:$3$ resonance, because higher-order resonances come closer the covering the entire $r$--$\theta$ like nonresonant orbits. The jump magnitude is sensitive to the orbital parameters on resonance, and decreases for lower eccentricity or lower inclination orbits~\cite{Flanagan2012a}: there can be no resonances for circular orbits.

%\section{EMRI population}\label{sec:sky}

To assess the impact of resonances on detectability, we studied an astrophysical population of EMRIs that could be detectable with eLISA, assuming a two-year mission~\cite{Amaro-Seoane2012a}. We considered $10 M_\odot$ black holes inspiralling into $10^4$--$10^7 M_\odot$ massive black holes. The initial eccentricity distribution followed Hopman and Alexander~\cite{Hopman2005}. As gravitational-wave emission circularises orbits~\cite{Peters1964}, we found that when important low-order resonances are encountered, the eccentricity is small and so are the resonant jumps. The match with adiabatic waveforms are good enough to still detect EMRIs in most cases~\cite{Berry2016a}. Considering the distribution of signal-to-noise ratios $\rho$ for the EMRIs before accounting for resonances and after including the reduction from imperfect waveforms, we find that $\sim4\%$ of signals drop below the assumed detection threshold $\rho_\mathrm{thres} = 15$.

%\begin{figure}
%  \centering
%   \subfigure[\label{fig:eccentricity}]{\includegraphics[width=0.45\columnwidth]{Fig_e_rms_p}} \quad
%   \subfigure[\label{fig:SNR}]{\includegraphics[width=0.45\columnwidth]{Fig_res_pop_SNR_overlap}}
%    \caption{. Figures adapted from \cite{Berry2016a}.} 
%    \label{fig:both}
%\end{figure}

%\section{Summary}

In summary, transient resonances are a generic feature of EMRIs. Passing through resonance causes a jump in the evolution: the jump depends upon the phase on resonance, its size depends upon the orbit and the resonance. Ignoring resonances leads to waveform mismatch, but jumps should be small for most EMRIs because of their eccentricity. Therefore, detectability is not significantly reduced. The effect of resonances on parameter estimation is still to be investigated.

\ack
This work was supported in part by STFC and the Cambridge Philosophical Society. This is LIGO document reference LIGO-P1600349.

\bibliography{Resonances.bib}
%\providecommand{\newblock}{}
%\begin{thebibliography}{9}
%\small
%\expandafter\ifx\csname url\endcsname\relax
%  \def\url#1{{\tt #1}}\fi
%\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
%\providecommand{\eprint}[2][]{\href{http://arxiv.org/abs/#2}{#2}}
%% Bibliography created with iopart-num v2.0
% /biblio/bibtex/contrib/iopart-num

%\end{thebibliography}

\end{document}

### 2017-01-12 Proceedings/Berry_LISA11.tex
Proceedings proof read

@@
 
-\address{$^1$ School of Physics \& Astronomy, University of Birmingham, Birmingham, B15 2TT, UK}
+\address{$^1$ School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, UK}
 \address{$^2$ Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, UK}
@@
 \begin{abstract}
-Extreme mass ratio inspirals (EMRIs) occur when a compact object orbits a much larger one, like a solar-mass black hole around a massive black hole.
+Extreme mass ratio inspirals (EMRIs) occur when a compact object orbits a much larger one, like a solar-mass black hole around a supermassive black hole.
 The orbit has $3$ frequencies which evolve through the inspiral.
@@
 We study these jumps and their impact on EMRI gravitational-wave detection.
-Jumps are smaller for lower eccentricity orbits; since most EMRIs have small eccentricities when passing through resonances, we expect that the impact of detection will be small.
-Neglecting the effects of transient resonances leads to a loss of $4\%$ of detectable signals for an astrophysically motivated population of EMRIs.
+Jumps are smaller for lower eccentricity orbits; since most EMRIs have small eccentricities when passing through resonances, we expect that the impact on detection will be small.
+Neglecting the effects of transient resonances leads to a loss of $\sim4\%$ of detectable signals for an astrophysically motivated population of EMRIs.
 \end{abstract}
@@
 For previous EMRI studies, adiabatic waveforms have been used.
-These average the self-force over the $r$--$\theta$ plane (or equivalently the $r$--$\theta$ phase) and do not include resonances.
+These average the self-force over the $r$--$\theta$ plane and do not include resonances.
 Adiabatic waveforms quickly dephase from true EMRI waveforms if passing through resonance causes a jump in the orbital parameters.
@@
 In \cite{Berry2016a}, we investigated jumps in the orbital parameters from passing through a resonance.
-We used the same approximate self-force model as Flanagan and Hinderer~\cite{Flanagan2012}.
-The jumps depend upon the magnitude of the self-force, the resonance time and the relative $r$--$\theta$ phase~\cite{Kevorkian1987,VanDeMeent2013,Berry2016a}.
-Jumps are most significant for low-order resonances, like the $2$:$3$ resonance, because higher-order resonances come closer the covering the entire $r$--$\theta$ like nonresonant orbits.
+We used the same approximate self-force as Flanagan and Hinderer~\cite{Flanagan2012}.
+The size of jumps depends upon the magnitude of the self-force, the resonance time and the relative $r$--$\theta$ phase~\cite{Kevorkian1987,VanDeMeent2013,Berry2016a}.
+Jumps are most significant for low-order resonances, like the $2$:$3$ resonance, because higher-order resonances come closer the covering the entire $r$--$\theta$ plane (like nonresonant orbits).
 The jump magnitude is sensitive to the orbital parameters on resonance, and decreases for lower eccentricity or lower inclination orbits~\cite{Flanagan2012a}: there can be no resonances for circular orbits.
@@
 To assess the impact of resonances on detectability, we studied an astrophysical population of EMRIs that could be detectable with eLISA, assuming a two-year mission~\cite{Amaro-Seoane2012a}.
-We considered $10 M_\odot$ black holes inspiralling into $10^4$--$10^7 M_\odot$ massive black holes.
+We considered $10 M_\odot$ black holes inspiralling into $10^4$--$10^7 M_\odot$ supermassive black holes with a uniform distribution of spins.
 The initial eccentricity distribution followed Hopman and Alexander~\cite{Hopman2005}.

### 2017-01-16 Proceedings/Berry_LISA11.tex
PnP comments

@@
 On resonance $\Omega = 0$ so the first term, which normally dominates, goes to zero; the second term then governs dynamics for a duration set by the resonance time $\tau_\mathrm{res}$~\cite{Ruangsri2014,Berry2016a}.
-The evolution of the inspiral is determined by the gravitational self-force~\cite{Barack2009}, across a resonance terms that usually average to zero can combine coherently, significantly impacting the orbital motion~\cite{Flanagan2012}.
+The evolution of the inspiral is determined by the gravitational self-force~\cite{Barack2009}; across a resonance, terms that usually average to zero can combine coherently, significantly impacting the orbital motion~\cite{Flanagan2012}.
 
@@
 The initial eccentricity distribution followed Hopman and Alexander~\cite{Hopman2005}.
-As gravitational-wave emission circularises orbits~\cite{Peters1964}, we found that when important low-order resonances are encountered, the eccentricity is small and so are the resonant jumps.
+As gravitational-wave emission circularises orbits~\cite{Peters1964}, we found that when important low-order resonances are encountered the eccentricity is small.
+Accordingly, the resonant jumps are also small.
 The match with adiabatic waveforms are good enough to still detect EMRIs in most cases~\cite{Berry2016a}.

### 2017-01-16 assemble-things.sh

@@ -10,25 +10,46 @@ cp res_char.bbl ./submit/
 
 cp Resonances.bib ./submit/
 
-cp Fig_good_traj.* ./submit/
-cp Fig_overlap.* ./submit/
-cp Fig_res_diff_traj.* ./submit/
-cp Fig_res_jump_calc.* ./submit/
-cp Fig_q_jump.* ./submit/
-cp Fig_e_rms_p.* ./submit/
-cp Fig_nu_rms_p.* ./submit/
-cp Fig_res_pop_M.* ./submit/
-cp Fig_res_pop_a.* ./submit/
-cp Fig_res_pop_z.* ./submit/
-cp Fig_res_pop_p.* ./submit/
-cp Fig_res_pop_e.* ./submit/
-cp Fig_res_pop_iota.* ./submit/
-cp Fig_res_pop_t.* ./submit/
-cp Fig_res_pop_SNR.* ./submit/
-cp Fig_res_overlap_t.* ./submit/
-cp Fig_res_pop_SNR_overlap.* ./submit/
-cp Fig_res-2-5-95-plane.* ./submit/
-cp Fig_fit-error-max-plane.* ./submit/
-cp Fig_fit-error-RMS-plane.* ./submit/
+#cp Fig_good_traj.pdf ./submit/
+#cp Fig_overlap.pdf ./submit/
+#cp Fig_res_diff_traj.pdf ./submit/
+#cp Fig_res_jump_calc.pdf ./submit/
+#cp Fig_q_jump.pdf ./submit/
+#cp Fig_e_rms_p.pdf ./submit/
+#cp Fig_nu_rms_p.pdf ./submit/
+#cp Fig_res_pop_M.pdf ./submit/
+#cp Fig_res_pop_a.pdf ./submit/
+#cp Fig_res_pop_z.pdf ./submit/
+#cp Fig_res_pop_p.pdf ./submit/
+#cp Fig_res_pop_e.pdf ./submit/
+#cp Fig_res_pop_iota.pdf ./submit/
+#cp Fig_res_pop_t.pdf ./submit/
+#cp Fig_res_pop_SNR.pdf ./submit/
+#cp Fig_res_overlap_t.pdf ./submit/
+#cp Fig_res_pop_SNR_overlap.pdf ./submit/
+#cp Fig_res-2-5-95-plane.pdf ./submit/
+#cp Fig_fit-error-max-plane.pdf ./submit/
[…]
